### 2020-05-07 analysis/PRT-Appendix-1.Rmd
Draft PRT template #9

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: ELise Gould 
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
library(here)
memos <- readr::read_csv(here("./data/PRT-WorkshopCoding - Memo Manager.csv"))%>% 
  select(Name, Groups, Comment) %>% 
  mutate(phase = substr(Name, 1,1) %>%  as.numeric(),
         step = substr(Name, 3,3) %>% as.numeric())

#@TODO Alternative 2: Make new Groups in Atlas that match with the Phase and step names

#@TODO: for appendix, we want justifications and also quotations of sources.
#For the latter, we can filter out quotes that come from documents in the PDF group only, and then join those in.

codes <- readr::read_csv(here("data/PRT-WorkshopCoding - Code Manager.csv")) %>%  
  select(Name, Comment, Groups) %>% 
  mutate(phase = substr(Name, 1,1) %>%  as.numeric(),
         step = ifelse(str_detect(Groups, "step"), substr(Name, 3,3),"0") %>% as.numeric())

PRT_items <- bind_rows(codes, memos) %>% select(phase, step, Name, Comment) %>% arrange(phase, step)

knitr::kable(PRT_items) #@TODO how to nest

PRT_items <- 
  PRT_items %>% mutate(response = "")

```

```{r write-draft-template, echo = FALSE, eval = FALSE}
write.csv(PRT_items, file = here("./outputs/draft_template.csv"))
```

### 2020-05-10 analysis/preregistration_template.Rmd
Close #9 Complete first draft template

---
title: "Ecology and Conservation Modelling Preregistration Template -- DRAFT"
author: "Elise Gould, Add your name here if you edit or make comments on this document, TBD:workshop collaborators"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document
bibliography: "../data/references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Problem Formulation

*Specifies the decision-making context in which the model will be used, the clients driving model development or stakeholders addressed by the model. It also includes specification of model outputs, statements of the domain of applicability of the model, as well as the extent by which model outputs may be acceptably extrapolated (Grimm et al. 2014).*

## 1.1 Model Purpose

*Defining the purpose of the model is critical because the model purpose influences choices at later stages of model development [@Jakeman:2006ii]. Common model purposes in ecology include: gaining a better qualitative understanding of the target system, knowledge synthesis and review, and providing guidance for management and decision-making [@Jakeman:2006ii] (could also see Addison paper on the use of models). Note that modelling objectives are distinct from the analytical objectives of the model.*

### 1.1.1 Define model purpose and problem context

*Provide a clear statement of the modelling objectives and problem that the model seeks to illuminate. What is the purpose of the model/s? Ensure that you specify any focal taxa and study objectives, as well as any clients for whom the model is developed for. Briefly outline the ecological problem and the decision-problem, including the decision-trigger and any regulatory frameworks relevant to the problem.*

## 1.2 Specify modelling context

*The scope of the model, including temporal and spatial resolutions are defined here [@Mahmoud2009], and any limitations on model development analysis and flexibility should be outlined here [@Jakeman:2006ii]. Note that the modelling context is different from the problem context.*

### 1.2.1 Identify model interest group

*Identify interest groups in the modelling context. This includes clients, and end-users of the model. Who is the model for, who is involved in formulating the model? Who needs buy in? Describe the decision-making context in which the model will be used.*

### 1.2.2 Model Scope and Scale

*Determine the temporal and spatial scope of the model. Where is the boundary of the modelled system? Everything beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al. 2006).The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately? (Should this last step go here or within define conceptual framework?).What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)? This is relevant to model transferability... (similar to a COG statement..)*

### 1.2.3 Logistical Constraints

*Time-frame - When must the model by completed, e.g. to help make a decision?*

*What effort and resources are available for both developing the model and operating the model?*

*What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?*

## 1.3.1 Explain analytical objectives

*How will the model be analysed, what analytical questions will the model be used to answer? Examples from ecological decision-making include: to compare the performance of alternative management actions under budget constraint [@Fraser:2017jf] to search for robust decisions under uncertainty [@McDonald-Madden2008], to choose the conservation policy that minimises uncertainty [insert ref]. See other examples in Moallemi et al. 2019 & ask Libby and others.*

### 1.3.2 - Identify outcomes to evaluate decisions under scenarios

*The outcomes should speak directly to the analytical objectives identified in 1.3.1. Outcomes could be qualitative or quantitative, and include descriptive statistic measures, satisficing measures or regret measures (Moallemi et al. 2019). For example, a model outcome in the top 10th percentile of worst conditions, or some minimum performance threshold for meeting a performance objective, or choosing a decision with the minimum regret or least amount of uncertainty.*

## 1.4 Define candidate actions decisions

*Candidate decisions should be investigated and are specified a priori. Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019]. Describe the method used to identify relevant management actions.*

## 1.5 Specify scenarios

*Describe what processes you will use to elicit and identify relevant scenarios, e.g. literature review, structured workshops with stakeholders or decision-makers. Specify scenarios under which decisions are investigated. Scenarios should be set a priori (i.e. before the model is built, [@Moallemi2019]) and may be stakeholder-defined or expert judgment-driven [@Mahmoud2009].*

# 2.0	Define Conceptual Model

*Conceptual models underpin the formal or quantitative model (Cartwright et al. 2016). The conceptual model describes the biological mechanisms relevant to the ecological problem and should capture basic premises about how the target system works, including any prior knowledge and assumptions about system processes. Conceptual models may be representeed in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables (Jakeman, Letcher, and Norton, 2006).*

## 2.1 Choose elicitation and representation method

*Describe what method you will use to elicit or identify the conceptual model. Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks (Moon et al. 2019). (Libby, to provide link to any structured expert elicitation methods?).While it is difficult to decide and justify which method is most appropriate, however Moon et al. (2019) provide guidance addressing this methodological question. Finally, how do you intend on representing the final conceptual model? This will likely depend on the method chosen to elicit the conceptual model.*

## 2.2 Specify key assumptions uncertainties

*This step should list and explain the critical conceptual design decisions, including: spatial and temporal scales, selection of entities and processes, representation of stochasticity and heterogeneity, consideration of local versus global interactions, environmental drivers, etc. [@Grimm:2014es]. The influence of particular theories, concepts, or importantly, earlier models, should be explained and justified against alternative conceptual design decisions that might lead to alternative model structures [@Grimm:2014es]. Specify key assumptions and uncertainties underlying the model's design. Describe how uncertainty and variation will be represented in this model. This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].*

##2.3-define_conceptual_model: identify predictor response variables

*The identification and definition of primary model input variables should be driven by scenario definitions, and by the scope of the model described in the problem formulation phase [@Mahmoud2009] -- Identify and define system variables structures:*

- *What variables will support this decision or action (things we can control).* 
- *What additional variables may interact with this system (things we can't control, but can hopefully measure).* 
- *In what ways do we expect these variables to interact (model structures).*

## 2.4 Define prior knowledge, data specification and evaluation

*Collect, process and prepare data available for parameterisation, determining model structure, and for scenario analysis.*

### 2.4.1 Collate available data sources that could be used to parameterise or structure the model

*Document the identity, quantity and provenance of any data used to develop, identify and test the model. Describe how the data is arranged, in terms of replication and covariates. Explain how you will summarise what data and knowledge will be used to designa and parameterise the model, including references, data sources, and information about whom, when and where those data were collected [@Grimm:2014es].*

### 2.4.2 Data Processing and Preparation

*Describe any data preparation and processing steps, including manipulation of environmental layers, e.g. standardisatino and geographic projection.*

### 2.4.3 Describe any data exploration or preliminary data analyses.

*In most modelling cases, it is necessary to perform preliminary analyses to understand the data and check that assumptions and requirements of the chosen modelling procedures are met. Data exploration prior to model fitting or development may include exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers, or the need for transformations [@Yates2018]. Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.*

### 2.4.4 Data Evaluation

*Describe how you will evaluate the quality and sources of both numerical and qualitative data that available for model parameterisation, and for determining the overall model structure [@Grimm:2014es]. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009]. Document any issues with data reliability. This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].*

## 2.5 Conceptual model evaluation

*Describe how the model will be critically evaluated. Evaluation includes both the overall model structure, and any submodels if relevant. How will any simplifying assumptiosn be critically assessed[@Augusiak:2014gz]? Explain whether this process will include consultation or feedback from a client, manager, or model user.*

# 3.0	Formalise and Specify Model

*Critical decisions are made here about the type of model, modelling framework and approach to be used. The model is formalised into a mathematical / statistical description of the system and respective changes it can experience  [@Mahmoud2009]. In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.*

## 3.1 Choose model class, framework and approach

*Describe what class or approach of model you will use and explain how the choice of model class was informed by the analytical objectives of the model. Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based  [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].*

##  3.2 Choose model features and family

*All modelling approaches require the selection of model features, which conform with the conceptual model and data specified in previous steps (Jakeman 2006). The choice of model are determined in conjunction with features are selected.Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006). However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.*

### 3.2.1 Explain how you will operationalise response variable(s)

*Specify how you will operationalise response variables in the model. This should relate directly to the analytical and or management objectives specified during the problem formulation phase. Specifications could include:- qualitative change, such as a direction of change- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, or- the frequency, location, or probability of some event occuring.Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).*

### 3.2.2 Choose model family

*Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions. Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase. Include in your rationalse for selection detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*

### 3.2.3 Choose model features

*Specify which types of variables are covered in the model, and the nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006). Specify model structural features, including:- the functional form of interactions,- data structures- measures used to specify links, spatio temporal scales and processes as well as their interactions- any bins or discretisation of continuous variables.*

## 3.3 Choose approach for identifying model structure and parameters

*This refers to the iterative process of determining the most efficient or parsimonious representation of the system at the appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives specified in the problem formulation phase. Approaches to finding model structure and parameters may be knowledge-supported, or data-driven [@Boets:2015gl]. Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods (Barnard et al. 2019). Ensemble modelling procedures might aim to derive a single model, or a multi-model average [@Yates2018]. Refining actions to develop a model could include iteratively dropping parameters or adding them, or aggregating / disaggregating system descriptors, such as dimensionality and processes (Jakeman, 2006). Specify what approach and methods you will use to identify model structure and parameters.*

## 3.4 Choose estimation technique and performance criteria

*Before calibrating the model to the data, the performance criteria on which the calibration is judged are chosen. These criteria and their underlying assumptions should reflect the desired properties of the parameter estimates / structure [@Jakeman:2006ii]. For example, modellers might seek that parameter estimates are robust to outliers, unbiased, and yield appropriate predictive performance. Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are appropriate to the problem at hand. For integrated or sub-divided models, other considerations might include choices about where to disaggregate the model for parameter estimation; e.g. spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii]. Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.This is important - because it involves pre-specifying how we will interpret the fitted or quantitative model to avoid cognitive biases like confirmation bias, HARKing or other questionable research practices.*

### 3.4.1 Choice of estimation performance criteria

*Specify which suite of performance criteria you will use to judge the performance of the model. Examples include correlation scores, coefficient of determination, specificity, sensitivity, AUC, etcetera (Yates et al. 2018). Relate any underlying assumptions of each criterion to the desired properties of the model, and justify the choice of performance metric in relation into whether it is sensitive to the problem at hand. Explain how you will identify which model features or components are significant or meaningful.*

### 3.4.2 Choice of structure and parameter estimation technique

*Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fi of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006). Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge.Specify which algorithm(s) you will use for any data-driven structure / parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*

## 3.5 Specify model assumptions and uncertainties

*Specify all assumptions and key uncertainties in the formal model. Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpertation of the model outputs, and what implications are there on evaluating model-output to inform decisions?*

## 3.6 Specify formal model

*Once critical decisions have been made about the approach and method of model specification, translate the conceptual model into the quantitative model. For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters. (Should this go here or in model calibration?)*

# 4.0 Model Calibration, Model Fitting & Checking

*This phase involves fitting the formally defined model to data and checking and testing the fitted model (what should go in thisphase if using a numerical or mathematical, but not correlative model or how could we change the wording to make this optional if not relevant?)*

## 4.1 Model calibration and validation scheme

*Describe the validation scheme you will use for model testing and evaluation.  Please explain your reasoning for your choice of model calibration and validation scheme. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;[@Barnard2019]. Although using independent data is the gold-standard or idealised practice for model evaluation, Typically testing on independent datasets is not possible for large and/or integrated models, especially when they are being used to generate anticipatory predictions beyond the conditions on which they were calibrated [@Jakeman:2006ii]. Describe and justify the method for your calibration and validation schema, including how many folds and what methods you will use to partition the data set into training/testing data. (What other items, esp. re. partitioning are missing?)*

## 4.2 Implementation verification

*What Quality Assurance measures will you take to verify the modle has been correctly implemented as per the formal model specification ? This could be as simple as an informal ‘model walkthrough' to examine the model's behaviour qualitatively, or it could include more sophisticated approaches such as formal unit testing approaches using continuous integration platforms (see [@Yenni2019]). Qualitative tests could include syntax checking of code, and code reviews by peers. Checks for verification implementation should include i) thoroughy checking for bugs or programming errors,  and ii) whether the implemented model performs as dictated by the model description  [@Grimm:2014es]. Specifying up front QA tests for implementation verification may help to avoid selective debugging.*

## 4.3 Model checking

*"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification" <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd]. During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences. Model checking therefore heps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd]. Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].*

### 4.3.1 Quantitative tests

*Specify any diagnostics or tests you will use during model checking. For each test, specify the numerical threshold or criteria that will you use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.*

### 4.3.2 Qualitative Model Checking

*This step is largely informal and case-specific, but requires ‘face validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an ddefensible  [@Grimm:2014es]. This process could be called a “laugh test" or a “pub test, and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii]. Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment. Briefly explain how you will qualitatively check the model, and whether + how you will include users and clients in the process.*

# 5.0 Model Validation and Evaluation

*Model Validation & Evaluation - This phase consists of a suite of analyses that collectively inform the decision and whether and when a model is suitable to meet its intended purpose (Augusiak, Van den Brink and Grimm 2014). Errors in design and implementation of the model and their implication on the model output are assessed. Ideally independent data is used against the model outputs to assess whether the model output behaviour exhibits the required accuracy for the model's intended purpose. The outcomes of these analyses build confidence in the model applications and increase understandingof model strengths and limitations.Model evaluation including model analysis and other testing should complement model checking. It should evaluate model checking, and consider over-fitting and extrapolation. “The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly […] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm).Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al. 2014).*

## 5.1 Model output corroboration

*Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model. Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step <2.4>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames — hence, why ecologists often model in the first place (Grimm, et al. 2014. Independent predictions can instead be tested on submodels. Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014). If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.*

## 5.2 Choose performance metrics and criteria

*Model performance can be quantified by a range of tests, including measures of agreement between predictions and independent observations, or estimates of accuracy, bias, calibration, discrimination refinement, resolution and skill [@Araujo2019]. Specify what performance measures you will use to evaluate the model and briefly explain how each test relates to different desired properties of a model's performance.Spatial, temporal and environmental pattern of errors and variance can change the interpretation of model predictions and conservation decisions [@Araujo2019], where relevant and possible, describe how you will characterise and report the spatial, temporal and environmental pattern of errors and variance. If comparing alternative models, specify what measures of model comparison or out-of-sample performance metrics will you use to find support for alternative models or else to optimise predictive ability. State what numerical threshold or qualities you will use for each of these metrics.*

## 5.3 Evaluate model assumptions

*Explain how you will demonstrate a lack of violation of or robustness to model assumptions. If you cannot perform quantitative tests, specify what theoretical justifications would justify a lack of violation of or robustness to model assumptions. If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.(Complete in follow-up iterations of the pre-registration, only if there are departures from assumptions as demonstrated in the planned tests above). The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019]. Specify how you will assess departures from assumptions. If assumption violations cannot be avoided, explain describe how you will explore their consequences.*

## 5.4 Model analysis

*Uncertainty in models arises due to incomplete system understanding (which processes to include, or which interact), from imprecise, finite and sparese data measurements, and from uncertainty in input conditions and scenarios for model simulations or runs [@Jakeman:2006ii]. Non-technical uncertainties can also be introduced throughout the modellign process, such as uncertainties arising from issues in problem-framing, indeterminicies, and modeller / client values [@Jakeman:2006ii].The purpose of model analysis is to prevent blind trust in the model by understanding how model outputs have emerged, and to ‘challenge' the model by verifying whether the model is still believeable and fit for purpose if one or more parameters are changed [@Grimm:2014es]. Model analysis should increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system. Model analysis typically consists of sensitivity analyses preceded by uncertainty analyses [@Saltelli2019], and a suite of other simulation or other computational experiments. The aim of such computational experiments is to increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system [@Grimm:2014es].Uncertainty analyses and sensitivity analyses augment one another to draw conclusions about model uncertainty — the overall model will be very uncertain if the parameters that are the most uncertain are the ones the model is most uncertain to., for instance.Because the results from a full suite of sensitivity analysis and uncertainty analysis can be difficult to interpret due to the number and complexity of causal relations examined [@Jakeman:2006ii], it is useful for the analyst to relate the choice of analysis to the modelling context, purpose and analytical objectives defined in the problem formulation phase, in tandem with any critical uncertainties that have emerged during model development and testing prior to this point.*

### 5.4.1 Uncertainty Analyses

*Uncertainty can arise from different modeling techniques, response data and predictor variables [@Araujo2019]. Uncertainty analyses characterise the uncertainty in model outputs, and identifies how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es;@Saltelli2019]. Uncertainty analyses can include propagating known uncertainties through the model, or by investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], for example. It also includes characterising the output distribution, for example "by constructing it empirically from model output data points". It could also include extracting summary statistics like the mean, median and variance from this distribution, and perhaps constructing confidence intervals on the mean [@Saltelli2019].Please describe how you will characterise any suspected critical model and data uncertainties, as well as how you plan on summarising the results of these experiments with graphical, tabular or other devices.Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase. For instance “discrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impcats of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic paraameter covariance, [@Jakeman:2006ii]), specify which measure of uncertainty you will measure.*

### 5.4.2 Sensitivity Analyses

*Sensitivity Analyses	Sensitivity analysis may improve understanding processes within models, and therefore on the natural system, or else as a quality assurance tool - sensitivity analyses may reveal logical or implementation errors in the model Sensitivity analysis examines how uncertainty in model outputs can be apportioned to different sources of uncertainty in model input [@Saltelli2019]. Model sensitivity can be evaluated using 3 approaches: 1. deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).Describe any sensitivity analyses you will conduct by specifying the initial conditions and input data for each analysis - i.e. which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameterrs will be varied. Where there were critical uncertainties or multiple plausible representations of a submodel, sensitivity analyses can be used to constract alternative submodels [@Grimm:2014es]. This analysis can be used for ‘pattern-oriented theory development' which analyses which submodel causes the full model to most closely reproduce a set of observed patterns? [@Grimm:2014es]State the primary objective of each sensitvity analysis, for example, a common objective is to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or altertantively to identify which variables or factors contribute little to overall model outputs, and so can be ‘dropped' from future iterations of the model [@Saltelli2019].What information should be reported to reproduce a sensitvity analysis should guide us as to what information we should get people to report now.*

### 5.4.3 Other simulation experiments and robustness analyses

*Please describe anyother simulation experiments and / or robustness experiments you will perform on the model.*

# 6.0 Model Application

*Explain how you will operationalise any scenarios identified during the problem formulation phase, if not already done so. For example, how will you operationalise any qualitative changes of interest, such as ‘deterioration' or ‘improvement'?Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario. Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections. Specify what raw and transformed model outputs will you extract from the model simulations or projections.*

## 6.1 Choose performance criteria for scenario analysis

*Describe how you will evaluate and distinguish the performance of alternative scenario outcomes. These measures should link back to the analytical objectives specified during the problem formulation phase, and might include for example, descriptive statistic measures, such as a model in the top 10th percentile of worst conditions, satisficing measures such as a minimum performance threshold for achieving some management objective, or perhaps regret measures, such as choosing a decision with minimum regret [@Moallemi2019]."Justify or otherwise explain how you chose these measures, and determined performance criteria in relation to the analytical objectives, modelling context, such as the risk attitudes of decision-makers and stakeholders within this system.*

## 6.2 Synthesis of model application results

*Specify how you will map, plot or otherwise display and synthesise the results of scenario and model analyses. Explain how will they support meaningful analysis and interpretation to aid in answering the analytical objectives and modelling context specified in the problem formulation phase.Explain how you will analyse the model analysis output to answer your analytical objectives. For instance, describe any trade-off or robustness analyses you will undertake to help evaluate and choose between different alternatives in consultation with experts or decision-makers.*

## 6.3 Consultation with Experts, Clients, Stakeholders

*How will you use the results of model application analysis to provide advice. and support to the clients / managers / stakeholders based on these findings?Describe how these results will be characterised and presented to experts, clients and stakeholders once the results from the modelling exercises have been synthesised. What role will these consultations have in determining the selection of management interventions, or future iterations of model development? For example, what key outliers or strongest examples of best or worst case scenarios should be presented to experts or decision-makers?(Do we need an item about how do we interpret the model outputs? Should there be some a priori decisions about how we interpret the model findings, and how does case 49 about the risk attitude inform how the analyst interprets the model findings?)*

# References

### 2020-05-10 analysis/preregistration_template.Rmd
#9 update draft template with context to the template

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
+# About this template
+
+This template has been collated from a combination of literature reviews of the 'Good Modelling Practice' and other modelling literature, as well as a collaborative workshop at DELWP.
+Chris Jones has kindly volunteered one component of the VEFMAP project investigating vegetation responses to environmental flows as a testing ground for the use of preregistration templates in applied ecological modelling research.
+This template will be used to preregister this research, and we will use the VEFMAP project to both develop and evaluate the template 'in situ'.
+
+The template is largely structured around 6 "phases" in the process of developing, evaluating and analysing a model or models in ecological applications: `Problem Formulation`, `Define Conceptual Model`, `Formalise and Specify Model`,  `Model Validation and Evaluation`, `Model Analysis`, `Model Application`.
+These have been defined and described to some extent throughout the template.
+
+## iterative model development and a 'living preregistration'
+
+One of the major challenges in designing a template for modelling is that the process of model development is inherently iterative.
+Typically, the process of preregistration is completely distinct from and precedes the implementation of the analysis plan laid out in the preregistration.
+However, translating the preregistration process to a modelling research context is not so clear cut because some decisions are inherently dependent on the outcomes and results of previous decisions and analyses.
+For example, some items at the later stages of modelling probably can't be answered until the model itself is fully or close to fully developed, especially for more complex process models (e.g.
+specifying exactly what sensitivity analyses or uncertainty analyses will be conducted).
+Thus deciding on the inner logic of the template, and what is in and out of scope has been particularly difficult.
+
+One method around this is to accept that the current model of preregistration is not going to work for iterative model development.
+Again, the more complex the model, the more likely this is to be the case.
+Consequently I propose a different model of preregistration that I think will work for modeling research contexts.
+At the moment I'm calling it 'living preregistration' but please suggest another name.
+Others have termed this 'adaptive preregistration', and this is maybe a better name.
+
+In this model of preregistration, all questions that can be answered, should be answered before any data analysis or modelling proceeds.
+For some items, particularly during model evaluation and analysis, perhaps only responses of a general nature can be provided.
+That is OK.
+As the modelling proceeds.
+Further detail can be provided and the next phase of analysis can be preregistered.
+Where the results of model evaluation and analysis reveal that there are problems with the model, or perhaps assumption violations, plans can be changed, and again the next phase of analysis can be preregistered.
+
+A static document for this type of process doesn't really make sense then.
+I propose leveraging the code and project management tools within the GitHub platform as a way of hosting a 'living' or 'adaptive' preregistration.
+Because GitHub documents are 'time-stamped', the genesis of the preregistration from one version to the next can be traced and tagged with versions.
+Since the document would exist in the repository with the code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated plan.
+
+--------------------------------
 

### 2020-05-10 analysis/preregistration_template.Rmd
#9 update draft template with about section

@@
 title: "Ecology and Conservation Modelling Preregistration Template -- DRAFT"
-author: "Elise Gould, Add your name here if you edit or make comments on this document, TBD:workshop collaborators"
+author: "Elise Gould, Add your name here if you edit or make comments on this document,TBD:workshop collaborators"
+
 date: "`r format(Sys.time(), '%d %B, %Y')`"
 output:
-  github_document
-bibliography: "../data/references.bib"
+  github_document: default
+bibliography: ../data/references.bib
 ---

### 2020-05-13 analysis/preregistration_template.Rmd
Update preregistration_template.Rmd

--- 2020-05-10 analysis/preregistration_template.Rmd
@@
 title: "Ecology and Conservation Modelling Preregistration Template -- DRAFT"
-author: "Elise Gould, Add your name here if you edit or make comments on this document, TBD:workshop collaborators"
+author: "Elise Gould, Megan Good, TBD:workshop collaborators"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-A static document for this type of process doesn't really make sense then.
+A static document for this type of process doesn't really make sense.
 I propose leveraging the code and project management tools within the GitHub platform as a way of hosting a 'living' or 'adaptive' preregistration.
@@
 What is the purpose of the model/s?
-Ensure that you specify any focal taxa and study objectives, as well as any clients for whom the model is developed for.
+Ensure that you specify any focal taxa and study objectives, as well as any clients for whom the model is developed.
 Briefly outline the ecological problem and the decision-problem, including the decision-trigger and any regulatory frameworks relevant to the problem.*
@@
 Everything beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al.
-2006).The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
-(Should this last step go here or within define conceptual framework?).What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)?
+2006).The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?
+Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
+(Should this last step go here or within define conceptual framework?).
+What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)?
 This is relevant to model transferability...
 (similar to a COG statement..)*
+A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation.
+I'm just thinking about how I might fill this out for an imagined study.
+The literal boundaries of the study are usually straightforward, but the theoretic boundaries are a different and also big thing to try to conceptualise in one section.
+You could have a section for the spatial/temporal boundaries of the study, and another section on how far the results can be extrapolated based on the study design.
+This makes sense to me in how I design experiments, if I have too many confounding variables and not enough spatial and environmental replication then I'm less likely to make any broad claims.
+However, if I have controlled or experimentally manipulated factors then I feel safer making bigger generalisations about results.
 
@@
 For example, a model outcome in the top 10th percentile of worst conditions, or some minimum performance threshold for meeting a performance objective, or choosing a decision with the minimum regret or least amount of uncertainty.*
+To be honest, I don't know how I might apply this to my work or any future works.
+Maybe I do not have enough knowledge of this type of research to apply it to my basic ecology research.
+But this isn't to say it's not important, I just wonder if there's a way to make this accessible to non-desicion or modelling peeps.
+Like, how would I apply it to the woodland condition analysis?
 
@@
 
-##2.3-define_conceptual_model: identify predictor response variables
+## 2.3-define_conceptual_model: identify predictor response variables
 
@@
 
-- *What variables will support this decision or action (things we can control).* 
+- *What variables would support taking this action or making this decision?* 
 - *What additional variables may interact with this system (things we can't control, but can hopefully measure).* 
 - *In what ways do we expect these variables to interact (model structures).* 
+Should you include variables that haven't been measured and that might also interact with the system (not ideal, but this happens in field studies).
+Possibly also variables that are assumed to be a surrogate measure of variables that we can't or haven't measured.
 
@@
 
-*Describe any data preparation and processing steps, including manipulation of environmental layers, e.g. standardisatino and geographic projection.*
+*Describe any data preparation and processing steps, including manipulation of environmental layers, e.g.
+standardisatino and geographic projection.* Something about always retaining the raw data in a safe place, no manipulations to the raw file.
 
@@
 *All modelling approaches require the selection of model features, which conform with the conceptual model and data specified in previous steps (Jakeman 2006).
-The choice of model are determined in conjunction with features are selected.Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006).
+The choice of model are determined in conjunction with features are selected.
+Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006).
 However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.*
@@
 Ideally independent data is used against the model outputs to assess whether the model output behaviour exhibits the required accuracy for the model's intended purpose.
-The outcomes of these analyses build confidence in the model applications and increase understandingof model strengths and limitations.Model evaluation including model analysis and other testing should complement model checking.
+The outcomes of these analyses build confidence in the model applications and increase understandingof model strengths and limitations.
+Model evaluation including model analysis and other testing should complement model checking.
 It should evaluate model checking, and consider over-fitting and extrapolation.
-“The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly […] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm).Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al.
+The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly […] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm).
+Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al.
 2014).*

### 2020-05-19 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
Fix typo
Fix #13 Respond to Meg's comments on variables

@@
 
-## 2.3-define_conceptual_model: identify predictor response variables
+## 2.3 Define conceptual model: identify predictor response variables
 
@@
 - *What variables would support taking this action or making this decision?* 
-- *What additional variables may interact with this system (things we can't control, but can hopefully measure).* 
-- *In what ways do we expect these variables to interact (model structures).* 
-Should you include variables that haven't been measured and that might also interact with the system (not ideal, but this happens in field studies).
-Possibly also variables that are assumed to be a surrogate measure of variables that we can't or haven't measured.
+- *What additional variables may interact with this system (things we can't control, but can hopefully measure)?*
+- *What variables have not been measured, but may interact with the system (often occurs in field or observational studies)?*
+- *What variables are indice or surrogate measures of variables that we cannot or have not measured?*
+- *In what ways do we expect these variables to interact (model structures)?*
 
@@
 *Describe any data preparation and processing steps, including manipulation of environmental layers, e.g.
-standardisatino and geographic projection.* Something about always retaining the raw data in a safe place, no manipulations to the raw file.
+standardisation and geographic projection.* Something about always retaining the raw data in a safe place, no manipulations to the raw file.
 

### 2020-05-19 analysis/preregistration_template.Rmd
Fix #15 update data processing items
Respond to Meg's comment in PR11 and look back at [https://osf.io/cte9k/](https://osf.io/cte9k/) to update data processing items.

@@
 
-*Describe any data preparation and processing steps, including manipulation of environmental layers, e.g.
-standardisation and geographic projection.* Something about always retaining the raw data in a safe place, no manipulations to the raw file.
+*Describe any data preparation and processing steps, including manipulation of environmental layers, e.g. standardisation and geographic projection.*
+
+*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.*
 
@@
 
-*Describe how you will evaluate the quality and sources of both numerical and qualitative data that available for model parameterisation, and for determining the overall model structure [@Grimm:2014es]. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].
+*Describe how you will evaluate the quality and sources of both numerical and qualitative data that available for model parameterisation, and for determining the overall model structure [@Grimm:2014es].
+Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].
 Document any issues with data reliability.
 This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].*
+
+*Data Exclusion: How will you determine what data, if any, will be excluded from your analyses?
+How will outliers be handled?
+Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.*
+
+*Missing data: How will you identify and deal with incomplete or missing data?*
 
@@
 
-*Describe the validation scheme you will use for model testing and evaluation.  Please explain your reasoning for your choice of model calibration and validation scheme.
+*Describe the validation scheme you will use for model testing and evaluation.
+Please explain your reasoning for your choice of model calibration and validation scheme.
 The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;[@Barnard2019].
 Although using independent data is the gold-standard or idealised practice for model evaluation, Typically testing on independent datasets is not possible for large and/or integrated models, especially when they are being used to generate anticipatory predictions beyond the conditions on which they were calibrated [@Jakeman:2006ii].
-Describe and justify the method for your calibration and validation schema, including how many folds and what methods you will use to partition the data set into training/testing data.
-(What other items, esp.
-re.
-partitioning are missing?)*
+Describe and justify the method for your calibration and validation schema.*
+
+## 4.1.2 Data partitioning
+
+*Describe the approach for partitioning the data in support of the calibration and validation schema described above.
+For example, specify the number of folds that will be created and the relative size of each.*
+
+*How will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced?*
 

### 2020-05-19 analysis/preregistration_template.Rmd
Fix #17 Add pre-existing data & data sampling
Add items from Data Description for Pre-existing data and own data collection [https://osf.io/cte9k/](https://osf.io/cte9k/).

@@
 *Document the identity, quantity and provenance of any data used to develop, identify and test the model.
-Describe how the data is arranged, in terms of replication and covariates.
-Explain how you will summarise what data and knowledge will be used to designa and parameterise the model, including references, data sources, and information about whom, when and where those data were collected [@Grimm:2014es].*
+Describe how the data is arranged, in terms of replication and covariates.Explain how you will summarise what data and knowledge will be used to design and parameterise the model, including references, data sources, and information about whom, when and where those data were collected [@Grimm:2014es].*
+
+**For pre-existing data: -- as appropriate.**
+
+- *For each dataset, is the data open or publically available? (Y/N)*
+- *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrctions on the use of data.*
+- *Date of download, access, or future access:*
+- *Describe the source of the data - what entity originally collected this data?
+(National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
+- *Codebook. If a Codebook is available, please link to it here and / or upload the document.*
+- *Prior work based on this dataset - Have you published / presented any previous work based on this dataset?
+Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont his dataset you have worked on.*
+- *Prior Research Activity - Have you worked with these data before? Describe any prior but unpublished research activity using these data. Be specific and transparent.*
+- *Prior knowledge of the current dataset - Describe any prior knowledge of the dataset. Be specific.*
+
+**Sampling Plan (For data you will collect) -- delete as appropriate.**
+
+- Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected.
+Describe any inclusion or exclusion rules, and the study timeline.
+- Sample Size - Describe the sample size of your study.
+- Sample Size Rationale - Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel. 
+- If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?
 

### 2020-05-19 analysis/preregistration_template.Rmd

@@
 output:
-  github_document
-bibliography: "../data/references.bib"
+  github_document: default
+bibliography: ../data/references.bib
 ---

### 2020-05-23 index.Rmd
Add empty manuscript paper index.Rmd and no jekyll for publish to GH pages

---
title: "Untitled"
description: |
  A new article created using the Distill format.
author:
  - name: Elise Gould 
    affiliation: School of BioSciences, University of Melbourne
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.

# Methods

# Structured Workshop Activities

Workshop materials, and outputs are accessible on the OSF @TODO link here.

### 2020-05-24 analysis/preregistration_template.Rmd
Incorporate @LibbyR's changes #10
Add @LibbyR 's changes to raw Rmd instead of to .md file.

@@
 
-*Specifies the decision-making context in which the model will be used, the clients driving model development or stakeholders addressed by the model.
-It also includes specification of model outputs, statements of the domain of applicability of the model, as well as the extent by which model outputs may be acceptably extrapolated (Grimm et al.
+*Specifies the decision-making context in which the model will be used.
+Important components include the decision maker and stakeholders (including experts) and their view on: i) the nature of the problem or decision addressed and how the scope of the modelling tool fits within the (broader) context (i.e.
+model purpose; ii) the relevant spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they forsee unnacceptable outcomes that need to be represented in the model (i.e.
+as constraints), and; vi) what future scenarios does the model need to account for (noting this may be revised later).
+It should also provide a summary of the domain of applicability of the model, and reasonable extrapolation limits (Grimm et al.
 2014).*
@@
 
-## 1.3.1 Explain analytical objectives
+## 1.3.1 Explain analytical objectives (MAKE THIS A SEPARATE SECTION WHICH INCLUDES PERF MEASURES)
 

### 2020-05-24 analysis/preregistration_template.Rmd
Add Libby as a template author
Add @LibbyR as an author after incorporating her changes on the .md file.

@@
 title: "Ecology and Conservation Modelling Preregistration Template -- DRAFT"
-author: "Elise Gould, Megan Good, TBD:workshop collaborators"
+author: "Elise Gould, Megan Good, Libby Rumpff, TBD:workshop collaborators"
 date: "`r format(Sys.time(), '%d %B, %Y')`"

### 2020-05-25 index.Rmd
Export Manuscript from Scrivener -> Rmd

@@
 ---
-title: "Untitled"
+title: "EcoConsPreReg"
 description: |
@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
-```
+## Introduction ##
 
-Distill is a publication format for scientific and technical writing, native to the web.
+### Problem ###
 
-Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.
+Problem: "Conceptual barriers to adopting preregistration, including users from different disciplines feeling that preregistration is not for them" Pu et al. (2019).
 
-# Methods
+"Preregistration considered alienating.
+Researchers who do non-experimental studies expressed a "not for me" sentiment." We have experienced this first-hand by our peers, and even by ourselves, when opening and looking at preregistration templates.
 
-# Structured Workshop Activities
+Questions about whether preregistration is useful for exploratory work because it 'limits flexibility' --- and moreover, modelling is, by many, considered exploratory.
 
-Workshop materials, and outputs are accessible on the OSF @TODO link here.
+### Research Aims, and Template Objectives ###
+
+Research Aim: To develop a Template for Ecology and Conservation.
+
+Objectives of the Template itself & internal model of preregistration (this will build upon previous work of mine, namely the QRP chapter.
+See also Pu et al.
+for the four 'purposes' of preregistration that they identified in their analysis.
+
+----
+
+## Methods ##
+
+### User-centred design approach ###
+
+As Pu et al.
+(2019) identify, a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings.
+In order to bridge this mismatch, we have taken a user-centred design approach to the problem of preregistration.
+We do this in two ways, by first using a series of structured and collaborative activities in a workshop setting to reflect on our own research experiences in order to capture scientific workflows as they are done in practice.
+And secondly to, use a real-world research problem as a case-study to test and evaluate the templates for utility and feasibility in practice.
+
+The workshop activities aimed to help us:
+
+a) understand how science is done in practice so that we can use this to inform and tailor the preregistration templates to adequately capture key decision-points in the scientific process;
+
+b) understand how science in practice, whether due to logistical, technical feasibility or normative constraints, differs from some idealised practice;
+
+c) identify challenges and roadblocks to the use and uptake of preregistration in among ecologists and conservation scientists, both in academic and applied research contexts; and to
+
+c) Use this knowledge to inform the design of the templates so they meet their intended purpose and are taken up by researchers.
+
+**Case Study: Environmental Flows Management**
+
+\<TEXT HERE DESCRIBING THE ENVIRONMENTAL FLOWS MANAGEMENT PROBLEM + RESEARCH PROBLEM\>
+
+### Designing Preregistration Templates Workshop ###
+
+On the 4^th^ March 2020, we ran a series of structured activities designed to inform the development of two pre-registration templates, one for field-work and sampling design, the other for modelling and data-analysis.
+We split the activities into two groups, to better focus discussion for each of these templates.
+All workshop materials and workshop outputs can be found on the OSF component here \<LINK\>.
+
+**Activity 1 - Capturing Workflows and Decision Steps**
+
+The objective of this activity was to capture critical decision-steps in the design of both field sampling efforts and modelling exercises.
+This activity was intended to identify key decision-steps in the scientific process that should be included in a preregistration template.
+The first part of this activity was done as individuals, where we reflected on our scientific process for a recent or memorable research project.
+We listed the steps in the workflow from beginning to end.
+We attempted to document  critical decision-steps in this process that should be present on a preregistration template.
+By "critical" we mean that without these decision-steps the template would not a) adequately constrain researcher degrees of freedom and b) transparently document the analytic decision-process.
+
+In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto a group workflow template This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase.
+The development of the draft modelling workflow was derived from previous work by Gould et al.
+*in prep.,* whilst the field sampling workflow was created specifically for this exercise.
+We sorted and categorised each decision-step as belonging to a particular workflow phase.
+Next we grouped similar decision-steps that could be included as a single item on a preregistration template.
+At the end of this process we reviewed the suggested phases on the workflow to assess whether phases needed further discretisation, aggregation, removal, or amendment.
+
+**Activity 2 - Challenges and Solutions to Preregistration in Ecology**
+
+This second activity aimed to identify critical challenges and roadblocks to the implementation and uptake of preregistration in ecology and conservation, as well as their solutions.
+We conducted two lots of structured discussions, one for each of the field and modelling template groups.
+Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature.
+These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in activity 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.
+
+Questions included: should we and how can we change the medium of the template and/or archiving platform to accomodate any roadblocks?
+How could we change the medium of both the template and the archiving platform to accommodate these issues?
+What procedural and or technical solutions could we implement to address these roadblocks to preregistration?
+What decision-tools could be used to aid in completing a preregistration?
+What sort of parsimony should we aim for in terms of the resolution of the templates and therefore their applicability across different methodologies?
+
+### Workshop Data Collation and Analysis - Identifying Scientific Workflows and Decision Steps ###
+
+All workshop outputs can be found here (\<Link to Workshop Outputs OSF repository\>).
+
+After the workshop, data sheets from activity 1 were collated and coded using atlas.TI.
+
+The point of this activity is to build on the suggested workflows for both fieldwork and modelling to provide structure and content to the preregistration templates.
+
+At the end of this exercise we expect to:
+
+Have a complete draft workflow for both fieldwork and modelling and data-analysis that will form the basis of the final preregistration templates.
+
+### Turning the Workflows into Preregistration Templates ###
+
+Part I: Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.
+
+Objective: Identify common workflow.
+
+Step 1: As a group, reflect on personal experience to identify unique decision-steps. Can a generalised / common workflow across individuals be derived?
+
+Step 2: Search literature to to identify an idealised scientific workflow for ecological modelling research.
+
+Step 3: Code workflows identified in step 1 using codes identified in step 3. Revise common workflow.
+
+Part II: Translating Workflows into Preregistration Templates
+
+We translated the workflows into the preregistration template by using the following ontology:
+
+1\. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases.
+
+2\. Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps'. Corresponds to an activity or set of activities in a scientific workflow.
+
+3\.
+Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
+The style should be in the form of a directive or a question.
+E.g.
+"explain how you will do xyzzy".
+Or "What performance measure will you use for assessing 'goodness of fit'?" In addition, we turned to existing preregistration templates to guide the wording  of some items.
+
+### In-situ Evaluation and Testing of Templates ###
+
+Rather than 'digging back' through the version history of a completed preregistration to ask researchers about the process and how it had been developed in detail, we chose to leverage the version-control and collaborative project management features of GitHub \<cite\> as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to 'live-develop' the template template.
+For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration, the suggested change, and its justification would be both recorded in GitHub and linked to an actual preregistration.
+
+How we did this using GitHub:
+
+Instruction to
+
+(As Pu et al did, they used qualitative interviews to 'walk through the version history of the preregistration, and discuss how it had developed in detail' (p.9) --- we can capture some of that thinking in the moment, by using comments and GitHub issues to capture the reasoning in the moment, as to why a particular decision was made or needs to change.
+However, I do think we should supplement this process with a structured reflection session.
+
+The reason why tried to capture this as we went, is because "the process of creating preregistration  is difficult to reconstruct with interview questions alone" p16.
+
+### Pu et al. ###
+
+What is the purpose of preregistration? Specifically in modelling.
+
+The original purpose of preregistration is "the action of confirming an unalterable version of one's research plan prior to collecting data", such that the act distinguishes preregistered, confirmatory findings from exploratory findings (Nosek et al.
+2018 + Kidwell et al.
+2016) by way of constraining 'researcher degrees of freedom' or flexibility in both experimental design and data analysis, which has been posited as one of the key reasons contributing to the replication crisis (Ioannidis, 2005, Wagenmakers, 2012).
+
+Pu et al.
+treat preregistration as a user-centred design problem to ask whether it achieves its proposed goals *in practice*, and how the process might be improved.
+They ask whether the current implementation of preregistration in practice, in terms of its design of the form, and how the process integrates into research practice and the peer review practice, effectively supports its intended goals.
+By interviewing academic psychologists, they attempt to characterise current understanding of practice, and both the norms and purposes for preregistration that actual users have.
+The keybindings are:
+
+\- Norms around how people use preregistration are uncertain, and vary among users
+
+\- different users have "different purposes for adopting preregistration, and these differ from the original purpose of distinguishing exploratory and confirmatory findings"
+
+\- Depending on what role you assume, there are different 'use-cases' for authors, reviewers/editors. BUT, these different use cases are not explicitly designed for.
+
+\- There are "conceptual barriers to adopting preregistration, including users from different disciplines feeling that preregistration is not for them"
+
+Pu et al found that because different users ascribed different purposes to preregistration, it resulted in conflicting designs of preregistration formats, which they argue may ultimately undermine the overarching goal of preregistration: "improving the reliability of scientific findings"
+
+What do we define as the *purposes* of preregistration, specifically within ecological modelling contexts?
+
+\- P**rimary purpose:** "Delimiting flexibility" Following from Pu et al. we define the primary purpose of preregistration for modelling exercises so as to
+
+We take emphasis *off* of the dichotomy of exploratory vs.
+confirmatory research, because I think these categories are confused and confusing in modelling, and actually unhelpful.
+Instead I want to focus on delimiting researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and the potential for resultant QRPs, such as, cherry-picking, HARKing, etc.
+etc.
+Link to my other manuscript.
+
+\- **Secondary purpose: increasing transparency** \<insert refs to modelling reporting literature - it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced\> Thus, want to improve the transparency of decision-making and methodology.
+What is "research transparency" 'Lupia and Elman define *research transparency* to be production transparency (open access data or data collection procedures) and analytic transparency ("a full account of how they draw their analytic conclusions from the data")' --- Lupia and Elman 2014.
+Pu et al.
+page 13.
+
+\- **Possibly:** As a way of providing better decision-making to authors.
+By shifting decision-making to the front.
+Front-heavy research process.
+Delays the start of analysis, but results in better decision-making, because the decision-making is more carefully reasoned.
+But not sure want to set this out as a 'purpose' of the template, or just have this as a section in the reflection towards the end of the manuscript (will depend on how I set up the angle, and the structure of the paper).
+At the moment, this is my hypothesis, is that it can improve the process, but we'd need to reflect on this.
+And I think we could do that at the end of the manuscript.
+
+This relates to what Pu et al.
+found where participants 'felt their research workflow improved by having preregistration as a "forcing function" \[...\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers' degree of freedmen advance.' (Page 13).
+For reviewers 'participant reports that pregistration makes the review process "easier".'
+
+So what are *our* use-cases?
+
+The shiny-app, allows us to tailor the form firstly to authors and analysts.
+With help-text, and descriptions of the modelling phase, and step under consideration, as well as definitions and references to supporting resources ---  such as decision-tools for choosing and justifying methodological choices.
+
+By being able to 'generate a report' - we aim to target reviewers/editors, to allow for a condensed version of the template to be exported.
+Importantly, Pu et al show that "not all preregistered studies confirm to their preregistrations, or are inconsistent in reporting deviations", thus by making the preregistration available to reviewers, and by providing a platform and method for linking the actual analysis under-taken, we can allow reviewers to properly evaluate the preregistration against the final paper.
+(AND, must talk back to original aim of bad reporting practices in modelling).
+Moreover, if you combine formatting changes with our proposed method with GitHub, we hope to tailor the presentation of the completed preregistration + analysis, such that actual analyses and decisions can be linked directly to both files and outcomes.
+
+Preregistration *Norms*:
+
+Pu et al.'s working definition of *norms* is 'as a set of ideal behaviours'.
+A key finding was that norms around both creating and using a preregistration are uncertain.
+*How will my work address this confusion among researchers about what the norms are, how will it address uncertain norms?*
+
+Creating:
+
+\- E.g. what sorts of decisions should be preregistered? Conflicting ideas about whether a power analysis, should for example be included in a preregistration.
+
+Using:
+
+\- How to mark the preregistered from the non-preregistered parts of the paper?
+
+\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
+as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
+"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
+And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
+Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
+How can we facilitate the process?
+They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
+Maybe we could do something else, rather than get the paper to refer to the parts of the preregistration (but I think we should do this with Chris anyway).
+If we propose the norm of using GitHub to create and check the preregistration., I think this could be a good way of getting around this checking problem.
+At the very end we have a 'how to box' targeted at authors.
+And then maybe we have a box targeted at reviewers?
+Not sure, maybe just focus this paper on authors and the process of creating and using a preregistration.
+
+In order to address tis confusion among norms, I state the expected norms that the template has been designed around here.
+However, I don't expect users to agree on this process.
+Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
+
+*Problems and Challenges in designing preregistration templates*
+
+***TIME***
+
+\- "A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
+This trade-off was something we considered important during the design phase, and deciding what was in and out of scope for the preregistration was particularly challenging.
+Some authors agreed that it did add extra time, but that it was worth it.
+
+\- However, as Pu et al.
+note, some participants who were both authors and reviewers didn't think that the time-costs of preregistration in research were significantly greater than not preregistering research.
+Instead they thought of the preregistration process invoking a 'time shift', that brought 'decision points' before data collection instead of during analysis." P.19.
+In addition, "some reviewers thought that preregistration saves them the time they might have spent wondering about undisclosed flexibilities when working on an un-preregistered paper."
+
+**PURPOSE AND DESIGN IMPLEMENTATION**
+
+\- Competing objectives?
+Although the purposes of preregistration are not mutually exclusive - as we have decided the purpose of preregistration we have defined here is to both delimit/state flexibility, *and* be flexible.
+However, "the preference of one purpose over another can have tangible effects on the design of preregistration" p15.
+
+----
+
+## PRT Conceptual Model and Philosophy underpinning model development ##
+
+### Model Scientific Workflow ###
+
+What model workflow and set of activities does the template presume?
+
+This section describes the 'model scientific workflow' underpinning  the Preregistration Template.
+
+I want to include a figure / diagram here visualising and describing some usual / common or even idealised workflow to accompany both the PRT for
+
+A) PRT users.
+
+B) To Explain the inner conceptual model driving the PRT (as per Pu et al.).
+
+*How does this inform the PRT? These broad phases are used to structure the preregistration template.*
+
+### Iterative Model Development & "Living" Preregistrations ###
+
+Paragraph or two describing iterative nature in model development. This was a common theme arising in our discussions and one of the key 'roadblocks' identified.
+
+*How do we accommodate this structurally and procedurally into the PRT?*
+
+*1\. Instruction to PRT users:*
+
+"If you find yourself needing to check something with the data before committing to a particular decision about the model or analysis (And you hadn't anticipated needing to perform this check), record:
+
+A) the relevant decision in the template.
+
+B) What do you need to know to move forward?
+
+C) How you will test this? What test or analysis will you perform?
+
+D) What values or outputs of the test or analysis map onto particular decisions (they can be qualitative as well)."
+
+I will create a template on GitHub based on the above to be used by Chris + Lyndsey for when this occurs, to be stored on the repository wiki.
+They can insert this template into the relevant decision step / PRT item and fill out.
+
+*2\. "Living" or "Timestamped" or "Adaptive" Preregistrations ---Preregistration in the research process:*
+
+See Figure 1 in Pu et al.
+And [*cos.io/rr*](http://cos.io/rr) - Traditionally the process of preregistration and the 'doing' or execution of the plan is a distinct and sequential process.
+However, this isn't going to work for modelling (What are the *norms* about the process of modelling - stated that it's iterative.
+And how does this look in practice --- use our in situ evaluation and testing to capture this.
+
+This obviously breaks with the convention of preregistrations as 'static' one-off documents... does it break with this convention so far as to no longer be a 'preregistration'?
+
+Pu et al: tendency for preregistrations to go from vague / general to specific over time.
+
+*\<div style="width: 960px; height: 720px; margin: 10px; position: relative;"\>\<iframe allowfullscreen frameborder="0" style="width:960px; height:720px" src="https://app.lucidchart.com/documents/embeddedchart/9c4480bc-8465-4731-83e6-d24d7d2b7818" id="K4J1KZqFLi3u"\>\</iframe\>\</div\>*
+--------------------------
+# References

### 2020-05-25 index.Rmd
restructure draft manuscript

@@
 ---
+The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements:
+
+Point 1: set the context for and purpose of the work;
+
+Point 2: indicate the approach and methods;
+
+Point 3: outline the main results;
+
+Point 4: identify the conclusions and the wider implications.
+
+**Keywords:**
+
+----
 
@@
 
+\*\*About the Approach\*\*
+
 As Pu et al.
@@
 
-**Case Study: Environmental Flows Management**
+\*\*
 
-\<TEXT HERE DESCRIBING THE ENVIRONMENTAL FLOWS MANAGEMENT PROBLEM + RESEARCH PROBLEM\>
-
-### Designing Preregistration Templates Workshop ###
+### Identifying Ecological Modelling Workflows in Practice ###
 
@@
 
-### Workshop Data Collation and Analysis - Identifying Scientific Workflows and Decision Steps ###
+### Turning the Workflows into Preregistration Templates ###
 
@@
 Have a complete draft workflow for both fieldwork and modelling and data-analysis that will form the basis of the final preregistration templates.
-
-### Turning the Workflows into Preregistration Templates ###
 
@@
 
-### In-situ Evaluation and Testing of Templates ###
+### 'In-situ' Evaluation and Testing of Templates ###
 
@@
 
+**Case Study: Environmental Flows Management**
+
+\<TEXT HERE DESCRIBING THE ENVIRONMENTAL FLOWS MANAGEMENT PROBLEM + RESEARCH PROBLEM\>
+
+----
+
+## PRT Conceptual Model and Philosophy underpinning model development ##
+
+### Model Scientific Workflow ###
+
+What model workflow and set of activities does the template presume?
+
+This section describes the 'model scientific workflow' underpinning  the Preregistration Template.
+
+I want to include a figure / diagram here visualising and describing some usual / common or even idealised workflow to accompany both the PRT for
+
+A) PRT users.
+
+B) To Explain the inner conceptual model driving the PRT (as per Pu et al.).
+
+*How does this inform the PRT? These broad phases are used to structure the preregistration template.*
+
+### Iterative Model Development & "Living" Preregistrations ###
+
+Paragraph or two describing iterative nature in model development. This was a common theme arising in our discussions and one of the key 'roadblocks' identified.
+
+*How do we accommodate this structurally and procedurally into the PRT?*
+
+*1\. Instruction to PRT users:*
+
+"If you find yourself needing to check something with the data before committing to a particular decision about the model or analysis (And you hadn't anticipated needing to perform this check), record:
+
+A) the relevant decision in the template.
+
+B) What do you need to know to move forward?
+
+C) How you will test this? What test or analysis will you perform?
+
+D) What values or outputs of the test or analysis map onto particular decisions (they can be qualitative as well)."
+
+I will create a template on GitHub based on the above to be used by Chris + Lyndsey for when this occurs, to be stored on the repository wiki.
+They can insert this template into the relevant decision step / PRT item and fill out.
+
+*2\. "Living" or "Timestamped" or "Adaptive" Preregistrations ---Preregistration in the research process:*
+
+See Figure 1 in Pu et al.
+And [*cos.io/rr*](http://cos.io/rr) - Traditionally the process of preregistration and the 'doing' or execution of the plan is a distinct and sequential process.
+However, this isn't going to work for modelling (What are the *norms* about the process of modelling - stated that it's iterative.
+And how does this look in practice --- use our in situ evaluation and testing to capture this.
+
+This obviously breaks with the convention of preregistrations as 'static' one-off documents... does it break with this convention so far as to no longer be a 'preregistration'?
+
+Pu et al: tendency for preregistrations to go from vague / general to specific over time.
+
+*\<div style="width: 960px; height: 720px; margin: 10px; position: relative;"\>\<iframe allowfullscreen frameborder="0" style="width:960px; height:720px" src="https://app.lucidchart.com/documents/embeddedchart/9c4480bc-8465-4731-83e6-d24d7d2b7818" id="K4J1KZqFLi3u"\>\</iframe\>\</div\>*
+
 ### Pu et al. ###
@@
 By interviewing academic psychologists, they attempt to characterise current understanding of practice, and both the norms and purposes for preregistration that actual users have.
-The keybindings are:
+The key findings are:
 
@@
 
+----
+
+## Discussion ##
+
+## talking points ##
+
+*We really didn't know what to expect when we began completing these preregistrations...*
+
 *Problems and Challenges in designing preregistration templates*
@@
 
-----
+\- **Possibly:** As a way of providing better decision-making to authors.
+By shifting decision-making to the front.
+Front-heavy research process.
+Delays the start of analysis, but results in better decision-making, because the decision-making is more carefully reasoned.
+But not sure want to set this out as a 'purpose' of the template, or just have this as a section in the reflection towards the end of the manuscript (will depend on how I set up the angle, and the structure of the paper).
+At the moment, this is my hypothesis, is that it can improve the process, but we'd need to reflect on this.
+And I think we could do that at the end of the manuscript.
 
-## PRT Conceptual Model and Philosophy underpinning model development ##
-
-### Model Scientific Workflow ###
-
-What model workflow and set of activities does the template presume?
-
-This section describes the 'model scientific workflow' underpinning  the Preregistration Template.
-
-I want to include a figure / diagram here visualising and describing some usual / common or even idealised workflow to accompany both the PRT for
-
-A) PRT users.
-
-B) To Explain the inner conceptual model driving the PRT (as per Pu et al.).
-
-*How does this inform the PRT? These broad phases are used to structure the preregistration template.*
-
-### Iterative Model Development & "Living" Preregistrations ###
-
-Paragraph or two describing iterative nature in model development. This was a common theme arising in our discussions and one of the key 'roadblocks' identified.
-
-*How do we accommodate this structurally and procedurally into the PRT?*
-
-*1\. Instruction to PRT users:*
-
-"If you find yourself needing to check something with the data before committing to a particular decision about the model or analysis (And you hadn't anticipated needing to perform this check), record:
-
-A) the relevant decision in the template.
-
-B) What do you need to know to move forward?
-
-C) How you will test this? What test or analysis will you perform?
-
-D) What values or outputs of the test or analysis map onto particular decisions (they can be qualitative as well)."
-
-I will create a template on GitHub based on the above to be used by Chris + Lyndsey for when this occurs, to be stored on the repository wiki.
-They can insert this template into the relevant decision step / PRT item and fill out.
-
-*2\. "Living" or "Timestamped" or "Adaptive" Preregistrations ---Preregistration in the research process:*
-
-See Figure 1 in Pu et al.
-And [*cos.io/rr*](http://cos.io/rr) - Traditionally the process of preregistration and the 'doing' or execution of the plan is a distinct and sequential process.
-However, this isn't going to work for modelling (What are the *norms* about the process of modelling - stated that it's iterative.
-And how does this look in practice --- use our in situ evaluation and testing to capture this.
-
-This obviously breaks with the convention of preregistrations as 'static' one-off documents... does it break with this convention so far as to no longer be a 'preregistration'?
-
-Pu et al: tendency for preregistrations to go from vague / general to specific over time.
-
-*\<div style="width: 960px; height: 720px; margin: 10px; position: relative;"\>\<iframe allowfullscreen frameborder="0" style="width:960px; height:720px" src="https://app.lucidchart.com/documents/embeddedchart/9c4480bc-8465-4731-83e6-d24d7d2b7818" id="K4J1KZqFLi3u"\>\</iframe\>\</div\>*
+This relates to what Pu et al.
+found where participants 'felt their research workflow improved by having preregistration as a "forcing function" \[...\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers' degree of freedmen advance.' (Page 13).
+For reviewers 'participant reports that pregistration makes the review process "easier".'
 --------------------------

### 2020-05-25 index.Rmd
Add image to manuscript, add shiny app design info

@@
 
+But not just a pre-registration template, we take into account the iterative nature of model development
+
 Objectives of the Template itself & internal model of preregistration (this will build upon previous work of mine, namely the QRP chapter.
@@
 (2019) identify, a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings.
-In order to bridge this mismatch, we have taken a user-centred design approach to the problem of preregistration.
-We do this in two ways, by first using a series of structured and collaborative activities in a workshop setting to reflect on our own research experiences in order to capture scientific workflows as they are done in practice.
+In order to avert this mismatch, we have taken a user-centred design approach to the problem of preregistration.
+We do this in two ways, firstly, by running a series of structured and collaborative activities in a workshop setting to reflect on our own research experiences in order to capture scientific workflows as they occur practice.
 And secondly to, use a real-world research problem as a case-study to test and evaluate the templates for utility and feasibility in practice.
 
-The workshop activities aimed to help us:
+\*\*Defining the purpose of preregistration\*\*
+
+Pu et al's key finding:
+
+\- different users have "different purposes for adopting preregistration, and these differ from the original purpose of distinguishing exploratory and confirmatory findings"
+
+Pu et al found that because different users ascribed different purposes to preregistration, it resulted in conflicting designs of preregistration formats, which they argue may ultimately undermine the overarching goal of preregistration: "improving the reliability of scientific findings".
+
+We state the intended purposes of the preregistration template up front:
+
+\- P**rimary purpose:** "Delimiting flexibility" Following from Pu et al. we define the primary purpose of preregistration for modelling exercises so as to
+
+We take emphasis *off* of the dichotomy of exploratory vs.
+confirmatory research, because I think these categories are confused and confusing in modelling, and actually unhelpful.
+Instead I want to focus on delimiting researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and the potential for resultant QRPs, such as, cherry-picking, HARKing, etc.
+etc.
+Link to my other manuscript.
+
+\- **Secondary purpose: increasing transparency** \<insert refs to modelling reporting literature - it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced\> Thus, want to improve the transparency of decision-making and methodology.
+What is "research transparency" 'Lupia and Elman define *research transparency* to be production transparency (open access data or data collection procedures) and analytic transparency ("a full account of how they draw their analytic conclusions from the data")' --- Lupia and Elman 2014.
+Pu et al.
+page 13.
+
+\*\*Acknowledging different use-cases\*\*
+
+Pu et al:
+
+\- norms around how people use preregistration are uncertain, and vary among users.
+
+\- Depending on what role you assume, there are different 'use-cases' for authors, reviewers/editors. BUT, these different use cases are not explicitly designed for.
+
+Our use-cases:
+
+1\. As creators
+
+2\. As users:
+
+### Identifying Ecological Modelling Workflows in Practice ###
+
+On the 4^th^ March 2020, we ran a series of structured activities designed to inform the development of two pre-registration templates, one for field-work and sampling design, the other for modelling and data-analysis.
+We split the activities into two groups, to better focus discussion for each of these templates.
+All workshop materials and workshop outputs can be found on the OSF component here \<LINK\>.
+
+The collaborative workshop activities aimed to help us:
 
@@
 c) Use this knowledge to inform the design of the templates so they meet their intended purpose and are taken up by researchers.
-
-\*\*
-
-### Identifying Ecological Modelling Workflows in Practice ###
-
-On the 4^th^ March 2020, we ran a series of structured activities designed to inform the development of two pre-registration templates, one for field-work and sampling design, the other for modelling and data-analysis.
-We split the activities into two groups, to better focus discussion for each of these templates.
-All workshop materials and workshop outputs can be found on the OSF component here \<LINK\>.
 
@@
 
+### Preregistration Items ###
+
+This will be a box or a table with numbered items, with just the preregistration item directives or question.
+The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
+
 ### Iterative Model Development & "Living" Preregistrations ###
@@
 
-*\<div style="width: 960px; height: 720px; margin: 10px; position: relative;"\>\<iframe allowfullscreen frameborder="0" style="width:960px; height:720px" src="https://app.lucidchart.com/documents/embeddedchart/9c4480bc-8465-4731-83e6-d24d7d2b7818" id="K4J1KZqFLi3u"\>\</iframe\>\</div\>*
+\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
+as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
+"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
+And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
+Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
+How can we facilitate the process?
+They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
+Maybe we could do something else, rather than get the paper to refer to the parts of the preregistration (but I think we should do this with Chris anyway).
+If we propose the norm of using GitHub to create and check the preregistration., I think this could be a good way of getting around this checking problem.
+At the very end we have a 'how to box' targeted at authors.
+And then maybe we have a box targeted at reviewers?
+Not sure, maybe just focus this paper on authors and the process of creating and using a preregistration.
+
+In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
+However, I don't expect users to agree on this process.
+Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
+
+![Living Preregistration Workflow](./figures/LivingPreregistration.png)
+
+### Shiny App ###
+
+Speak back to the the predefined use-cases, how does our proposed methodology and its template address the intended use-cases and purposes?
+
+The shiny-app, allows us to tailor the form firstly to authors and analysts.
+With help-text, and descriptions of the modelling phase, and step under consideration, as well as definitions and references to supporting resources ---  such as decision-tools for choosing and justifying methodological choices.
+
+By being able to 'generate a report' - we aim to target reviewers/editors, to allow for a condensed version of the template to be exported.
+Importantly, Pu et al show that "not all preregistered studies confirm to their preregistrations, or are inconsistent in reporting deviations", thus by making the preregistration available to reviewers, and by providing a platform and method for linking the actual analysis under-taken, we can allow reviewers to properly evaluate the preregistration against the final paper.
+(AND, must talk back to original aim of bad reporting practices in modelling).
+Moreover, if you combine formatting changes with our proposed method with GitHub, we hope to tailor the presentation of the completed preregistration + analysis, such that actual analyses and decisions can be linked directly to both files and outcomes.
 
@@
 
-In order to address tis confusion among norms, I state the expected norms that the template has been designed around here.
+In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
 However, I don't expect users to agree on this process.
@@
 
-\- **Possibly:** As a way of providing better decision-making to authors.
+\- **Front-heavy research timeline:** As a way of providing better decision-making to authors.
 By shifting decision-making to the front.
@@
 For reviewers 'participant reports that pregistration makes the review process "easier".'
+
+**DEALING WITH CONFLICTING NORMS IN CREATING & USING the PREREGISTRATION TEMPLATE**
 --------------------------

### 2020-05-25 index.Rmd
Update manuscript #20 -- inner conceptual model of template, iterative cycle of model dev.

@@
 
-**Keywords:**
+**Keywords: preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?)**
 
@@
 \- different users have "different purposes for adopting preregistration, and these differ from the original purpose of distinguishing exploratory and confirmatory findings"
+
+Pu et al. "Current designs represent a failure of designers to make their internal model of preregistration clear to users: the purpose of each question, and indeed the entire form, is often opaque."
 
@@
 
-## PRT Conceptual Model and Philosophy underpinning model development ##
+## A methodology for 'Living Preregistrations' ##
 
-### Model Scientific Workflow ###
+Failing to make the 'internal model of preregistration' clear to users is a critical source of confusion among preregistration template users that may impede preregistration uptake, and undermine the ability of the template to achieve its objectives (Pu et al.
+2020).
+Consequently, in this section we articulate the inner conceptual model of preregistration driving the content, structure and our proposed procedure for implementing the template.
 
-What model workflow and set of activities does the template presume?
+\*\*Modelling workflows for Ecology and Conservation\*\*
 
-This section describes the 'model scientific workflow' underpinning  the Preregistration Template.
+Table X, below, describes the 'model scientific workflow' underpinning the template, which was identified through the structured collaborative workshop, literature review and qualitative coding.
+By articulating this workflow, we explicitly state the workflow and sets of activities that the template presumes.
+We identified and coded decision-points into 6 'phases' in the process of developing, evaluating and analysing a model or models in ecological application: 'problem formulation', 'define conceptual model', 'formalise and specify model', 'Model Validation and Evaluation', 'Model Analysis', 'Model Application'.
+These phases are comprised of a series of smaller 'steps' and 'substeps'.
+These are defined and described in detail within Table X.
+In terms of how this translates into the format of the preregistration template, these broad phases and steps are used to structure the template, and the descriptions and definitions of the phases and steps are used as 'help-text' in the template interface.
 
-I want to include a figure / diagram here visualising and describing some usual / common or even idealised workflow to accompany both the PRT for
+@TODO - anything about *applied contexts??* Aligning with management? (In terms of the structure and steps above).
 
-A) PRT users.
+\*\*Iterative Cycle of Model Development\*\*
 
-B) To Explain the inner conceptual model driving the PRT (as per Pu et al.).
+The iterative nature of model development was identified in our structured discussions as a key barrier to adopting preregistration.
+\<Insert refs to literature describing this\>.
+\<Explain what I mean by iterative model development and provide examples\>
 
-*How does this inform the PRT? These broad phases are used to structure the preregistration template.*
+Typically, the process of preregistration is completely distinct from and precedes the implementation of the analysis plan laid out in the preregistration.	In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" (Nosek et al.
+2018) or prior to analysing the data (Ref, or link to secondary data preregistration template).
+This results in a linear research workflow, whereby a researcher shifts from the process of ideation and design, documentation of that design in a preregistration, to the process of action, where the researcher executes the plan specified in the preregistration \<insert link to this fig..\>
 
-### Preregistration Items ###
+Translating the preregistration process to a modelling research context is thus challenging because the iterative nature of model development conflicts with the inner logic of existing preregistration templates that presume a linear research workflow.
+For instance, some modelling decisions are inherently dependent on the outcomes and results of previous decisions and analyses, which involves conducting some preliminary or investigatory analyses before being able to specify future decision-steps in an analysis plan, contravening the critical(?adj) feature of preregistration of specifying the analysis plan prior to seeing the data.
+Moreover, some items in the template, particularly at later stages of modelling, might not be able to be answered by the researcher until the model is fully or close to fully specified on a first-pass (e.g. specifying exactly what sensitivity analyses or uncertainty analyses will be conducted).
+This might especially hold true for more complex process models.
+\<Mention that this is not a flaw in preregistration, but rather reflects the inner logic of preregistration for hypothesis-testing contexts\>
 
-This will be a box or a table with numbered items, with just the preregistration item directives or question.
-The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
+\*\*The \*checking\* problem\*\*
+
+Model Checking:
+
+"Existing preregistration formats are designed primarily for authors to quickly input the desired information, but may verge on being write-only media: they are ill-suited to the checking task that reviewers, paper readers, and in some cases even authors themselves might perform.
+Thus, we recommend exploring designs that (1) make relevant preregistration content easier to *query* and (2) encourage wider *coverage* of preregistration content during review." Pu et al.
+
+\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
+as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
+"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
+And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
+Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
+How can we facilitate the process?
+They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
+Maybe we could do something else, rather than get the paper to refer to the parts of the preregistration (but I think we should do this with Chris anyway).
+If we propose the norm of using GitHub to create and check the preregistration., I think this could be a good way of getting around this checking problem.
+At the very end we have a 'how to box' targeted at authors.
+And then maybe we have a box targeted at reviewers?
+Not sure, maybe just focus this paper on authors and the process of creating and using a preregistration.
+
+In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
+However, I don't expect users to agree on this process.
+Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
+
+@TODO insert table from atlas TI.
 
@@
 
-Paragraph or two describing iterative nature in model development. This was a common theme arising in our discussions and one of the key 'roadblocks' identified.
+Given this defining feature of the modelling process, it is critical that the design of the interface and procedure for implementing the preregistration inherently capture this iterative nature of development.
+We attempt to do this by...
+
+Capturing this inner logic has repercussions for
+
+Thus deciding on the inner logic of the template, and what is in and out of scope has been particularly difficult.
+
+One method around this is to accept that the current model of preregistration is not going to work for iterative model development.
+Again, the more complex the model, the more likely this is to be the case.
+Consequently I propose a different model of preregistration that I think will work for modeling research contexts.
+At the moment I'm calling it 'living preregistration' but please suggest another name.
+Others have termed this 'adaptive preregistration', and this is maybe a better name.
+
+In this model of preregistration, all questions that can be answered, should be answered before any data analysis or modelling proceeds.
+For some items, particularly during model evaluation and analysis, perhaps only responses of a general nature can be provided.
+That is OK.
+As the modelling proceeds.
+Further detail can be provided and the next phase of analysis can be preregistered.
+Where the results of model evaluation and analysis reveal that there are problems with the model, or perhaps assumption violations, plans can be changed, and again the next phase of analysis can be preregistered.
+
+A static document for this type of process doesn't really make sense.
+I propose leveraging the code and project management tools within the GitHub platform as a way of hosting a 'living' or 'adaptive' preregistration.
+Because GitHub documents are 'time-stamped', the genesis of the preregistration from one version to the next can be traced and tagged with versions.
+Since the document would exist in the repository with the code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated plan.
 
@@
 
-\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
-as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
-"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
-And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
-Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
-How can we facilitate the process?
-They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
-Maybe we could do something else, rather than get the paper to refer to the parts of the preregistration (but I think we should do this with Chris anyway).
-If we propose the norm of using GitHub to create and check the preregistration., I think this could be a good way of getting around this checking problem.
-At the very end we have a 'how to box' targeted at authors.
-And then maybe we have a box targeted at reviewers?
-Not sure, maybe just focus this paper on authors and the process of creating and using a preregistration.
+![Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
 
-In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
-However, I don't expect users to agree on this process.
-Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
+### Preregistration Items ###
 
-![Living Preregistration Workflow](./figures/LivingPreregistration.png)
+This will be a box or a table with numbered items, with just the preregistration item directives or question.
+The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
 
@@
 Moreover, if you combine formatting changes with our proposed method with GitHub, we hope to tailor the presentation of the completed preregistration + analysis, such that actual analyses and decisions can be linked directly to both files and outcomes.
-
-### Pu et al. ###
 

### 2020-05-26 analysis/preregistration_template.Rmd
#10 HANNAH feedback
Feedback up to section 1.2.3

@@
 However, translating the preregistration process to a modelling research context is not so clear cut because some decisions are inherently dependent on the outcomes and results of previous decisions and analyses.
-For example, some items at the later stages of modelling probably can't be answered until the model itself is fully or close to fully developed, especially for more complex process models (e.g.
-specifying exactly what sensitivity analyses or uncertainty analyses will be conducted).
-Thus deciding on the inner logic of the template, and what is in and out of scope has been particularly difficult.
+For example, which sensitivity analyses will be conducted will depend on the relationships between variables found during earlier earlier stages in the modelling process.
 
@@
 Again, the more complex the model, the more likely this is to be the case.
-Consequently I propose a different model of preregistration that I think will work for modeling research contexts.
-At the moment I'm calling it 'living preregistration' but please suggest another name.
-Others have termed this 'adaptive preregistration', and this is maybe a better name.
+Consequently I propose an 'adaptive preregistration' model  that I think will work for modeling research contexts.
 
@@
 That is OK.
-As the modelling proceeds.
-Further detail can be provided and the next phase of analysis can be preregistered.
+As the modelling proceeds, further detail can be provided and the next phase of analysis can be preregistered.
 Where the results of model evaluation and analysis reveal that there are problems with the model, or perhaps assumption violations, plans can be changed, and again the next phase of analysis can be preregistered.
@@
 A static document for this type of process doesn't really make sense.
-I propose leveraging the code and project management tools within the GitHub platform as a way of hosting a 'living' or 'adaptive' preregistration.
+I propose leveraging the code and project management tools within the GitHub platform as a way of hosting an 'adaptive' preregistration.
 Because GitHub documents are 'time-stamped', the genesis of the preregistration from one version to the next can be traced and tagged with versions.
@@
 # 1. Problem Formulation
+***HF comment: I think this paragraph makes things more rather than less confusing - and I wonder if its necessary given that each point is stepped through below. I would recommend removing it***
 
-*Specifies the decision-making context in which the model will be used.
+*Specifies the decision-making context in which the model will be used or the intended scope and context of conclusions.
 Important components include the decision maker and stakeholders (including experts) and their view on: i) the nature of the problem or decision addressed and how the scope of the modelling tool fits within the (broader) context (i.e.
-model purpose; ii) the relevant spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they forsee unnacceptable outcomes that need to be represented in the model (i.e.
+model purpose; ii) the spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they forsee unnacceptable outcomes that need to be represented in the model (i.e.
 as constraints), and; vi) what future scenarios does the model need to account for (noting this may be revised later).
@@
 *Defining the purpose of the model is critical because the model purpose influences choices at later stages of model development [@Jakeman:2006ii].
-Common model purposes in ecology include: gaining a better qualitative understanding of the target system, knowledge synthesis and review, and providing guidance for management and decision-making [@Jakeman:2006ii] (could also see Addison paper on the use of models).
+Common model purposes in ecology include: gaining a better qualitative understanding of the target system, synthesising and reviewing knowledge, and providing guidance for management and decision-making [@Jakeman:2006ii] (could also see Addison paper on the use of models).
 Note that modelling objectives are distinct from the analytical objectives of the model.*
@@
 *The scope of the model, including temporal and spatial resolutions are defined here [@Mahmoud2009], and any limitations on model development analysis and flexibility should be outlined here [@Jakeman:2006ii].
-Note that the modelling context is different from the problem context.*
+Note that the modelling context is different from the problem context which are described in 1.3.1.*
 
@@
 
+***HF Comment: this is lovely, concise and informative***
+
 ### 1.2.2 Model Scope and Scale
@@
 Where is the boundary of the modelled system?
-Everything beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al.
-2006).The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?
+Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al.
+2006) ***HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me.
+Also, is this question more to do with describing the scope and limitations of the data rather then the model?***.
+The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?
+
 Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
-(Should this last step go here or within define conceptual framework?).
-What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)?
-This is relevant to model transferability...
-(similar to a COG statement..)*
+(Should this last step go here or within define conceptual framework?.
+***HF comment.
+I think it belongs with the conceptual framework***).
+
+What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)? This is relevant to model transferability... (similar to a COG statement..)* 
 A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation.
@@
 
-*What effort and resources are available for both developing the model and operating the model?*
+*What effort and resources are available for both developing the model and operating the model?* ***HF comment; as written this might be slightly confronting because you could interpret this to include writing explictly about budget constraints which might be confidential.
+Maybe this could be alleviated with with an example?***
 
-*What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?*
+*What degree of flexibility is required from the model?
+Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?* ***HF comment: love the inclusion of this question***
 

### 2020-05-26 index.Rmd
#20 update manuscript - write UX design section

@@
 ---
-title: "EcoConsPreReg"
+title: "Preregistration for modelling in ecolgy and conservation"
 description: |
-  A new article created using the Distill format.
+  Working Draft.
 author:
@@
 ---
+
+Intended journal: Methods in Ecology and Evolution.
+
 The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements:
@@
 
-**Keywords: preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?)**
+**Keywords: preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?) (no more than 8 for MEE).**
 
@@
 
-\*\*About the Approach\*\*
+**Designing Preregistration Templates with a User-Centred Design Approach**
 
@@
 We do this in two ways, firstly, by running a series of structured and collaborative activities in a workshop setting to reflect on our own research experiences in order to capture scientific workflows as they occur practice.
-And secondly to, use a real-world research problem as a case-study to test and evaluate the templates for utility and feasibility in practice.
+And secondly, by using a real-world research problem as a case-study to test and evaluate the templates for utility and feasibility in practice.
 
-\*\*Defining the purpose of preregistration\*\*
+@TODO what does a UX design approach entail... stating purpose, use-case and contexts a priori? Explain.
 
-Pu et al's key finding:
+**Intended Purpose of this Preregistration**
 
-\- different users have "different purposes for adopting preregistration, and these differ from the original purpose of distinguishing exploratory and confirmatory findings"
+Defining the purpose and intended use-cases of a preregistration template has important implications for the design of the template --- from its structure, content, user-interface and use within the broader research workflow.
+We define the primary purpose of the template as "to delimit researcher degrees of freedom" when conducting model-based research in applied ecology and conservation.
+Although others have defined the purpose of preregistration to include 'delineating between *exploratory* from *confirmatory* research' \<refs\>, this is not the intended purpose of the template here.
+Beyond hypothesis-testing contexts, this is a simplifying dichotomy that, in practice, is more blurred and nuanced than the ideal.
+For instance researchers might move back and forth between different research-modes throughout the modelling cycle.
+Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, perhaps with the aid of some philosophers of science.
+Consequently, we wish to deemphasise the purpose of distinguishing between exploratory and confirmatory research, and focus the design of the templates in being able to delimit researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and other questionable research practices, like cherry-picking, HARKing etc.
+\<insert ref to my QRP manuscript\>.
 
-Pu et al. "Current designs represent a failure of designers to make their internal model of preregistration clear to users: the purpose of each question, and indeed the entire form, is often opaque."
+The secondary purpose of the model is to increase *research transparency* of model-based research in ecology and conservation.
+*Research transparency* refers to both production transparency --- which includes research artefacts like open-access data and materials, or data collection procedures --- and analytic transparency --- which describes a complete account of how the analytic conclusions are drawn from the data (Lupia and Elman, 2014).
+\<refs to model reporting literature- it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced\>
 
-Pu et al found that because different users ascribed different purposes to preregistration, it resulted in conflicting designs of preregistration formats, which they argue may ultimately undermine the overarching goal of preregistration: "improving the reliability of scientific findings".
+**Preregistration template users and research-contexts**
 
-We state the intended purposes of the preregistration template up front:
+Given the stated purposes of the preregistration templates we designed, it implies two types of users: creators of preregistrations using the template, and reviewers of a completed template, including reviewers, journal editors and any readership.
+The content and structure of the items on the template will be designed with the preregistration-creator and the preregistration readership in mind - it will capture critical decision-points that may influence the overall model and analysis and recommendations or decisions that the model informs, thus delimiting researcher degrees of freedom at these junctures in the modelling decision process.
+It should also enhance the research transparency of the study, ensuring that these choices are documented with sufficient transparency to let the reader make informed decisions about the credibility and reliability of the model and associated analysis and interventions.
 
-\- P**rimary purpose:** "Delimiting flexibility" Following from Pu et al. we define the primary purpose of preregistration for modelling exercises so as to
+More specifically, we intend this template to be suitable not just in 'pure-research'-only contexts, but to be suitable for ecologists and conservation decision-makers who are working in the 'actionable sciences' \<which guy was this again..
+Sutherland??
+define\>.
+The template should be adequate for use by scientists and modellers working in applied ecological contexts, but also for management and conservation decision-making contest where there are real-world decisions hinging on the model-based research.
 
-We take emphasis *off* of the dichotomy of exploratory vs.
-confirmatory research, because I think these categories are confused and confusing in modelling, and actually unhelpful.
-Instead I want to focus on delimiting researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and the potential for resultant QRPs, such as, cherry-picking, HARKing, etc.
-etc.
-Link to my other manuscript.
+\<Research Types: Modelling - what do we mean by modelling??
+I think if we don't specify this then people will just think 'classical linear regression' as modelling, and then will say, well why don't use existing hypothesis testing templates, like AsPredicted.
+I need to specify some particularities about the type of modelling that occurs in in these research contexts, and why existing templates don't suit, and that we need to design specifically for these research contexts.\>
 
-\- **Secondary purpose: increasing transparency** \<insert refs to modelling reporting literature - it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced\> Thus, want to improve the transparency of decision-making and methodology.
-What is "research transparency" 'Lupia and Elman define *research transparency* to be production transparency (open access data or data collection procedures) and analytic transparency ("a full account of how they draw their analytic conclusions from the data")' --- Lupia and Elman 2014.
-Pu et al.
-page 13.
+**Use-Cases**
 
-\*\*Acknowledging different use-cases\*\*
+Depending on what role is assumed, there are different 'use-cases' for authors, reviewers/editors *Pu et al.* However, these use-cases are not typically explicitly designed for.
+For example, reviewers and editors (and sometimes authors) face the problem of 'checking' --- that is, 'the process of comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified.' Existing preregistration formats are designed with the role of the author in mind --- they are text-based documents, that are ill-suited to checking the reported study against the preregistration.
 
-Pu et al:
-
-\- norms around how people use preregistration are uncertain, and vary among users.
-
-\- Depending on what role you assume, there are different 'use-cases' for authors, reviewers/editors. BUT, these different use cases are not explicitly designed for.
-
-Our use-cases:
-
-1\. As creators
-
-2\. As users:
+Consequently, we aim not just to simply design a new template, deciding on its structure and content, but we propose a workflow and methodology to be implemented by the author when completing a preregistration of their model-based research \<see section - inner logic of the model - discuss features model-based research, like iterative development?\>
 
 ### Identifying Ecological Modelling Workflows in Practice ###
+
+What attributes do we aim for the template
+
+A) an interface that ensures ease of use for the creator of a preregistration
+
+B) structure and content that is necessarily opinionated..
+forces user to a particular workflow, so getting that workflow right, that reflects both norms and practice of model-based research is pivotal.
 
@@
 
+@TODO Insert screenshot of app.
+
+@TODO link to shiny app.
+
+As *Pu et al* find, norms around preregistration are both uncertain and vary among users.
+By norms, we mean \<insert definition from Pu\>.
+Although the design of the template in terms of the content and structure is *opinionated*, in that it forces a particular view of the modelling research workflow, people are free to modify the template according to their needs.
+\<The preregistration infrastructure we provide here allows for people to modify the template easily - set up the infrastructure of the shiny app to take in a CSV file, which is then exported to a JSON file for creating the shiny app (alt, just modify the JSON directly)\>
+
 What is the purpose of preregistration? Specifically in modelling.

### 2020-05-26 index.Rmd
#20 Add authors, add text from proposal to introduction

@@
   - name: Elise Gould 
+    affiliation: School of BioSciences, University of Melbourne
+  - name: Chris Jones
+    affiliation: Department of Environment Land and Water
+  - name: Libby Rumpff
+    affiliation: School of BioSciences, University of Melbourne
+  - name: Hannah Fraser
+    affiliation: School of BioSciences, University of Melbourne
+  - name: Fiona Fidler
     affiliation: School of BioSciences, University of Melbourne
@@
 
+Failure to reproduce a large proportion of published studies in psychology and medicine has received considerable attention and provoked heated discussion among researchers in the broader scientific community.
+Initial meta-research in ecology has revealed that the conditions found to foster irreproducibility are present (Fidler et al.
+2017), evidencing that the discipline is also at risk of a "reproducibility crisis," with rates of Questionable Research Practices in ecology similar to those in other fields (Fraser et al.
+2018).
+Ecology has seen recent improvements in journal policy concerning reporting guidelines and data/code archiving, but there has been slow uptake of other important tools, like preregistration and registered reports (Parker, Fraser, and Nakagawa 2019).
+
+Preregistration is designed to mitigate the opportunistic misuse of researcher degrees of freedom, or "Questionable Research Practices" (QRPs, Wicherts et al.
+(2016)).
+Researcher degrees of freedom refer to the many analytic decisions a researcher makes during the planning, data collection, data analysis and reporting of results in scientific studies (Wicherts et al.
+2016; Simmons, Nelson, and Simonsohn 2011).
+Examples of these analytic choices include whether to exclude observations, whether more data should be collected, which set of conditions should be combined and compared, which model evaluation metric should be used, etc.
+Coupled with a lack of transparency -- or "undisclosed flexibility" -- researcher degrees of freedom are vulnerable to being exploited by researchers because of cognitive biases.
+
+Within basic ecological research employing null hypothesis significance testing, QRPs may increase the probability of false-positive findings (Seaman and Weber 2015; Earp and Trafimow 2015), and may spuriously increase effect size estimates (Wicherts et al.
+2016; Parker et al.
+2016).
+QRPs in model-based research may "artificially increase the accuracy or precision of a model, its predictions, and or/evaluation tests to the effect that the model is perceived to be more credible than it would be if the QRP did not occur" (Gould et al.
+2019).
+They are "practices that lead to consumers of the model placing false belief in the reliability, validity and utility of the model than would be warranted without the QRP" (Gould et al.
+2019).
+Some examples of questionable research practices in model-based research include cherry-picking and 'statistic-hacking' model evaluation and test results in order to overemphasise the model's performance and tweaking model priors after fitting the model to data and checking results (Gould, n.d.).
+
+Preregistration requires the methods and analysis plan of a future study are registered in a secure and accessible platform prior to any data collection, and after submission it can no longer be altered.
+Preregistration guards against questionable research practices, by making clear what is a genuine a priori planned analysis, and what is not (Parker, Fraser, and Nakagawa 2019).
+Different research methodologies, and their specific and nuanced application across different domains and fields of study follow different analytical choices and are therefore susceptible to different QRPs.
+Consequently, domain-specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom unique to a particular domain-specific methodology.
+Moreover, we believe that a lack of discipline-specific preregistration templates is a key impediment to the uptake of these Open Science tools.
+
+------
+
 Problem: "Conceptual barriers to adopting preregistration, including users from different disciplines feeling that preregistration is not for them" Pu et al. (2019).
@@
 
-Research Aim: To develop a Template for Ecology and Conservation.
-
-But not just a pre-registration template, we take into account the iterative nature of model development
-
-Objectives of the Template itself & internal model of preregistration (this will build upon previous work of mine, namely the QRP chapter.
-See also Pu et al.
-for the four 'purposes' of preregistration that they identified in their analysis.
-
-----
-
-## Methods ##
+There are many templates for preregistering experimental research, particularly for studies employing null-hypothesis significance testing.
+However, to date there is only a single published template for pre-registering model-based studies, and it is for cognitive modelling in psychology (Crüwell and Evans 2019).
+Consequently, this project aims to develop a template for preregistering ecological modelling research.
+In this project we aim to develop a template and new methodology for implementing preregistration of ecological models in applied ecology.
+This research builds on recently developed "QRP roadmaps" (Gould et al.
+2019) and research by Cruwell et al.
+(2019) showing that existing templates are too general, and do not account for domain-specific and research category-specific nuances and questionable research practices.
+Template development and testing will be driven by a real-world case study of environmental-flow modelling.
 
@@
 Consequently, we aim not just to simply design a new template, deciding on its structure and content, but we propose a workflow and methodology to be implemented by the author when completing a preregistration of their model-based research \<see section - inner logic of the model - discuss features model-based research, like iterative development?\>
+
+----
+
+## Methods ##
 

### 2020-05-27 index.Rmd
Update Manuscript #20 - write section on user interface design and draft shiny app box

@@
 ---
+**Intended journal:** Methods in Ecology and Evolution.
 
-Intended journal: Methods in Ecology and Evolution.
+*The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements:*
 
-The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements:
+*Point 1: set the context for and purpose of the work;*
 
-Point 1: set the context for and purpose of the work;
+*Point 2: indicate the approach and methods;*
 
-Point 2: indicate the approach and methods;
+*Point 3: outline the main results;*
 
-Point 3: outline the main results;
+*Point 4: identify the conclusions and the wider implications.*
 
-Point 4: identify the conclusions and the wider implications.
-
-**Keywords: preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?) (no more than 8 for MEE).**
+**Keywords** *(no more than 8 for MEE).***:** *preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?).*
 
@@
 Given this defining feature of the modelling process, it is critical that the design of the interface and procedure for implementing the preregistration inherently capture this iterative nature of development.
-We attempt to do this by...
+Current formats for preregistration are inappropriate given the iterative nature of the model development cycle, and the need for checking the preregistration against the reported model and analysis.
+Existing templates are text-based documents wherein authors respond to template items to complete the preregistration.
+They are standalone documents, with no way of linking the specified decision points to the actual analysis - thereby making model checking extremely tedious.
+
+**Preregistration Interface**
+
+We felt that the length and presence of large chunks of help-text within the template we developed would make using the template to create a preregistration for a study cumbersome, and might act as a barrier for model-based researchers to adopt preregistration.
+We considered removing in-template text describing and defining key concepts, phases and steps, as well as reducing the number of decision-points in a model-based research workflow captured by the template by removing items within the template.
+\<@TODO - did we end up removing any items by the end of the preregistration and evaluation with Chris' study?\> While we 'tweaked' the template items and descriptive text during the in situ evaluation of the templates for the environmental flows preregistration, we decided not to alter the template structure and content significantly, but to instead implement a new format and interface for working with the template.
+We outline these reasons below.
+
+1\. Opinionated Design - norms around modelling practice.  'Good Modelling Practice' Literature. This also has the benefit of aiding new modellers.
+
+2\.
+A related point: linguistic ambiguity and other problems with modelling concepts and terms.
+Contested terms in the ecological modelling literature about what each different step is, their activities, etc.
+
+3\.
+*research transparency*.
+We felt that the decision-points captured in the template are critical to increasing reporting transparency sufficiently for others to properly evaluate the modelling and analysis.
+
+Consequently, we instead decided to implement a user-friendly interface that makes the completion of a preregistration with our template much more *approachable\<?adj\>*.
+
+**Living Preregistration Workflow**
+
+2\. Workflow - allows for iterative model development AND
+
+Why, what is the current form? Existing templates are read-only documents.
 
@@
 
-![Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
-
-### Preregistration Items ###
-
-This will be a box or a table with numbered items, with just the preregistration item directives or question.
-The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
-
-### Shiny App ###
-
-Speak back to the the predefined use-cases, how does our proposed methodology and its template address the intended use-cases and purposes?
-
-The shiny-app, allows us to tailor the form firstly to authors and analysts.
-With help-text, and descriptions of the modelling phase, and step under consideration, as well as definitions and references to supporting resources ---  such as decision-tools for choosing and justifying methodological choices.
-
-By being able to 'generate a report' - we aim to target reviewers/editors, to allow for a condensed version of the template to be exported.
 Importantly, Pu et al show that "not all preregistered studies confirm to their preregistrations, or are inconsistent in reporting deviations", thus by making the preregistration available to reviewers, and by providing a platform and method for linking the actual analysis under-taken, we can allow reviewers to properly evaluate the preregistration against the final paper.
@@
 Moreover, if you combine formatting changes with our proposed method with GitHub, we hope to tailor the presentation of the completed preregistration + analysis, such that actual analyses and decisions can be linked directly to both files and outcomes.
-
-@TODO Insert screenshot of app.
-
-@TODO link to shiny app.
-
-As *Pu et al* find, norms around preregistration are both uncertain and vary among users.
-By norms, we mean \<insert definition from Pu\>.
-Although the design of the template in terms of the content and structure is *opinionated*, in that it forces a particular view of the modelling research workflow, people are free to modify the template according to their needs.
-\<The preregistration infrastructure we provide here allows for people to modify the template easily - set up the infrastructure of the shiny app to take in a CSV file, which is then exported to a JSON file for creating the shiny app (alt, just modify the JSON directly)\>
-
-What is the purpose of preregistration? Specifically in modelling.
-
-The original purpose of preregistration is "the action of confirming an unalterable version of one's research plan prior to collecting data", such that the act distinguishes preregistered, confirmatory findings from exploratory findings (Nosek et al.
-2018 + Kidwell et al.
-2016) by way of constraining 'researcher degrees of freedom' or flexibility in both experimental design and data analysis, which has been posited as one of the key reasons contributing to the replication crisis (Ioannidis, 2005, Wagenmakers, 2012).
-
-Pu et al.
-treat preregistration as a user-centred design problem to ask whether it achieves its proposed goals *in practice*, and how the process might be improved.
-They ask whether the current implementation of preregistration in practice, in terms of its design of the form, and how the process integrates into research practice and the peer review practice, effectively supports its intended goals.
-By interviewing academic psychologists, they attempt to characterise current understanding of practice, and both the norms and purposes for preregistration that actual users have.
-The key findings are:
-
-\- Norms around how people use preregistration are uncertain, and vary among users
-
-\- different users have "different purposes for adopting preregistration, and these differ from the original purpose of distinguishing exploratory and confirmatory findings"
-
-\- Depending on what role you assume, there are different 'use-cases' for authors, reviewers/editors. BUT, these different use cases are not explicitly designed for.
-
-\- There are "conceptual barriers to adopting preregistration, including users from different disciplines feeling that preregistration is not for them"
-
-Pu et al found that because different users ascribed different purposes to preregistration, it resulted in conflicting designs of preregistration formats, which they argue may ultimately undermine the overarching goal of preregistration: "improving the reliability of scientific findings"
-
-What do we define as the *purposes* of preregistration, specifically within ecological modelling contexts?
-
-\- P**rimary purpose:** "Delimiting flexibility" Following from Pu et al. we define the primary purpose of preregistration for modelling exercises so as to
-
-We take emphasis *off* of the dichotomy of exploratory vs.
-confirmatory research, because I think these categories are confused and confusing in modelling, and actually unhelpful.
-Instead I want to focus on delimiting researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and the potential for resultant QRPs, such as, cherry-picking, HARKing, etc.
-etc.
-Link to my other manuscript.
-
-\- **Secondary purpose: increasing transparency** \<insert refs to modelling reporting literature - it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced\> Thus, want to improve the transparency of decision-making and methodology.
-What is "research transparency" 'Lupia and Elman define *research transparency* to be production transparency (open access data or data collection procedures) and analytic transparency ("a full account of how they draw their analytic conclusions from the data")' --- Lupia and Elman 2014.
-Pu et al.
-page 13.
-
-\- **Possibly:** As a way of providing better decision-making to authors.
-By shifting decision-making to the front.
-Front-heavy research process.
-Delays the start of analysis, but results in better decision-making, because the decision-making is more carefully reasoned.
-But not sure want to set this out as a 'purpose' of the template, or just have this as a section in the reflection towards the end of the manuscript (will depend on how I set up the angle, and the structure of the paper).
-At the moment, this is my hypothesis, is that it can improve the process, but we'd need to reflect on this.
-And I think we could do that at the end of the manuscript.
-
-This relates to what Pu et al.
-found where participants 'felt their research workflow improved by having preregistration as a "forcing function" \[...\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers' degree of freedmen advance.' (Page 13).
-For reviewers 'participant reports that pregistration makes the review process "easier".'
-
-So what are *our* use-cases?
-
-The shiny-app, allows us to tailor the form firstly to authors and analysts.
-With help-text, and descriptions of the modelling phase, and step under consideration, as well as definitions and references to supporting resources ---  such as decision-tools for choosing and justifying methodological choices.
-
-By being able to 'generate a report' - we aim to target reviewers/editors, to allow for a condensed version of the template to be exported.
-Importantly, Pu et al show that "not all preregistered studies confirm to their preregistrations, or are inconsistent in reporting deviations", thus by making the preregistration available to reviewers, and by providing a platform and method for linking the actual analysis under-taken, we can allow reviewers to properly evaluate the preregistration against the final paper.
-(AND, must talk back to original aim of bad reporting practices in modelling).
-Moreover, if you combine formatting changes with our proposed method with GitHub, we hope to tailor the presentation of the completed preregistration + analysis, such that actual analyses and decisions can be linked directly to both files and outcomes.
-
-Preregistration *Norms*:
-
-Pu et al.'s working definition of *norms* is 'as a set of ideal behaviours'.
-A key finding was that norms around both creating and using a preregistration are uncertain.
-*How will my work address this confusion among researchers about what the norms are, how will it address uncertain norms?*
-
-Creating:
-
-\- E.g. what sorts of decisions should be preregistered? Conflicting ideas about whether a power analysis, should for example be included in a preregistration.
 
@@
 
+This will be a box or a table with numbered items, with just the preregistration item directives or question.
+The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
+
+### Shiny Preregistration ###
+
+The shiny app is intended as an interface for authors and analysts to respond to template items and input their planned decisions to complete a preregistration for their study.
+In addition to having directive text and questions for each item, there is also 'help-text' consisting of descriptions of the modelling phases and steps, as well as definitions and references to supporting resources, such as decision-tools for choosing and justifying methodological choices \<reference to image with shiny app\>
+
+Once the preregistration has been completed \<for the time being, it doesn't actually have to be complete, see 'living preregistration' section\>, the author is able to 'generate a report'.
+The report removes any supporting text in the user interface, and generates a text-based document, containing the preregistration template items and any responses to them by the author or analyst.
+This feature is targeted towards reviewers and editors, providing a more condensed version of the preregistration for greater ease of reading.
+
+Norms around preregistration are both uncertain and vary among preregistration users (Pu et al.) By norms, we mean \<insert definition from Pu\>.
+Given that the preregistration template is *opinionated* in that it presumes particular norms around the modelling research workflow and sets of activities, and decision-points to be reported, we have designed both the underlying infrastructure of the shiny app, and chosen to host the template and interface on GitHub to allow users to adapt any aspect of the template itself to be congruent with their personal needs and norms around model-based practice.
+
+The GitHub repository storing the template and infrastructure implementing the interface is open-source and able to be *forked* --- this means that a user is able to copy the entire contents of the repository to their own user account on GitHub, to be changed in any way they see fit.
+We have included design features in the infrastructure of the shiny app that allow for people to very easily change the structure and content of the template, that will then be rendered into the user-interface.
+\<Probably won't describe here, but could link to a vignette on the app that describes the following: easy modification - set up the infrastructure of the shiny app to take in a CSV file, which is then exported to a JSON file for creating the shiny app (alt, just modify the JSON directly)\>
+
+@TODO Insert screenshot of app.
+
+@TODO link to shiny app (make a doi for it).
+
+@TODO make an issue in the shiny repository for being able to 'load' a partially completed preregistration... Maybe we need to export in JSON?
+
+![Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
+
 ----
@@
 
-## talking points ##
+### talking points ###
 
@@
 
-\- "A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
+- "A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
 This trade-off was something we considered important during the design phase, and deciding what was in and out of scope for the preregistration was particularly challenging.
@@
 
-\- However, as Pu et al.
+- However, as Pu et al.
 note, some participants who were both authors and reviewers didn't think that the time-costs of preregistration in research were significantly greater than not preregistering research.
@@
 
-\- Competing objectives?
+- Competing objectives?
 Although the purposes of preregistration are not mutually exclusive - as we have decided the purpose of preregistration we have defined here is to both delimit/state flexibility, *and* be flexible.
@@
 
-\- **Front-heavy research timeline:** As a way of providing better decision-making to authors.
+- **Front-heavy research timeline:** As a way of providing better decision-making to authors.
 By shifting decision-making to the front.
@@
 **DEALING WITH CONFLICTING NORMS IN CREATING & USING the PREREGISTRATION TEMPLATE**
+
 --------------------------

### 2020-05-27 analysis/preregistration_template.Rmd
#10<Hannah>feedback
Rest of Hannah's feedback.
Overall, the prereg template looks fantastic! My main points are 

- remove the intro paragraph to sections that have subheading (e.g. no intro paragraph for section 3.4 if there is a 3.4.1.
- I am not sure about the bit about corroborating on a different dataset in addition to validation described in section 5

--- 2020-05-24 analysis/preregistration_template.Rmd
@@
 
-### 1.3.2 - Identify outcomes to evaluate decisions under scenarios
+### 1.3.2 - Identify how the model will tell you about your analytic objectives
 
 *The outcomes should speak directly to the analytical objectives identified in 1.3.1.
-Outcomes could be qualitative or quantitative, and include descriptive statistic measures, satisficing measures or regret measures (Moallemi et al.
-2019).
-For example, a model outcome in the top 10th percentile of worst conditions, or some minimum performance threshold for meeting a performance objective, or choosing a decision with the minimum regret or least amount of uncertainty.*
+Outcomes could be qualitative or quantitative.
+For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.*
 To be honest, I don't know how I might apply this to my work or any future works.
@@
 Like, how would I apply it to the woodland condition analysis?
+HF - Made the changes above in response to the comment.... do they help? If not, I'm struggling to understand exactly what this step is achieving.
 
-## 1.4 Define candidate actions decisions
+## 1.4 Define candidate decisions and method of identifying relevant management actions
 
 *Candidate decisions should be investigated and are specified a priori.
-Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019]. Describe the method used to identify relevant management actions.*
+Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019].¬†Describe the method used to identify relevant management actions.*
 
@@
 Scenarios should be set a priori (i.e.
-before the model is built, [@Moallemi2019]) and may be stakeholder-defined or expert judgment-driven [@Mahmoud2009].*
+before the model is built, [@Moallemi2019]) and may be stakeholder-defined or driven by the judgement of the modeller or other experts [@Mahmoud2009].*
 
@@
 Conceptual models may be representeed in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables (Jakeman, Letcher, and Norton, 2006).*
+HF- Love this step
 
@@
 This will likely depend on the method chosen to elicit the conceptual model.*
+HF- not sure if this step is required - I think we should expect people to have already done this bit before they start their preregistration...
+Or we could put it into 2.0 as an addendum - "if you haven't defined your conceptual model yet describe how you plan to do so."
 
-## 2.2 Specify key assumptions uncertainties
+## 2.2 Specify key assumptions and uncertainties
 
@@
 [@Grimm:2014es].
-The influence of particular theories, concepts, or importantly, earlier models, should be explained and justified against alternative conceptual design decisions that might lead to alternative model structures [@Grimm:2014es].
-Specify key assumptions and uncertainties underlying the model's design.
+The influence of particular theories, concepts, or importantly, earlier models, should be explained and justified against alternative conceptual design decisions that might lead to alternative model structures [@Grimm:2014es].*
+
+HF comment: the above part of 2.2 seems like a big ask. I think we want to achieve that we will need to break it down further. Would it be sufficient to just have the bit below?
+
+*Specify key assumptions and uncertainties underlying the model's design.
 Describe how uncertainty and variation will be represented in this model.
@@
 
-## 2.3 Define conceptual model: identify predictor response variables
+## 2.3 Identify predictor and response variables
 
@@
 
+HF- love this step
+
 ## 2.4 Define prior knowledge, data specification and evaluation
@@
 *Collect, process and prepare data available for parameterisation, determining model structure, and for scenario analysis.*
+HF - this isn't super clear to me. Maybe we need more clarification here or an example
 
@@
 
-*Document the identity, quantity and provenance of any data used to develop, identify and test the model.
-Describe how the data is arranged, in terms of replication and covariates.Explain how you will summarise what data and knowledge will be used to design and parameterise the model, including references, data sources, and information about whom, when and where those data were collected [@Grimm:2014es].*
+*Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model. Describe how the data is arranged, in terms of replicates and covariates.*
 
-**For pre-existing data: -- as appropriate.**
+**For pre-existing data: -- delete as appropriate.**
 
@@
 (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
-- *Codebook. If a Codebook is available, please link to it here and / or upload the document.*
+- *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
 - *Prior work based on this dataset - Have you published / presented any previous work based on this dataset?
-Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont his dataset you have worked on.*
-- *Prior Research Activity - Have you worked with these data before? Describe any prior but unpublished research activity using these data. Be specific and transparent.*
-- *Prior knowledge of the current dataset - Describe any prior knowledge of the dataset. Be specific.*
+Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont this dataset you have worked on.* HF- Love this question
+- *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
+- *Prior knowledge of the current dataset - Describe any prior knowledge of the dataset.
+Be specific.* HF- not sure how someone would answer this question.
+Maybe we need a more specific Q or an example
 
@@
 
-- Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected.
+- Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged.
 Describe any inclusion or exclusion rules, and the study timeline.
@@
 
-*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.*
+*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.* HF - Not sure exactly what this means
 
@@
 Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.*
+HF- should we include something here asking people to specify what they intend to find out from the data exploration?
 
@@
 
-*Describe how you will evaluate the quality and sources of both numerical and qualitative data that available for model parameterisation, and for determining the overall model structure [@Grimm:2014es].
-Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].
-Document any issues with data reliability.
+*Describe how you will determine how reliable the data is for the given purpose.
+Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].*
+
+*Document any issues with data reliability.
 This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].*
@@
 
-*Describe how the model will be critically evaluated.
-Evaluation includes both the overall model structure, and any submodels if relevant.
-How will any simplifying assumptiosn be critically assessed[@Augusiak:2014gz]? Explain whether this process will include consultation or feedback from a client, manager, or model user.*
+*Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.*
+
+*How will any simplifying assumptions be critically assessed [@Augusiak:2014gz]?* HF- I'm not sure how to answer this one. Maybe an example would help?
+
+*†Explain whether this process will include consultation or feedback from a client, manager, or model user.*
 
@@
 
-*Critical decisions are made here about the type of model, modelling framework and approach to be used.
-The model is formalised into a mathematical / statistical description of the system and respective changes it can experience  [@Mahmoud2009].
-In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.*
+*In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.*
 
@@
 
-*Describe what class or approach of model you will use and explain how the choice of model class was informed by the analytical objectives of the model.
-Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based  [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].*
+*Describe what class or approach of model you will use and¬†explain how the choice of model class was informed by the analytical objectives of the model.
+Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based¬† [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].*
 
@@
 However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.*
+HF- it's not clear to me what a model feature is and I find this a little confusing. I wonder if we need the text describing the 3.2 heading when it's stepped out in the subsections.
 
@@
 
-*Specify how you will operationalise response variables in the model.
+*Specify how you will operationalise the response variables in the model.
 This should relate directly to the analytical and or management objectives specified during the problem formulation phase.
-Specifications could include:- qualitative change, such as a direction of change- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, or- the frequency, location, or probability of some event occuring.Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).*
+Operationalisations could include:- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, - qualitative change, such as a direction of change or- the frequency, location, or probability of some event occuring.
+Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).*
+HF- this is great. Only slight concern is the deceptively difficult task that last sentence represents
 
@@
 *Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions.
-Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.
-Include in your rationalse for selection detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*
+Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.* HF- great
+
+*Include in your rational for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.* HF - I feel like this belongs somewhere else.
+I'm not sure where thinking about which variables the model outputs are sensitive to would relate to me choosing between logistic and cloglog models etc
 
@@
 
-*Specify which types of variables are covered in the model, and the nature of their treatment (e.g.
-lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).
-Specify model structural features, including:- the functional form of interactions,- data structures- measures used to specify links, spatio temporal scales and processes as well as their interactions- any bins or discretisation of continuous variables.*
+*Specify which variables are included in the model, and the nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).*
+
+*Specify model structural features, including:- the functional form of interactions,- data structures- measures used to specify links,- any bins or discretisation of continuous variables.*
 
@@
 
-*This refers to the iterative process of determining the most efficient or parsimonious representation of the system at the appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives specified in the problem formulation phase.
+*This refers to the process of determining the best/most efficient/parsimonious representation of the system at the appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives specified in the problem formulation phase.
 Approaches to finding model structure and parameters may be knowledge-supported, or data-driven [@Boets:2015gl].
@@
 These criteria and their underlying assumptions should reflect the desired properties of the parameter estimates / structure [@Jakeman:2006ii].
-For example, modellers might seek that parameter estimates are robust to outliers, unbiased, and yield appropriate predictive performance.
+For example, modellers might seek parameter estimates that are robust to outliers, unbiased, and yield appropriate predictive performance.
 Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are appropriate to the problem at hand.
@@
 spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii].
-Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.This is important - because it involves pre-specifying how we will interpret the fitted or quantitative model to avoid cognitive biases like confirmation bias, HARKing or other questionable research practices.*
+Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.*
+HF- I think this paragraph is only relevant to performance criteria- doesn't really explore estimation technique.
+As above though, I'm not sure that this paragraph is neessary given that it's stepped out in the subheadings below
 
-### 3.4.1 Choice of estimation performance criteria
+### 3.4.1 Choice of performance criteria
 
@@
 
-### 3.4.2 Choice of structure and parameter estimation technique
-
+### 3.4.2 Choice of parameter estimation technique	
+HF- I think that it might be double dipping to include structure here as well as in 3.3
 *Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g.
 distributed boundary conditions).
-For example, will you calibrate all variables simultaneously by optimising fi of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006).
-Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge.Specify which algorithm(s) you will use for any data-driven structure / parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*
+For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006).
+Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge.
+Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*
 
@@
 
-*Once critical decisions have been made about the approach and method of model specification, translate the conceptual model into the quantitative model.
+*Once critical decisions have been made about the approach and method of model specification,¬†translate the conceptual model into the quantitative model.
 For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters.
 (Should this go here or in model calibration?)*
+HF- I thought that this had already been covered in 3.2 and 3.3. Help me understand how this is distinct?
 
@@
 *This phase involves fitting the formally defined model to data and checking and testing the fitted model (what should go in thisphase if using a numerical or mathematical, but not correlative model or how could we change the wording to make this optional if not relevant?)*
+HF- I dont think this paragraph is necessary
 
@@
 
-*Describe the validation scheme you will use for model testing and evaluation.
-Please explain your reasoning for your choice of model calibration and validation scheme.
+*Describe how you will validate and check the calibration of the model.
 The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;[@Barnard2019].
-Although using independent data is the gold-standard or idealised practice for model evaluation, Typically testing on independent datasets is not possible for large and/or integrated models, especially when they are being used to generate anticipatory predictions beyond the conditions on which they were calibrated [@Jakeman:2006ii].
-Describe and justify the method for your calibration and validation schema.*
+Justify your choice*
 
-## 4.1.2 Data partitioning
+## 4.1.2 Describe validation data
 
-*Describe the approach for partitioning the data in support of the calibration and validation schema described above.
-For example, specify the number of folds that will be created and the relative size of each.*
+*If partitioning data for cross-validation: describe the approach specifying the number of folds that will be created and the relative size of each.
+Describe how will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced.*
 
-*How will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced?*
+*If using external data: describe any known differences between the training and validation datasets and the size of the validation dataset.
+Describe how will you document and/or share the external data such that any subsequent modelling can be computationally reproduced.*
 
@@
 
-*What Quality Assurance measures will you take to verify the modle has been correctly implemented as per the formal model specification ?
-This could be as simple as an informal ‘model walkthrough' to examine the model's behaviour qualitatively, or it could include more sophisticated approaches such as formal unit testing approaches using continuous integration platforms (see [@Yenni2019]).
-Qualitative tests could include syntax checking of code, and code reviews by peers.
-Checks for verification implementation should include i) thoroughy checking for bugs or programming errors,  and ii) whether the implemented model performs as dictated by the model description  [@Grimm:2014es].
-Specifying up front QA tests for implementation verification may help to avoid selective debugging.*
+*What Quality Assurance measures will you take to verify the model has been correctly implemented?
+Tests could include syntax checking of code,¬†and code reviews by peers.
+Checks for verification implementation should include i) thoroughy checking for bugs or programming errors, ¬†and ii) whether the implemented model performs as dictated by the model description  [@Grimm:2014es].
+Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging.*
+HF- I've tweaked this section to make it more distinct from 4.3
 
@@
 *"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification" <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd].
-During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences.
 Model checking therefore heps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd].
 Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].*
+HF- I've oved some of the information otu of 4.3 and put it in the sub headings so that this paragraph isn't necessary
 
-### 4.3.1 Quantitative tests
+### 4.3.1 Quantitative Model Checking
 
-*Specify any diagnostics or tests you will use during model checking.
-For each test, specify the numerical threshold or criteria that will you use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.*
+*Specify any diagnostics or tests you will use during model checking to assess a model's performance in representing the system of interest.
+During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences.
+For each test, specify the criteria that will you use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.*
 
@@
 
-*This step is largely informal and case-specific, but requires ‘face validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an ddefensible  [@Grimm:2014es].
-This process could be called a “laugh test" or a “pub test, and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii]. Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.
+*This step is largely informal and case-specific, but requires ‚Äòface validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an ddefensible  [@Grimm:2014es].
+This process could be called a ‚Äúlaugh test" or a ‚Äúpub test, and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii].¬†Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.
 Briefly explain how you will qualitatively check the model, and whether + how you will include users and clients in the process.*
+HF- great section
 
 # 5.0 Model Validation and Evaluation
+HF- I think that this is covered in section 4 with descriptions of validation and checking. I'm not sure how this section is different.
 
@@
 It should evaluate model checking, and consider over-fitting and extrapolation.
-The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly […] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm).
+The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly [‚Ä¶] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm).
 Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al.
@@
 ## 5.1 Model output corroboration	
-
+HF- This sounds like a second round of validation. I dont think I've ever seen this happen.
 *Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model.
-Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step <2.4>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames — hence, why ecologists often model in the first place (Grimm, et al.
+Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step <2.4>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al.
 2014.
@@
 ## 5.3 Evaluate model assumptions	
-
+HF- I would put this up in section 4 and get rid of section 5
 *Explain how you will demonstrate a lack of violation of or robustness to model assumptions.
@@
 ## 5.4 Model analysis	
-
+HF- I would put 5.4 and it's sbsections up in section 4 and get rid of section 5. Also, as above, remove the paragraph under 5.4
 *Uncertainty in models arises due to incomplete system understanding (which processes to include, or which interact), from imprecise, finite and sparese data measurements, and from uncertainty in input conditions and scenarios for model simulations or runs [@Jakeman:2006ii].
-Non-technical uncertainties can also be introduced throughout the modellign process, such as uncertainties arising from issues in problem-framing, indeterminicies, and modeller / client values [@Jakeman:2006ii].The purpose of model analysis is to prevent blind trust in the model by understanding how model outputs have emerged, and to ‘challenge' the model by verifying whether the model is still believeable and fit for purpose if one or more parameters are changed [@Grimm:2014es].
+Non-technical uncertainties can also be introduced throughout the modellign process, such as uncertainties arising from issues in problem-framing, indeterminicies, and modeller / client values [@Jakeman:2006ii].The purpose of model analysis is to prevent blind trust in the model by understanding how model outputs have emerged, and to ‚Äòchallenge' the model by verifying whether the model is still believeable and fit for purpose if one or more parameters are changed [@Grimm:2014es].
 Model analysis should increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system.
 Model analysis typically consists of sensitivity analyses preceded by uncertainty analyses [@Saltelli2019], and a suite of other simulation or other computational experiments.
-The aim of such computational experiments is to increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system [@Grimm:2014es].Uncertainty analyses and sensitivity analyses augment one another to draw conclusions about model uncertainty — the overall model will be very uncertain if the parameters that are the most uncertain are the ones the model is most uncertain to., for instance.Because the results from a full suite of sensitivity analysis and uncertainty analysis can be difficult to interpret due to the number and complexity of causal relations examined [@Jakeman:2006ii], it is useful for the analyst to relate the choice of analysis to the modelling context, purpose and analytical objectives defined in the problem formulation phase, in tandem with any critical uncertainties that have emerged during model development and testing prior to this point.*
+The aim of such computational experiments is to increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system [@Grimm:2014es].Uncertainty analyses and sensitivity analyses augment one another to draw conclusions about model uncertainty ‚Äî the overall model will be very uncertain if the parameters that are the most uncertain are the ones the model is most uncertain to., for instance.Because the results from a full suite of sensitivity analysis and uncertainty analysis can be difficult to interpret due to the number and complexity of causal relations examined [@Jakeman:2006ii], it is useful for the analyst to relate the choice of analysis to the modelling context, purpose and analytical objectives defined in the problem formulation phase, in tandem with any critical uncertainties that have emerged during model development and testing prior to this point.*
 
@@
 *Uncertainty can arise from different modeling techniques, response data and predictor variables [@Araujo2019].
-Uncertainty analyses characterise the uncertainty in model outputs, and identifies how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es;@Saltelli2019].
+Uncertainty analyses characterise the uncertainty in model outputs, and identify how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es;@Saltelli2019].
 Uncertainty analyses can include propagating known uncertainties through the model, or by investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], for example.
 It also includes characterising the output distribution, for example "by constructing it empirically from model output data points".
-It could also include extracting summary statistics like the mean, median and variance from this distribution, and perhaps constructing confidence intervals on the mean [@Saltelli2019].Please describe how you will characterise any suspected critical model and data uncertainties, as well as how you plan on summarising the results of these experiments with graphical, tabular or other devices.Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase.
-For instance “discrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impcats of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic paraameter covariance, [@Jakeman:2006ii]), specify which measure of uncertainty you will measure.*
+It could also include extracting summary statistics like the mean, median and variance from this distribution, and perhaps constructing confidence intervals on the mean [@Saltelli2019].Please describe how you will characterise model and data uncertainties, as well as how you plan on summarising the results of these experiments with graphical, tabular or other devices.
+Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase.
+For instance ‚Äúdiscrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impcats of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic paraameter covariance, [@Jakeman:2006ii]), specify which measure of uncertainty you will measure.*
 
@@
 
-*Sensitivity Analyses	Sensitivity analysis may improve understanding processes within models, and therefore on the natural system, or else as a quality assurance tool - sensitivity analyses may reveal logical or implementation errors in the model Sensitivity analysis examines how uncertainty in model outputs can be apportioned to different sources of uncertainty in model input [@Saltelli2019].
+*Sensitivity analysis examines how uncertainty in model outputs can be apportioned to different sources of uncertainty in model input [@Saltelli2019].
 Model sensitivity can be evaluated using 3 approaches: 1.
-deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).Describe any sensitivity analyses you will conduct by specifying the initial conditions and input data for each analysis - i.e.
-which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameterrs will be varied.
-Where there were critical uncertainties or multiple plausible representations of a submodel, sensitivity analyses can be used to constract alternative submodels [@Grimm:2014es].
-This analysis can be used for ‘pattern-oriented theory development' which analyses which submodel causes the full model to most closely reproduce a set of observed patterns?
-[@Grimm:2014es]State the primary objective of each sensitvity analysis, for example, a common objective is to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or altertantively to identify which variables or factors contribute little to overall model outputs, and so can be ‘dropped' from future iterations of the model [@Saltelli2019].What information should be reported to reproduce a sensitvity analysis should guide us as to what information we should get people to report now.*
+deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).
+Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameterrs will be varied.
+State the primary objective of each sensitvity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or altertantively to identify which variables or factors contribute little to overall model outputs, and so can be ‚Äòdropped' from future iterations of the model [@Saltelli2019].*
 
@@
 
-*Please describe anyother simulation experiments and / or robustness experiments you will perform on the model.*
+*Please describe any other simulation experiments and / or robustness experiments you will perform on the model.*
 
 # 6.0 Model Application
-
-*Explain how you will operationalise any scenarios identified during the problem formulation phase, if not already done so. For example, how will you operationalise any qualitative changes of interest, such as ‘deterioration' or ‘improvement'?Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario.
+HF- I feel that this belongs at the end of section 1
+*Explain how you will operationalise any scenarios identified during the problem formulation phase, if not already done so.¬†For example, how will you operationalise any qualitative changes of interest, such as ‚Äòdeterioration' or ‚Äòimprovement'?Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario.
 Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections.
@@
 ## 6.1 Choose performance criteria for scenario analysis
-
+HF- I feel that this belongs at the end of section 1
 *Describe how you will evaluate and distinguish the performance of alternative scenario outcomes.
@@
 ## 6.2 Synthesis of model application results	
-
+HF- I think this should stay here at the end
 *Specify how you will map, plot or otherwise display and synthesise the results of scenario and model analyses.
@@
 ## 6.3 Consultation with Experts, Clients, Stakeholders	
-
+HF- I'm not sure how possible people will find doing this...
+and whether it is potentially problematic to ask people to because their clients will amost certainly change their mind about how they deal with the information at this point in a way that the researcher cant control.
 *How will you use the results of model application analysis to provide advice.
-and support to the clients / managers / stakeholders based on these findings?Describe how these results will be characterised and presented to experts, clients and stakeholders once the results from the modelling exercises have been synthesised.
+and support to the clients / managers / stakeholders based on these findings?
+Describe how these results will be characterised and presented to experts, clients and stakeholders once the results from the modelling exercises have been synthesised.
 What role will these consultations have in determining the selection of management interventions, or future iterations of model development?

### 2020-05-31 index.Rmd
dummy changes to test autoccommit & knit ghaction
#20 update manuscript, first draft evaluation & testing criteria, rationale and methods

@@
 
-Rather than 'digging back' through the version history of a completed preregistration to ask researchers about the process and how it had been developed in detail, we chose to leverage the version-control and collaborative project management features of GitHub \<cite\> as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to 'live-develop' the template template.
-For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration, the suggested change, and its justification would be both recorded in GitHub and linked to an actual preregistration.
+*Evaluation Criteria and Rationale*
 
-How we did this using GitHub:
+\<which other template design papers have included evaluation criteria and rationale, Sophia's?\>
 
-Instruction to
+Of primary concern was the structure and content of the preregistration template.
+Much like variable selection during model-building, the template should be 'parsimonious' - without the inclusion of any given item on the template, it would not adequately a) constrain researcher degrees of freedom and b) transparently document the decision-process of the modeller during model development.
+The template items were evaluated based on these criteria.
 
-(As Pu et al did, they used qualitative interviews to 'walk through the version history of the preregistration, and discuss how it had developed in detail' (p.9) --- we can capture some of that thinking in the moment, by using comments and GitHub issues to capture the reasoning in the moment, as to why a particular decision was made or needs to change.
-However, I do think we should supplement this process with a structured reflection session.
+Modellers \<here, here, and here\> have expressed concern about preregistration making modelling a more difficult task than it already is, and hampering the ability to meet project deadlines, especially in decision-problems.
+Many modellers already struggle with sufficiently documenting models during publication and reporting \<ref, model reporting and transparency review\>, and modellers working in government or other agencies have expressed concern that preregistration will simply add to the list of  'red-tape' expected of them.
+Given these concerns, and considering the concerns of many with regard to the iterative nature of model development precluding making preregistration impossible or inappropriate, we figured it was of utmost importance that the templates be evaluated on their ease-of-use and feasibility.
 
-The reason why tried to capture this as we went, is because "the process of creating preregistration  is difficult to reconstruct with interview questions alone" p16.
+*Evaluation Method*
 
-**Case Study: Environmental Flows Management**
+We tested and evaluated the templates on the ability to meet these criteria by using a model-based research case-study \<see box\>.
+Rather than 'digging back' through the version history of a completed preregistration to ask researchers about the process and how it had been developed in detail \<like Pu et al\>, we chose to leverage the version-control and collaborative project management features of GitHub \<cite\> as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to 'live-develop' the template template.
+For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration, the suggested change, and its justification would be both recorded in GitHub and linked to the actual preregistration.
+Since *Pu et al* (p16).
+Found that the process of creating a preregistration is 'difficult to reconstruct with interview questions alone', we we aimed to capture the analysts 'thinking in the moment' by using comments and GitHub issues as to why a particular decision about the template item inclusion and specification should change.
+\<How we did this using GitHub:How to present this material?
+Link to the wiki issue and instructions to users?\>
 
-\<TEXT HERE DESCRIBING THE ENVIRONMENTAL FLOWS MANAGEMENT PROBLEM + RESEARCH PROBLEM\>
+Finally, we supplemented this 'in situ' testing and evaluation process with a semi-structured interviews in order to capture more detailed reflections of the analysts.
+This was because analysts were often under strict time-constraints, trying to meet their deliverables for the project.
+In addition, the process of preregistration was a new and unfamiliar task.
+
+\<TEXT HERE DESCRIBING THE ENVIRONMENTAL FLOWS MANAGEMENT PROBLEM + RESEARCH PROBLEM, to be completed by Chris / Lyndsey\>.
 
@@
 2020).
-Consequently, in this section we articulate the inner conceptual model of preregistration driving the content, structure and our proposed procedure for implementing the template.
+Consequently, in this section we articulate the inner logic of preregistration the preregistration template that informs the the content, structure of the template, and our proposed procedure and interface for implementing the template.
 
-\*\*Modelling workflows for Ecology and Conservation\*\*
+**Modelling workflows for Ecology and Conservation**
 
@@
 
-\*\*Iterative Cycle of Model Development\*\*
+**Iterative Cycle of Model Development**
 
@@
 
-\*\*The \*checking\* problem\*\*
+__The *checking* problem__
 
-Model Checking:
+The iterative nature of model construction and development complicates what is known as the preregistration *checking problem* even further.
+The process of checking a preregistered study involves comparing a manuscript to its preregistration, to verify that the study and analyses are conducted as specified in the preregistration Pu *et al.* Typically journal editors and reviewers might engage in checking, however, sometimes authors might too.
+Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
+
+The two main issues with checking are;  *And the 'checking problem' --- how do we evaluate studies that have used preregistrations, when should they be compared against their preregistration, and*
+
+*to what detail should a reviewer or editor check the preregistration?*
+
+\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\>
+
+*How do we  facilitate checking? And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
 
@@
 
-\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
-as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
-"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
-And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
-Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
+*The question we face in terms of designing the format of the template is, how do we facilitate checking?*
+
 How can we facilitate the process?
@@
 
-![Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
+!\[Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
 
@@
 
-- "A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
+\- "A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
 This trade-off was something we considered important during the design phase, and deciding what was in and out of scope for the preregistration was particularly challenging.
@@
 
-- However, as Pu et al.
+\- However, as Pu et al.
 note, some participants who were both authors and reviewers didn't think that the time-costs of preregistration in research were significantly greater than not preregistering research.
@@
 
-- Competing objectives?
+\- Competing objectives?
 Although the purposes of preregistration are not mutually exclusive - as we have decided the purpose of preregistration we have defined here is to both delimit/state flexibility, *and* be flexible.
@@
 
-- **Front-heavy research timeline:** As a way of providing better decision-making to authors.
+\- **Front-heavy research timeline:** As a way of providing better decision-making to authors.
 By shifting decision-making to the front.
@@
 
+**------**
+
+**"**We fear that preregistration will lessen the engagement of analysts with data and that it represents a step back toward the punch card style of analysis." MacEachern, S.
+N., & Van Zandt, T.
+(2019).
+Preregistration of Modeling Exercises May Not Be Useful.
+Comput Brain Behav, 2(3-4), 179-182.
+doi:10.1007/s42113-019-00038-x
+
+Ask Chris to reflect on the preregistration process and whether this is true or not.
+I don't think it is true.
+I think the opposite is true.
+This "Punchcard" style analysis they describe is probably actually an artefact of the way statistical analyses are done in science these days - science is a sausage factory of p-values.
+
 --------------------------

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
Remove characters due to encoding error from atlas

--- 2020-05-27 analysis/preregistration_template.Rmd
@@
 *Candidate decisions should be investigated and are specified a priori.
-Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019].¬†Describe the method used to identify relevant management actions.*
+Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019].
+Describe the method used to identify relevant management actions.*
 

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
Split Item 2.2
Update analysis/preregistration_template.Rmd
fix typos
remove character encoding errors

@@
 
-## 2.2 Specify key assumptions and uncertainties
+## 2.2.1 Explain Critical Conceptual Design Decisions
 
@@
 deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).
-Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameterrs will be varied.
-State the primary objective of each sensitvity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or altertantively to identify which variables or factors contribute little to overall model outputs, and so can be ‚Äòdropped' from future iterations of the model [@Saltelli2019].*
+Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameters will be varied.
+State the primary objective of each sensitivity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or alternatively to identify which variables or factors contribute little to overall model outputs, and so can be ‚'dropped' from future iterations of the model [@Saltelli2019].*
 

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
move in-text comments to GH comments

@@
 ## 6.2 Synthesis of model application results	
-HF- I think this should stay here at the end
 *Specify how you will map, plot or otherwise display and synthesise the results of scenario and model analyses.

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
convert in-text to GH comment

@@
 Conceptual models may be representeed in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables (Jakeman, Letcher, and Norton, 2006).*
-HF- Love this step
 

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
shift in-text comment to GH
Update analysis/preregistration_template.Rmd
Split 2.2 as per Hannah's comments

@@
 
-HF comment: the above part of 2.2 seems like a big ask. I think we want to achieve that we will need to break it down further. Would it be sufficient to just have the bit below?
+## 2.2.2 Specify key assumptions and uncertainties underlying the model's design.
 
-*Specify key assumptions and uncertainties underlying the model's design.
-Describe how uncertainty and variation will be represented in this model.
+* Describe how uncertainty and variation will be represented in this model.
 This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].*

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
shift in-text comment to GH

@@
 Like, how would I apply it to the woodland condition analysis?
-HF - Made the changes above in response to the comment.... do they help? If not, I'm struggling to understand exactly what this step is achieving.
 
@@
 This will likely depend on the method chosen to elicit the conceptual model.*
-HF- not sure if this step is required - I think we should expect people to have already done this bit before they start their preregistration...
-Or we could put it into 2.0 as an addendum - "if you haven't defined your conceptual model yet describe how you plan to do so."
 
@@
 - *In what ways do we expect these variables to interact (model structures)?*
-
-HF- love this step
 

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
move in-text comment to GH comment

@@
 *Collect, process and prepare data available for parameterisation, determining model structure, and for scenario analysis.*
-HF - this isn't super clear to me. Maybe we need more clarification here or an example
 

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
move in-text comment to GH comment

@@
 
-*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.* HF - Not sure exactly what this means
+*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.*
 

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
Update item as per Hannah's feedback about stating what you intend to learn from Exploratory analyses.

@@
 Data exploration prior to model fitting or development may include exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers, or the need for transformations [@Yates2018].
-Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.*
+Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.
+Specify what you intend to find out or hope to learn from each data exploration analysis.*
 HF- should we include something here asking people to specify what they intend to find out from the data exploration?

### 2020-06-01 analysis/preregistration_template.Rmd
Apply suggestions from code review
Fix typos and character encoding, move Hannah's in-text comments

@@
 Specify what you intend to find out or hope to learn from each data exploration analysis.*
-HF- should we include something here asking people to specify what they intend to find out from the data exploration?
 
@@
 
-*How will any simplifying assumptions be critically assessed [@Augusiak:2014gz]?* HF- I'm not sure how to answer this one. Maybe an example would help?
+*How will any simplifying assumptions be critically assessed [@Augusiak:2014gz]?*
 
@@
 *Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions.
-Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.* HF- great
+Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.*
 
-*Include in your rational for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.* HF - I feel like this belongs somewhere else.
-I'm not sure where thinking about which variables the model outputs are sensitive to would relate to me choosing between logistic and cloglog models etc
+*Include in your rational for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*
 
@@
 
-*Specify model structural features, including:- the functional form of interactions,- data structures- measures used to specify links,- any bins or discretisation of continuous variables.*
+*Specify model structural features, including:
+- the functional form of interactions,
+- data structures
+- measures used to specify links,
+- any bins or discretisation of continuous variables.*
 
@@
 (Should this go here or in model calibration?)*
-HF- I thought that this had already been covered in 3.2 and 3.3. Help me understand how this is distinct?
 
 # 4.0 Model Calibration, Model Fitting & Checking
-
-*This phase involves fitting the formally defined model to data and checking and testing the fitted model (what should go in thisphase if using a numerical or mathematical, but not correlative model or how could we change the wording to make this optional if not relevant?)*
-HF- I dont think this paragraph is necessary
 
@@
 *Describe how you will validate and check the calibration of the model.
-The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;[@Barnard2019].
+The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;@Barnard2019].
 Justify your choice*
@@
 *What Quality Assurance measures will you take to verify the model has been correctly implemented?
-Tests could include syntax checking of code,¬†and code reviews by peers.
-Checks for verification implementation should include i) thoroughy checking for bugs or programming errors, ¬†and ii) whether the implemented model performs as dictated by the model description  [@Grimm:2014es].
+Tests could include syntax checking of code, and code reviews by peers.
+Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as dictated by the model description  [@Grimm:2014es].
 Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging.*
-HF- I've tweaked this section to make it more distinct from 4.3
 
@@
 Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].*
-HF- I've oved some of the information otu of 4.3 and put it in the sub headings so that this paragraph isn't necessary
 
@@
 
-*This step is largely informal and case-specific, but requires ‚Äòface validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an ddefensible  [@Grimm:2014es].
-This process could be called a ‚Äúlaugh test" or a ‚Äúpub test, and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii].¬†Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.
-Briefly explain how you will qualitatively check the model, and whether + how you will include users and clients in the process.*
+*This step is largely informal and case-specific, but requires ‚ 'face validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an defensible  [@Grimm:2014es].
+This process could be called a ‚"laugh test" or a "pub test" and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii].
+Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.
+Briefly explain how you will qualitatively check the model, and whether and how you will include users and clients in the process.*
 HF- great section

### 2020-06-01 analysis/preregistration_template.Rmd
Apply suggestions from code review
Fix typos and move HF's comments out of text

@@
 - *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
-- *Prior knowledge of the current dataset - Describe any prior knowledge of the dataset.
-Be specific.* HF- not sure how someone would answer this question.
-Maybe we need a more specific Q or an example
+- *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study.
+For example, have you read any reports or publications about this data?
+*
 
@@
 
-*†Explain whether this process will include consultation or feedback from a client, manager, or model user.*
+*Explain whether this process will include consultation or feedback from a client, manager, or model user.*
 
@@
 Briefly explain how you will qualitatively check the model, and whether and how you will include users and clients in the process.*
-HF- great section
 
@@
 ## 5.1 Model output corroboration	
-HF- This sounds like a second round of validation. I dont think I've ever seen this happen.
 *Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model.

### 2020-06-01 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd
Add new item number and heading after HF's suggestion to split 2.4.5

@@
 Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].*
-
+### 2.4.5 Data Evaluation: Data reliability reporting
 *Document any issues with data reliability.

### 2020-06-01 index.Rmd
#20 update manuscript - first draft workshop analysis methods and add new todos to in situ eval and interface / shiny app.

@@
 
+**Identifying**
+
 What attributes do we aim for the template
@@
 
-**Activity 1 - Capturing Workflows and Decision Steps**
+**Activity 1 - Capturing Scientific Workflows and Decision Points**
 
@@
 
+**1\. Synthesise Workshop Outputs: Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
+
+The goal of this analysis task was to generate a complete draft workflow for both fieldwork and modelling and data-analysis that will form the basis of the final preregistration templates.
+We used a combination of inductive and deductive coding \<is this what it is called\> to analyse the workshop outputs and generate a model scientific workflow for model-based research that would constitute the presumed scientific workflow that would be used to guide the structure of the preregistration template.
+We collated and coded the workshop outputs using \<atlas.TI, insert software citation and version\>, \<a qualitative coding software\>.
 All workshop outputs can be found here (\<Link to Workshop Outputs OSF repository\>).
 
-After the workshop, data sheets from activity 1 were collated and coded using atlas.TI.
+A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management.
+Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substep' \<insert link to appendices or just present the final?\>.
+Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps.
+We revised the coding structure \<inductively (is this the word)\>, according to patterns and themes identified across multiple personal workflows.
+For example, we added the final phase \<name for last phase\>, which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions.
+In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both \<idealised norms and norms of practice, Pu et al., see other philosophy of sci refs\>.
+This workflow is used to provide structure and content to the preregistration template.
 
-The point of this activity is to build on the suggested workflows for both fieldwork and modelling to provide structure and content to the preregistration templates.
+**2\. Translating Workflows into Preregistration Templates**
 
-At the end of this exercise we expect to:
+We translated the model-based workflow identified in part I of workshop analysis into a text-based preregistration template.
+Reflecting on existing templates \<link to OSF repository of templates \>, we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis.
+We synthesised the following ontologies:
 
-Have a complete draft workflow for both fieldwork and modelling and data-analysis that will form the basis of the final preregistration templates.
+1\. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. \<Note, that these are being scrapped from the template and are turned into tool-tips\>
 
-Part I: Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.
-
-Objective: Identify common workflow.
-
-Step 1: As a group, reflect on personal experience to identify unique decision-steps. Can a generalised / common workflow across individuals be derived?
-
-Step 2: Search literature to to identify an idealised scientific workflow for ecological modelling research.
-
-Step 3: Code workflows identified in step 1 using codes identified in step 3. Revise common workflow.
-
-Part II: Translating Workflows into Preregistration Templates
-
-We translated the workflows into the preregistration template by using the following ontology:
-
-1\. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases.
-
-2\. Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps'. Corresponds to an activity or set of activities in a scientific workflow.
+2\.
+Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps'.
+Corresponds to an activity or set of activities in a scientific workflow.
+\<again, these are being scrapped and turned into tool-tips.\>
 
@@
 "explain how you will do xyzzy".
-Or "What performance measure will you use for assessing 'goodness of fit'?" In addition, we turned to existing preregistration templates to guide the wording  of some items.
+Or "What performance measure will you use for assessing 'goodness of fit'?"
+
+**3\. Designing a new methodology for implementing preregistration in model-based research**
+
+We analysed the outputs of Activity 2 in the collaborative workshop identifying barriers and solutions to undertaking preregistration in model-based research and to broader adoption of preregistration in the applied ecology and modeller communities of practice to design a preregistration procedure appropriate for this research domain.
+We considered such questions as:
+
+\- What should a research workflow look like for a modeller in an applied ecology setting who wishes to preregister their work?
+
+\- How do we accommodate the iterative cycle of model development into the preregistration process?
+
+\- At what stage in a modelling or structured decision-making process should preregistration begin?
+
+We then used 'in situ evaluation' to test and evaluate both the template and procedure design.
 
@@
 In addition, the process of preregistration was a new and unfamiliar task.
+
+\<WHAT DID WE EVALUATE\>
+
+NOT JUST THE TEMPLATES THEMSELVES BUT ALSO THE PROCEDURE, and the INTERFACE. I.e. bc *in situ evaluation is used to inform the design of a new* interface *- the shiny app...*
 
@@
 
+Reflecting on the in situ evaluation - shiny app
+
+*And the in situ evaluation is used to inform the design of a new* interface *- the shiny app.*
+
 The shiny app is intended as an interface for authors and analysts to respond to template items and input their planned decisions to complete a preregistration for their study.
@@
 
-!\[Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
+![Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
 

### 2020-06-03 analysis/preregistration_template.Rmd

--- 2020-06-01 analysis/preregistration_template.Rmd
@@
 Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].*
+### 2.4.5 Data Evaluation: Data reliability reporting
+### 2.4.5. Data Evaluation: data reliability reporting
 
@@
 
-*Data Exclusion: How will you determine what data, if any, will be excluded from your analyses?
+### 2.4.6. Data evaluation: data exclusion
+
+*How will you determine what data, if any, will be excluded from your analyses?
 How will outliers be handled?
@@
 
-*Missing data: How will you identify and deal with incomplete or missing data?*
+### 2.4.7 Data Evaluation: Missing Data
+
+*How will you identify and deal with incomplete or missing data?*
 

### 2020-06-03 analysis/preregistration_template.Rmd
Update analysis/preregistration_template.Rmd

@@
 
-*What degree of flexibility is required from the model?
-Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?* ***HF comment: love the inclusion of this question***
+*What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?*
 

### 2020-06-05 index.Rmd
#20 restructure and write adaptive preregistation section

@@
 **Intended journal:** Methods in Ecology and Evolution.
+
+**Article Type:** Research Article
 
@@
 
-## A methodology for 'Living Preregistrations' ##
+## Results: A methodology for Preregistration in Ecology and Conservation ##
+
+This will be a box or a table with numbered items, with just the preregistration item directives or question.
+The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
+Could possibly go into the appendix, but I think given that it's the primary research output, it should go into the main-text of the manuscript.
+
+----
+
+### Internal Model of Preregistration ###
 
@@
 @TODO - anything about *applied contexts??* Aligning with management? (In terms of the structure and steps above).
+
+@TODO insert table from atlas TI.
 
@@
 
-@TODO insert table from atlas TI.
+\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
+as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
+"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
+And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
+Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
+How can we facilitate the process?
+They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
 
-### Iterative Model Development & "Living" Preregistrations ###
+----
 
-Given this defining feature of the modelling process, it is critical that the design of the interface and procedure for implementing the preregistration inherently capture this iterative nature of development.
-Current formats for preregistration are inappropriate given the iterative nature of the model development cycle, and the need for checking the preregistration against the reported model and analysis.
-Existing templates are text-based documents wherein authors respond to template items to complete the preregistration.
-They are standalone documents, with no way of linking the specified decision points to the actual analysis - thereby making model checking extremely tedious.
+### Adaptive Preregistration ###
+
+The process of preregistration as it is currently implemented is tailored towards the hypothetic-deductive model of the scientific method, and thus follows a linear process, where the process of planning and preregistering is distinct and completely precedes the execution of that plan.
+The researcher shifts from ideation, to study/analysis design and preregistration, to collecting and analysing the data, writing the report or manuscript, to publishing ([*cos.io/rr*](http://cos.io/rr)) \<See also Figure 1 in Pu et al.\> The iterative nature of model development, however, inherently precludes this model of preregistration from being applied to model-based research in ecology and conservation.
+
+Firstly, as Srivastava notes, there are some decisions that are either difficult to anticipate, or simply cannot be made in advance \<Srivastava\>.
+This is particularly true for decisions occurring at the later phases of the modelling construction and development process (Fig X), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses.
+For example, in the modelling process we have described (previous section), some decisions cannot be made unless you have a fully specified model, such as \<sensitivity analysis, and other model evaluation tasks and test\>.
+
+Secondly, very often in modelling, preliminary or exploratory analyses must be implemented before decisions can be made about how to proceed.
+For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation.
+Sometimes the results of these analyses and checks mean that the modeller is forced to return to earlier decision-points and change the analysis.
+For instance, the model might fail to converge, or perhaps there is insfufficient data to fit the desired model such that the model runs out of variation to apportion to all the different variables \<there's a name for this\>.
+
+Consequently, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by \<Srivastava\>.
+This view of preregistration breaks with the current model of preregistration, wherein the author writes a single deterministic preregistration containing a rule for every decision, and shifts from ideation, preregistration to execution of the plan.
+Two features define adaptive preregistration.
+It contains 'plans to deploy flexible strategies' \<Srivastava\> wherein the author can supply a heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses.
+Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed \<Srivastava\>.
+Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, <x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?id=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0>\>.
+
+Given the iterative nature of adaptive preregistration then, the current convention of the static text-based document format of preregistration is inadequate.
+If there are interim versions of a preregistration, each version must be time-stamped and *version-controlled*, such that it is clear not just that two or more versions differ from each other, but it is apparent in exactly *how* they differ.
+Moreover, the *checking problem*, is exacerbated --- for flexible modelling and analysis strategies, we need to be able to link the results of preliminary analyses back to the specified strategy: the link between a particular triggered decision, and the analysis outcome that triggered it must be explicit.
+In addition, if the results of some preliminary analyses cause revisions to the model itself or a different modelling procedure or approach to be selected, then the trigger for this decision must be explicitly linked to the new version of the preregistration.
+
+We propose leveraging both the *version-control* and the project management features of *GitHub* \<link to GitHub\> as a tool for implementing adaptive preregistration \<link to box\>.
+When a document is version-controlled using git and GitHub, each version of a document is time-stamped, and is assigned its own unique document identifier, or *commit-hash.* Because GitHub documents are 'time-stamped' the genesis of the preregistration from one version to the next is apparent.
+Moreover, using the *diff* view of a document on *GitHub*, exactly *how* a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted \<insert picture\>.
+Since the preregistration document would exist within the project *repository* containing the data and modelling and analysis code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated version of the plan --- thus facilitating the process of model checking.
+By providing a method for contrasting the actual analysis undertaken against the registered analysis, this is a mechanism for allowing reviewers to properly evaluate  the preregistration against the reported analysis.
+It thus provides a way of marking preregistered parts of a report from non-preregistered parts of the analysis.
+
+**How it works....**
+
+GitHub is used to track amendments to the preregistration --- each time an additional question is answered, or an alteration is made to the preregistration, the change should be *committed* to the *repository*.
+*Semantic versioning \<insert link\>* is used to document each version of the preregistration.
+\<explain semantic versioning in one sentence\>.
+Using *GitHub's* tag and release feature, each interim preregistration is tagged and released with its semantic version number.
+
+All questions that can be answered should be completed before any data analysis or modelling proceeds.
+The preregistration should be completed until a) no further detail can be supplied or b) no additional template item can be answered until preliminary investigations or analyses are undertaken (Fig X, steps 1 - 3).
+
+**Shifting from planning to execution**
+
+Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding *GitHub issue* created (Figure X, step 3, document).
+\<define GitHub issue, <https://guides.github.com/features/issues/>\>.
+Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an *issue*'s thread.
+Each issue is assigned a unique number.
+Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the *issue's* thread (Fig X, step 4, Do) \<screen-cap?\>.
+The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new *release*.
+
+... **And back again.**
+
+As the modelling proceeds, further detail on later phases of the analysis can be iteratively updated and preregistered.
+Where the results of model evaluation and analysis reveal that there are problems with the model, plans can be changed, and again the next phase can be preregistered.
+For example, if a researcher finds that assumptions are violated or other unexpected results occur that force a change to the planned analysis, the preregistration can be updated based on the findings of those analyses.
+When this occurs, the *major* version number of the preregistration document should be incremented (Fig X steps 5 - 6).
+Since the findings of the analyses are tracked within the relevant issue for that template item, and the issue is recorded in the preregistration itself, the trigger for the change in the plan is made explicitly clear.
+
+For each altered or new preregistration item, a the issue thread should either be updated, or a new issue created respectively (Fig X, step 6).
+Those analyses are then either reconnected based on the revised plan or the new investigatory analyses are then conducted (Fig X, step 7).
+
+This process continues until the preregistration is complete, and the researcher can continue to execute the plan as it has been fully described in the final version of the preregistration document (Fig X, step 8 - 9).
+
+**Flexible Strategies, Preliminary Analyses or Investigations**
+
+When a researcher needs to conduct preliminary analyses or check parts of the data before committing to a particular decision about the model or analysis, they can specify a flexible strategy in their response to that preregistration item by:
+
+- stating what needs to be known to move forward with the modelling and analysis
+
+- explaining how the researcher will test this, and what parts of the data will be used in that analysis
+
+- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.
+
+![Adaptive Preregistration Workflow - Research Workflow for implementing a 'adaptive' preregistration in a modelling context](./figures/LivingPreregistration.png)
+
+----
+
+### Preregistration Interface - Shiny App ###
 
@@
 Consequently, we instead decided to implement a user-friendly interface that makes the completion of a preregistration with our template much more *approachable\<?adj\>*.
-
-**Living Preregistration Workflow**
-
-2\. Workflow - allows for iterative model development AND
-
-Why, what is the current form? Existing templates are read-only documents.
-
-Capturing this inner logic has repercussions for
-
-Thus deciding on the inner logic of the template, and what is in and out of scope has been particularly difficult.
-
-One method around this is to accept that the current model of preregistration is not going to work for iterative model development.
-Again, the more complex the model, the more likely this is to be the case.
-Consequently I propose a different model of preregistration that I think will work for modeling research contexts.
-At the moment I'm calling it 'living preregistration' but please suggest another name.
-Others have termed this 'adaptive preregistration', and this is maybe a better name.
-
-In this model of preregistration, all questions that can be answered, should be answered before any data analysis or modelling proceeds.
-For some items, particularly during model evaluation and analysis, perhaps only responses of a general nature can be provided.
-That is OK.
-As the modelling proceeds.
-Further detail can be provided and the next phase of analysis can be preregistered.
-Where the results of model evaluation and analysis reveal that there are problems with the model, or perhaps assumption violations, plans can be changed, and again the next phase of analysis can be preregistered.
-
-A static document for this type of process doesn't really make sense.
-I propose leveraging the code and project management tools within the GitHub platform as a way of hosting a 'living' or 'adaptive' preregistration.
-Because GitHub documents are 'time-stamped', the genesis of the preregistration from one version to the next can be traced and tagged with versions.
-Since the document would exist in the repository with the code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated plan.
-
-*How do we accommodate this structurally and procedurally into the PRT?*
-
-*1\. Instruction to PRT users:*
-
-"If you find yourself needing to check something with the data before committing to a particular decision about the model or analysis (And you hadn't anticipated needing to perform this check), record:
-
-A) the relevant decision in the template.
-
-B) What do you need to know to move forward?
-
-C) How you will test this? What test or analysis will you perform?
-
-D) What values or outputs of the test or analysis map onto particular decisions (they can be qualitative as well)."
-
-I will create a template on GitHub based on the above to be used by Chris + Lyndsey for when this occurs, to be stored on the repository wiki.
-They can insert this template into the relevant decision step / PRT item and fill out.
-
-*2\. "Living" or "Timestamped" or "Adaptive" Preregistrations ---Preregistration in the research process:*
-
-See Figure 1 in Pu et al.
-And [*cos.io/rr*](http://cos.io/rr) - Traditionally the process of preregistration and the 'doing' or execution of the plan is a distinct and sequential process.
-However, this isn't going to work for modelling (What are the *norms* about the process of modelling - stated that it's iterative.
-And how does this look in practice --- use our in situ evaluation and testing to capture this.
-
-This obviously breaks with the convention of preregistrations as 'static' one-off documents... does it break with this convention so far as to no longer be a 'preregistration'?
-
-Pu et al: tendency for preregistrations to go from vague / general to specific over time.
-
-Importantly, Pu et al show that "not all preregistered studies confirm to their preregistrations, or are inconsistent in reporting deviations", thus by making the preregistration available to reviewers, and by providing a platform and method for linking the actual analysis under-taken, we can allow reviewers to properly evaluate the preregistration against the final paper.
-(AND, must talk back to original aim of bad reporting practices in modelling).
-Moreover, if you combine formatting changes with our proposed method with GitHub, we hope to tailor the presentation of the completed preregistration + analysis, such that actual analyses and decisions can be linked directly to both files and outcomes.
-
-Using:
-
-\- How to mark the preregistered from the non-preregistered parts of the paper?
-
-\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
-as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
-"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
-And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
-Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
-How can we facilitate the process?
-They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
-Maybe we could do something else, rather than get the paper to refer to the parts of the preregistration (but I think we should do this with Chris anyway).
-If we propose the norm of using GitHub to create and check the preregistration., I think this could be a good way of getting around this checking problem.
-At the very end we have a 'how to box' targeted at authors.
-And then maybe we have a box targeted at reviewers?
-Not sure, maybe just focus this paper on authors and the process of creating and using a preregistration.
-
-In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
-However, I don't expect users to agree on this process.
-Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
-
-This will be a box or a table with numbered items, with just the preregistration item directives or question.
-The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
-
-### Shiny Preregistration ###
 
@@
 @TODO make an issue in the shiny repository for being able to 'load' a partially completed preregistration... Maybe we need to export in JSON?
-
-![Living Preregistration Workflow - Research Workflow for implementing a 'living' preregistration in a modelling context](./figures/LivingPreregistration.png)
 
@@
 
+Hosting the preregistration template on GitHub, has the added benefit of providing a mechanism for dealing with *uncertain* *norms* in modelling practice \<Pu et al\>, and for following modelling methods and decision contexts that might not fit the presumed modelling development cycle underpinning the preregistration template we designed.
+Anyone is free to *fork* our GitHub repository, meaning that the entire contents of the repository can be copied to a users own repository, and the template is free to be altered as the user sees fit.
+This provides the opportunity for creating domain- and/or modelling-approach-specific templates \<Cruwell\>.
+
 **------**

### 2020-06-05 index.Rmd
#20 update manuscript - evaluation & checking problem sections

@@
 
-Of primary concern was the structure and content of the preregistration template.
+Of primary concern was the structure and content of the preregistration template, however we also aimed to evaluate both the procedure and the interface for completing the preregistration.
 Much like variable selection during model-building, the template should be 'parsimonious' - without the inclusion of any given item on the template, it would not adequately a) constrain researcher degrees of freedom and b) transparently document the decision-process of the modeller during model development.
-The template items were evaluated based on these criteria.
+Yet, with too many items, or with too high a resolution of the template items, the preregistration would be too difficult and time-consuming to complete.
+We evaluated the template items based on these two criteria.
 
@@
 Many modellers already struggle with sufficiently documenting models during publication and reporting \<ref, model reporting and transparency review\>, and modellers working in government or other agencies have expressed concern that preregistration will simply add to the list of  'red-tape' expected of them.
-Given these concerns, and considering the concerns of many with regard to the iterative nature of model development precluding making preregistration impossible or inappropriate, we figured it was of utmost importance that the templates be evaluated on their ease-of-use and feasibility.
+Given these concerns, and considering the concerns of many with regard to the iterative nature of model development precluding making preregistration impossible or inappropriate, we figured it was of utmost importance that the templates be also evaluated on their ease-of-use and feasibility.
 
@@
 In addition, the process of preregistration was a new and unfamiliar task.
-
-\<WHAT DID WE EVALUATE\>
-
-NOT JUST THE TEMPLATES THEMSELVES BUT ALSO THE PROCEDURE, and the INTERFACE. I.e. bc *in situ evaluation is used to inform the design of a new* interface *- the shiny app...*
 
@@
 2020).
-Consequently, in this section we articulate the inner logic of preregistration the preregistration template that informs the the content, structure of the template, and our proposed procedure and interface for implementing the template.
+Consequently, in this section we articulate the inner logic of our preregistration templates that informs the the content, structure of the template, and our proposed procedure and interface for implementing the template.
 
@@
 
-The iterative nature of model development was identified in our structured discussions as a key barrier to adopting preregistration.
+During our structured discussions, we identified the iterative nature of model development as a key barrier to adopting preregistration.
 \<Insert refs to literature describing this\>.
@@
 
-Typically, the process of preregistration is completely distinct from and precedes the implementation of the analysis plan laid out in the preregistration.	In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" (Nosek et al.
+Typically, the process of preregistration is completely distinct from and precedes the implementation of the analysis plan laid out in the preregistration.
+In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" (Nosek et al.
 2018) or prior to analysing the data (Ref, or link to secondary data preregistration template).
@@
 
-The iterative nature of model construction and development complicates what is known as the preregistration *checking problem* even further.
-The process of checking a preregistered study involves comparing a manuscript to its preregistration, to verify that the study and analyses are conducted as specified in the preregistration Pu *et al.* Typically journal editors and reviewers might engage in checking, however, sometimes authors might too.
-Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
+Authors, reviewers, editors and readers may all engage in the process of checking a report or manuscript against its preregistration to verify that the study and analyses were conducted as specified in the preregistration.
+However, existing preregistration formats \<do not encourage/make the process of checking difficult\> --- static text-based documents are designed for authors to quickly input the desired information.
+In fact, *Pu et al* go so far as to say that this format "verge\[s\] on being write-only media" and are therefore completely ill-suited to the checking task.
 
-The two main issues with checking are;  *And the 'checking problem' --- how do we evaluate studies that have used preregistrations, when should they be compared against their preregistration, and*
+The task of model checking is further complicated by the non-linear and iterative nature of model development.
+It is difficult to verify the conducted study and analysis against the preregistered analysis for existing preregistration formats used in hypothetico-deductive research .
+However, when the modelling and preregistration is iterative, such that there might be interim preregistrations, then this means that there are *multiple* versions of the preregistration to check against the completed analysis.
 
-*to what detail should a reviewer or editor check the preregistration?*
+The question we faced then is, how do we design both the format of the template, and its procedure for completion by the author, so that we facilitate preregistration checking --- especially when this task is complicated by the iterative nature of model development?
+*Pu et al* put forth two design elements that may facilitate checking.
+They recommend designs that "make relevant preregistration easier to *query*" and that "encourage wider *coverage* of preregistration content during review".
 
-\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\>
+For *Pu et al.* The process of *querying* a preregistration consists of "paper-to-preregistration" and "preregistration-to-paper" lookup whereby the reader searches for relevant information from the paper in the preregistration and *vice-versa*.
+This requires that the report of paper must be specific in referring to parts of the preregistration, *and*, that the author has tagged the relevant part of the preregistration in their report or manuscript.
 
-*How do we  facilitate checking? And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
+Consequently we propose a methodology for managing the preregistration process that leverages features of GitHub that implement the task of linking the actual study and analysis that was conducted to the pre-specfied plan in the preregistration \<link to part of GH methodology that explains how checking is accounted for in the procedural design\>.
+In addition, we designed our shiny interface for completing the preregistration to include a 'checking mode' that easily allows the author to state how and why the the analysis might have diverged from the planned \<link to part of shiny app interface that discusses checking mode\>.
 
-"Existing preregistration formats are designed primarily for authors to quickly input the desired information, but may verge on being write-only media: they are ill-suited to the checking task that reviewers, paper readers, and in some cases even authors themselves might perform.
-Thus, we recommend exploring designs that (1) make relevant preregistration content easier to *query* and (2) encourage wider *coverage* of preregistration content during review." Pu et al.
+---------
 
-*The question we face in terms of designing the format of the template is, how do we facilitate checking?*
+*I'm not exactly sure whether this content should go here in a summarised form or whether it should go into its respective places coming up in the manuscript and we just link to it.*
 
-How can we facilitate the process?
-They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
-Maybe we could do something else, rather than get the paper to refer to the parts of the preregistration (but I think we should do this with Chris anyway).
-If we propose the norm of using GitHub to create and check the preregistration., I think this could be a good way of getting around this checking problem.
-At the very end we have a 'how to box' targeted at authors.
-And then maybe we have a box targeted at reviewers?
-Not sure, maybe just focus this paper on authors and the process of creating and using a preregistration.
+*How does our GH procedure incorporates checking:*
 
-In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
-However, I don't expect users to agree on this process.
-Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
+1\. Using issues to track committed work on the same analysis task, and that issue is spawned from a template item so the conducted work is inherently linked to that initial plan.
 
-\- how to evaluate studies that have used preregistrations and when they should be compared against their preregistration\> This process of "checking" a preregistered study is provisionally defined by Pu et al.
-as "comparing a manuscript to its preregistration and verifying that the study and the analyses are conducted as the preregistration has specified." Who engages in 'checking' a preregistration?
-"Authors, reviewers and editors may all engage in checking." *How do we  facilitate checking?
-And the "checking problem"* "to what detail should or does an author or reviewer / editor check?
-Pu et al found that the authors engaged in a linear and thorough checking process, but reviewers on editors were not nearly "as exhaustive with checking".
-How can we facilitate the process?
-They found that if the "paper is very specific to referring to parts of the preregistration" then this made it easier for reviewers and editors to check.
+2\. In the preregistration document
+
+*How does our shiny-preregistration Interface incorporates checking:*
+
+\- Checking mode - The author can switch to checking mode after having conducted the analysis for a given PRT item. *Or* after the preregistration is complete.
+
+What do you think about this process?
+What if we allowed check boxes in our preregistration software design so that the author could switch to 'checking mode' and they can tick some boxes Yes / No / Partial match.
+For No and Partial, they briefly describe \_how\_ it varies from the preregistration, and why (could link to the GitHub issue thread where conversation and decisions were logged during the process of conducting the analyses).
+
+This is going to take some further thought because there might be \_interim\_ versions of the document.
+Our process means that the author keeps adapting the preregistration until they actually go ahead and implement it...
+so doesn't this mean that they will \_always\_ follow their latest version of the preregistration??
+I guess then checking entails a comparison of first and interim preregistrations against the actual conducted analysis, not just the final.
+
+AND it is possible that in other cases the author deviated from their plan for some reason, and didn't follow our proposed process using GH.
+So we want to allow for this to occur, but for the author to be able to state \_that\_ this happened, \_how\_ it differs, and why.
+
+---------
 
@@
 
-- stating what needs to be known to move forward with the modelling and analysis
+\- stating what needs to be known to move forward with the modelling and analysis
 
-- explaining how the researcher will test this, and what parts of the data will be used in that analysis
+\- explaining how the researcher will test this, and what parts of the data will be used in that analysis
 
-- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.
+\- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.
 
@@
 
+In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
+However, I don't expect users to agree on this process.
+Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
+
 **------**

### 2020-09-11 index.Rmd
Restructure outline - add UCD info.

@@
 author:
-  - name: Elise Gould 
+  - name: Elliot Gould 
     affiliation: School of BioSciences, University of Melbourne
@@
 
-Failure to reproduce a large proportion of published studies in psychology and medicine has received considerable attention and provoked heated discussion among researchers in the broader scientific community.
-Initial meta-research in ecology has revealed that the conditions found to foster irreproducibility are present (Fidler et al.
-2017), evidencing that the discipline is also at risk of a "reproducibility crisis," with rates of Questionable Research Practices in ecology similar to those in other fields (Fraser et al.
-2018).
-Ecology has seen recent improvements in journal policy concerning reporting guidelines and data/code archiving, but there has been slow uptake of other important tools, like preregistration and registered reports (Parker, Fraser, and Nakagawa 2019).
+Ecology as a discipline is at risk of a "reproducibility crisis" (Fidler et al.
+2017), with rates of Questionable Research Practices (QRPs) in ecology similar to those in other fields (Fraser et al.
+2018), such as psychology and medicine, where the crisis has received considerable attention, and provoked heated discussion among researchers within the broader scientific community.
+In particular, there are serious problems with the way ecological models and the modelling process are reported in the published literature.
+However, concerns about poor model reporting practices are not new, and have been raised repeatedly --- nearly a decade ago researchers concluded that most model reporting was neither transparent or comprehensive [@Schmolke:2010fd; @deVos:2011], and that a culture of model reporting was yet to be established [@Grimm:2014es].
+Moreover, most models were assessed predominantly on the basis of scientific originality during peer-review  [@Schmolke:2010fd] to the effect that the burden of evaluating models for credibility is placed onto the reader or model-user, rather than the peer-review process, especially in the case of extremely complex models [@Alexandrov:2011iv].
 
-Preregistration is designed to mitigate the opportunistic misuse of researcher degrees of freedom, or "Questionable Research Practices" (QRPs, Wicherts et al.
-(2016)).
+Despite attempted to promote thorough and transparent model documentation using 'Good Modelling Practice' (e.g.
+[@Jakeman:2006ii;@Alexandrov:2011iv]), more recent quantitative assessments of transparency and reproducibility demonstrate that the transparent model reporting is still an ongoing problem.
+Just as Jakeman argued that many models in the published literature lack a clear statement of model objectives and purpose, lack descriptions of model assumptions and implications, and also fail to report model results, including validation and evaluation in 2006, [-@Jakeman:2006ii], [@Feng2019]'s quantitative assessment of Ecological Niche Models found that reporting of model evaluation and model results was less than adequate.
+
+We should be concerned about research transparency because, a lack of complete and transparent reporting practices provide the conditions in which QRPs are likely to arise [@Fidler:2017he].
+Moreover, as Morrison [-@Morrison:2016cd] demonstrated, when models are not reported transparently with sufficient detail, they cannot be reproduced.
+This is particularly problematic because in ecology, models are typically used in decision-support, often directly underpinning policy and management decisions.
+Incomplete and non-transparent reporting of the model and the modelling process risk the model being used in inappropriate applications or decision contexts [@Grimm:2014es] because the underlying modelling choices and assumptions are not communicated [@deVos:2011].
+Alternatively, a lack of transparency may mask poor model design or serious flaws in the model, with the risk that they be used in decision support resulting in wrong, or irreversible decisions [@Grimm:2014es;@Schmolke:2010fd].
+A lack of transparency in reporting also risks undermining the credibility of the model, such that the decision-maker rejects the use of the model and instead turns to unstructured and subjective judgments [@Grimm:2014es;@Addison:2013wa], thus contributing further to "decision complacency", or the "knowing-doing" gap [@Gardner:2018dm].
+
+Preregistration is a practice that has been adopted in disciplines that have confronted the 'reproducibility crisis' head-on, and can improve the transparency and reproducibility of ecological models.
+Preregistration requires the methods and analysis plan of a future study are registered in a secure and accessible platform prior to any data collection, and after submission it can no longer be altered.
+Preregistration distinguishes between what is a genuine a priori planned analysis, and what is not (Parker, Fraser, and Nakagawa 2019), and therefore prevents the opportunistic misuse of researcher degrees of freedom or 'Questionable Research Practices' (QRPs, Wicherts et al., 2016).
 Researcher degrees of freedom refer to the many analytic decisions a researcher makes during the planning, data collection, data analysis and reporting of results in scientific studies (Wicherts et al.
@@
 Examples of these analytic choices include whether to exclude observations, whether more data should be collected, which set of conditions should be combined and compared, which model evaluation metric should be used, etc.
-Coupled with a lack of transparency -- or "undisclosed flexibility" -- researcher degrees of freedom are vulnerable to being exploited by researchers because of cognitive biases.
+Coupled with a lack of transparency -- or "undisclosed flexibility" -- researcher degrees of freedom are vulnerable to being exploited by researchers because of cognitive biases and incentive structures within science.
+\<lead in sentence to next para??\>
 
-Within basic ecological research employing null hypothesis significance testing, QRPs may increase the probability of false-positive findings (Seaman and Weber 2015; Earp and Trafimow 2015), and may spuriously increase effect size estimates (Wicherts et al.
-2016; Parker et al.
-2016).
-QRPs in model-based research may "artificially increase the accuracy or precision of a model, its predictions, and or/evaluation tests to the effect that the model is perceived to be more credible than it would be if the QRP did not occur" (Gould et al.
-2019).
-They are "practices that lead to consumers of the model placing false belief in the reliability, validity and utility of the model than would be warranted without the QRP" (Gould et al.
-2019).
-Some examples of questionable research practices in model-based research include cherry-picking and 'statistic-hacking' model evaluation and test results in order to overemphasise the model's performance and tweaking model priors after fitting the model to data and checking results (Gould, n.d.).
+Ecology has seen recent improvements in journal policy concerning reporting guidelines and data/code archiving, but there has been slow uptake of other important tools, like preregistration and registered reports (Parker, Fraser, and Nakagawa 2019).
+Pu et al.
+2019 found that conceptual barriers may prevent users from different disciplines to think that preregistration is "not for them", this was a particularly salient sentiment for researchers engaged in non-experimental studies.
+We have experienced this sentiment ourselves when looking at preregistration templates to use for model-based research --- nearly all preregistration templates are geared towards experimental hypothesis-testing research.
+Such templates contain items that are simply irrelevant to model-based and/or applied ecological research.
+More importantly, they do not reflect the iterative nature of model development [@MacEachern2019] and fail to capture critical analytic decisions in the modelling process.
+Presently, we are aware of only a single preregistration template for modelling [@Cruwell2019b], however it is steeped heavily in an an academic psychology context, is focussed on the use of models for hypothesis-testing, neglects the entire model development process, and approaches modelling from a very narrow epistemic function, in a way that is not commonly applied in ecology.
 
-Preregistration requires the methods and analysis plan of a future study are registered in a secure and accessible platform prior to any data collection, and after submission it can no longer be altered.
-Preregistration guards against questionable research practices, by making clear what is a genuine a priori planned analysis, and what is not (Parker, Fraser, and Nakagawa 2019).
+Preregistration cannot simply be ported over to other disciplines and adopted as-is.
+If we wish to adopt preregistration in ecology, specifically within model-based research, both the content required to be preregistered, as well as the procedure for creating a preregistration must be adapted to suit model-based research as it is undertaken in ecology.
 Different research methodologies, and their specific and nuanced application across different domains and fields of study follow different analytical choices and are therefore susceptible to different QRPs.
-Consequently, domain-specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom unique to a particular domain-specific methodology.
-Moreover, we believe that a lack of discipline-specific preregistration templates is a key impediment to the uptake of these Open Science tools.
-
-------
-
-Problem: "Conceptual barriers to adopting preregistration, including users from different disciplines feeling that preregistration is not for them" Pu et al. (2019).
-
-"Preregistration considered alienating.
-Researchers who do non-experimental studies expressed a "not for me" sentiment." We have experienced this first-hand by our peers, and even by ourselves, when opening and looking at preregistration templates.
-
-Questions about whether preregistration is useful for exploratory work because it 'limits flexibility' --- and moreover, modelling is, by many, considered exploratory.
+Consequently, domain-specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom as they uniquely emerge within a domain or field-specific application of a particular research methodology.
 
@@
 
-There are many templates for preregistering experimental research, particularly for studies employing null-hypothesis significance testing.
-However, to date there is only a single published template for pre-registering model-based studies, and it is for cognitive modelling in psychology (Crüwell and Evans 2019).
-Consequently, this project aims to develop a template for preregistering ecological modelling research.
-In this project we aim to develop a template and new methodology for implementing preregistration of ecological models in applied ecology.
+The goal of this study is to \<rethink\> preregistration to better suit the domain-specific challenges and methodological particularities of model-based research in ecology.
 This research builds on recently developed "QRP roadmaps" (Gould et al.
@@
 (2019) showing that existing templates are too general, and do not account for domain-specific and research category-specific nuances and questionable research practices.
-Template development and testing will be driven by a real-world case study of environmental-flow modelling.
+We aim to develop a template for preregistering ecological models, and devise a methodology for implementing an expanded view of the preregistration process, 'adaptive preregistration' (*sensu* Srivastava), that better accounts for the inherently iterative nature of model development than static, deterministic document-based preregistration.
+
+We aim to consider questions such as... (open questions we had about preregistration)...
+
+Finally, we develop and evaluate both the template and proposed adaptive preregistration process with a case-study of environmental-flow modelling in Victoria, Australia.
+
+----
+
+## Methods ##
 
@@
 
-As Pu et al.
-(2019) identify, a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings.
-In order to avert this mismatch, we have taken a user-centred design approach to the problem of preregistration.
-We do this in two ways, firstly, by running a series of structured and collaborative activities in a workshop setting to reflect on our own research experiences in order to capture scientific workflows as they occur practice.
-And secondly, by using a real-world research problem as a case-study to test and evaluate the templates for utility and feasibility in practice.
+We took a user-centred design (UCD) approach to the task of translating preregistration into model-based research contexts for ecology and conservation.
+UCD is a philosophy for designing services or products that accounts for user requirements at all stages of the design process --- the goal of which is to generate both an effective and usable tool for specific people in specific contexts [@Pavelin2012].
+As Pavelin et al illustrate,  it is critically important that users be involved in the design process because it is difficult to predict how users will actually use a resource.
+This mirrors Pu et al's qualitative analysis into preregistration design, that a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings.
+Consequently, although user-centered design is typically used within software engineering domains, we heed Pu's advice that "It is crucial to treat preregistration [...] as a user-centered design problem to ensure preregistration achieves its intended goal."
 
-@TODO what does a UX design approach entail... stating purpose, use-case and contexts a priori? Explain.
+We borrow from the UCD methodology proposed by Pavelin et al.
+[-@Pavelin2012], and follow the core steps ![](./figures/user-centered-design-preregistration.png): 1.
+Define the goals of the resource under development and identify audience, 2.
+Characterise users and their needs, 3.
+Design mock-ups of your product, 4.
+Build prototypes of selected mock-ups and test with users, 5.
+Implement the product, writing technical implementations and building the product.
+In the case of this project, the 'product' consists of a preregistration template, specification of the internal logic of the template, including an updated domain-specific procedure for using the template, a user-interface for completing the preregistration, and a user-guide to accompany the template and interface.
+
+#### Define Vision & Objectives ####
 
@@
 Beyond hypothesis-testing contexts, this is a simplifying dichotomy that, in practice, is more blurred and nuanced than the ideal.
-For instance researchers might move back and forth between different research-modes throughout the modelling cycle.
-Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, perhaps with the aid of some philosophers of science.
+For instance researchers might move back and forth between different reasoning-modes throughout the modelling cycle.
+Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, and consideration of research from philosophy of science.
 Consequently, we wish to deemphasise the purpose of distinguishing between exploratory and confirmatory research, and focus the design of the templates in being able to delimit researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and other questionable research practices, like cherry-picking, HARKing etc.
 \<insert ref to my QRP manuscript\>.
+
+See margin notes from Mac eachern... for this section.
+
+Moallemi I described the querying of he model in,y as exploratory.
+
+Brief mention of abduction.
 
@@
 
-----
+Questions included: should we and how can we change the medium of the template and/or archiving platform to accomodate any roadblocks?
+How could we change the medium of both the template and the archiving platform to accommodate these issues?
+What procedural and or technical solutions could we implement to address these roadblocks to preregistration?
+What decision-tools could be used to aid in completing a preregistration?
+What sort of parsimony should we aim for in terms of the resolution of the templates and therefore their applicability across different methodologies?
 
-## Methods ##
-
-### Identifying Ecological Modelling Workflows in Practice ###
-
-**Identifying**
+**Identifying aspirations and desirable attribute for the template**
 
@@
 forces user to a particular workflow, so getting that workflow right, that reflects both norms and practice of model-based research is pivotal.
+
+#### User Research ####
 
@@
 
-**Activity 2 - Challenges and Solutions to Preregistration in Ecology**
+## Workshop Activity 2: Understanding ecology- and modelling-specific challenges to adopting preregistration in research practice
 
@@
 
-Questions included: should we and how can we change the medium of the template and/or archiving platform to accomodate any roadblocks?
-How could we change the medium of both the template and the archiving platform to accommodate these issues?
-What procedural and or technical solutions could we implement to address these roadblocks to preregistration?
-What decision-tools could be used to aid in completing a preregistration?
-What sort of parsimony should we aim for in terms of the resolution of the templates and therefore their applicability across different methodologies?
+##### Workshop Analysis #####
 
-### Turning the Workflows into Preregistration Templates ###
+#### Design ####
+
+##### Translating Workflows into a Preregistration Template #####
 
@@
 
+[@Moallemi2019] featuresof a good guideline:
+
+a) A standard step-wise structure relevant to a variety of (exploratory modelling) methods in robust decision
+
+support.
+
+(b) The enumeration of (methodological and mixing design) choices available at each step to reflect on alternative
+
+options available to researchers.
+
+(c) Examples/illustrations to clarify the step of the guideline and choices at each step.
+
+(d) An easy to follow checklist to promote wider adoption.
+
 **3\. Designing a new methodology for implementing preregistration in model-based research**
@@
 
-### 'In-situ' Evaluation and Testing of Templates ###
+##### Redesigning a new methodology for preregistering model-based ecology and conservation research #####
+
+##### 'In-situ' Evaluation and Testing of Templates #####
 
@@
 We tested and evaluated the templates on the ability to meet these criteria by using a model-based research case-study \<see box\>.
-Rather than 'digging back' through the version history of a completed preregistration to ask researchers about the process and how it had been developed in detail \<like Pu et al\>, we chose to leverage the version-control and collaborative project management features of GitHub \<cite\> as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to 'live-develop' the template template.
+Rather than 'digging back' through the version history of a completed preregistration to ask researchers about the process and how it had been developed in detail \<like Pu et al\>, we chose to leverage the version-control and collaborative project management features of Git and GitHub [@BenBondLamberty:2018by; @Blischak:2016kr] as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to 'live-develop' the template template.
 For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration, the suggested change, and its justification would be both recorded in GitHub and linked to the actual preregistration.
@@
 
-This will be a box or a table with numbered items, with just the preregistration item directives or question.
-The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
-Could possibly go into the appendix, but I think given that it's the primary research output, it should go into the main-text of the manuscript.
-
 ----
@@
 By articulating this workflow, we explicitly state the workflow and sets of activities that the template presumes.
-We identified and coded decision-points into 6 'phases' in the process of developing, evaluating and analysing a model or models in ecological application: 'problem formulation', 'define conceptual model', 'formalise and specify model', 'Model Validation and Evaluation', 'Model Analysis', 'Model Application'.
-These phases are comprised of a series of smaller 'steps' and 'substeps'.
-These are defined and described in detail within Table X.
-In terms of how this translates into the format of the preregistration template, these broad phases and steps are used to structure the template, and the descriptions and definitions of the phases and steps are used as 'help-text' in the template interface.
+We identified and coded decision-points into 6 'phases' in the process of developing, evaluating and analysing a model or models in ecological application: 'problem formulation', 'define conceptual model', 'formalise and specify model', 'model validation and evaluation', 'model analysis', 'model application'.
+These phases are comprised of a series of smaller 'steps' and 'substeps', which are defined and described in detail within Table X.
+The broad phases and steps are used to structure the template by grouping similar modelling activities together, and the descriptions and definitions of the phases and steps are used as 'help-text' in the template interface.
+
+We synthesised the literature on many different environmental modelling
+
+We synthesised modelling workflows from various
+
+Although there are many different environmental modelling frameworks
+
+We synthesised the literature on many different environmental modeling frameworks with the outputs of our workshop.
+
+\-
 
@@
 During our structured discussions, we identified the iterative nature of model development as a key barrier to adopting preregistration.
-\<Insert refs to literature describing this\>.
-\<Explain what I mean by iterative model development and provide examples\>
+
+\<Insert refs to literature describing this\>.  \<Explain what I mean by iterative model development and provide examples\>
+
+\<Explain why this is a barrier to preregistration\>
 
@@
 However, existing preregistration formats \<do not encourage/make the process of checking difficult\> --- static text-based documents are designed for authors to quickly input the desired information.
-In fact, *Pu et al* go so far as to say that this format "verge\[s\] on being write-only media" and are therefore completely ill-suited to the checking task.
+In fact, *Pu et al* go so far as to say that this format "verge[s] on being write-only media" and are therefore completely ill-suited to the checking task.
 
@@
 
+This will be a box or a table with numbered items, with just the preregistration item directives or question.
+The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
+Could possibly go into the appendix, but I think given that it's the primary research output, it should go into the main-text of the manuscript.
+
 ----
 
-### Adaptive Preregistration ###
+### Adaptive Preregistration for model-based research using GitHub ###
 
@@
 GitHub is used to track amendments to the preregistration --- each time an additional question is answered, or an alteration is made to the preregistration, the change should be *committed* to the *repository*.
-*Semantic versioning \<insert link\>* is used to document each version of the preregistration.
-\<explain semantic versioning in one sentence\>.
-Using *GitHub's* tag and release feature, each interim preregistration is tagged and released with its semantic version number.
+*Semantic versioning* is used to document each version of the preregistration.
+*Semantic versioning* is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it [@Kitzes2018vn;@Raemaekers2017].
+We have created our own *semantic versioning* scheme to be used to specifically document the incremental version of the preregistration, in tandem with *GitHub's* tag and release feature, each interim preregistration is tagged and released with its semantic version number.
 
@@
 
-### Preregistration Interface - Shiny App ###
+### A Shiny Preregistration Interface ###
 
@@
 
+*Problem*
+
+\> *Lack of discipline specific preregistration templates impedes the uptake of open-science tools, such as preregistration within ecology.*
+
+\> *Domain specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom unique to a particular domain-specific methodology.*
+
+*Overarching aim*
+
+*We assumed a UX or user-centred design approach to develop a preregistration template and methodology that accounts for the domain and methodology specific nuances in ecology.*
+
+------ *We did this by:*
+
+*1\.
+Identifying norms and practices to develop 'idealised' / 'generalised' research workflows that describe sets of commonly undertaken tasks (and some sort of order to reflect the workflow), within an ecological modelling research analysis (within applied settings??).
+(idealised + practical).*
+
+*2\.
+We then transformed these workflows into a 'preregistration template', by distilling a set of decision-items that, we think, should a) adequately constrain researcher degrees of freedom to mitigate the risk of Questionable Research Practices in an ecological modelling exercise and b) adequately facilitates* research/analytic transparency.
+We transformed the decision-steps (the highest resolution) into template items posed as directives or questions for preregistration users to complete, as appropriate, for their study's preregistration.
+
+\> we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both idealised norms and norms of practice.
+
+3\.
+Our UX approach revealed critical features of the model development process, namely the inherent iterative nature of model development, and requirements for model checking that forced us to reconsider the procedure for completing a preregistration suitable to ecological modelling.
+Consequently, we took an 'expanded view of preregistration', "Adaptive Preregistration", as proposed and first described by Srivastava.
+Adaptive preregistration
+
+Plans to deploy flexible strategies, and iterative interim preregistrations
+
+4\.
+
+\> We proposed a methodology for managing the preregistration process that leverages features of GitHub that implement .....
+This methodology (and interface?) deviates from the current/traditional deterministic, single-use write-once read-only document currently used in preregistration.
+
+5\.
+Furthermore, we created an interface for completing and amending interim preregistrations using shinyApps to facilitate uptake of preregistration in both pure and applied ecological modelling research.
+
+*consisting of a series of*
+
+*(What is a workflow)?? I think one of the CI papers defined this.*
+
+*research 'workflows'*
+
+*scientific norms and practices specific to ecological modelling*
+
+*---------*
+
 *We really didn't know what to expect when we began completing these preregistrations...*
@@
 This relates to what Pu et al.
-found where participants 'felt their research workflow improved by having preregistration as a "forcing function" \[...\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers' degree of freedmen advance.' (Page 13).
+found where participants 'felt their research workflow improved by having preregistration as a "forcing function" [...] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers' degree of freedmen advance.' (Page 13).
 For reviewers 'participant reports that pregistration makes the review process "easier".'
@@
 
+All questions are on the table. We approached this goal with a 'proof is in the pudding' mentality - what are the limits of preregistration? Does it slow down the research process?
+
 **------**
+
+**Pu et al: found researchers thought preregistration would delimit creativity and flexibility.**
+
+**[@MacEachern2019]:**
 
@@
 
+Combining with other open science tools.
+
+I thinkthe preregistration can easily be usd with a reporting check
+
+It's, such as t proposed by Feng et al. The decisions and details are recorded within the preregistration. Along with any deviation from the initial plan.  
 --------------------------

### 2020-09-12 index.Rmd
revise working draft - restructure user design section

@@
 date: "`r Sys.Date()`"
+bibliography: ./data/references.bib
 output: 
@@
 
-### User-centred design approach ###
-
-**Designing Preregistration Templates with a User-Centred Design Approach**
+### Designing Preregistration Templates with a User-Centred Design Approach ###
 
 We took a user-centred design (UCD) approach to the task of translating preregistration into model-based research contexts for ecology and conservation.
-UCD is a philosophy for designing services or products that accounts for user requirements at all stages of the design process --- the goal of which is to generate both an effective and usable tool for specific people in specific contexts [@Pavelin2012].
+UCD is a philosophy for designing services or products that accounts for user requirements at all stages of the design process --- the goal of which is to generate both an effective and usable resource for specific people in specific contexts [@Pavelin2012].
 As Pavelin et al illustrate,  it is critically important that users be involved in the design process because it is difficult to predict how users will actually use a resource.
-This mirrors Pu et al's qualitative analysis into preregistration design, that a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings.
+This mirrors Pu et al's qualitative analysis of preregistration template design, finding that a designer's intention for preregistration and users' purposes for preregistration may be in conflict and undermine the overarching goal of preregistration --- to improve the reliability of scientific findings.
 Consequently, although user-centered design is typically used within software engineering domains, we heed Pu's advice that "It is crucial to treat preregistration [...] as a user-centered design problem to ensure preregistration achieves its intended goal."
 
-We borrow from the UCD methodology proposed by Pavelin et al.
-[-@Pavelin2012], and follow the core steps ![](./figures/user-centered-design-preregistration.png): 1.
-Define the goals of the resource under development and identify audience, 2.
-Characterise users and their needs, 3.
-Design mock-ups of your product, 4.
-Build prototypes of selected mock-ups and test with users, 5.
-Implement the product, writing technical implementations and building the product.
-In the case of this project, the 'product' consists of a preregistration template, specification of the internal logic of the template, including an updated domain-specific procedure for using the template, a user-interface for completing the preregistration, and a user-guide to accompany the template and interface.
+We borrow from the UCD methodology proposed by Pavelin et al. [-@Pavelin2012], and follow the core steps ![](./figures/user-centered-design-preregistration.png):
 
-#### Define Vision & Objectives ####
+1\. Define the goals of the resource under development and identify the audience,
+
+2\. Characterise users and their needs,
+
+3\. Design mock-ups of your product,
+
+4\. Build prototypes of selected mock-ups and test with users,
+
+5\. Implement the product, write technical implementations and build the product.
+
+In this case, the 'product' so-to-speak, consists of a preregistration template, specification of the internal logic of the template, including an updated domain-specific procedure for using the template, a user-interface for completing the preregistration, and a user-guide to accompany the template and interface.
+
+The first step in a user-centred design approach is to articulate the goals, objectives and target audience of the tool, consequently we define the intended purpose of the preregistration, the intended users of the templates, and their research contexts.
+These aspirations for the preregistration template are the guiding principles for the development and evaluation of ecology-specific preregistration, and have important implications for the design of the template - from its structure, content, user-interface and usage within the broader research workflow.
 
@@
 
-Defining the purpose and intended use-cases of a preregistration template has important implications for the design of the template --- from its structure, content, user-interface and use within the broader research workflow.
 We define the primary purpose of the template as "to delimit researcher degrees of freedom" when conducting model-based research in applied ecology and conservation.
 Although others have defined the purpose of preregistration to include 'delineating between *exploratory* from *confirmatory* research' \<refs\>, this is not the intended purpose of the template here.
-Beyond hypothesis-testing contexts, this is a simplifying dichotomy that, in practice, is more blurred and nuanced than the ideal.
+Beyond hypothesis-testing contexts, this is an idealised and simplifying dichotomy that, in practice, is more complicated and nuanced.
 For instance researchers might move back and forth between different reasoning-modes throughout the modelling cycle.
-Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, and consideration of research from philosophy of science.
+Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, and consideration of perspectives from philosophy of science.
 Consequently, we wish to deemphasise the purpose of distinguishing between exploratory and confirmatory research, and focus the design of the templates in being able to delimit researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and other questionable research practices, like cherry-picking, HARKing etc.
@@
 It should also enhance the research transparency of the study, ensuring that these choices are documented with sufficient transparency to let the reader make informed decisions about the credibility and reliability of the model and associated analysis and interventions.
+(MOVE RESEARCH TRANSPARENCY QUOTES FROM BELOW TO HERE).
 
@@
 
-**Use-Cases**
+**Preregistration use-cases**
 
@@
 
-#### User Research ####
-
-On the 4^th^ March 2020, we ran a series of structured activities designed to inform the development of two pre-registration templates, one for field-work and sampling design, the other for modelling and data-analysis.
+On the 4^th^ March 2020, we ran a series of structured activities designed to inform the development of two preregistration templates, one for field-work and sampling design, the other for modelling and data-analysis.
 We split the activities into two groups, to better focus discussion for each of these templates.
@@
 
-**Activity 1 - Capturing Scientific Workflows and Decision Points**
+Failing to make the 'internal model of preregistration' clear to users is a critical source of confusion among preregistration template users that may impede preregistration uptake, and undermine the ability of the template to achieve its objectives (Pu et al.
+2020).
+Consequently, we aim to articulate a generalised modelling workflow that is used to structure the preregistration template, and to guide the selection of items for inclusion in the template.
 
-The objective of this activity was to capture critical decision-steps in the design of both field sampling efforts and modelling exercises.
-This activity was intended to identify key decision-steps in the scientific process that should be included in a preregistration template.
-The first part of this activity was done as individuals, where we reflected on our scientific process for a recent or memorable research project.
-We listed the steps in the workflow from beginning to end.
-We attempted to document  critical decision-steps in this process that should be present on a preregistration template.
+## Workshop Activity 1: Identifying Ecological Modelling Workflows Norms and Practice
+
+The objective of this activity was to identify modelling workflows that capture critical decision-steps in the process of field sampling efforts and modelling exercises that should be included in a preregistration template.
 By "critical" we mean that without these decision-steps the template would not a) adequately constrain researcher degrees of freedom and b) transparently document the analytic decision-process.
-
-In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto a group workflow template This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase.
+The activity consisted of two-parts, first, we individually reflected on our scientific process for a recent or memorable research project, listing the steps we undertook from the beginning to end.
+In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto a group workflow template.
+This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase.
 The development of the draft modelling workflow was derived from previous work by Gould et al.
@@
 
-This second activity aimed to identify critical challenges and roadblocks to the implementation and uptake of preregistration in ecology and conservation, as well as their solutions.
+This second activity aimed to identify critical challenges and roadblocks to the implementation and uptake of preregistration in ecology and conservation, as well as brainstorming solutions to these challenges.
 We conducted two lots of structured discussions, one for each of the field and modelling template groups.
-Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature.
+Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature \<cite / link to OSF, or even discuss??\>.
 These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in activity 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.
 
-##### Workshop Analysis #####
+#### Analysis and Design ####
 
-#### Design ####
-
-##### Translating Workflows into a Preregistration Template #####
-
-**1\. Synthesise Workshop Outputs: Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
+**1\. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
 
@@
 A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management.
-Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substep' \<insert link to appendices or just present the final?\>.
+Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], \<insert link to appendices or just present the final?\>.
 Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps.
@@
 We translated the model-based workflow identified in part I of workshop analysis into a text-based preregistration template.
-Reflecting on existing templates \<link to OSF repository of templates \>, we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis.
-We synthesised the following ontologies:
+Reflecting on existing templates \<link to OSF repository of templates page\>, we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis.
+Following the features of a good guideline identified by Moallemi et al.
+[-@Moallemi2019], we provide examples of choices or illustrations to clarify each step of the preregistration, for parts of the modelling process that are semantically, epistemically and/or philosophically contentious or ambiguous \<link to paper describing contentious issues in modelling history\>.
+We synthesised the following ontology to build our preregistration template:
 
@@
 2\.
-Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps'.
-Corresponds to an activity or set of activities in a scientific workflow.
+Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow.
 \<again, these are being scrapped and turned into tool-tips.\>
@@
 
-[@Moallemi2019] featuresof a good guideline:
+**3\. Designing a new methodology for implementing preregistration in model-based research**
 
-a) A standard step-wise structure relevant to a variety of (exploratory modelling) methods in robust decision
-
-support.
-
-(b) The enumeration of (methodological and mixing design) choices available at each step to reflect on alternative
-
-options available to researchers.
-
-(c) Examples/illustrations to clarify the step of the guideline and choices at each step.
-
-(d) An easy to follow checklist to promote wider adoption.
-
-**3\. Designing a new methodology for implementing preregistration in model-based research**
+**What about the challenges and solutions part + the group discussion? Make sure to discuss the atlas TI part for this...**
 
@@
 We then used 'in situ evaluation' to test and evaluate both the template and procedure design.
-
-##### Redesigning a new methodology for preregistering model-based ecology and conservation research #####
-
-##### 'In-situ' Evaluation and Testing of Templates #####
 
@@
 
-Failing to make the 'internal model of preregistration' clear to users is a critical source of confusion among preregistration template users that may impede preregistration uptake, and undermine the ability of the template to achieve its objectives (Pu et al.
-2020).
-Consequently, in this section we articulate the inner logic of our preregistration templates that informs the the content, structure of the template, and our proposed procedure and interface for implementing the template.
+The modelling workflow + adaptive preregistration process described below articulates the inner logic of the preregistration templates that underpins the content, structure of the template, as well as our proposed procedure and interface for implementing the template.
 
@@
 
-This will be a box or a table with numbered items, with just the preregistration item directives or question.
-The help-text describing each phase or key steps in the workflow is removed, as this goes in the previous table under 'model scientific workflow'.
-Could possibly go into the appendix, but I think given that it's the primary research output, it should go into the main-text of the manuscript.
+The template is designed to be used within an adaptive preregistration methodology \<link to section\>.
 
@@
 
-I thinkthe preregistration can easily be usd with a reporting check
+I think the preregistration can easily be usd with a reporting check
 

### 2020-09-16 index.Rmd
Complete first draft abstract. Rework iterative model development section. rework adaptive preregistration and model checking sections.

@@
 date: "`r Sys.Date()`"
-bibliography: ./data/references.bib
 output: 
@@
 
-*The Abstract must not exceed 350 words and should list the main results and conclusions, using simple, factual, numbered statements:*
+**Title:** "But I can't preregister my research": Improving the reproducibility and transparency of ecology and conservation with preregistration for model-based research
 
-*Point 1: set the context for and purpose of the work;*
+**Abstract:** *Ecology is at risk of a 'reproducibility crisis', in particular, there are serious transparency problems in the reporting of ecological models, which is concerning because a lack of complete and transparent reporting practices provide the conditions in which questionable research practices are likely to arise, leading to irreproducible findings.
+Irreproducibility of model-based research threatens the progress of both scientific discovery in 'pure' research, and risks wrong or sub-optimal conservation and management decisions being implemented when models are used in decision-support.
+Preregistration is an open-science practice adopted in scientific disciplines that have confronted the reproducibility crisis, which aims to improve research transparency and mitigate questionable research practices.
+There has been little uptake of preregistration in ecology and conservation, and within disciplines wherein preregistration has been adopted and become common practice, many modellers and researchers engaged in non-hypothesis testing research have vehemently eschewed preregistration on the grounds that existing preregistration templates are irrelevant to their research because they are focussed on null-hypothesis significance testing.
+We took a user-centred design approach to the task of translating preregistration into model-based research contexts for ecology and conservation.
+Structured discussion and collaborative activities underpinned the development of a preregistration template that captures critical analysis decisions in the process of the model development cycle.
+The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the linear procedure of preregistration in which a researcher shifts from ideation to analysis planning, to analysis and then reporting.
+We propose a methodology for implementing 'adaptive preregistration' leveraging git and GitHub, which accommodates common practices in modelling where some analysis decisions are difficult or cannot be made in advance, where downstream decisions might be dependent on the outcome of earlier decisions, and where preliminary analyses might need to be conducted before proceeding with further modelling.
+We tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
+This research provides an extensible template and methodology for implementing adaptive preregistration of ecological models, to be used in either pure or applied settings.
+We hope that this research will facilitate wider uptake of preregistration in ecology and that our proposed methodology for implementing adaptive preregistration might be applied in model-based research within science more broadly.
+Modellers need no longer cry "but, I can't preregister my research."*
 
-*Point 2: indicate the approach and methods;*
-
-*Point 3: outline the main results;*
-
-*Point 4: identify the conclusions and the wider implications.*
-
-**Keywords** *(no more than 8 for MEE).***:** *preregistration, preregistration template, modelling, applied ecology, conservation decision-making, metaresearch, reproducibility, non-hypothesis testing(?).*
+**Keywords** *(no more than 8 for MEE).***:** *preregistration, transparency, reproducibility, modelling, ecology, conservation, open-science, research software engineering.*
 
@@
 However, concerns about poor model reporting practices are not new, and have been raised repeatedly --- nearly a decade ago researchers concluded that most model reporting was neither transparent or comprehensive [@Schmolke:2010fd; @deVos:2011], and that a culture of model reporting was yet to be established [@Grimm:2014es].
-Moreover, most models were assessed predominantly on the basis of scientific originality during peer-review  [@Schmolke:2010fd] to the effect that the burden of evaluating models for credibility is placed onto the reader or model-user, rather than the peer-review process, especially in the case of extremely complex models [@Alexandrov:2011iv].
+Moreover, most models are assessed predominantly on the basis of scientific originality during peer-review  [@Schmolke:2010fd] to the effect that the burden of evaluating models for credibility is placed onto the reader or model-user, rather than the peer-review process, especially in the case of extremely complex models [@Alexandrov:2011iv].
 
-Despite attempted to promote thorough and transparent model documentation using 'Good Modelling Practice' (e.g.
-[@Jakeman:2006ii;@Alexandrov:2011iv]), more recent quantitative assessments of transparency and reproducibility demonstrate that the transparent model reporting is still an ongoing problem.
+Despite attempts to promote thorough and transparent model documentation using 'Good Modelling Practice' (e.g.
+[@Jakeman:2006ii;@Alexandrov:2011iv]), more recent quantitative assessments of transparency and reproducibility demonstrate that transparent model reporting is an ongoing problem.
 Just as Jakeman argued that many models in the published literature lack a clear statement of model objectives and purpose, lack descriptions of model assumptions and implications, and also fail to report model results, including validation and evaluation in 2006, [-@Jakeman:2006ii], [@Feng2019]'s quantitative assessment of Ecological Niche Models found that reporting of model evaluation and model results was less than adequate.
@@
 
-The modelling workflow + adaptive preregistration process described below articulates the inner logic of the preregistration templates that underpins the content, structure of the template, as well as our proposed procedure and interface for implementing the template.
-
 **Modelling workflows for Ecology and Conservation**
 
-Table X, below, describes the 'model scientific workflow' underpinning the template, which was identified through the structured collaborative workshop, literature review and qualitative coding.
+Figure X and Table X, below, describes the 'model scientific workflow' underpinning the template, which was identified through the structured collaborative workshop, literature review and qualitative coding.
 By articulating this workflow, we explicitly state the workflow and sets of activities that the template presumes.
@@
 
-We synthesised the literature on many different environmental modelling
+@TODO - anything about *applied contexts??* Aligning with management? (In terms of the structure and steps above).
 
-We synthesised modelling workflows from various
-
-Although there are many different environmental modelling frameworks
-
-We synthesised the literature on many different environmental modeling frameworks with the outputs of our workshop.
-
-\-
-
-@TODO - anything about *applied contexts??* Aligning with management? (In terms of the structure and steps above).
+One of the key concerns raised at the workshop by those who work in applied contexts, is that preregistration could be a useful tool, but needed to consist of items that were useful to applied contexts.
+Another thing is that they felt that it could a) slot into existing paper-work and b) be used to make internal data analysis plan requirements more useful.
 
@@
 
-During our structured discussions, we identified the iterative nature of model development as a key barrier to adopting preregistration.
+During our structured discussions, we identified the iterative nature of model development as a significant barrier to adopting preregistration.
+ Translating the preregistration process to a modelling research context is thus challenging because the iterative nature of model development conflicts with the inner logic of existing preregistration templates that presume a linear research workflow.
+The process of preregistration as it is currently implemented is tailored towards the hypothetico-deductive model of the scientific method.
+As such, the process of preregistration is completely distinct from, and precedes the implementation of the analysis plan laid out in the preregistration.
+In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" (Nosek et al.
+2018) or prior to analysing the data [@Mertens2019].
+The researcher shifts from ideation, to study/analysis design and preregistration, to collecting and analysing the data, writing the report or manuscript, to documenting the research and publishing ([*cos.io/rr*](http://cos.io/rr)) [See also Figure 1 in @Pu2019].
+The iterative nature of model development, however, inherently precludes this model of preregistration from being applied to model-based research in ecology and conservation.
 
-\<Insert refs to literature describing this\>.  \<Explain what I mean by iterative model development and provide examples\>
+Firstly, as Srivastava notes [-@Srivastava2018], there are some decisions that are either difficult to anticipate, or simply cannot be made in advance.
+This is particularly true for decisions occurring at the later phases of the modelling construction and development process (Fig X), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses.
+Some items on the template might not be able to be answered on the first go by the researcher until the model is fully or at least close to fully specified, for example, specifying precisely what and how sensitivity analyses or uncertainty analyses will be conducted.
 
-\<Explain why this is a barrier to preregistration\>
+Secondly, very often in modelling, preliminary or investigatory analyses might need to be conducted before being able to specify future decision-steps in the analysis plan.
+For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation.
+In addition, sometimes the results of these analyses and checks mean that the modeller is forced to return to earlier decision-points and change the analysis.
+For instance, the model might fail to converge, or is saturated and runs out of variation to apportion, such that the model must be respecified.
+Other times, the modelling process itself may generate knowledge and learning about the system, and the model structure changes throughout the process of model development.
 
-Typically, the process of preregistration is completely distinct from and precedes the implementation of the analysis plan laid out in the preregistration.
-In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" (Nosek et al.
-2018) or prior to analysing the data (Ref, or link to secondary data preregistration template).
-This results in a linear research workflow, whereby a researcher shifts from the process of ideation and design, documentation of that design in a preregistration, to the process of action, where the researcher executes the plan specified in the preregistration \<insert link to this fig..\>
-
-Translating the preregistration process to a modelling research context is thus challenging because the iterative nature of model development conflicts with the inner logic of existing preregistration templates that presume a linear research workflow.
-For instance, some modelling decisions are inherently dependent on the outcomes and results of previous decisions and analyses, which involves conducting some preliminary or investigatory analyses before being able to specify future decision-steps in an analysis plan, contravening the critical(?adj) feature of preregistration of specifying the analysis plan prior to seeing the data.
-Moreover, some items in the template, particularly at later stages of modelling, might not be able to be answered by the researcher until the model is fully or close to fully specified on a first-pass (e.g. specifying exactly what sensitivity analyses or uncertainty analyses will be conducted).
-This might especially hold true for more complex process models.
-\<Mention that this is not a flaw in preregistration, but rather reflects the inner logic of preregistration for hypothesis-testing contexts\>
-
-__The *checking* problem__
-
-Authors, reviewers, editors and readers may all engage in the process of checking a report or manuscript against its preregistration to verify that the study and analyses were conducted as specified in the preregistration.
-However, existing preregistration formats \<do not encourage/make the process of checking difficult\> --- static text-based documents are designed for authors to quickly input the desired information.
-In fact, *Pu et al* go so far as to say that this format "verge[s] on being write-only media" and are therefore completely ill-suited to the checking task.
-
-The task of model checking is further complicated by the non-linear and iterative nature of model development.
-It is difficult to verify the conducted study and analysis against the preregistered analysis for existing preregistration formats used in hypothetico-deductive research .
-However, when the modelling and preregistration is iterative, such that there might be interim preregistrations, then this means that there are *multiple* versions of the preregistration to check against the completed analysis.
-
-The question we faced then is, how do we design both the format of the template, and its procedure for completion by the author, so that we facilitate preregistration checking --- especially when this task is complicated by the iterative nature of model development?
-*Pu et al* put forth two design elements that may facilitate checking.
-They recommend designs that "make relevant preregistration easier to *query*" and that "encourage wider *coverage* of preregistration content during review".
-
-For *Pu et al.* The process of *querying* a preregistration consists of "paper-to-preregistration" and "preregistration-to-paper" lookup whereby the reader searches for relevant information from the paper in the preregistration and *vice-versa*.
-This requires that the report of paper must be specific in referring to parts of the preregistration, *and*, that the author has tagged the relevant part of the preregistration in their report or manuscript.
-
-Consequently we propose a methodology for managing the preregistration process that leverages features of GitHub that implement the task of linking the actual study and analysis that was conducted to the pre-specfied plan in the preregistration \<link to part of GH methodology that explains how checking is accounted for in the procedural design\>.
-In addition, we designed our shiny interface for completing the preregistration to include a 'checking mode' that easily allows the author to state how and why the the analysis might have diverged from the planned \<link to part of shiny app interface that discusses checking mode\>.
-
----------
-
-*I'm not exactly sure whether this content should go here in a summarised form or whether it should go into its respective places coming up in the manuscript and we just link to it.*
-
-*How does our GH procedure incorporates checking:*
-
-1\. Using issues to track committed work on the same analysis task, and that issue is spawned from a template item so the conducted work is inherently linked to that initial plan.
-
-2\. In the preregistration document
-
-*How does our shiny-preregistration Interface incorporates checking:*
-
-\- Checking mode - The author can switch to checking mode after having conducted the analysis for a given PRT item. *Or* after the preregistration is complete.
-
-What do you think about this process?
-What if we allowed check boxes in our preregistration software design so that the author could switch to 'checking mode' and they can tick some boxes Yes / No / Partial match.
-For No and Partial, they briefly describe \_how\_ it varies from the preregistration, and why (could link to the GitHub issue thread where conversation and decisions were logged during the process of conducting the analyses).
-
-This is going to take some further thought because there might be \_interim\_ versions of the document.
-Our process means that the author keeps adapting the preregistration until they actually go ahead and implement it...
-so doesn't this mean that they will \_always\_ follow their latest version of the preregistration??
-I guess then checking entails a comparison of first and interim preregistrations against the actual conducted analysis, not just the final.
-
-AND it is possible that in other cases the author deviated from their plan for some reason, and didn't follow our proposed process using GH.
-So we want to allow for this to occur, but for the author to be able to state \_that\_ this happened, \_how\_ it differs, and why.
-
----------
+In all of these examples, during the course of model-based preregistration a researcher would switch back and forth between preregistration and analysis, implementing parts of the analysis plan that were specified in the previous preregistration step, and letting the outcomes in that analysis step inform the next specification of the analysis.
+This contravenes a critical feature of preregistration as implemented in hypothetico-deductive frameworks, where the analysis plan must be fully specified prior to observing and analysing the data.
+This is not a flaw of existing preregistration, but rather reflects the inner-logic of preregistration for hypothesis-testing contexts.
+The procedure for implementing preregistration in modelling must therefore account for this iterative and non-linear feature of model development, should we wish to preregister model-based research.
 
@@
 
-----
-
-### Adaptive Preregistration for model-based research using GitHub ###
-
-The process of preregistration as it is currently implemented is tailored towards the hypothetic-deductive model of the scientific method, and thus follows a linear process, where the process of planning and preregistering is distinct and completely precedes the execution of that plan.
-The researcher shifts from ideation, to study/analysis design and preregistration, to collecting and analysing the data, writing the report or manuscript, to publishing ([*cos.io/rr*](http://cos.io/rr)) \<See also Figure 1 in Pu et al.\> The iterative nature of model development, however, inherently precludes this model of preregistration from being applied to model-based research in ecology and conservation.
-
-Firstly, as Srivastava notes, there are some decisions that are either difficult to anticipate, or simply cannot be made in advance \<Srivastava\>.
-This is particularly true for decisions occurring at the later phases of the modelling construction and development process (Fig X), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses.
-For example, in the modelling process we have described (previous section), some decisions cannot be made unless you have a fully specified model, such as \<sensitivity analysis, and other model evaluation tasks and test\>.
-
-Secondly, very often in modelling, preliminary or exploratory analyses must be implemented before decisions can be made about how to proceed.
-For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation.
-Sometimes the results of these analyses and checks mean that the modeller is forced to return to earlier decision-points and change the analysis.
-For instance, the model might fail to converge, or perhaps there is insfufficient data to fit the desired model such that the model runs out of variation to apportion to all the different variables \<there's a name for this\>.
-
-Consequently, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by \<Srivastava\>.
+In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by \<Srivastava\>.
 This view of preregistration breaks with the current model of preregistration, wherein the author writes a single deterministic preregistration containing a rule for every decision, and shifts from ideation, preregistration to execution of the plan.
@@
 
-Given the iterative nature of adaptive preregistration then, the current convention of the static text-based document format of preregistration is inadequate.
-If there are interim versions of a preregistration, each version must be time-stamped and *version-controlled*, such that it is clear not just that two or more versions differ from each other, but it is apparent in exactly *how* they differ.
-Moreover, the *checking problem*, is exacerbated --- for flexible modelling and analysis strategies, we need to be able to link the results of preliminary analyses back to the specified strategy: the link between a particular triggered decision, and the analysis outcome that triggered it must be explicit.
-In addition, if the results of some preliminary analyses cause revisions to the model itself or a different modelling procedure or approach to be selected, then the trigger for this decision must be explicitly linked to the new version of the preregistration.
+The iterative nature of model construction and development complicates what is known as the preregistration *checking problem* even further.
+Although there are several aspects to the *checking problem*, here we focus on the part of the checking problem that includes *how do we facilitate checking?* Authors, reviewers, editors and readers may all engage in the process of checking a manuscript or published paper against its preregistration to verify that the study and analyses were conducted as specified in the preregistration.
+Existing pre-registration formats hinder checking, because the prevailing format of a single static text-based document is designed for authors to very quickly input the required information, rather than to facilitate a comparison of the reported analysis and results against the planned analysis within the preregistration.
+In fact, *Pu et al* go so far as to say that this format "verge[s] on being write-only media" and are therefore completely ill-suited to the checking task.
+The task of model checking is further complicated by the non-linear and iterative nature of model development, and by our proposed adaptive preregistration methodology because in most modelling studies there will likely be interim preregistrations, such that there are *multiple* versions of the preregistration to check against the completed analysis.
+
+In redesigning the procedure and interface for completing a preregistration using an adaptive preregistration process, we followed the advice by *Pu et al.
+[-@Pu2019]* who recommended two design elements that may facilitate checking.
+Firstly, that the design should "make relevant preregistration easier to *query"* and secondly it should "encourage wider *coverage* of preregistration content during review." The process of *querying* a preregistration consists of "paper-to-preregistration" and "preregistration-to-paper" lookup whereby the reader searches for relevant information from the paper in the preregistration and *vice-versa*.
+This requires that the report or paper must be specific in referring to parts of the preregistration, *and*, that the author has tagged the relevant part of the preregistration in their report or manuscript.
+
+Given the iterative nature of adaptive preregistration then, the current convention of the single-use, deterministic and static text-based document format of preregistration is inadequate.
+Interim versions of a preregistration must be time-stamped and ideally *version-controlled*, such that it is clear not just that two or more versions differ from each other, and in what order they were created, but so that it is explicitly clear in how they differ.
+To facilitate preregistration checking of flexible modelling and analysis strategies, we must link the results of both final and preliminary analyses back to the specified strategy: the link between a particular triggered decision, and the analysis outcome that triggered the decision must be explicit.
+Additionally, if the results of some preliminary analyses cause revisions to the model itself or a different modelling procedure or approach to be selected than was originally or previously planned, then the trigger for this decision must be explicitly linked to the next version of the preregistration.
+
+Consequently we propose a methodology for implementing adaptive preregistration that leverages features of GitHub to also facilitate the task of linking the actual study and analysis conducted to the pre-specified plan in the preregistration \<link to part of GH methodology that explains how checking is accounted for in the procedural design\>.
+In addition, we designed a shiny interface for completing the preregistration to include a 'checking mode' that easily allows the author to state how and why the the analysis might have diverged from the planned \<link to part of shiny app interface that discusses checking mode\>.
+
+----
+
+### Adaptive Preregistration for model-based research using GitHub ###
 
@@
 
-I think the preregistration can easily be usd with a reporting check
+I thinkthe preregistration can easily be usd with a reporting check
 
 It's, such as t proposed by Feng et al. The decisions and details are recorded within the preregistration. Along with any deviation from the initial plan.
+
+**Continuous Improvement and Feedback**
+
+One of the key features of user-centred design is the notion of 'continuous feedback'.
+
+We encourage modellers to preregister their research using our template and methodology here, and to provide feedback and suggestions for improvement.
+Our template + shiny app is hosted on \<insert GitHub link\>, suggested changes may be submitted in the form of a pull-request to the repository \<link to how to make a pull request\>, or by posting a new issue \<link to creating a new issue on the repo\>.
+The repository and issues are publicly available, so any discussion will be transparent and open to the ecological modelling community.
 --------------------------

### 2020-09-21 analysis/preregistration_template.Rmd
Update preface to preregistration for template review by Libby

@@
 ---
-title: "Ecology and Conservation Modelling Preregistration Template -- DRAFT"
+title: "Preregistration Template for Modelling in Ecology and Conservation"
 author: "Elise Gould, Megan Good, Libby Rumpff, TBD:workshop collaborators"
@@
 output:
-  github_document: default
+  word_document
 bibliography: ../data/references.bib
@@
 
-This template has been collated from a combination of literature reviews of the 'Good Modelling Practice' and other modelling literature, as well as a collaborative workshop at DELWP.
-Chris Jones has kindly volunteered one component of the VEFMAP project investigating vegetation responses to environmental flows as a testing ground for the use of preregistration templates in applied ecological modelling research.
-This template will be used to preregister this research, and we will use the VEFMAP project to both develop and evaluate the template 'in situ'.
+This template has been collated from a combination of literature reviews of the 'Good Modeling Practice' and other modeling literature, as well as a collaborative workshop at DELWP/ARI.
+Chris Jones has kindly volunteered one component of the VEFMAP project investigating vegetation responses to environmental flows as a testing ground for the use of preregistration templates in applied ecological modeling research.
 
@@
 These have been defined and described to some extent throughout the template.
+This text (usually italicised) will eventually be converted to 'help-text', in the form-of mouse-over pop-ups, or else prompt-text within a text-box.
+Other parts of this text will be pulled into a supporting diagram describing the modelling workflow stages.
 
-## iterative model development and a 'living preregistration'
+## Adaptive Preregistration
 
-One of the major challenges in designing a template for modelling is that the process of model development is inherently iterative.
-Typically, the process of preregistration is completely distinct from and precedes the implementation of the analysis plan laid out in the preregistration.
-However, translating the preregistration process to a modelling research context is not so clear cut because some decisions are inherently dependent on the outcomes and results of previous decisions and analyses.
-For example, which sensitivity analyses will be conducted will depend on the relationships between variables found during earlier earlier stages in the modelling process.
+The following is an excerpt from my manuscript as at 21 September 2020.
+It describes the procedure for how this template is meant to be used within a modeling context.
+I hope to write this up as a guide (including screen-shots) that will accompany the template.
 
-One method around this is to accept that the current model of preregistration is not going to work for iterative model development.
-Again, the more complex the model, the more likely this is to be the case.
-Consequently I propose an 'adaptive preregistration' model  that I think will work for modeling research contexts.
+Adaptive Preregistration for model-based research using GitHub
+We propose leveraging both the version-control and the project management features of GitHub <link to GitHub> as a tool for implementing adaptive preregistration <link to box>.
+When a document is version-controlled using git and GitHub, each version of a document is time-stamped, and is assigned its own unique document identifier, or commit-hash.
+Because GitHub documents are ‘time-stamped’ the genesis of the preregistration from one version to the next is apparent.
+Moreover, using the diff view of a document on GitHub, exactly how a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted <insert picture>.
+Since the preregistration document would exist within the project repository containing the data and modelling and analysis code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated version of the plan — thus facilitating the process of model checking.
+By providing a method for contrasting the actual analysis undertaken against the registered analysis, this is a mechanism for allowing reviewers to properly evaluate the preregistration against the reported analysis.
+It thus provides a way of marking preregistered parts of a report from non-preregistered parts of the analysis.
 
-In this model of preregistration, all questions that can be answered, should be answered before any data analysis or modelling proceeds.
-For some items, particularly during model evaluation and analysis, perhaps only responses of a general nature can be provided.
-That is OK.
-As the modelling proceeds, further detail can be provided and the next phase of analysis can be preregistered.
-Where the results of model evaluation and analysis reveal that there are problems with the model, or perhaps assumption violations, plans can be changed, and again the next phase of analysis can be preregistered.
+How it works….
 
-A static document for this type of process doesn't really make sense.
-I propose leveraging the code and project management tools within the GitHub platform as a way of hosting an 'adaptive' preregistration.
-Because GitHub documents are 'time-stamped', the genesis of the preregistration from one version to the next can be traced and tagged with versions.
-Since the document would exist in the repository with the code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated plan.
+GitHub is used to track amendments to the preregistration — each time an additional question is answered, or an alteration is made to the preregistration, the change should be committed to the repository.
+Semantic versioning is used to document each version of the preregistration.
+Semantic versioning is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it  .
+We have created our own semantic versioning scheme to be used to specifically document the incremental version of the preregistration, in tandem with GitHub’s tag and release feature, each interim preregistration is tagged and released with its semantic version number.
+
+All questions that can be answered should be completed before any data analysis or modelling proceeds.
+The preregistration should be completed until a) no further detail can be supplied or b) no additional template item can be answered until preliminary investigations or analyses are undertaken (Fig X, steps 1 - 3).
+
+Shifting from planning to execution
+
+Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding GitHub issue created (Figure X, step 3, document).
+<define GitHub issue, https://guides.github.com/features/issues/>.
+Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an issue’s thread.
+Each issue is assigned a unique number.
+Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the issue’s thread (Fig X, step 4, Do) <screen-cap?>.
+The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new release.
+
+… And back again.
+
+As the modelling proceeds, further detail on later phases of the analysis can be iteratively updated and preregistered.
+Where the results of model evaluation and analysis reveal that there are problems with the model, plans can be changed, and again the next phase can be preregistered.
+For example, if a researcher finds that assumptions are violated or other unexpected results occur that force a change to the planned analysis, the preregistration can be updated based on the findings of those analyses.
+When this occurs, the major version number of the preregistration document should be incremented (Fig X steps 5 - 6).
+Since the findings of the analyses are tracked within the relevant issue for that template item, and the issue is recorded in the preregistration itself, the trigger for the change in the plan is made explicitly clear.
+
+For each altered or new preregistration item, a the issue thread should either be updated, or a new issue created respectively (Fig X, step 6).
+Those analyses are then either reconnected based on the revised plan or the new investigatory analyses are then conducted (Fig X, step 7).
+
+This process continues until the preregistration is complete, and the researcher can continue to execute the plan as it has been fully described in the final version of the preregistration document (Fig X, step 8 - 9).
+
+Flexible Strategies, Preliminary Analyses or Investigations
+
+When a researcher needs to conduct preliminary analyses or check parts of the data before committing to a particular decision about the model or analysis, they can specify a flexible strategy in their response to that preregistration item by:
+
+- stating what needs to be known to move forward with the modeling and analysis
+
+- explaining how the researcher will test this, and what parts of the data will be used in that analysis
+
+- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.
+
+![Adaptive Preregistration Workflow - Research Workflow for implementing a ‘adaptive’ preregistration in a modelling context](../figures/LivingPreregistration.png)
 
@@
 # 1. Problem Formulation
-***HF comment: I think this paragraph makes things more rather than less confusing - and I wonder if its necessary given that each point is stepped through below. I would recommend removing it***
 
 *Specifies the decision-making context in which the model will be used or the intended scope and context of conclusions.
-Important components include the decision maker and stakeholders (including experts) and their view on: i) the nature of the problem or decision addressed and how the scope of the modelling tool fits within the (broader) context (i.e.
-model purpose; ii) the spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they forsee unnacceptable outcomes that need to be represented in the model (i.e.
+Important components include the decision maker and stakeholders (including experts) and their view on: i) the nature of the problem or decision addressed and how the scope of the modeling tool fits within the (broader) context (i.e.
+model purpose; ii) the spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they foresee unacceptable outcomes that need to be represented in the model (i.e.
 as constraints), and; vi) what future scenarios does the model need to account for (noting this may be revised later).

### 2020-10-13 supp-mat/AdaptivePreregistration.Rmd
fix #52 Add Adaptive Preregistration using GitHub vignette

---
title: "Adaptive Preregistration for Model-Based Research with GitHub"
author: "Elliot Gould"
date: "`r Sys.Date()`"
output: tint::tintHtml
link-citations: yes
latexfonts: 
  - package: lato
    options: default
  - package: FiraMono
linkcolor: "0.3,0.3,0.6"
bibliography: ../data/references.bib
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

# Background & Motivation

```{r, fig.margin = TRUE, fig.width=10, fig.height = 5, echo = FALSE, fig.cap= "Figure 1: Conventional Preregistration Process, as illustrated by Pu et al (2019)."}
magick::image_read(path = "./Pu_et_al_fig_prereg.png")
```

The iterative nature of model development is a significant barrier to adopting preregistration for model-based research. Translating the preregistration process to a modelling research context is challenging because the iterative nature of model development conflicts with the inner logic of existing preregistration templates that presume a linear research workflow. The process of preregistration as it is currently implemented is tailored towards the hypothetico-deductive model of the scientific method. As such, the process of preregistration is completely distinct from, and precedes the implementation of the analysis plan laid out in the preregistration. In fact, preregistration has been defined as “the action of confirming an unalterable version of one’s research plan prior to collecting data” (Nosek et al. 2018) or prior to analysing the data [@Mertens2019]. In this process, researcher sequentially shifts from ideation, to study/analysis design and preregistration, to collecting and analysing the data, writing the report or manuscript, to documenting the research and publishing (Figure 1). The iterative nature of model development, however, inherently precludes this model of preregistration from being applied to model-based research in ecology and conservation.

## Iterative Cycle of Model Development

Firstly, as Srivastava notes [-@Srivastava2018], there are some decisions that are either difficult to anticipate, or simply cannot be made in advance. This is particularly true for decisions occurring at the later phases of the modelling construction and development process (Figure 2, TBC), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses. Some items on the template might not be able to be answered on the first go by the researcher until the model is fully or at least close to fully specified, for example, specifying precisely what and how sensitivity analyses or uncertainty analyses will be conducted.

Secondly, very often in modelling, preliminary or investigatory analyses might need to be conducted before being able to specify future decision-steps in the analysis plan. For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation. In addition, sometimes the results of these analyses and checks mean that the modeller is forced to return to earlier decision-points and change the analysis. For instance, the model might fail to converge, or is saturated and runs out of variation to apportion, such that the model must be respecified. Other times, the modelling process itself may generate knowledge and learning about the system, and the model structure changes throughout the process of model development.

In these examples, if an analyst were to try and preregister their research, they would switch back and forth between preregistration and analysis, implementing parts of the analysis plan that were specified in the previous preregistration step, and letting the outcomes in that analysis step inform the next specification of the analysis. This contravenes a critical feature of preregistration as currently implemented in hypothetico-deductive frameworks, where the analysis plan must be fully specified prior to observing and analysing the data. This is not a flaw of existing preregistration, but rather reflects the inner-logic of preregistration for hypothesis-testing contexts. The procedure for implementing preregistration in modelling must therefore account for this iterative and non-linear feature of model development, should we wish to preregister model-based research.

## The *Checking Problem*

The iterative nature of model construction and development complicates what is known as the preregistration checking problem even further. Authors, reviewers, editors and readers may all engage in the process of checking a manuscript or published paper against its preregistration to verify that the study and analyses were conducted as specified in the preregistration. Existing pre-registration formats hinder checking, because the prevailing format of a single static text-based document is designed for authors to very quickly input the required information, rather than to facilitate a comparison of the reported analysis and results against the planned analysis within the preregistration. In fact, Pu et al. [-@Pu2019] go so far as to say that this format “verge[s] on being write-only media” and are therefore completely ill-suited to the checking task. The task of model checking is further complicated by the non-linear and iterative nature of model development, and by our proposed adaptive preregistration methodology because in most modelling studies there will likely be interim preregistrations, such that there are multiple versions of the preregistration to check against the completed analysis.

## Preregistration Template Format & Procedure

Given the iterative nature of adaptive preregistration then, the current convention of the single-use, deterministic and static text-based document format of preregistration is inadequate. Interim versions of a preregistration must be time-stamped and ideally version-controlled, such that it is clear not just that two or more versions differ from each other, and in what order they were created, but so that it is explicitly clear in how they differ. To facilitate preregistration checking of flexible modelling and analysis strategies, we must link the results of both final and preliminary analyses back to the specified strategy: the link between a particular triggered analysis decision, and the analysis outcome that triggered the decision must be explicit. Additionally, if the results of some preliminary analyses cause revisions to the model itself or a different modelling procedure or approach to be selected than was originally or previously planned, then the trigger for this decision must be explicitly linked to the next version of the preregistration.

# Adaptive Preregistration

In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template format, we take an ‘expanded view of preregistration’, that fits with the concept of ‘adaptive preregistration’ proposed by Srivastava [-@Srivastava2018]. This view of preregistration breaks with the current model of preregistration, wherein the author writes a single deterministic preregistration containing a rule for every decision, and sequentially shifts from ideation, preregistration to execution of the plan. Two features define adaptive preregistration. It contains ‘plans to deploy flexible strategies’ [@Srivastava2018] wherein the author can supply a heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses. Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed [@Srivastava2018]. Thus, as the modeller proceeds through the model development process, they will shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed (Figure 3). Consequently we propose a methodology for implementing adaptive preregistration that leverages features of GitHub to also facilitate the task of linking the actual study and analysis conducted to the pre-specified plan in the preregistration.

![Figure 3: Adaptive Preregistration for model-based research using GitHub](../figures/LivingPreregistration.png)

## Adaptive Preregistration using GitHub

We propose leveraging both the version-control and the project management features of GitHub ([www.github.com](www.github.com)) as a tool for implementing adaptive preregistration. When a document is version-controlled using git and GitHub, each version of a document is time-stamped, and is assigned its own unique identifier, or commit-hash ^[https://docs.github.com/en/free-pro-team@latest/github/getting-started-with-github/github-glossary#commit-id]. Because GitHub documents are ‘time-stamped’ the genesis of the preregistration from one version to the next is apparent. Moreover, using the diff view of a document on GitHub, exactly how and where a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted.

```{r, fig.margin = TRUE, fig.width=10, fig.height = 5, echo = FALSE, fig.cap= "'Diff View' of changes made to a preregistration in process."}
magick::image_read(path = "./DiffView-PRT.png")
```

The preregistration document should be saved within the project repository that contains the data and modelling and analysis code, such that any results from analyses that cause the analysis plan to change can be linked and referenced in the updated version of the plan -- thus facilitating the process of model checking. By providing a method for contrasting the actual analysis undertaken against the preregistered analysis, this provides a mechanism for allowing reviewers to properly evaluate the preregistration against the reported analysis. It thus provides a way of marking preregistered parts of a report from non-preregistered parts of the analysis.

## How it works...

GitHub is used to track amendments to the preregistration — each time an additional template item is answered, or an alteration is made to the preregistration, the change should be committed to the repository. *Semantic versioning* is used to document each version of the preregistration. Semantic versioning^[The semantic version is a 3-part numeric identifier separated by points, e.g.: `0.0.1`, wherein the first number sequences a major version of the preregistration, the second number sequences a minor change, and the third number sequences small changes or patches, yielding: `MAJOR.MINOR.PATCH`. See https://semver.org/ for further information. We provide more details on how to use semantic versioning with adaptive preregistration below.] is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it [@Kitzes2018vn]. We have created our own semantic versioning scheme to be used to specifically document the incremental version of the preregistration, in tandem with GitHub’s tag and release feature, each interim preregistration is tagged and released with its semantic version number. All questions that can be answered should be completed before any data analysis or modelling proceeds. The preregistration should be completed until a) no further detail can be supplied or b) no additional template item can be answered until preliminary investigations or analyses are undertaken (Fig 3, steps 1 - 3).

```{r, fig.margin = TRUE, fig.width=10, fig.height = 5, echo = FALSE, fig.cap= "Initial Preregistration: All questions that can be answered should be completed prior to data analysis or modelling."}
magick::image_read(path = "./initial_prereg-closeup.png")
```

### …Shifting from planning to execution…

```{r, fig.margin = TRUE, fig.width=10, fig.height = 5, echo = FALSE, fig.cap= "Document: After the initial preregistration, document the initial preregistration using GitHub issues."}
magick::image_read(path = "./document-initial-prereg.png")
```

Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding GitHub issue^[https://guides.github.com/features/issues/] created (Figure 3, step 3, ‘Document’). Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an issue’s thread. Each issue is automatically assigned a unique number by GitHub. Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the issue’s thread (Fig 3, step 4, ‘Do’). The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new release.

### … And back again.

```{r, fig.margin = TRUE, fig.width=10, fig.height = 5, echo = FALSE, fig.cap= "Adaptive Preregistration - When Plans Change"}
magick::image_read(path = "./when-plans-change.png")
```

As the modelling proceeds, further detail on later phases of the analysis can be iteratively updated and preregistered. Where the results of model evaluation and analysis reveal that there are problems with the model, plans can be changed, and again the next phase can be preregistered. For example, if a researcher finds that assumptions are violated or other unexpected results occur that force a change to the planned analysis, the preregistration can be updated based on the findings of those analyses. When this occurs, the major version number of the preregistration document should be incremented (Fig 3 steps 5 - 6). Since the findings of the analyses are tracked within the relevant issue for that template item, and the issue is recorded in the preregistration itself, the trigger for the change in the plan is made explicitly clear.

For each altered or new preregistration item, the issue thread should either be updated, or a new issue created respectively (Fig 3, step 6). Those analyses are then either reconnected based on the revised plan or the new investigatory analyses are then conducted (Fig 3, step 7).

This process continues until the preregistration is complete, and the researcher can continue to execute the plan as it has been fully described in the final version of the preregistration document (Fig 3, step 8 - 9).

### Flexible Strategies, Preliminary Analyses or Investigations

When a researcher needs to conduct preliminary analyses or check parts of the data before committing to a particular decision about the model or analysis, they can specify a flexible strategy in their response to that preregistration item by:

- stating what needs to be known to move forward with the modelling and analysis and why the analysis is necessary
- explaining how the researcher will test this, and what parts of the data will be used in that analysis
- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.

# A step-by-step guide to implementing Adaptive Preregistration with GitHub

## The initial preregistration phase

### The first iteration of preregistration

1. Edit the document with your responses to the template items

- View the document in GitHub [here](https://github.com/egouldo/VEFMAP_VEG_Stage6/blob/master/analysis/preregistration_template.Rmd) 
- Click on The 'edit document' button:
![Screen Shot 2020-05-10 at 6 29 33 pm](https://user-images.githubusercontent.com/8400682/81494459-791b7080-92ec-11ea-979b-900317b4a9b6.png)
- You will be able to edit the document directly in GitHub, Edit away!

2. Commit changes

- If you are finished, or ready to save your changes by 'committing' them, commit by scrolling underneath the document editing window. There will be a box called "commit changes", enter in the commit message box the following:
- Your message should contain the following: `Initial Preregistration #IssueNumber`. Tag the Preregistration GitHub issue number in the commit message (For this repository, it's `#17`).
- Fill out the additional description box briefly describing the changes you just made
- Click "commit"

![Screen Shot 2020-05-18 at 12 19 29 pm](https://user-images.githubusercontent.com/8400682/82168463-2b14f700-9902-11ea-95a0-21fe385629bb.png)

## Incrementally Updating the Preregistration Plan

### Moving from general to specific: Incrementally updating the preregistration

You will likely find that some items on the template cannot be answered just yet.
One reason being that the level of detail required to adequately answer the template item just isn't clear yet, and more thought and time is needed, or you need to be further along the modelling process to be able to adequately answer the question. Another reason is that downstream decisions might depend on some preliminary investigations of the data -- for example, you might need to explore the shape of a measured variable before you can properly specify the full model.

In the latter situation, where possible, you can preregister your preliminary analysis: describing what the preliminary analysis entails, providing a short and relevant list of plausible outcomes, and describing for each of those possible outcomes how the dependent or down-stream decision-point will change depending on the outcome.

### Using GitHub to update the plan

We will use GitHub to track amendments to the preregistration for the two situations described above ( a) adding more detail, answering incomplete questions; b) Responding to findings of a preliminary analysis already described in your preregistration).

To make the changes to the preregistration document, follow the same instructions as for creating the initial preregistration above - simply edit the preregistration in GitHub, and commit the changes.

As for the situation where you are responding to findings of a preliminary analysis that you have already described in your preregistration, see the below sections on `Moving from planning to doing` and `Shifting back to planning from doing`.

## Moving from planning to doing

Before you begin to do or start anything you've planned within your preregistration, you **must** create a GitHub issue for that item, phase, or step in the modelling process, follow these steps:

1. View the preregistration document in Github, and click on the line-numbers on the left to highlight the PRT item text, as well as your response to the item. For example, let's say you want to start working the model checking, you would highlight the following in the document. Note that once you've highlighted some lines, a little box with an ellipsis appears on top of the lines numbers.
![Screen Shot 2020-05-13 at 1 06 25 pm](https://user-images.githubusercontent.com/8400682/81767208-2d9ad980-951b-11ea-8f59-6d6f583eb155.png) 
2. Click on the tiny box with the ellipsis, and a pop-up dialogue box will appear with 4 options: `Copy lines`, `Copy permalink`, `View git blame`, `References in new issue`. Select the last option `Reference in new issue.`
![Screen Shot 2020-05-13 at 1 06 58 pm](https://user-images.githubusercontent.com/8400682/81767344-80749100-951b-11ea-8b76-1e829528973a.png)
3. This will open up a new window with a new GitHub Issue for you to complete. Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted: 
![Screen Shot 2020-05-13 at 1 07 30 pm](https://user-images.githubusercontent.com/8400682/81767518-ef51ea00-951b-11ea-9398-18fffdcbbeae.png)
Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this:
![Screen Shot 2020-05-13 at 1 17 21 pm](https://user-images.githubusercontent.com/8400682/81767585-1c9e9800-951c-11ea-9d2c-46d2faf6368a.png)

Add any further information or technical details about going forward and working on this task, as necessary. Submit the issue to save. Any work, and subsequent commits on this topic should be tagged with the issue number. We now have a mechanism for linking your actual analysis and actions to the planned analysis in the preregistration document.

### Dealing with unexpected outcomes, and changing plans: Shifting back to planning from doing

Sometimes assumptions are violated, or there are other unexpected outcomes of preliminary analyses, and you will need to change plans. How do we deal with this?

- Add the URL to the relevant GitHub issue under the corresponding Preregistration Item.
- Add any the URL for any linked commits to the Preregistration Item. 
- revise the Preregistration, where necessary by editing the Preregistration document in GitHub, as per previous instructions.
- Commit the changes with the following commit message header:
- Describe the trigger for the Preregistration revisions in brief detail within the commit message body, adding the link to the relevant GitHub issue containing the tagged commits that triggered the revision.

## 'Living preregistration' - tracking changes to the Preregistration Draft with GitHub

We will leverage [GitHub's tagging and release feature](https://help.github.com/en/github/administering-a-repository/releasing-projects-on-github) in conjunction with [Semantic Versioning]((https://semver.org/)) to document and track changes to the preregistration document.

GitHub releases mark software iterations that you can release and deploy^[Instructions for creating releases are here: [https://help.github.com/en/github/administering-a-repository/managing-releases-in-a-repository](https://help.github.com/en/github/administering-a-repository/managing-releases-in-a-repository)]. Within the context of adaptive preregistration, we can use releases to mark interim versions of the preregistration, as well as to take a snapshot of the entire repository at that moment in time.

Note that anything up until the complete preregistration should be additionally tagged as a 'pre-release'

Preregistration Change or Action | GitHub Mechanism | Semantic Versioning|
------|-------|--------|
**Changes preceding initial completion of preregistration** | Tag and Mark as 'pre-release' | `Minor` increment. E.g. `0.2.0` The `Major` value should not exceed `0`. |
**Incremental Change**, e.g. respond to new question. | Tag and Release | `Minor` increment. E.g. `0.2.0` The `Major` value should not exceed `0`. |
**Initial Preregistration complete**. Note that 'complete' needs to be decided. Is complete after all preliminary and investigatory analyses have been done and the final Preregistration is complete, or is it the first iteration, where no more items can be filled out until some initial analyses are undertaken? | Tag and Release | `Major` increment to `1.0.0` |
**Major change to a complete Preregistration** -- no looking or analysing at data | Tag and Release | `Major` increment `>1.0.0` |
**Linking implemented work to the Preregistration Item** | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit. Tag and Release that preregistration commit. | `Minor` increment. |
**When plans change** Adapting the Preregistration on the basis of results from preliminary or investigatory analyses |  Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit. Respond to dependent incomplete Preregistration Items or Update existing / already completed Preregistration Items + Commit. Tag and Release | `Major` increment.
**Minor mistakes** | Changes including small typos, or minor elaborations on existing specified plans | `Patch` increment. E.g. `0.1.2`. |

### 2020-10-25 analysis/Preregistration_Template_Oct.Rmd
#46 update preregistration template after LR and HF feedback

---
title: "Preregistration Template"
output: github_document
author: "Elliot Gould, Libby Rumpff, Hannah Fraser, Megan Good, co-authors TBD."
---

# 1. Problem Formulation

Includes a clear specification of the problem context in which the model will be used or designed to inform. Defining the purpose of the model is critical because the model purpose influences choices at later stages of model development (Jakeman, Letcher, and Norton 2006). This includes engaging with the decision maker and/or key stakeholders (if relevant).

## 1.1 Model Context and Purpose

### 1.1.1. Key Stakeholders

First, identify relevant interest groups:

- Who is the model for?
-	Who is involved in formulating the model? 
-	Who needs buy in? 
-	How key stakeholders will/wish to be involved in model development and testing

### 1.1.2 The problem, and subsequent purpose of the model

Briefly outline:

- the problem (including the decision to be made, if relevant). 
- outline how your model is to address the problem, being clear about the scope i.e. is your model addressing the whole problem, or part of it? Are there any linked problems that your model needs to cognisant of?. 
- The scope of the model, including temporal and spatial resolutions (Mahmoud et al. 2009)
- Finish with a clear statement of the purpose of your model.
Common model purposes in ecology can include: gaining a better qualitative understanding of the target system, synthesising and reviewing knowledge, providing guidance for management and decision-making (Jakeman, Letcher, and Norton 2006), comparing the performance of alternative management actions under budget constraint (Fraser et al. 2017) to search for robust decisions under uncertainty (McDonald-Madden, Baxter, and Possingham 2008), to choose the conservation policy that minimises uncertainty (Tulloch et al. 2015).

### 1.1.3 Analytical purpose and scope, including scenario analysis

Provide more detail on the analytical purpose and scope of the model:

- How will the model be analysed?
- What analytical questions will the model be used to answer? 
- What is the intended domain in which the model is to be applied? Are there any reasonable extrapolation limits beyond which you expect the model should not be applied? (Grimm et al. 2014)
- Are there potentially unacceptable management or policy outcomes identified by stakeholders that should be captured in the model, i.e. as constraints?
- Are there scenarios that model outputs have to accommodate? Scenarios should be set a priori, i.e. before the model is built, (Moallemi, Elsawah, and Ryan 2019).
- Describe what processes you will use to elicit and identify relevant scenarios, e.g. literature review, structured workshops with stakeholders or decision-makers. 
- What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?
- Last, are there any limitations on model development analysis and flexibility, such as time or budget constraints, for example, does a model need to be deployed rapidly?

### 1.1.4 Output (or response) variable(s)

After defining the problem context, you will need to understand from key stakeholders what the outputs of the model are, and how they will be measured/modelled at the appropriate spatial and temporal scale (see 1.1.2, but to be addressed in 3.4.1).

In a decision context, the model outputs relate directly to the fundamental objectives (what is driving the decision, what is to be fundamentally achieved, or avoided? Gregory et al 2012).   These are differentiated from means objectives, which specify means to achieving the fundamental, or end objectives.  For example, improving the condition of an area of habitat, could be considered a means to improving the status of the fauna that utilise the habitat.   As such, the status of the given fauna is an output (or response) variable for the model. Questions you might ask of key stakeholders to elicit the output variable(s):

- What variables would support taking a particular course of action (i.e. a decision), or help define the problem at hand?
- What are the values to be addressed in this problem context? Are they all relevant to this model?
There may be multiple objectives, which may not all be relevant in your model.  It is important to capture this in the description of model scope (1.1.2). 
Any model variable may be qualitative or quantitative. It is the role of the modeller to ensure all variables, particularly the output variable(s) are measured at the appropriate scale. If the models are to be used in a decision context, this step is particularly crucial, as the choice of performance measure can result in substantial differences in model outputs (Giljohan et al, 2014).

### 1.1.5 Scenario analysis operationalization (delete as appropriate)

- How will you operationalize any scenarios identified in 1.1.3? For example, how will you operationalise any qualitative changes of interest, such as ‚ ‘deterioration’ or ‘improvement’?
- Describe how you will evaluate and distinguish the performance of alternative scenario outcomes. Examples include: descriptive statistic measures, such as a model in the top 10th percentile of worst conditions, satisficing measures such as a minimum performance threshold for achieving some management objective, or perhaps regret measures, such as choosing a decision with minimum regret (Moallemi, Elsawah, and Ryan 2019). 
- Justify or otherwise explain how you chose these measures and determined performance criteria in relation to the analytical objectives, model purpose and modelling context, such as the risk attitudes of decision-makers and stakeholders within this system.

# 2.0 Define Conceptual Model

Conceptual models underpin the formal or quantitative model (Cartwright et al. 2016). The conceptual model describes the biological mechanisms relevant to the ecological problem and should capture basic premises about how the target system works, including any prior knowledge and assumptions about system processes. Conceptual models may be represented in a variety of formats – such as influence diagrams, linguistic model block diagram or bond graphs – illustrating how model drivers are linked to both outputs or observed responses, and internal (state) variables (Jakeman, Letcher, and Norton, 2006).

## 2.1 Conceptual Model Representation and Elicitation Method

- Describe what method you will use to elicit or identify the conceptual model. Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks (see Moon et al. (2019) for guidance on selecting an appropriate elicitation method). 
- How will you represent the final conceptual model? This will likely depend on the method chosen to elicit the conceptual model.

### 2.2.1 	Explain Critical Conceptual Design Decisions

List and explain critical conceptual design decisions, including: 
- spatial and temporal scales,
-  selection of entities and processes,
-  representation of stochasticity and heterogeneity, 
- consideration of local versus global interactions, environmental drivers, etc. (Grimm et al. 2014).

Explain and justify the influence of particular theories, concepts, or earlier models against alternative conceptual design decisions that might lead to alternative model structures (Grimm et al. 2014).

### 2.2.2	Model assumptions and uncertainties

Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be will be represented in the model. Sources of uncertainty may include:

- exogenous uncertainties affecting the system,
- parametric uncertainty in input data and 
- structural / conceptual nonparametric uncertainty in the model (Moallemi, Elsawah, and Ryan 2019).

## 2.3 Predictor Variables

Identify and define system system variables and structures, referencing scenario definitions, and the scope of the model as described within problem formulation (1.1.3):

- What variables would support taking this action or making this decision?
- What additional variables may interact with this system (things we can’t control, but can hopefully measure)?
- What variables have not been measured, but may interact with the system (often occurs in field or observational studies)?
- What variables are indice or surrogate measures of variables that we cannot or have not measured?
- In what ways do we expect these variables to interact (model structures)?

## 2.4 Define prior knowledge, data specification and evaluation

This section should specify the plan for collecting, processing and preparing data available for parameterisation, determining model structure, and for scenario analysis.

### 2.4.1	Document data sources

Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model. Describe how the data is arranged, in terms of replicates and covariates.

**For pre-existing data (delete as appropriate):**

- For each dataset, is the data open or publicly available? (Y/N)
- How can the data be accessed? Provide a link or contact as appropriate, indicating any restrictions on the use of data.
- Date of download, access, or future access:
- Describe the source of the data - what entity originally collected this data? (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).
- Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).
- Prior work based on this dataset - Have you published / presented any previous work based on this dataset? Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont this dataset you have worked on. 
- Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.
- Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study. For example, have you read any reports or publications about this data?

**Sampling Plan (For data you will collect) – delete as appropriate:**

- Data collection procedures
  - Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged. 
  - Describe any inclusion or exclusion rules, and the study timeline.
- Sample Size: Describe the sample size of your study.
- Sample Size Rationale: Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel.
- If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?

### 2.4.2	Data Processing and Preparation

Describe any data preparation and processing steps, including manipulation of environmental layers (e.g. standardisation and geographic projection) or variable construction (e.g. Principal Component Analysis).

### 2.4.3 Data Exploration or preliminary data analyses

It is often necessary when modeling to perform preliminary analyses to check assumptions and ensure requirements of the chosen modelling procedure are met, for instance, exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers or the need for transformations (Yates et al. 2018). Because each preliminary analysis increases the risk of researcher degrees of freedom, flexible analysis strategies may be specified. For each separate preliminary or investigatory analysis:

- State what needs to be known to proceed with further decision-making about the modelling procedure, and why the analysis is necessary. 
- Explain how you will implement this analysis, as well as any techniques you will use to summarise and explore your data.
- What method will you use to represent this analysis (graphical, tabular, or otherwise)
- Specify exactly which parts of the data will be used.
- Describe how the results will be interpreted, listing each potential analytic decision, as well as the analysis finding that will trigger each decision, where possible.

### 2.4.4 Data evaluation, exclusion, and missing data

- Describe how you will determine how reliable the data is for the given model purpose. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand (Mahmoud et al. 2009).
- Document any issues with data reliability. This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems (Grimm et al. 2014).
- How will you determine what data, if any, will be excluded from your analyses? How will outliers be handled? Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.
- How will you identify and deal with incomplete or missing data?

## 2.5 Conceptual Model Evaluation

- Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.
- How will you critically assess any simplifying assumptions?
- Will the conceptual model evaluation process include consultation or feedback from a client, manager, or model user?

# 3.0 Formalise and Specify Model

## 3.1 Model Class, Framework and Approach

Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based (Yates et al. 2018); where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data (White and Marshall 2019). A model ‘class,’ ‘family’’ or ‘type’ is often used to describe a set of models each of which has a distinct but related sampling distribution (Liu & Aitken, 2008). The model family is driven by choices about the types of variables covered and the nature of their treatment, as well as structural features of the model, such as link functions, spatial and temporal scales of processes and their interactions (Jakeman et al. 2006).

- Describe what modelling framework, approach or class of model you will use to implement your model and relate your choice to the model purpose and analytical objectives described in 1.1.2 and 1.1.3.

### 3.1.1 Operationalising Model Variables

- Specify how you will operationalise the response variables in the model. This should relate directly to the analytical and or management objectives specified during the problem formulation phase. Operationalisations could include:
  - the extent of a response,
  - an extreme value,
  - a trend,
  - a long-term mean,
  - a probability distribution,
  - a spatial pattern,
  - a time-series,
  - qualitative change, such as a direction of change or
  - the frequency, location, or probability of some event occurring (Jakeman, 2006).
- Provide a rationale for your choices, including why plausible alternatives were not chosen.
- For the remaining model variables, specify each variable and their nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).

### 3.1.2	Model Structure

Specify model structural features, including:

- the functional form of interactions, 
- data structures,
- measures used to specify links,
- any bins or discretisation of continuous variables (Jakeman, 2006).

### 3.1.3 	Model Class / Family

- Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions. 
- Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.
- Include in your rational for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.

### 3.2	Model Structure and Parameter Estimation

This section refers to the process of determining the best/most efficient/parsimonious representation of the system at a scale relevant to the model purpose, context and scope appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives (as specified in 1.1.3). Approaches to finding model structure and parameters may be knowledge-supported, or data-driven (Boets et al. 2015). Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods (Barnard et al. 2019). Ensemble modelling procedures might aim to derive a single model, or a multi-model average (Yates et al. 2018). Refining actions to develop a model could include iteratively dropping parameters or adding them, or aggregating / disaggregating system descriptors, such as dimensionality and processes (Jakeman, 2006).
- Specify what approach and methods you will use to identify model structure and parameters.

### 3.2.1 	Parameter/Structure estimation technique

- Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006).
- Identify which variables were parameterised directly (if any), such as by expert elicitation or prior knowledge. 
- Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms (Liu et al. 2018).
- Repeat the above bullet points but for model structure estimation, if you are using quantitative methods for structure estimation rather than knowledge or expert-judgment driven methods.

### 3.2.2	Estimation performance criteria

- Specify which suite of performance criteria you will use to judge the performance of the model. Examples include correlation scores, coefficient of determination, specificity, sensitivity, AUC, etcetera (Yates et al. 2018). State any numeric thresholds and their qualitative interpretation for judging adequate estimation.
- Relate any underlying assumptions of each criterion to the desired properties of the model and justify the choice of performance metric in relation into whether it is sensitive to the problem at hand. 
- Explain how you will identify which model features or components are significant or meaningful.

## 3.3 Model assumptions and uncertainties

- Specify all assumptions and key uncertainties in the formal model. 
- Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpertation of the model outputs, and 
- what implications are there on evaluating model-output to inform decisions?

## 3.4	Specify formal model

- Once all decisions have been made about the approach and method of model specification, translate the conceptual model into the formal or quantitative model. 
- For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters.

# 4.0 Model Calibration, Fitting & Checking

## 4.1 Model calibration and validation scheme

- Describe how you will validate and check the calibration of the model. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) (Yates et al. 2018; Barnard et al. 2019).
- Describe any model tuning procedures for estimating hyper parameters (if any).
- Justify your choices

### 4.1.2 	Describe calibration/validation data
- If partitioning data for cross-validation: 
  - Describe the approach specifying the number of folds that will be created and the relative size of each. 
  - Describe how will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced.
- If using external / independent holdout data:
  - Describe any known differences between the training and validation datasets and the size of the validation dataset. 
  - Describe how will you document and/or share the external data such that any subsequent modelling can be computationally reproduced.
  - It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modellers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?

## 4.2 Implementation verification

- What Quality Assurance measures will you take to verify the model has been correctly implemented as specified? Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging and silent errors.
  - Tests could include syntax checking of code, and code reviews by peers. 
  - Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as dictated by the model description (Grimm et al. 2014). 
  - Technical measures include using unit tests, or in-built checks within functions to prevent potential errors.

## 4.3	Model Checking

"Model Checking" goes by many names ("calibration testing or validation", "conditional verification", "quantitative verification", "model output verification"), and refers to a series of analyses that assess a model’s performance in representing the system of interest (Conn et al. 2018). Model checking therefore helps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system (Conn et al. 2018). Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation (Jakeman, Letcher, and Norton 2006).

### 4.3.1 Quantitative model checking

During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences.
- Specify any diagnostics or tests you will use during model checking to assess a model’s performance in representing the system of interest. 
- For each test, specify the criteria that will you use to interpret the outcome of the test in assessing the model’s ability to sufficiently represent the gathered data used to develop and parameterise the model.

### 4.3.2 Qualitative model checking

This step is largely informal and case-specific but requires ‘face validation’ with model users / clients / managers who aren’t involved in the development of the model to assess whether the interactions and outcomes of the model are feasible and defensible (Grimm et al. 2014). Other terms for qualitative checking include the "laugh test" or the "pub test," and in addition to checking the model’s believability, they build the client or users’ confidence in the model (Jakeman, Letcher, and Norton 2006).

- Explain how you will qualitatively check the model, and whether and how you will include users and clients in the process. Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.

### 4.3.3 Assumption violation checks

- Explain how you will demonstrate robustness to model assumptions and check for violations of model assumptions.
- If you cannot perform quantitative assumption checks, describe what theoretical justifications would justify a lack of violation of or robustness to model assumptions. 
- If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.
- The consequences of assumption violations on the interpretation of results should be assessed (Araújo et al. 2019). If assumption violations cannot be avoided, explain describe how you will explore their consequences (To be completed in interim iterations of the preregistration, only if there are departures from assumptions as demonstrated in the planned tests above).

# 5.0 Model Validation and Evaluation

Model evaluation including model analysis and other testing complements model checking. By evaluating the model against independent test or holdout data the chance that the model fits the calibration or fitting data well, but has not captured the relevant ecological mechanisms of the system, can be ruled out (Grimm et al. 2014). Collectively, modelling activities in section 4 and 5 provide provide information about the structural realism and therefore credibility of the model (Grimm et al. 2014).

## 5.1 Model output corroboration

- State whether you will corroborate the model outputs on external test data, and ensure that any independent validation data is documented in 4.1.2. 
- Testing the model against data that are collected spatially and/or temporally independent from the training dataset minimizes the risk of artificially inflating model performance measures (Araújo et al. 2019). Although the corroboration of model outputs against an independent validation dataset is considered the ‘gold standard’ for showing that a model properly represents the internal organisation of the system), model validation is not always possible (Grimm, et al. 2014). If you are unable to evaluate the model outputs against independent data, explain why not, and describe what steps you will take to interrogate the model. For example, Independent predictions can instead be tested on submodels, alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model’s output is (Grimm et al.2014).

### 5.1.1 Model corroboration performance metrics and criteria

- Specify what performance measures you will use to evaluate the model and briefly explain how each test relates to different desired properties of a model’s performance. Examples include measures of agreement between predictions and independent observations, or estimates of accuracy, bias, calibration, discrimination refinement, resolution and skill (Araújo et al. 2019).
- Spatial, temporal and environmental pattern of errors and variance can change the interpretation of model predictions and conservation decisions (Araújo et al. 2019), where relevant and possible, describe how you will characterise and report the spatial, temporal and environmental pattern of errors and variance.
- If comparing alternative models, specify what measures of model comparison or out-of-sample performance metrics will you use to find support for alternative models or else to optimise predictive ability.
- State what numerical threshold or qualities you will use for each of these metrics.

## 5.2	Model Analysis

### 5.2.1 Uncertainty Analyses

Uncertainty can arise from different modeling techniques, response data and predictor variables (Araújo et al. 2019). Uncertainty analyses characterise the uncertainty in model outputs, and identify how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour (Grimm et al. 2014; Saltelli et al. 2019).

- Please describe how you will characterise model and data uncertainties. E.g. propagating known uncertainties through the model, investigating the effect of different model scenarios with different parameters and modelling technique combinations (Araújo et al. 2019), or empirically constructing model distributions from model output datapoints, and extracting summary statistics, including the mean, median, variance, and constructing confidence intervals (Saltelli et al. 2019).
- Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase. For instance, high model precision may be important for predictive modelling when cost, benefit and risk must be adequately gauged, but for decision-making contexts where only the predicted ranked order of management options is necessary, model accuracy and precision might be less important (Jakeman, Letcher, and Norton 2006). 
- Describe how you will summarise the results of these in silico experiments with graphical, tabular, or other devices, such as summary statistics.
- If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic parameter covariance, (Jakeman, Letcher, and Norton 2006)), specify which measure of uncertainty you will measure.

### 5.2.2 Sensitivity Analyses

Sensitivity analysis examines how uncertainty in model outputs can be apportioned to different sources of uncertainty in model input (Saltelli et al. 2019).

- Describe the sensitivity analysis approach you will take: deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).
- Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameters will be varied. 
- State the primary objective of each sensitivity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or alternatively to identify which variables or factors contribute little to overall model outputs, and so can be ‘dropped’ from future iterations of the model (Saltelli et al. 2019).

### 5.2.3	Model application / scenario analysis

- Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario specified in section 1. 
- Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections. 
- What raw and transformed model outputs will you extract from the model simulations or projections, and how will you map, plot, or otherwise display and synthesise the results of scenario and model analyses. 
- Explain how you will analyse the model analysis output to answer your analytical objectives. For instance, describe any trade-off or robustness analyses you will undertake to help evaluate and choose between different alternatives in consultation with experts or decision-makers.

### 5.2.4	Other in silico analyses

- Describe any other simulation experiments and or robustness analyses you will perform on the model, including any metrics and their criteria / thresholds for interpreting the results of the analysis.

# References

Araújo, MB, RP Anderson, A Márcia Barbosa, CM Beale, CF Dormann, R Early, RA Garcia, et al. 2019. "Standards for Distribution Models in Biodiversity Assessments." Sci Adv 5 (1): eaat4858.

Augusiak, Jacqueline, Paul J Van den Brink, and Volker Grimm. 2014. "Merging Validation and Evaluation of Ecological Models to ‘Evaludation’: A Review of Terminology and a Practical Approach." Ecological Modelling 280: 117–28.

Barnard, David M., Matthew J. Germino, David S. Pilliod, Robert S. Arkle, Cara Applestein, Bill E. Davidson, and Matthew R. Fisk. 2019. "Can’t See the Random Forest for the Decision Trees: Selecting Predictive Models for Restoration Ecology." Restoration Ecology.

Boets, Pieter, Dries Landuyt, Gert Everaert, Steven Broekx, and Peter L M Goethals. 2015. "Evaluation and Comparison of Data-Driven and Knowledge-Supported Bayesian Belief Networks to Assess the Habitat Suitability for Alien Macroinvertebrates" 74: 92–103.

Conn, Paul B, Devin S Johnson, Perry J Williams, Sharon R Melin, and Mevin B Hooten. 2018. "A Guide to Bayesian Model Checking for Ecologists." Ecological Monographs 9: 341–17.

Fraser, Hannah, Libby Rumpff, Jian D L Yen, Doug Robinson, and Brendan A Wintle. 2017. "Integrated Models to Support Multiobjective Ecological Restoration Decisions." Conservation Biology 31 (6): 1418–27.

Giljohann, K. M., McCarthy, M. A., McCarthy, M. A., Kelly, L. T., & Kelly, L. T. (2014). Choice of biodiversity index drives optimal fire management decisions. Ecological Modelling.

Gregory, R., Failing, L., Failing, L., Harstone, M., Harstone, M., Long, G., . . . Ohlson, D. (2012). Structured decision making: a practical guide to environmental management choices. John Wiley & Sons.

Grimm, Volker, Jacqueline Augusiak, Andreas Focks, Béatrice M Frank, Faten Gabsi, Alice S A Johnston, Chun Liu, et al. 2014. "Towards Better Modelling and Decision Support: Documenting Model Development, Testing, and Analysis Using Trace." Ecological Modelling 280: 129–39.

Jakeman, A J, R A Letcher, and J P Norton. 2006. "Ten Iterative Steps in Development and Evaluation of Environmental Models." Environmental Modelling & Software 21 (5): 602–14.

Liu, C. C., & Aitkin, M. (2008). Bayes factors: Prior sensitivity and model generalizability. Journal of Mathematical Psychology, 52(6), 362-375. doi:10.1016/j.jmp.2008.03.002

Liu, Zelin, Changhui Peng, Timothy Work, Jean-Noel Candau, Annie DesRochers, and Daniel Kneeshaw. 2018. "Application of Machine-Learning Methods in Forest Ecology: Recent Progress and Future Challenges." Environmental Reviews 26 (4): 339–50.

Mahmoud, Mohammed, Yuqiong Liu, Holly Hartmann, Steven Stewart, Thorsten Wagener, Darius Semmens, Robert Stewart, et al. 2009. "A Formal Framework for Scenario Development in Support of Environmental Decision-Making." Environmental Modelling & Software 24 (7): 798–808.

McDonald-Madden, Eve, Peter W. J. Baxter, and Hugh P. Possingham. 2008. "Making Robust Decisions for Conservation with Restricted Money and Knowledge." Journal of Applied Ecology 45 (6): 1630–8.

Moallemi, Enayat A., Sondoss Elsawah, and Michael J. Ryan. 2019. "Strengthening ‘Good’ Modelling Practices in Robust Decision Support: A Reporting Guideline for Combining Multiple Model-Based Methods." Mathematics and Computers in Simulation.

Saltelli, Andrea, Ksenia Aleksankina, William Becker, Pamela Fennell, Federico Ferretti, Niels Holst, Sushan Li, and Qiongli Wu. 2019. "Why so Many Published Sensitivity Analyses Are False: A Systematic Review of Sensitivity Analysis Practices." Environmental Modelling & Software 114: 29–39.

Tulloch, A. I. T., Maloney, R. F., Joseph, L. N., Bennett, J. R., Di Fonzo, M. M. I., Probert, W. J. M., . . . Possingham, H. P. (2015). Effect of risk aversion on prioritizing conservation projects. Conservation Biology.

White, Craig R, and Dustin J Marshall. 2019. "Should We Care If Models Are Phenomenological or Mechanistic." Trends in Ecology & Evolution 34 (4): 276–78.

Yates, KL, PJ Bouchet, MJ Caley, K Mengersen, CF Randin, S Parnell, AH Fielding, et al. 2018. "Outstanding Challenges in the Transferability of Ecological Models." Trends Ecol. Evol. (Amst.) 33 (10): 790–802.

### 2021-05-17 EcoConsPreReg-supervisor_draft-Sep-2020.Rmd
updates to ms and add supervisor draft from sep 2020

--- 2020-09-16 index.Rmd
@@
 author:
-  - name: Elliot Gould 
+  - name: Elise Gould 
     affiliation: School of BioSciences, University of Melbourne
@@
 date: "`r Sys.Date()`"
-output: 
-  distill::distill_article:
-    self_contained: false
+bibliography: ./data/references.bib
+output: word_document
+  #distill::distill_article:
+    #self_contained: false
 ---
@@
 
-**Abstract:** *Ecology is at risk of a 'reproducibility crisis', in particular, there are serious transparency problems in the reporting of ecological models, which is concerning because a lack of complete and transparent reporting practices provide the conditions in which questionable research practices are likely to arise, leading to irreproducible findings.
-Irreproducibility of model-based research threatens the progress of both scientific discovery in 'pure' research, and risks wrong or sub-optimal conservation and management decisions being implemented when models are used in decision-support.
-Preregistration is an open-science practice adopted in scientific disciplines that have confronted the reproducibility crisis, which aims to improve research transparency and mitigate questionable research practices.
-There has been little uptake of preregistration in ecology and conservation, and within disciplines wherein preregistration has been adopted and become common practice, many modellers and researchers engaged in non-hypothesis testing research have vehemently eschewed preregistration on the grounds that existing preregistration templates are irrelevant to their research because they are focussed on null-hypothesis significance testing.
-We took a user-centred design approach to the task of translating preregistration into model-based research contexts for ecology and conservation.
-Structured discussion and collaborative activities underpinned the development of a preregistration template that captures critical analysis decisions in the process of the model development cycle.
-The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the linear procedure of preregistration in which a researcher shifts from ideation to analysis planning, to analysis and then reporting.
-We propose a methodology for implementing 'adaptive preregistration' leveraging git and GitHub, which accommodates common practices in modelling where some analysis decisions are difficult or cannot be made in advance, where downstream decisions might be dependent on the outcome of earlier decisions, and where preliminary analyses might need to be conducted before proceeding with further modelling.
-We tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
-This research provides an extensible template and methodology for implementing adaptive preregistration of ecological models, to be used in either pure or applied settings.
-We hope that this research will facilitate wider uptake of preregistration in ecology and that our proposed methodology for implementing adaptive preregistration might be applied in model-based research within science more broadly.
+**Abstract:**
+
+**1.** *Preregistration is an open-science practice adopted in scientific disciplines that have begun to confront the 'reproducibility crisis', which aims to improve research transparency and mitigate questionable research practices within a study.
+There has been little uptake of preregistration in ecology and conservation, and many modellers and researchers engaged in non-hypothesis testing research both within and outside of ecology have eschewed preregistration on the grounds that existing templates are irrelevant because they are focussed on null-hypothesis significance testing.*
+
+*2.
+We advocate that preregistration can and should be used in model-based research in ecology and conservation but that preregistration should reflect the norms and practice of the research context in which it is applied in order to adequately restrict researcher degrees of freedom that are unique to a domain-specific methodology.
+We took a user-centred design approach to the task of translating preregistration into model-based research contexts for ecology and conservation.*
+
+*3.
+We identified a generalised 'modelling workflow' that captures critical analysis decisions in the process of the model development cycle.
+The workflow informed the content and structure of a preregistration template to be used in model-based research.
+The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the current format and linear process of preregistration.
+To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose a methodology for implementing 'adaptive preregistration' leveraging git and GitHub.
+We tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.*
+
+*4.
+This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings, and will facilitate the wider uptake of preregistration in ecology.
+Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model-based research within science more broadly.
 Modellers need no longer cry "but, I can't preregister my research."*
 
-**Keywords** *(no more than 8 for MEE).***:** *preregistration, transparency, reproducibility, modelling, ecology, conservation, open-science, research software engineering.*
+**Keywords:** *preregistration, transparency, reproducibility, modelling, ecology, conservation, open-science, research software engineering.*
 
@@
 
-We should be concerned about research transparency because, a lack of complete and transparent reporting practices provide the conditions in which QRPs are likely to arise [@Fidler:2017he].
+A lack of complete and transparent reporting practices is concerning because it  provides the conditions in which QRPs are likely to arise [@Fidler:2017he].
 Moreover, as Morrison [-@Morrison:2016cd] demonstrated, when models are not reported transparently with sufficient detail, they cannot be reproduced.
@@
 More importantly, they do not reflect the iterative nature of model development [@MacEachern2019] and fail to capture critical analytic decisions in the modelling process.
-Presently, we are aware of only a single preregistration template for modelling [@Cruwell2019b], however it is steeped heavily in an an academic psychology context, is focussed on the use of models for hypothesis-testing, neglects the entire model development process, and approaches modelling from a very narrow epistemic function, in a way that is not commonly applied in ecology.
+Presently, we are aware of only a single preregistration template for modelling [@Cruwell2019b], however it is steeped heavily in an an academic psychology context, is focussed on the use of models for hypothesis-testing, neglects the8 model development process, and approaches modelling from a very narrow epistemic function, in a way that is not commonly applied in ecology.
 
@@
 If we wish to adopt preregistration in ecology, specifically within model-based research, both the content required to be preregistered, as well as the procedure for creating a preregistration must be adapted to suit model-based research as it is undertaken in ecology.
-Different research methodologies, and their specific and nuanced application across different domains and fields of study follow different analytical choices and are therefore susceptible to different QRPs.
+Different research methodologies, and their domain-specific application across different domains and fields of study follow different analytical choices and are therefore susceptible to different QRPs.
 Consequently, domain-specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom as they uniquely emerge within a domain or field-specific application of a particular research methodology.
@@
 
+@TODO: add argument from Berna's formal theory paper about the need for testing and evaluating new reforms. While only a qualitative assessment... still useful information, such as XYZ.
+
+Mention testing / eval here? Yes need to mention a little bit, BUT perhaps could save Berna's quote for discussion.
+
 We borrow from the UCD methodology proposed by Pavelin et al. [-@Pavelin2012], and follow the core steps ![](./figures/user-centered-design-preregistration.png):
 
-1\. Define the goals of the resource under development and identify the audience,
+1. Define the goals of the resource under development and identify the audience,
 
-2\. Characterise users and their needs,
+2. Characterise users and their needs,
 
-3\. Design mock-ups of your product,
+3. Design mock-ups of your product,
 
-4\. Build prototypes of selected mock-ups and test with users,
+4. Build prototypes of selected mock-ups and test with users,
 
-5\. Implement the product, write technical implementations and build the product.
+5. Implement the product, write technical implementations and build the product.
 
@@
 We define the primary purpose of the template as "to delimit researcher degrees of freedom" when conducting model-based research in applied ecology and conservation.
-Although others have defined the purpose of preregistration to include 'delineating between *exploratory* from *confirmatory* research' \<refs\>, this is not the intended purpose of the template here.
-Beyond hypothesis-testing contexts, this is an idealised and simplifying dichotomy that, in practice, is more complicated and nuanced.
+Although the purpose of preregistration is predominantly conceptualised in terms of distinguishing between exploratory from confirmatory research [ e.g.
+@Nosek:2018eg], we wish to emphasise that this is not the intended purpose of this template.
+We adopt a more general stance towards the function of preregistration, i.e.
+that preregistration should mitigate against questionable research practices by reducing the potential for cognitive biases to cause opportunistic misuse of researcher degrees of freedom, such as cherry-picking and HARKing.
+
+Explanation of why we adopt this stance (problem with exploratory vs. confirmatory distinction):
+
+*Beyond hypothesis-testing contexts, this is an idealised and simplifying dichotomy that, in practice, is more complicated and nuanced.
 For instance researchers might move back and forth between different reasoning-modes throughout the modelling cycle.
-Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, and consideration of perspectives from philosophy of science.
-Consequently, we wish to deemphasise the purpose of distinguishing between exploratory and confirmatory research, and focus the design of the templates in being able to delimit researcher degrees of freedom, so as to reduce the potential for cognitive biases, such as confirmation bias, and other questionable research practices, like cherry-picking, HARKing etc.
-\<insert ref to my QRP manuscript\>.
+Thinking through how modes of inference might hold within applied ecological modelling contexts requires further, more careful thought, and consideration of perspectives from philosophy of science.*
 
@@
 
-**1\. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
+**1. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
 
@@
 
-**2\. Translating Workflows into Preregistration Templates**
+**2. Translating Workflows into Preregistration Templates**
 
@@
 
-1\. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. \<Note, that these are being scrapped from the template and are turned into tool-tips\>
+1. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. \<Note, that these are being scrapped from the template and are turned into tool-tips\>
 
-2\.
+2.
 Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow.
@@
 
-3\.
+3.
 Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
@@
 
-**3\. Designing a new methodology for implementing preregistration in model-based research**
+**3. Designing a new methodology for implementing preregistration in model-based research**
 
@@
 
-Modellers \<here, here, and here\> have expressed concern about preregistration making modelling a more difficult task than it already is, and hampering the ability to meet project deadlines, especially in decision-problems.
+Modellers [@MacEachern2019] \<here, here, and here\> have expressed concern about preregistration making modelling a more difficult task than it already is, and hampering the ability to meet project deadlines, especially in decision-problems.
 Many modellers already struggle with sufficiently documenting models during publication and reporting \<ref, model reporting and transparency review\>, and modellers working in government or other agencies have expressed concern that preregistration will simply add to the list of  'red-tape' expected of them.
@@
 *Evaluation Method*
+
+Berna - should evaluate open science reforms.
 
@@
 
-1\. Opinionated Design - norms around modelling practice.  'Good Modelling Practice' Literature. This also has the benefit of aiding new modellers.
+1. Opinionated Design - norms around modelling practice.  'Good Modelling Practice' Literature. This also has the benefit of aiding new modellers.
 
-2\.
+2.
 A related point: linguistic ambiguity and other problems with modelling concepts and terms.
@@
 
-3\.
+3.
 *research transparency*.
@@
 
+## Case Study ##
+
+Should speak to the following, which could be included in the supporting materials.
+
+\- Completed preregistration
+
+\- Interim Preregistration Documents
+
+\- link to repository
+
+\- screenshots illustrating  how the GitHub process is used.
+
+----
+
 ## Discussion ##
@@
 
------- *We did this by:*
+*------ We did this by (overview of what we did):*
 
-*1\.
+*1.
 Identifying norms and practices to develop 'idealised' / 'generalised' research workflows that describe sets of commonly undertaken tasks (and some sort of order to reflect the workflow), within an ecological modelling research analysis (within applied settings??).
@@
 
-*2\.
+*2.
 We then transformed these workflows into a 'preregistration template', by distilling a set of decision-items that, we think, should a) adequately constrain researcher degrees of freedom to mitigate the risk of Questionable Research Practices in an ecological modelling exercise and b) adequately facilitates* research/analytic transparency.
@@
 
-3\.
+3.
 Our UX approach revealed critical features of the model development process, namely the inherent iterative nature of model development, and requirements for model checking that forced us to reconsider the procedure for completing a preregistration suitable to ecological modelling.
@@
 
-4\.
+4.
 
@@
 
-5\.
+5.
 Furthermore, we created an interface for completing and amending interim preregistrations using shinyApps to facilitate uptake of preregistration in both pure and applied ecological modelling research.
 
-*consisting of a series of*
+***Evaluation and testing (I think this provides a nice lead on to the guts of the discussion, so it should be at the end of our summary of what we did).***
+
+Berna Devezer et al. argue that most open science reforms propose them and are never formally tested. In this research, we provide a qualitative assessment of our new method only.
+
+Thus preregistration is based on the working hypothesis that it is able to mitigate against QRPs.
+
+We need empirical testing of this hypothesis, also so we can empirically evaluate whether the new method meets its intended purpose (link back to intended purpose of the preregistration specified in the methods section).
+
+Then, following that, we need metaresearch, so empirical research at the level of the body of literature, perhaps examining transparency is the easiest starting point..
+but how would you empirically evaluate a series of models??)
+
+Either way alternative benefits of preregistration (pu et al. talks about this.. what about others).
 
@@
 The repository and issues are publicly available, so any discussion will be transparent and open to the ecological modelling community.
+
+**Putting a positive spin on the time-costs of recording all decisions. The result of completing an adaptive preregistration using our methodology is that you get a 'Lab Notebook' at the end of it.**
+
+People have described modelling as being a digital or computational computer lab.
+I think the term '*in silico* experiments' has been used to describe some computational modelling, especially for simulation modelling.
+I'm pretty sure that some authors have made arguments that we would expect rigorous note-taking and recording in the lab environment, so why should we expect any less in a computational modelling setting?
+
+**Preregistration will NOT lead to punch card style analysis (ala MacEachern).**
+
+In our preregistration testing with the case study, we found the opposite, that it prompted critical thinking about analytical choices during modelling.
+
+Moreover, our design template is *opinionated*. (See Elith's latest SDM reporting guideline) -\> it provides a *guide*. For those whom are at risk of 'punch card style analysis', ala MacEachern.  
 --------------------------

### 2023-05-15 EcoConsPreReg-supervisor_draft-Sep-2020.Rmd
update manuscript introduction

@@
     affiliation: School of BioSciences, University of Melbourne
+abstract: "Blah blah"
 date: "`r Sys.Date()`"
 bibliography: ./data/references.bib
-output: word_document
-  #distill::distill_article:
-    #self_contained: false
+output: #
+  distill::distill_article:
+  self_contained: false
 ---
+
 **Intended journal:** Methods in Ecology and Evolution.
@@
 
-----
+------------------------------------------------------------------------
 
-## Introduction ##
+## Introduction
 
-### Problem ###
-
-Ecology as a discipline is at risk of a "reproducibility crisis" (Fidler et al.
-2017), with rates of Questionable Research Practices (QRPs) in ecology similar to those in other fields (Fraser et al.
-2018), such as psychology and medicine, where the crisis has received considerable attention, and provoked heated discussion among researchers within the broader scientific community.
-In particular, there are serious problems with the way ecological models and the modelling process are reported in the published literature.
-However, concerns about poor model reporting practices are not new, and have been raised repeatedly --- nearly a decade ago researchers concluded that most model reporting was neither transparent or comprehensive [@Schmolke:2010fd; @deVos:2011], and that a culture of model reporting was yet to be established [@Grimm:2014es].
-Moreover, most models are assessed predominantly on the basis of scientific originality during peer-review  [@Schmolke:2010fd] to the effect that the burden of evaluating models for credibility is placed onto the reader or model-user, rather than the peer-review process, especially in the case of extremely complex models [@Alexandrov:2011iv].
-
-Despite attempts to promote thorough and transparent model documentation using 'Good Modelling Practice' (e.g.
-[@Jakeman:2006ii;@Alexandrov:2011iv]), more recent quantitative assessments of transparency and reproducibility demonstrate that transparent model reporting is an ongoing problem.
-Just as Jakeman argued that many models in the published literature lack a clear statement of model objectives and purpose, lack descriptions of model assumptions and implications, and also fail to report model results, including validation and evaluation in 2006, [-@Jakeman:2006ii], [@Feng2019]'s quantitative assessment of Ecological Niche Models found that reporting of model evaluation and model results was less than adequate.
-
-A lack of complete and transparent reporting practices is concerning because it  provides the conditions in which QRPs are likely to arise [@Fidler:2017he].
-Moreover, as Morrison [-@Morrison:2016cd] demonstrated, when models are not reported transparently with sufficient detail, they cannot be reproduced.
-This is particularly problematic because in ecology, models are typically used in decision-support, often directly underpinning policy and management decisions.
-Incomplete and non-transparent reporting of the model and the modelling process risk the model being used in inappropriate applications or decision contexts [@Grimm:2014es] because the underlying modelling choices and assumptions are not communicated [@deVos:2011].
-Alternatively, a lack of transparency may mask poor model design or serious flaws in the model, with the risk that they be used in decision support resulting in wrong, or irreversible decisions [@Grimm:2014es;@Schmolke:2010fd].
-A lack of transparency in reporting also risks undermining the credibility of the model, such that the decision-maker rejects the use of the model and instead turns to unstructured and subjective judgments [@Grimm:2014es;@Addison:2013wa], thus contributing further to "decision complacency", or the "knowing-doing" gap [@Gardner:2018dm].
-
-Preregistration is a practice that has been adopted in disciplines that have confronted the 'reproducibility crisis' head-on, and can improve the transparency and reproducibility of ecological models.
+Preregistration is a practice that has been adopted in disciplines that have confronted the 'reproducibility crisis' head-on, and aims to improve the transparency and reproducibility of ecological models.
 Preregistration requires the methods and analysis plan of a future study are registered in a secure and accessible platform prior to any data collection, and after submission it can no longer be altered.
@@
 Examples of these analytic choices include whether to exclude observations, whether more data should be collected, which set of conditions should be combined and compared, which model evaluation metric should be used, etc.
-Coupled with a lack of transparency -- or "undisclosed flexibility" -- researcher degrees of freedom are vulnerable to being exploited by researchers because of cognitive biases and incentive structures within science.
-\<lead in sentence to next para??\>
+Coupled with a lack of transparency -- or "undisclosed flexibility" (cite)-- researcher degrees of freedom are vulnerable to being exploited by researchers because of researchers' cognitive biases and incentive structures within science.
 
-Ecology has seen recent improvements in journal policy concerning reporting guidelines and data/code archiving, but there has been slow uptake of other important tools, like preregistration and registered reports (Parker, Fraser, and Nakagawa 2019).
+Ecology has seen recent improvements in journal policy encouraging or mandating open-scientific reforms that aim to increase transparency and reproducibility of published research, for instance, reporting guidelines and data/code archiving.
+However, there has been slow uptake of other important open-science tools, such as preregistration and registered reports (Parker, Fraser, and Nakagawa 2019).
 Pu et al.
-2019 found that conceptual barriers may prevent users from different disciplines to think that preregistration is "not for them", this was a particularly salient sentiment for researchers engaged in non-experimental studies.
-We have experienced this sentiment ourselves when looking at preregistration templates to use for model-based research --- nearly all preregistration templates are geared towards experimental hypothesis-testing research.
-Such templates contain items that are simply irrelevant to model-based and/or applied ecological research.
-More importantly, they do not reflect the iterative nature of model development [@MacEachern2019] and fail to capture critical analytic decisions in the modelling process.
-Presently, we are aware of only a single preregistration template for modelling [@Cruwell2019b], however it is steeped heavily in an an academic psychology context, is focussed on the use of models for hypothesis-testing, neglects the8 model development process, and approaches modelling from a very narrow epistemic function, in a way that is not commonly applied in ecology.
+2019 found that conceptual barriers may prevent users from different disciplines adopting preregistration in their scientific practice, because they consider preregistration is "not for them".
+We have experienced this sentiment ourselves when considering preregistering model-based research because nearly all preregistration templates are focussed on experimental hypothesis-testing research, and as such do not reflect the research life-cycle of model-based research; they fail to capture critical decisions commonly implemented in the model-development process, or include fields that relate to analytic decisions that are simply irrelevant to model-based research.
+Moreover,
 
+Presently, we are aware of only a single preregistration template for modelling [@Cruwell2019b], however the template explicitly excludes the construction and development phases of modelling, because the authors adopt a narrow but dominant view pervading the published literature on preregistration, 1.
+that scientific inquiry can be divided wholly into confirmatory and exploratory research, and that 2.
+that preregistration is intended to delimit researcher degrees of freedom within hypothesis-testing research only.
+The exclusion of preregistration from model-based research is premised on the idea, summarised by MacEachern and Van Zandt [-@MacEachern2019], that the modelling process can be dichotomised into two distinct phases that map onto the exploratory/confirmatory research divide, wherein model-development is inherently exploratory, 'creative', and 'confirmatory modelling' occurs within the context of experimental studies and the model is treated as a hypothesis to be tested when confronted with new data.
+
+In this paper, we take a radical departure from this perspective of preregistration, arguing that preregistration can and should be used for model-based research, including model development and construction, however, preregistration should reflect the norms and practices of the research context in which it is applied, in order for it to be broadly accepted and adopted by any given community of practice.
+Dichotomising model-based research into either exploratory or confirmatory research is reductive, and does not reflect the domain-specific uses of modelling in ecology and conservation in particular (For example, see [@Brudvig2017] for one epistimology describing the uses of modelling in applied and basic restoration) (**TODO elaborate further??**).
+
+Crucially, we believe it's problematic to preclude the model-development and construction phases of modelling from preregistration, because questionable research practices (QRPs) are possible (and likely just as prevalent for other types of ecological and scientific research), and preregistration is a valuable tool in mitigating the risk of QRPs in model-based research.
+Thirdly, preregistration provides additional benefits, and can improve the transparency of decision-making during modelling, which is notoriously opaque and often under-described (**TODO insert citation / refs on the lack of transparency in model based research).**
+
+The goal of this study is to rethink preregistration to better suit the domain-specific challenges and methodological particularities of model-based research in ecology and conservation.
+This research builds on recently developed "QRP roadmaps" (Gould et al.
+2019) and research by Cruwell et al.
+(2019) showing that existing templates are overly-focussed on hypothesis-testing research, and do not account for domain-specific and research category-specific nuances and questionable research practices.
 Preregistration cannot simply be ported over to other disciplines and adopted as-is.
 If we wish to adopt preregistration in ecology, specifically within model-based research, both the content required to be preregistered, as well as the procedure for creating a preregistration must be adapted to suit model-based research as it is undertaken in ecology.
+
+We first aim to devise a methodology for implementing an expanding view of the preregistration process, 'adaptive preregistration' [sensu @Srivastava2018], that better accounts for the inherently iterative nature of model development than the prevailing immutable, unalterable and deterministic view of preregistration.
+Adaptive preregistration can be applied to model-based research because it allows for the iterative cycle of model development to be captured in the preregistration.
+
 Different research methodologies, and their domain-specific application across different domains and fields of study follow different analytical choices and are therefore susceptible to different QRPs.
 Consequently, domain-specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom as they uniquely emerge within a domain or field-specific application of a particular research methodology.
+Secondly, we aim to develop a preregistration template for ecological modelling.
+Finally, we evaluate both the template and proposed procedure for adaptive preregistration using a case-study in which we preregister and undertake an environmental flows modelling study from Victoria, Australia.
 
-### Research Aims, and Template Objectives ###
+## Methods
 
-The goal of this study is to \<rethink\> preregistration to better suit the domain-specific challenges and methodological particularities of model-based research in ecology.
-This research builds on recently developed "QRP roadmaps" (Gould et al.
-2019) and research by Cruwell et al.
-(2019) showing that existing templates are too general, and do not account for domain-specific and research category-specific nuances and questionable research practices.
-We aim to develop a template for preregistering ecological models, and devise a methodology for implementing an expanded view of the preregistration process, 'adaptive preregistration' (*sensu* Srivastava), that better accounts for the inherently iterative nature of model development than static, deterministic document-based preregistration.
-
-We aim to consider questions such as... (open questions we had about preregistration)...
-
-Finally, we develop and evaluate both the template and proposed adaptive preregistration process with a case-study of environmental-flow modelling in Victoria, Australia.
-
-----
-
-## Methods ##
-
-### Designing Preregistration Templates with a User-Centred Design Approach ###
+### Designing Preregistration Templates with a User-Centred Design Approach
 
@@
 We define the primary purpose of the template as "to delimit researcher degrees of freedom" when conducting model-based research in applied ecology and conservation.
-Although the purpose of preregistration is predominantly conceptualised in terms of distinguishing between exploratory from confirmatory research [ e.g.
+Although the purpose of preregistration is predominantly conceptualised in terms of distinguishing between exploratory from confirmatory research [e.g.
 @Nosek:2018eg], we wish to emphasise that this is not the intended purpose of this template.
@@
 
-c) Use this knowledge to inform the design of the templates so they meet their intended purpose and are taken up by researchers.
+d)  Use this knowledge to inform the design of the templates so they meet their intended purpose and are taken up by researchers.
 
@@
 
-#### Analysis and Design ####
+#### Analysis and Design
 
@@
 
-----
+------------------------------------------------------------------------
 
-## Results: A methodology for Preregistration in Ecology and Conservation ##
+## Results: A methodology for Preregistration in Ecology and Conservation
 
-----
+------------------------------------------------------------------------
 
-### Internal Model of Preregistration ###
+### Internal Model of Preregistration
 
@@
 Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed \<Srivastava\>.
-Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, <x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?id=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0>\>.
+Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, \<x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?id=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0\>\>.
 
@@
 [-@Pu2019]* who recommended two design elements that may facilitate checking.
-Firstly, that the design should "make relevant preregistration easier to *query"* and secondly it should "encourage wider *coverage* of preregistration content during review." The process of *querying* a preregistration consists of "paper-to-preregistration" and "preregistration-to-paper" lookup whereby the reader searches for relevant information from the paper in the preregistration and *vice-versa*.
+Firstly, that the design should "make relevant preregistration easier to *query"* and secondly it should"encourage wider *coverage* of preregistration content during review." The process of *querying* a preregistration consists of "paper-to-preregistration" and "preregistration-to-paper" lookup whereby the reader searches for relevant information from the paper in the preregistration and *vice-versa*.
 This requires that the report or paper must be specific in referring to parts of the preregistration, *and*, that the author has tagged the relevant part of the preregistration in their report or manuscript.
@@
 
-----
+------------------------------------------------------------------------
 
-### Adaptive Preregistration for model-based research using GitHub ###
+### Adaptive Preregistration for model-based research using GitHub
 
@@
 *Semantic versioning* is used to document each version of the preregistration.
-*Semantic versioning* is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it [@Kitzes2018vn;@Raemaekers2017].
+*Semantic versioning* is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it [@Kitzes2018vn; @Raemaekers2017].
 We have created our own *semantic versioning* scheme to be used to specifically document the incremental version of the preregistration, in tandem with *GitHub's* tag and release feature, each interim preregistration is tagged and released with its semantic version number.
@@
 
-----
+------------------------------------------------------------------------
 
-### A Shiny Preregistration Interface ###
+### A Shiny Preregistration Interface
 
@@
 
-----
+------------------------------------------------------------------------
 
-## Case Study ##
+## Case Study
 
@@
 
-----
+------------------------------------------------------------------------
 
-## Discussion ##
+## Discussion
 
-### talking points ###
+### talking points
 
@@
 
-**[@MacEachern2019]:**
+[@MacEachern2019]**:**
 
@@
 Comput Brain Behav, 2(3-4), 179-182.
-doi:10.1007/s42113-019-00038-x
+<doi:10.1007/s42113-019-00038-x>
 
@@
 
-Moreover, our design template is *opinionated*. (See Elith's latest SDM reporting guideline) -\> it provides a *guide*. For those whom are at risk of 'punch card style analysis', ala MacEachern.  
---------------------------
+## Moreover, our design template is *opinionated*. (See Elith's latest SDM reporting guideline) -\> it provides a *guide*. For those whom are at risk of 'punch card style analysis', ala MacEachern.
+
 # References

### 2023-09-27 _analysis/EcoConsPreReg-supervisor_draft-Sep-2020.Rmd
rename _analysis

--- 2023-05-15 EcoConsPreReg-supervisor_draft-Sep-2020.Rmd
@@
 
-@TODO: add argument from Berna's formal theory paper about the need for testing and evaluating new reforms. While only a qualitative assessment... still useful information, such as XYZ.
+@TODO: add argument from Berna's formal theory paper about the need for testing and evaluating new reforms.
+While only a qualitative assessment... still useful information, such as XYZ.
 
-Mention testing / eval here? Yes need to mention a little bit, BUT perhaps could save Berna's quote for discussion.
+Mention testing / eval here?
+Yes need to mention a little bit, BUT perhaps could save Berna's quote for discussion.
 
@@
 This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase.
-The development of the draft modelling workflow was derived from previous work by Gould et al.
-*in prep.,* whilst the field sampling workflow was created specifically for this exercise.
+The development of the draft modelling workflow was derived from previous work by Gould et al. *in prep.,* whilst the field sampling workflow was created specifically for this exercise.
 We sorted and categorised each decision-step as belonging to a particular workflow phase.
@@
 We conducted two lots of structured discussions, one for each of the field and modelling template groups.
-Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature \<cite / link to OSF, or even discuss??\>.
+Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature \<cite / link to OSF, or even discuss??\>
+.
 These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in activity 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.
@@
 A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management.
-Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], \<insert link to appendices or just present the final?\>.
+Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], \<insert link to appendices or just present the final?\>
+.
 Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps.
@@
 
-1.  Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. \<Note, that these are being scrapped from the template and are turned into tool-tips\>
+1.  Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases.
+    \<Note, that these are being scrapped from the template and are turned into tool-tips\>
 
-2.
-Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow.
+2.  Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow.
     \<again, these are being scrapped and turned into tool-tips.\>
 
-3.
-Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
+3.  Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
     The style should be in the form of a directive or a question.
-E.g.
-"explain how you will do xyzzy".
+    E.g. "explain how you will do xyzzy".
     Or "What performance measure will you use for assessing 'goodness of fit'?"
@@
 
-@TODO - anything about *applied contexts??* Aligning with management? (In terms of the structure and steps above).
+@TODO - anything about *applied contexts??* Aligning with management?
+(In terms of the structure and steps above).
 
@@
 Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed \<Srivastava\>.
-Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, \<x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?id=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0\>\>.
+Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, \<x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?i
+d=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0\>\>.
 
@@
 Each issue is assigned a unique number.
-Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the *issue's* thread (Fig X, step 4, Do) \<screen-cap?\>.
+Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the *issue's* thread (Fig X, step 4, Do) \<screen-cap?\>
+.
 The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new *release*.
@@
 We considered removing in-template text describing and defining key concepts, phases and steps, as well as reducing the number of decision-points in a model-based research workflow captured by the template by removing items within the template.
-\<@TODO - did we end up removing any items by the end of the preregistration and evaluation with Chris' study?\> While we 'tweaked' the template items and descriptive text during the in situ evaluation of the templates for the environmental flows preregistration, we decided not to alter the template structure and content significantly, but to instead implement a new format and interface for working with the template.
+\<@TODO - did we end up removing any items by the end of the preregistration and evaluation with Chris' study?\>
+While we 'tweaked' the template items and descriptive text during the in situ evaluation of the templates for the environmental flows preregistration, we decided not to alter the template structure and content significantly, but to instead implement a new format and interface for working with the template.
 We outline these reasons below.
 
-1.  Opinionated Design - norms around modelling practice. 'Good Modelling Practice' Literature. This also has the benefit of aiding new modellers.
+1.  Opinionated Design - norms around modelling practice.
+    'Good Modelling Practice' Literature.
+    This also has the benefit of aiding new modellers.
 
-2.
-A related point: linguistic ambiguity and other problems with modelling concepts and terms.
+2.  A related point: linguistic ambiguity and other problems with modelling concepts and terms.
     Contested terms in the ecological modelling literature about what each different step is, their activities, etc.
 
-3.
-*research transparency*.
+3.  *research transparency*.
     We felt that the decision-points captured in the template are critical to increasing reporting transparency sufficiently for others to properly evaluate the modelling and analysis.
@@
 
-Berna Devezer et al. argue that most open science reforms propose them and are never formally tested. In this research, we provide a qualitative assessment of our new method only.
+Berna Devezer et al. argue that most open science reforms propose them and are never formally tested.
+In this research, we provide a qualitative assessment of our new method only.
 
@@
 
-All questions are on the table. We approached this goal with a 'proof is in the pudding' mentality - what are the limits of preregistration? Does it slow down the research process?
+All questions are on the table.
+We approached this goal with a 'proof is in the pudding' mentality - what are the limits of preregistration?
+Does it slow down the research process?
 
@@
 
-**"**We fear that preregistration will lessen the engagement of analysts with data and that it represents a step back toward the punch card style of analysis." MacEachern, S.
-N., & Van Zandt, T.
+**"**We fear that preregistration will lessen the engagement of analysts with data and that it represents a step back toward the punch card style of analysis." MacEachern, S. N., & Van Zandt, T.
 (2019).
@@
 
-It's, such as t proposed by Feng et al. The decisions and details are recorded within the preregistration. Along with any deviation from the initial plan.
+It's, such as t proposed by Feng et al.
+The decisions and details are recorded within the preregistration.
+Along with any deviation from the initial plan.
 

### 2023-09-27 _analysis/PRT-Appendix-1.Rmd
rename _analysis

--- 2020-05-07 analysis/PRT-Appendix-1.Rmd
@@
 ---
-title: "Untitled"
-description: |
-  A new article created using the Distill format.
-author:
-  - name: ELise Gould 
+title: "PRT Appendix - 1"
 date: "`r Sys.Date()`"
-output: distill::distill_article
+output: github_document
 ---
+
+Table specifying phases, steps, and item inclusion with justifications from workshop and literature in support.
 
@@
 ```
+
+```{r session-info, echo = TRUE, eval = TRUE}
+sessionInfo()
+```

### 2023-09-27 _analysis/PreReg_May_2021.Rmd
rename _analysis

---
title: "Ecology and Conservation Modelling Preregistration Template -- DRAFT"
author: "Chris Jones, Elliot Gould"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document: default
---

# 1. Problem Formulation

## 1.1 Model Context and Purpose

### 1.1.1. Key Stakeholders

*First, identify relevant interest groups:*
-   *Who is the model for?*
-   *Who is involved in formulating the model?*
-   *Who needs buy in?*
-   *How key stakeholders will/wish to be involved in model development and testing*

### 1.1.2 The problem, and subsequent purpose of the model
*Briefly outline:*

-   *The problem (including the decision to be made, if relevant).*
-   *How your model will address the problem, being clear about the scope i.e. is your model addressing the whole problem, or part of it? Are there any linked problems that your model needs to cognisant of?*
-   *The scope of the model, including temporal and spatial resolutions*
-   *Finish with a clear statement of the purpose of your model.*

### 1.1.3 Analytical purpose and scope, including scenario analysis

*Provide more detail on the analytical purpose and scope of the model:*
-   *How will the model be analysed?*
-   *What analytical questions will the model be used to answer?*
-   *What is the intended domain in which the model is to be applied? Are there any reasonable extrapolation limits beyond which you expect the model should not be applied? *
-   *Are there potentially unacceptable management or policy outcomes identified by stakeholders that should be captured in the model, i.e. as constraints?*
-   *Are there scenarios that model inputs or outputs must accommodate? Scenarios should be set a priori, i.e. ideally before the model is built..*
-   *If relevant, describe what processes you will use to elicit and identify relevant scenarios, e.g. literature review, structured workshops with stakeholders or decision-makers.*
-   *What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?*
-   *Last, are there any limitations on model development analysis and flexibility, such as time or budget constraints, for example, does a model need to be deployed rapidly?*

### 1.1.4 Output (or response) variable(s)

-   *What variables would support taking a particular course of action (i.e. a decision), or help define the problem at hand?*
-   *What are the values to be addressed in this problem context? Are they all relevant to this model?*
- *After answering these questions with relevant stakeholders, please describe and define your response variable(s).*

### 1.1.5 Scenario analysis operationalization (delete as appropriate)

-   *How will you operationalize any scenarios identified in 1.1.3? For example, how will you operationalise any qualitative changes of interest, such as ‚ 'deterioration' or 'improvement'?*
-   *Describe how you will evaluate and distinguish the performance of alternative scenario outcomes. *
-   *Justify or otherwise explain how you chose these measures and determined performance criteria in relation to the analytical objectives, model purpose and modelling context, such as the risk attitudes of decision-makers and stakeholders within this system.*

# 2.0 Conceptual Model Definition

## 2.1 Conceptual Model Representation and Elicitation Method

-   *Describe what method you will use to elicit or identify the conceptual model.*
-   *How will you represent the final conceptual model? This will likely depend on the method chosen to elicit the conceptual model.*

### 2.2.1 Explain Critical Conceptual Design Decisions

*List and explain critical conceptual design decisions, including:*

-   *spatial and temporal scales,*
-   *selection of entities and processes,*
-   *representation of stochasticity and heterogeneity,*
-   *consideration of local versus global interactions, environmental drivers, etc.*
- *Explain and justify the influence of particular theories, concepts, or earlier models against alternative conceptual design decisions that might lead to alternative model structures .*

### 2.2.2 Model assumptions and uncertainties
*Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be represented in the model. Sources of uncertainty may include:*

-   *exogenous uncertainties affecting the system,*
-   *parametric uncertainty in input data and*
-   *structural / conceptual nonparametric uncertainty in the model.*

## 2.3 Predictor Variables

*Identify and define system system variables and structures, referencing scenario definitions, and the scope of the model as described within problem formulation (1.1.3):*

-   *What variables would support taking this action or making this decision?*
-   *What additional variables may interact with this system (things we can't control, but can hopefully measure)?*
-   *What variables have not been measured, but may interact with the system (often occurs in field or observational studies)?*
-   *What variables are indice or surrogate measures of variables that we cannot or have not measured?*
-   *In what ways do we expect these variables to interact (model structures)?*

## 2.4 Define prior knowledge, data specification and evaluation

*This section should specify the plan for collecting, processing and preparing data available for parameterisation, determining model structure, and for scenario analysis.*

### 2.4.1 Document data sources

*Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model. Describe how the data is arranged, in terms of replicates and covariates.*

**For pre-existing data (delete as appropriate):**

-   *For each dataset, is the data open or publicly available? (Y/N)*
-   *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrictions on the use of data.*
-   *Date of download, access, or future access:*
-   *Describe the source of the data - what entity originally collected this data? (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
-   *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
-   *Prior work based on this dataset - Have you published / presented any previous work based on this dataset? Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont this dataset you have worked on.*
-   *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
-   *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study. For example, have you read any reports or publications about this data?*

**Sampling Plan (For data you will collect) – delete as appropriate:**

-   *Data collection procedures*
    -   *Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged.*
    -   *Describe any inclusion or exclusion rules, and the study timeline.*
-   *Sample Size: Describe the sample size of your study.*
-   *Sample Size Rationale: Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel.*
-   *If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?*

### 2.4.2 Data Processing and Preparation

*Describe any data preparation and processing steps, including manipulation of environmental layers (e.g. standardisation and geographic projection) or variable construction (e.g. Principal Component Analysis).*

### 2.4.3 Data Exploration or preliminary data analyses

*For each separate preliminary or investigatory analysis:*

-   *State what needs to be known to proceed with further decision-making about the modelling procedure, and why the analysis is necessary.*
-   *Explain how you will implement this analysis, as well as any techniques you will use to summarise and explore your data.*
-   *What method will you use to represent this analysis (graphical, tabular, or otherwise, describe)*
-   *Specify exactly which parts of the data will be used.*
-   *Describe how the results will be interpreted, listing each potential analytic decision, as well as the analysis finding that will trigger each decision, where possible.*

### 2.4.4 Data evaluation, exclusion, and missing data

-   *Describe how you will determine how reliable the data is for the given model purpose. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand*
-   *Document any issues with data reliability.*
-   *How will you determine what data, if any, will be excluded from your analyses? How will outliers be handled? Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.*
-   *How will you identify and deal with incomplete or missing data?*

## 2.5 Conceptual Model Evaluation

-   *Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.*
-   *How will you critically assess any simplifying assumptions?*
-   *Will the conceptual model evaluation process include consultation or feedback from a client, manager, or model user?*

# 3.0 Model Formalisation and Implementation

## 3.1 Model Class, Framework and Approach

-   *Describe what modelling framework, approach or class of model you will use to implement your model and relate your choice to the model purpose and analytical objectives described in 1.1.2 and 1.1.3.*

### 3.1.1 Operationalising Model Variables

-   *Specify how you will operationalise the response variables in the model. This should relate directly to the analytical and or management objectives specified during the problem formulation phase. Operationalisations could include:*
    -   *the extent of a response,*
    -   *an extreme value,*
    -   *a trend,*
    -   *a long-term mean,*
    -   *a probability distribution,*
    -   *a spatial pattern,*
    -   *a time-series,*
    -   *qualitative change, such as a direction of change or*
    -   *the frequency, location, or probability of some event occurring.*
-   *Provide a rationale for your choices, including why plausible alternatives were not chosen.*
-   *For the remaining model variables, specify each variable and their nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).*

### 3.1.2 Model Structure

*Specify model structural features, including:*
-   *the functional form of interactions,*
-   *data structures,*
-   *measures used to specify links,*
-   *any bins or discretisation of continuous variables (Jakeman, 2006).*

### 3.1.3 Model Class / Family

-   *Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions.*
-   *Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints specified in the problem formulation phase.*
-   *Include in your rationale for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*

### 3.2 Model Structure and Parameter Estimation

-   *Specify what approach and methods you will use to identify model structure and parameters.*

### 3.2.1 Parameter/Structure estimation technique

-   *Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006).*
-   *Identify which variables were parameterised directly (if any), such as by expert elicitation or prior knowledge.*
-   *Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms (Liu et al. 2018).*
-   *Repeat the above bullet points but for model structure estimation, if you are using quantitative methods for structure estimation rather than knowledge or expert-judgment driven methods.*

### 3.2.2 Estimation performance criteria

-   *Specify which suite of performance criteria you will use to judge the performance of the model. Examples include correlation scores, coefficient of determination, specificity, sensitivity, AUC, etcetera (Yates et al. 2018). State any numeric thresholds and their qualitative interpretation for judging adequate estimation.*
-   *Explain how you will identify which model features or components are significant or meaningful.*

## 3.3 Model assumptions and uncertainties

-   *Specify all assumptions and key uncertainties in the formal model.*
-   *Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpretation of the model outputs, and*
- *How do these choices about what is included in the model and the way it has been operationalised in the formal model influence any inferences or conclusions you might draw? If the model is being used in a decision-making context, how might these assumptions influence what decisions you can make?*

## 3.4 Specify formal model(s)

-   *Once all decisions have been made about the approach and method of model specification, translate the conceptual model into the formal or quantitative model, i.e. provide mathematical, statistical equations, OR the model code for the models you will fit.*
-   *Some modelling approaches require you to specify initial starting values within the call to the software function, especially when using data-driven approaches to model structure definition and parameterisation, or when using formal model selection procedures -- describe any initial model specifications and parameterisations, including for any tune-in parameters.*

# 4.0 Model Calibration, Fitting & Checking

## 4.1 Model calibration and validation scheme

-   *Describe how you will validate and check the calibration of the model. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) .*
-   *Some models are sensitive to the hyper-parameters you select, e.g. min buckets etc., therefore model tuning using cross-validation needs to be done before the final model can be parameterised / estimated. Please describe any model tuning procedures for estimating hyper-parameters, if any.*
-   *Justify your choices*

### 4.1.2 Describe calibration/validation data

-   *If partitioning data for cross-validation:*
    -   *Describe the approach specifying the number of folds that will be created and the relative size of each.*
    -   *Describe how will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced.*

-   *If using external / independent holdout data:*
    -   *Describe any known differences between the training and validation datasets and the size of the validation dataset.*
    -   *Describe how will you document and/or share the external data such that any subsequent modelling can be computationally reproduced.*
    -   *It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modellers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?*

## 4.2 Implementation verification

-   *What Quality Assurance measures will you take to verify the model has been correctly implemented as specified? Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging and silent errors.*
    -   *Tests could include syntax checking of code, and code reviews by peers or in-house code-reviewers at the journal / publisher.*
    -   *Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as dictated by the model description (Grimm et al. 2014).*
    -   *Technical measures include using unit tests, or in-built checks within functions to prevent potential errors.*

## 4.3 Model Checking

*"Model Checking" goes by many names ("calibration testing or validation", "conditional verification", "quantitative verification", "model output verification"), and refers to a series of analyses that assess a model's performance in representing the system of interest (Conn et al. 2018). Model checking therefore helps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system (Conn et al. 2018). Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation (Jakeman, Letcher, and Norton 2006).*

### 4.3.1 Quantitative model checking

*During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences.*

-   *Specify any diagnostics or tests you will use during model checking to assess a model's performance in representing the system of interest.*
-   *For each test, specify the criteria that you will use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.*

### 4.3.2 Qualitative model checking

-   *Explain how you will qualitatively check the model, and whether and how you will include users and clients in the process. Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.*

### 4.3.3 Assumption violation checks

-   *Explain how you will demonstrate robustness to model assumptions and check for violations of model assumptions.*
-   *If you cannot perform quantitative assumption checks, describe what theoretical justifications would justify a lack of violation of or robustness to model assumptions.*
-   *If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.*
-  *If assumption violations cannot be avoided, explain how you will explore the consequences of assumption violations on the interpretation of results (To be completed in interim iterations of the preregistration, only if there are departures from assumptions as demonstrated in the planned tests above).*

# 5.0 Model Validation and Evaluation

## 5.1 Model output corroboration

-   *State whether you will corroborate the model outputs on external test data, and ensure that any independent validation data is documented in 4.1.2.*
-   *Testing the model against data that are collected spatially and/or temporally independent from the training dataset minimizes the risk of artificially inflating model performance measures (Araújo et al. 2019). Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system), model validation is not always possible (Grimm, et al. 2014). If you are unable to evaluate the model outputs against independent data, explain why not, and describe what steps you will take to interrogate the model. For example, Independent predictions can instead be tested on submodels, alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014).*

### 5.1.1 Model corroboration performance metrics and criteria

-   *Specify what performance measures you will use to evaluate the model and briefly explain how each test relates to different desired properties of a model's performance. Examples include measures of agreement between predictions and independent observations, or estimates of accuracy, bias, calibration, discrimination refinement, resolution and skill (Araújo et al. 2019).*
-   *Spatial, temporal and environmental pattern of errors and variance can change the interpretation of model predictions and conservation decisions (Araújo et al. 2019), where relevant and possible, describe how you will characterise and report the spatial, temporal and environmental pattern of errors and variance.*
-   *If comparing alternative models, specify what measures of model comparison or out-of-sample performance metrics will you use to find support for alternative models or else to optimise predictive ability.*
-   *State what numerical threshold or qualities you will use for each of these metrics.*

## 5.2 Model Analysis

### 5.2.1 Uncertainty Analyses

-   *Please describe how you will characterise model and data uncertainties. E.g. propagating known uncertainties through the model, investigating the effect of different model scenarios with different parameters and modelling technique combinations (Araújo et al. 2019), or empirically constructing model distributions from model output data points, and extracting summary statistics, including the mean, median, variance, and constructing confidence intervals (Saltelli et al. 2019).*
-   *Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase. For instance, high model precision may be important for predictive modelling when cost, benefit and risk must be adequately gauged, but for decision-making contexts where only the predicted ranked order of management options is necessary, model accuracy and precision might be less important (Jakeman, Letcher, and Norton 2006).*
-   *Describe how you will summarise the results of these in silico experiments with graphical, tabular, or other devices, such as summary statistics.*
-   *If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic parameter covariance, (Jakeman, Letcher, and Norton 2006)), specify which measure of uncertainty you will measure.*

### 5.2.2 Sensitivity Analyses

*Sensitivity analysis examines how uncertainty in model outputs can be apportioned to different sources of uncertainty in model input (Saltelli et al. 2019).*

-   *Describe the sensitivity analysis approach you will take: deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).*
-   *Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameters will be varied.*
-   *State the primary objective of each sensitivity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or alternatively to identify which variables or factors contribute little to overall model outputs, and so can be 'dropped' from future iterations of the model (Saltelli et al. 2019).*

### 5.2.3 Model application / scenario analysis

-   *Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario specified in section 1.*
-   *Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections.*
-   *What raw and transformed model outputs will you extract from the model simulations or projections, and how will you map, plot, or otherwise display and synthesise the results of scenario and model analyses.*
-   *Explain how you will analyse the analysis output to answer your analytical objectives. For instance, describe any trade-off or robustness analyses you will undertake to help evaluate and choose between different alternatives in consultation with experts or decision-makers.*

### 5.2.4 Other in silico analyses

-   *Describe any other simulation experiments and or robustness analyses you will perform on the model, including any metrics and their criteria / thresholds for interpreting the results of the analysis.*

# References

Araújo, MB, RP Anderson, A Márcia Barbosa, CM Beale, CF Dormann, R Early, RA Garcia, et al. 2019. "Standards for Distribution Models in Biodiversity Assessments." Sci Adv 5 (1): eaat4858.

Augusiak, Jacqueline, Paul J Van den Brink, and Volker Grimm. 2014. "Merging Validation and Evaluation of Ecological Models to 'Evaludation': A Review of Terminology and a Practical Approach." Ecological Modelling 280: 117–28.

Barnard, David M., Matthew J. Germino, David S. Pilliod, Robert S. Arkle, Cara Applestein, Bill E. Davidson, and Matthew R. Fisk. 2019. "Can't See the Random Forest for the Decision Trees: Selecting Predictive Models for Restoration Ecology." Restoration Ecology.

Boets, Pieter, Dries Landuyt, Gert Everaert, Steven Broekx, and Peter L M Goethals. 2015. "Evaluation and Comparison of Data-Driven and Knowledge-Supported Bayesian Belief Networks to Assess the Habitat Suitability for Alien Macroinvertebrates" 74: 92–103.

Conn, Paul B, Devin S Johnson, Perry J Williams, Sharon R Melin, and Mevin B Hooten. 2018. "A Guide to Bayesian Model Checking for Ecologists." Ecological Monographs 9: 341–17.

Fraser, Hannah, Libby Rumpff, Jian D L Yen, Doug Robinson, and Brendan A Wintle. 2017. "Integrated Models to Support Multiobjective Ecological Restoration Decisions." Conservation Biology 31 (6): 1418–27.

Giljohann, K. M., McCarthy, M. A., McCarthy, M. A., Kelly, L. T., & Kelly, L. T. (2014). Choice of biodiversity index drives optimal fire management decisions. Ecological Modelling.

Gregory, R., Failing, L., Failing, L., Harstone, M., Harstone, M., Long, G., . . . Ohlson, D. (2012). Structured decision making: a practical guide to environmental management choices. John Wiley & Sons.

Grimm, Volker, Jacqueline Augusiak, Andreas Focks, Béatrice M Frank, Faten Gabsi, Alice S A Johnston, Chun Liu, et al. 2014. "Towards Better Modelling and Decision Support: Documenting Model Development, Testing, and Analysis Using Trace." Ecological Modelling 280: 129–39.

Jakeman, A J, R A Letcher, and J P Norton. 2006. "Ten Iterative Steps in Development and Evaluation of Environmental Models." Environmental Modelling & Software 21 (5): 602–14.

Liu, C. C., & Aitkin, M. (2008). Bayes factors: Prior sensitivity and model generalizability. Journal of Mathematical Psychology, 52(6), 362-375. <doi:10.1016/j.jmp.2008.03.002>

Liu, Zelin, Changhui Peng, Timothy Work, Jean-Noel Candau, Annie DesRochers, and Daniel Kneeshaw. 2018. "Application of Machine-Learning Methods in Forest Ecology: Recent Progress and Future Challenges." Environmental Reviews 26 (4): 339–50.

Mahmoud, Mohammed, Yuqiong Liu, Holly Hartmann, Steven Stewart, Thorsten Wagener, Darius Semmens, Robert Stewart, et al. 2009. "A Formal Framework for Scenario Development in Support of Environmental Decision-Making." Environmental Modelling & Software 24 (7): 798–808.

McDonald-Madden, Eve, Peter W. J. Baxter, and Hugh P. Possingham. 2008. "Making Robust Decisions for Conservation with Restricted Money and Knowledge." Journal of Applied Ecology 45 (6): 1630–8.

Moallemi, Enayat A., Sondoss Elsawah, and Michael J. Ryan. 2019. "Strengthening 'Good' Modelling Practices in Robust Decision Support: A Reporting Guideline for Combining Multiple Model-Based Methods." Mathematics and Computers in Simulation.

Saltelli, Andrea, Ksenia Aleksankina, William Becker, Pamela Fennell, Federico Ferretti, Niels Holst, Sushan Li, and Qiongli Wu. 2019. "Why so Many Published Sensitivity Analyses Are False: A Systematic Review of Sensitivity Analysis Practices." Environmental Modelling & Software 114: 29–39.

Tulloch, A. I. T., Maloney, R. F., Joseph, L. N., Bennett, J. R., Di Fonzo, M. M. I., Probert, W. J. M., . . . Possingham, H. P. (2015). Effect of risk aversion on prioritizing conservation projects. Conservation Biology.

White, Craig R, and Dustin J Marshall. 2019. "Should We Care If Models Are Phenomenological or Mechanistic." Trends in Ecology & Evolution 34 (4): 276–78.

Yates, KL, PJ Bouchet, MJ Caley, K Mengersen, CF Randin, S Parnell, AH Fielding, et al. 2018. "Outstanding Challenges in the Transferability of Ecological Models." Trends Ecol. Evol. (Amst.) 33 (10): 790–802.

### 2023-09-27 _analysis/index.Rmd
rename _analysis

--- 2020-09-16 index.Rmd
@@
 Coupled with a lack of transparency -- or "undisclosed flexibility" -- researcher degrees of freedom are vulnerable to being exploited by researchers because of cognitive biases and incentive structures within science.
-\<lead in sentence to next para??\>
+(lead in sentence to next para??)
 
@@
 
-The goal of this study is to \<rethink\> preregistration to better suit the domain-specific challenges and methodological particularities of model-based research in ecology.
+The goal of this study is to (rethink) preregistration to better suit the domain-specific challenges and methodological particularities of model-based research in ecology.
 This research builds on recently developed "QRP roadmaps" (Gould et al.
@@
 *Research transparency* refers to both production transparency --- which includes research artefacts like open-access data and materials, or data collection procedures --- and analytic transparency --- which describes a complete account of how the analytic conclusions are drawn from the data (Lupia and Elman, 2014).
-\<refs to model reporting literature- it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced\>
+(refs to model reporting literature- it's dire, can't properly be evaluated by a reviewer, moreover can't be reproduced)
 
@@
 
-More specifically, we intend this template to be suitable not just in 'pure-research'-only contexts, but to be suitable for ecologists and conservation decision-makers who are working in the 'actionable sciences' \<which guy was this again..
+More specifically, we intend this template to be suitable not just in 'pure-research'-only contexts, but to be suitable for ecologists and conservation decision-makers who are working in the 'actionable sciences' (which guy was this again..
 Sutherland??
-define\>.
+define).
 The template should be adequate for use by scientists and modellers working in applied ecological contexts, but also for management and conservation decision-making contest where there are real-world decisions hinging on the model-based research.
 
-\<Research Types: Modelling - what do we mean by modelling??
+(Research Types: Modelling - what do we mean by modelling??
 I think if we don't specify this then people will just think 'classical linear regression' as modelling, and then will say, well why don't use existing hypothesis testing templates, like AsPredicted.
-I need to specify some particularities about the type of modelling that occurs in in these research contexts, and why existing templates don't suit, and that we need to design specifically for these research contexts.\>
+I need to specify some particularities about the type of modelling that occurs in in these research contexts, and why existing templates don't suit, and that we need to design specifically for these research contexts.)
 
@@
 
-Consequently, we aim not just to simply design a new template, deciding on its structure and content, but we propose a workflow and methodology to be implemented by the author when completing a preregistration of their model-based research \<see section - inner logic of the model - discuss features model-based research, like iterative development?\>
+Consequently, we aim not just to simply design a new template, deciding on its structure and content, but we propose a workflow and methodology to be implemented by the author when completing a preregistration of their model-based research (see section - inner logic of the model - discuss features model-based research, like iterative development?)
 
@@
 
-**Identifying aspirations and desirable attribute for the template**
+**Specifying template objectives and performance criteria**
 
-What attributes do we aim for the template
-
-A) an interface that ensures ease of use for the creator of a preregistration
-
-B) structure and content that is necessarily opinionated..
-forces user to a particular workflow, so getting that workflow right, that reflects both norms and practice of model-based research is pivotal.
+The overarching quality we wish the preregistration template to achieve is to adequately constrain researcher degrees of freedom during the model development process so as to reduce the the potential for questionable research practices and to transparently document the analytic decision process during modelling.
+Preregistration templates should be specific, precise and exhaustive [@Wicherts2016] if they are to adequately delimit researcher degrees of freedom.
+They should be *specific*, such that the preregistration is sufficiently detailed in describing all phases of the research process; *precise* such that each pre-specified aspect of the analysis plan is unambiguous and has only one interpretation; and *exhaustive* such that the preregistered analysis excludes the 'possibility of deviations from the preregistered research plan'.
+Although quantitative evidence about the effectiveness of preregistration is largely lacking, Bakker et al.
+[-@Bakker2018] quantitatively compared structured and unstructured preregistrations using a Transparency Score protocol that encapsulated these performance criteria, and concluded that detailed instructions and guidance within a template yields preregistrations that are better at restricting researcher degrees of freedom than unstructured formats with minimal guidance.
+Consequently we adopt these properties as are the means objectives for the template and aim to develop an *opinionated* template that provides detailed guidance and illustration such that preregistered analyses are sufficiently *specific,* *precise*, and *exhaustive* in order to curtail researcher degrees of freedom.
+Preregistered analyses meeting these criteria should result in transparent and detailed records of any analysis decisions.
 
@@
 We split the activities into two groups, to better focus discussion for each of these templates.
-All workshop materials and workshop outputs can be found on the OSF component here \<LINK\>.
+All workshop materials and workshop outputs can be found on the OSF component here (LINK).
 
@@
 In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto a group workflow template.
-This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase.
+This template was prepared earlier ((specific OSF link to the document)), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase.
 The development of the draft modelling workflow was derived from previous work by Gould et al.
@@
 We conducted two lots of structured discussions, one for each of the field and modelling template groups.
-Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature \<cite / link to OSF, or even discuss??\>.
+Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature (cite / link to OSF, or even discuss??).
 These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in activity 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.
@@
 
-**1\. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
+**1. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
 
 The goal of this analysis task was to generate a complete draft workflow for both fieldwork and modelling and data-analysis that will form the basis of the final preregistration templates.
-We used a combination of inductive and deductive coding \<is this what it is called\> to analyse the workshop outputs and generate a model scientific workflow for model-based research that would constitute the presumed scientific workflow that would be used to guide the structure of the preregistration template.
-We collated and coded the workshop outputs using \<atlas.TI, insert software citation and version\>, \<a qualitative coding software\>.
-All workshop outputs can be found here (\<Link to Workshop Outputs OSF repository\>).
+We used a combination of inductive and deductive coding (is this what it is called) to analyse the workshop outputs and generate a model scientific workflow for model-based research that would constitute the presumed scientific workflow that would be used to guide the structure of the preregistration template.
+We collated and coded the workshop outputs using (atlas.TI, insert software citation and version), (a qualitative coding software).
+All workshop outputs can be found here ((Link to Workshop Outputs OSF repository)).
 
 A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management.
-Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], \<insert link to appendices or just present the final?\>.
+Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], (insert link to appendices or just present the final?).
 Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps.
-We revised the coding structure \<inductively (is this the word)\>, according to patterns and themes identified across multiple personal workflows.
-For example, we added the final phase \<name for last phase\>, which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions.
-In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both \<idealised norms and norms of practice, Pu et al., see other philosophy of sci refs\>.
+We revised the coding structure (inductively (is this the word)), according to patterns and themes identified across multiple personal workflows.
+For example, we added the final phase (name for last phase), which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions.
+In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both (idealised norms and norms of practice, Pu et al., see other philosophy of sci refs).
 This workflow is used to provide structure and content to the preregistration template.
 
-**2\. Translating Workflows into Preregistration Templates**
+**2. Translating Workflows into Preregistration Templates**
 
 We translated the model-based workflow identified in part I of workshop analysis into a text-based preregistration template.
-Reflecting on existing templates \<link to OSF repository of templates page\>, we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis.
+Reflecting on existing templates (link to OSF repository of templates page), we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part I of workshop output analysis.
 Following the features of a good guideline identified by Moallemi et al.
-[-@Moallemi2019], we provide examples of choices or illustrations to clarify each step of the preregistration, for parts of the modelling process that are semantically, epistemically and/or philosophically contentious or ambiguous \<link to paper describing contentious issues in modelling history\>.
+[-@Moallemi2019], we provide examples of choices or illustrations to clarify each step of the preregistration, for parts of the modelling process that are semantically, epistemically and/or philosophically contentious or ambiguous (link to paper describing contentious issues in modelling history).
 We synthesised the following ontology to build our preregistration template:
 
-1\. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. \<Note, that these are being scrapped from the template and are turned into tool-tips\>
+1. Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. (Note, that these are being scrapped from the template and are turned into tool-tips)
 
-2\.
+2.
 Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow.
-\<again, these are being scrapped and turned into tool-tips.\>
+(again, these are being scrapped and turned into tool-tips.)
 
-3\.
-Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
+3.
+Choices ---) each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
 The style should be in the form of a directive or a question.
@@
 
-\- What should a research workflow look like for a modeller in an applied ecology setting who wishes to preregister their work?
+- What should a research workflow look like for a modeller in an applied ecology setting who wishes to preregister their work?
 
-\- How do we accommodate the iterative cycle of model development into the preregistration process?
+- How do we accommodate the iterative cycle of model development into the preregistration process?
 
-\- At what stage in a modelling or structured decision-making process should preregistration begin?
+- At what stage in a modelling or structured decision-making process should preregistration begin?
 
@@
 
-\<which other template design papers have included evaluation criteria and rationale, Sophia's?\>
+(which other template design papers have included evaluation criteria and rationale, Sophia's?)
 
@@
 
-Modellers \<here, here, and here\> have expressed concern about preregistration making modelling a more difficult task than it already is, and hampering the ability to meet project deadlines, especially in decision-problems.
-Many modellers already struggle with sufficiently documenting models during publication and reporting \<ref, model reporting and transparency review\>, and modellers working in government or other agencies have expressed concern that preregistration will simply add to the list of  'red-tape' expected of them.
+Modellers [@MacEachern2019] (here, here, and here) have expressed concern about preregistration making modelling a more difficult task than it already is, and hampering the ability to meet project deadlines, especially in decision-problems.
+Many modellers already struggle with sufficiently documenting models during publication and reporting (ref, model reporting and transparency review), and modellers working in government or other agencies have expressed concern that preregistration will simply add to the list of  'red-tape' expected of them.
 Given these concerns, and considering the concerns of many with regard to the iterative nature of model development precluding making preregistration impossible or inappropriate, we figured it was of utmost importance that the templates be also evaluated on their ease-of-use and feasibility.
@@
 
-We tested and evaluated the templates on the ability to meet these criteria by using a model-based research case-study \<see box\>.
-Rather than 'digging back' through the version history of a completed preregistration to ask researchers about the process and how it had been developed in detail \<like Pu et al\>, we chose to leverage the version-control and collaborative project management features of Git and GitHub [@BenBondLamberty:2018by; @Blischak:2016kr] as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to 'live-develop' the template template.
-For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration, the suggested change, and its justification would be both recorded in GitHub and linked to the actual preregistration.
-Since *Pu et al* (p16).
-Found that the process of creating a preregistration is 'difficult to reconstruct with interview questions alone', we we aimed to capture the analysts 'thinking in the moment' by using comments and GitHub issues as to why a particular decision about the template item inclusion and specification should change.
-\<How we did this using GitHub:How to present this material?
-Link to the wiki issue and instructions to users?\>
+We tested and evaluated the templates on the ability to meet these criteria by using a model-based research case-study.
+The user-centered design procedure we implemented follows two phases of evaluation.
+Firstly, the template is iteratively evaluated and refined 'in situ' -- throughout the process of preregistering the VEFMAP analysis with the template, any required changes to make the template more *specific*, *precise*, and *exhaustive* are implemented.
+After the preregistration is complete, the preregistration was subjected to a final round of quantitative evaluation in which we evaluated the ability of the template to meet its objectives using a Transparency Score protocol.
+We also subjectively reflected on the template and the process of preregistration.
+
+*Case-study*
+
+Chris / Lyndsey to complete -- 1 paragraph or 2 short paragraphs describing the decision problem / trigger and the modelling research question / problem.
+
+*Iterative refinement*
+
+Rather than 'digging back' through the version history of a completed preregistration to ask researchers about the process and how it had been developed in detail, we chose to leverage the version-control and collaborative project management features of Git and GitHub (Ben, Smith, and Bailey 2018; Blischak, Davenport, and Wilson 2016)  to 'live-develop' the template throughout the process of writing the preregistration for the case-study.
+Since the process of creating a preregistration is hard to reconstruct using *post hoc* interviews (Pu et al.
+2019), we aimed to capture the analysts 'thinking in the moment' by using comments and GitHub issues describing why a particular decision to refine the template was made.
+For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration; the suggested change, and its justification was recorded in GitHub and simultaneously linked to the actual preregistration document change in the version-control record.
+
+*Quantitative Evaluation using a Transparency Score Protocol*
+
+We created a Transparency Score protocol following the method used by Bakker et al.
+[-@Bakker2018] to assess and compare the quality of preregistrations created with structured *versus* unstructured templates.
+We first created a 'QRP coding protocol' based on QRP roadmaps (Gould et al.
+*in prep.*), which we used to code the final preregistered analysis for the case-study (Table X).
+This process identified the presence and/or possibility of QRP's to occur within the preregistered analysis.
+
+After coding for QRP's within the preregistration, we then quantitatively assessed the quality of the preregistration using a Transparency Score protocol (link to protocol in OSF).
+For each QRP in the QRP coding protocol, we assigned a score between 0 and 3 where:
+
+* 0 = the QRP was not restricted at all
+* 1 = the QRP was restricted to some degree
+* 2 = the QRP was completely restricted (the template was both *specific* and *precise*)
+* 3 = the QRP was completely restricted and the preregistration was *exhaustive*, i.e. it included explicit statements that no deviation from the way it was registered would occur.
+
+For every QRP, these scores are qualified in specific detail to aid in scoring the template.
+
+*Reflection*
 
@@
 In addition, the process of preregistration was a new and unfamiliar task.
-
-\<TEXT HERE DESCRIBING THE ENVIRONMENTAL FLOWS MANAGEMENT PROBLEM + RESEARCH PROBLEM, to be completed by Chris / Lyndsey\>.
 
@@
 
-The template is designed to be used within an adaptive preregistration methodology \<link to section\>.
+The template is designed to be used within an adaptive preregistration methodology (link to section).
 
-In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by \<Srivastava\>.
+In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by (Srivastava).
 This view of preregistration breaks with the current model of preregistration, wherein the author writes a single deterministic preregistration containing a rule for every decision, and shifts from ideation, preregistration to execution of the plan.
 Two features define adaptive preregistration.
-It contains 'plans to deploy flexible strategies' \<Srivastava\> wherein the author can supply a heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses.
-Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed \<Srivastava\>.
-Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, <x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?id=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0>\>.
+It contains 'plans to deploy flexible strategies' (Srivastava) wherein the author can supply a heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses.
+Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed (Srivastava).
+Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed (Link to Figure, GitHub as a platform for adaptive preregistration, <x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?id=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0>).
 
@@
 
-Consequently we propose a methodology for implementing adaptive preregistration that leverages features of GitHub to also facilitate the task of linking the actual study and analysis conducted to the pre-specified plan in the preregistration \<link to part of GH methodology that explains how checking is accounted for in the procedural design\>.
-In addition, we designed a shiny interface for completing the preregistration to include a 'checking mode' that easily allows the author to state how and why the the analysis might have diverged from the planned \<link to part of shiny app interface that discusses checking mode\>.
+Consequently we propose a methodology for implementing adaptive preregistration that leverages features of GitHub to also facilitate the task of linking the actual study and analysis conducted to the pre-specified plan in the preregistration (link to part of GH methodology that explains how checking is accounted for in the procedural design).
+In addition, we designed a shiny interface for completing the preregistration to include a 'checking mode' that easily allows the author to state how and why the the analysis might have diverged from the planned (link to part of shiny app interface that discusses checking mode).
 
@@
 
-We propose leveraging both the *version-control* and the project management features of *GitHub* \<link to GitHub\> as a tool for implementing adaptive preregistration \<link to box\>.
+We propose leveraging both the *version-control* and the project management features of *GitHub* (link to GitHub) as a tool for implementing adaptive preregistration (link to box).
 When a document is version-controlled using git and GitHub, each version of a document is time-stamped, and is assigned its own unique document identifier, or *commit-hash.* Because GitHub documents are 'time-stamped' the genesis of the preregistration from one version to the next is apparent.
-Moreover, using the *diff* view of a document on *GitHub*, exactly *how* a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted \<insert picture\>.
+Moreover, using the *diff* view of a document on *GitHub*, exactly *how* a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted (insert picture).
 Since the preregistration document would exist within the project *repository* containing the data and modelling and analysis code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated version of the plan --- thus facilitating the process of model checking.
@@
 Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding *GitHub issue* created (Figure X, step 3, document).
-\<define GitHub issue, <https://guides.github.com/features/issues/>\>.
+(define GitHub issue, <https://guides.github.com/features/issues/>).
 Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an *issue*'s thread.
 Each issue is assigned a unique number.
-Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the *issue's* thread (Fig X, step 4, Do) \<screen-cap?\>.
+Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the *issue's* thread (Fig X, step 4, Do) (screen-cap?).
 The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new *release*.
@@
 
-\- stating what needs to be known to move forward with the modelling and analysis
+- stating what needs to be known to move forward with the modelling and analysis
 
-\- explaining how the researcher will test this, and what parts of the data will be used in that analysis
+- explaining how the researcher will test this, and what parts of the data will be used in that analysis
 
-\- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.
+- describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.
 
@@
 We considered removing in-template text describing and defining key concepts, phases and steps, as well as reducing the number of decision-points in a model-based research workflow captured by the template by removing items within the template.
-\<@TODO - did we end up removing any items by the end of the preregistration and evaluation with Chris' study?\> While we 'tweaked' the template items and descriptive text during the in situ evaluation of the templates for the environmental flows preregistration, we decided not to alter the template structure and content significantly, but to instead implement a new format and interface for working with the template.
+(@TODO - did we end up removing any items by the end of the preregistration and evaluation with Chris' study?) While we 'tweaked' the template items and descriptive text during the in situ evaluation of the templates for the environmental flows preregistration, we decided not to alter the template structure and content significantly, but to instead implement a new format and interface for working with the template.
 We outline these reasons below.
 
-1\. Opinionated Design - norms around modelling practice.  'Good Modelling Practice' Literature. This also has the benefit of aiding new modellers.
+1. Opinionated Design - norms around modelling practice.  'Good Modelling Practice' Literature. This also has the benefit of aiding new modellers.
 
-2\.
+2.
 A related point: linguistic ambiguity and other problems with modelling concepts and terms.
@@
 
-Consequently, we instead decided to implement a user-friendly interface that makes the completion of a preregistration with our template much more *approachable\<?adj\>*.
+Consequently, we instead decided to implement a user-friendly interface that makes the completion of a preregistration with our template much more *approachable(?adj)*.
 
@@
 The shiny app is intended as an interface for authors and analysts to respond to template items and input their planned decisions to complete a preregistration for their study.
-In addition to having directive text and questions for each item, there is also 'help-text' consisting of descriptions of the modelling phases and steps, as well as definitions and references to supporting resources, such as decision-tools for choosing and justifying methodological choices \<reference to image with shiny app\>
+In addition to having directive text and questions for each item, there is also 'help-text' consisting of descriptions of the modelling phases and steps, as well as definitions and references to supporting resources, such as decision-tools for choosing and justifying methodological choices (reference to image with shiny app)
 
-Once the preregistration has been completed \<for the time being, it doesn't actually have to be complete, see 'living preregistration' section\>, the author is able to 'generate a report'.
+Once the preregistration has been completed (for the time being, it doesn't actually have to be complete, see 'living preregistration' section), the author is able to 'generate a report'.
 The report removes any supporting text in the user interface, and generates a text-based document, containing the preregistration template items and any responses to them by the author or analyst.
@@
 
-Norms around preregistration are both uncertain and vary among preregistration users (Pu et al.) By norms, we mean \<insert definition from Pu\>.
+Norms around preregistration are both uncertain and vary among preregistration users (Pu et al.) By norms, we mean (insert definition from Pu).
 Given that the preregistration template is *opinionated* in that it presumes particular norms around the modelling research workflow and sets of activities, and decision-points to be reported, we have designed both the underlying infrastructure of the shiny app, and chosen to host the template and interface on GitHub to allow users to adapt any aspect of the template itself to be congruent with their personal needs and norms around model-based practice.
@@
 We have included design features in the infrastructure of the shiny app that allow for people to very easily change the structure and content of the template, that will then be rendered into the user-interface.
-\<Probably won't describe here, but could link to a vignette on the app that describes the following: easy modification - set up the infrastructure of the shiny app to take in a CSV file, which is then exported to a JSON file for creating the shiny app (alt, just modify the JSON directly)\>
+(Probably won't describe here, but could link to a vignette on the app that describes the following: easy modification - set up the infrastructure of the shiny app to take in a CSV file, which is then exported to a JSON file for creating the shiny app (alt, just modify the JSON directly))
 
@@
 
+## Case Study ##
+
+Should speak to the following, which could be included in the supporting materials.
+
+- Completed preregistration
+
+- Interim Preregistration Documents
+
+- link to repository
+
+- screenshots illustrating  how the GitHub process is used.
+
+----
+
 ## Discussion ##
@@
 
-\> *Lack of discipline specific preregistration templates impedes the uptake of open-science tools, such as preregistration within ecology.*
+) *Lack of discipline specific preregistration templates impedes the uptake of open-science tools, such as preregistration within ecology.*
 
-\> *Domain specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom unique to a particular domain-specific methodology.*
+) *Domain specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom unique to a particular domain-specific methodology.*
 
@@
 
-\> we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both idealised norms and norms of practice.
+) we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both idealised norms and norms of practice.
 
-3\.
+3.
 Our UX approach revealed critical features of the model development process, namely the inherent iterative nature of model development, and requirements for model checking that forced us to reconsider the procedure for completing a preregistration suitable to ecological modelling.
@@
 
-4\.
+4.
 
-\> We proposed a methodology for managing the preregistration process that leverages features of GitHub that implement .....
+) We proposed a methodology for managing the preregistration process that leverages features of GitHub that implement .....
 This methodology (and interface?) deviates from the current/traditional deterministic, single-use write-once read-only document currently used in preregistration.
 
-5\.
+5.
 Furthermore, we created an interface for completing and amending interim preregistrations using shinyApps to facilitate uptake of preregistration in both pure and applied ecological modelling research.
 
-*consisting of a series of*
+***Evaluation and testing (I think this provides a nice lead on to the guts of the discussion, so it should be at the end of our summary of what we did).***
+
+Berna Devezer et al. argue that most open science reforms propose them and are never formally tested. In this research, we provide a qualitative assessment of our new method only.
+
+Thus preregistration is based on the working hypothesis that it is able to mitigate against QRPs.
+
+We need empirical testing of this hypothesis, also so we can empirically evaluate whether the new method meets its intended purpose (link back to intended purpose of the preregistration specified in the methods section).
+
+Then, following that, we need metaresearch, so empirical research at the level of the body of literature, perhaps examining transparency is the easiest starting point..
+but how would you empirically evaluate a series of models??)
+
+Either way alternative benefits of preregistration (pu et al. talks about this.. what about others).
 
@@
 
-Hosting the preregistration template on GitHub, has the added benefit of providing a mechanism for dealing with *uncertain* *norms* in modelling practice \<Pu et al\>, and for following modelling methods and decision contexts that might not fit the presumed modelling development cycle underpinning the preregistration template we designed.
+Hosting the preregistration template on GitHub, has the added benefit of providing a mechanism for dealing with *uncertain* *norms* in modelling practice (Pu et al), and for following modelling methods and decision contexts that might not fit the presumed modelling development cycle underpinning the preregistration template we designed.
 Anyone is free to *fork* our GitHub repository, meaning that the entire contents of the repository can be copied to a users own repository, and the template is free to be altered as the user sees fit.
-This provides the opportunity for creating domain- and/or modelling-approach-specific templates \<Cruwell\>.
+This provides the opportunity for creating domain- and/or modelling-approach-specific templates (Cruwell).
 
@@
 We encourage modellers to preregister their research using our template and methodology here, and to provide feedback and suggestions for improvement.
-Our template + shiny app is hosted on \<insert GitHub link\>, suggested changes may be submitted in the form of a pull-request to the repository \<link to how to make a pull request\>, or by posting a new issue \<link to creating a new issue on the repo\>.
+Our template + shiny app is hosted on (insert GitHub link), suggested changes may be submitted in the form of a pull-request to the repository (link to how to make a pull request), or by posting a new issue (link to creating a new issue on the repo).
 The repository and issues are publicly available, so any discussion will be transparent and open to the ecological modelling community.
+
+**Putting a positive spin on the time-costs of recording all decisions. The result of completing an adaptive preregistration using our methodology is that you get a 'Lab Notebook' at the end of it.**
+
+People have described modelling as being a digital or computational computer lab.
+I think the term '*in silico* experiments' has been used to describe some computational modelling, especially for simulation modelling.
+I'm pretty sure that some authors have made arguments that we would expect rigorous note-taking and recording in the lab environment, so why should we expect any less in a computational modelling setting?
+
+**Preregistration will NOT lead to punch card style analysis (ala MacEachern).**
+
+In our preregistration testing with the case study, we found the opposite, that it prompted critical thinking about analytical choices during modelling.
+
+Moreover, our design template is *opinionated*. (See Elith's latest SDM reporting guideline) -) it provides a *guide*. For those whom are at risk of 'punch card style analysis', ala MacEachern.
+
+**"Preregistering research is an acquired skill" Veldkamp 2017.**
+
 --------------------------

### 2023-09-27 index.qmd
add quarto website scaffolding

---
title: "quarto"
---

This is a Quarto website.

To learn more about Quarto websites visit <https://quarto.org/docs/websites>.

```{r}
1 + 1
```

### 2023-09-27 index.qmd
Add introduction to site and overview of pages / topics

@@
 ---
-title: "quarto"
+title: "Adaptive Preregistration for Model-Based Research"
 ---
 
-This is a Quarto website.
+This site provides a guide to implementing *Adaptive Preregistration for Model-Based Research*, with a focus on ecological modelling as it is used in both basic and applied contexts.
+This site covers the following topics:
 
-To learn more about Quarto websites visit <https://quarto.org/docs/websites>.
-
-```{r}
-1 + 1
-```
+1.  Background on researcher degrees of freedom, the risk of questionable research practices (QRPs), and introduction to preregistration as a tool for mitigating QRPs.
+2.  Adaptive Preregistration
+    1.  *Background*: What is it and what are its key features
+    2.  *A practical guide*: How to implement adaptive preregistration with git and github.
+3.
+Preregistration of model-based research in ecology and conservation.
+Provides an overview of the sorts of things that we should preregister when modelling, what practices to look out for, what things should be registered as decision-trees.

### 2023-09-27 RDF_QRP.qmd
Add scaffolding structure for website and incorp into _quarto.yml

---
title: From Researcher Degrees of Freedom to Questionable Research Practices
---

### 2023-09-27 adaptive_preregistration_background.qmd
Add scaffolding structure for website and incorp into _quarto.yml

---
title: A primer on Adaptive Preregistration
---

### 2023-09-27 adaptive_preregistration_implementation.qmd
Add scaffolding structure for website and incorp into _quarto.yml

---
title: Implementing Adaptive Preregistration with Git and Github
---

### 2023-09-27 adaptive_preregistration_background.qmd
Add and adapt material from adaptive preregistration guide

@@
 title: A primer on Adaptive Preregistration
+author: Elliot Gould
+date: today
+link-citations: true
+bibliography: data/references.bib
 ---
+
+# Background & Motivation
+
+```{r}
+#| label: fig-traditional-prereg
+#| fig-width: 10
+#| fig-height: 5
+#| echo: false
+#| fig-cap: "Conventional Preregistration Process, as illustrated by [@Pu2019]"
+magick::image_read(path = "_supp-mat/Pu_et_al_fig_prereg.png")
+```
+
+The iterative nature of model development is a significant barrier to adopting preregistration for model-based research.
+Translating the preregistration process to a modelling research context is challenging because the iterative nature of model development conflicts with the inner logic of existing preregistration templates that presume a linear research workflow.
+The process of preregistration as it is currently implemented is tailored towards the hypothetico-deductive model of the scientific method.
+As such, the process of preregistration is completely distinct from, and precedes the implementation of the analysis plan laid out in the preregistration.
+In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" [@Pu2019] or prior to analysing the data [@Mertens2019].
+In this process, researcher sequentially shifts from ideation, to study/analysis design and preregistration, to collecting and analysing the data, writing the report or manuscript, to documenting the research and publishing (@fig-traditional-prereg).
+The iterative nature of model development, however, inherently precludes this model of preregistration from being applied to model-based research in ecology and conservation.
+
+## Iterative Cycle of Model Development
+
+Firstly, as Srivastava notes [-@Srivastava2018], there are some decisions that are either difficult to anticipate, or simply cannot be made in advance.
+This is particularly true for decisions occurring at the later phases of the modelling construction and development process (Figure 2, TBC), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses.
+Some items on the template might not be able to be answered on the first go by the researcher until the model is fully or at least close to fully specified, for example, specifying precisely what and how sensitivity analyses or uncertainty analyses will be conducted.
+
+Secondly, very often in modelling, preliminary or investigatory analyses might need to be conducted before being able to specify future decision-steps in the analysis plan.
+For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation.
+In addition, sometimes the results of these analyses and checks mean that the modeller is forced to return to earlier decision-points and change the analysis.
+For instance, the model might fail to converge, or is saturated and runs out of variation to apportion, such that the model must be respecified.
+Other times, the modelling process itself may generate knowledge and learning about the system, and the model structure changes throughout the process of model development.
+
+In these examples, if an analyst were to try and preregister their research, they would switch back and forth between preregistration and analysis, implementing parts of the analysis plan that were specified in the previous preregistration step, and letting the outcomes in that analysis step inform the next specification of the analysis.
+This contravenes a critical feature of preregistration as currently implemented in hypothetico-deductive frameworks, where the analysis plan must be fully specified prior to observing and analysing the data.
+This is not a flaw of existing preregistration, but rather reflects the inner-logic of preregistration for hypothesis-testing contexts.
+The procedure for implementing preregistration in modelling must therefore account for this iterative and non-linear feature of model development, should we wish to preregister model-based research.
+
+## The *Checking Problem*
+
+The iterative nature of model construction and development complicates what is known as the preregistration checking problem even further.
+Authors, reviewers, editors and readers may all engage in the process of checking a manuscript or published paper against its preregistration to verify that the study and analyses were conducted as specified in the preregistration.
+Existing pre-registration formats hinder checking, because the prevailing format of a single static text-based document is designed for authors to very quickly input the required information, rather than to facilitate a comparison of the reported analysis and results against the planned analysis within the preregistration.
+In fact, Pu et al. [-@Pu2019] go so far as to say that this format "verge\[s\] on being write-only media" and are therefore completely ill-suited to the checking task.
+The task of model checking is further complicated by the non-linear and iterative nature of model development, and by our proposed adaptive preregistration methodology because in most modelling studies there will likely be interim preregistrations, such that there are multiple versions of the preregistration to check against the completed analysis.
+
+## Preregistration Template Format & Procedure
+
+Given the iterative nature of adaptive preregistration then, the current convention of the single-use, deterministic and static text-based document format of preregistration is inadequate.
+Interim versions of a preregistration must be time-stamped and ideally version-controlled, such that it is clear not just that two or more versions differ from each other, and in what order they were created, but so that it is explicitly clear in how and why they differ.
+To facilitate preregistration checking of flexible modelling and analysis strategies, we must link the results of both final and preliminary analyses back to the specified strategy: the link between a particular triggered analysis decision, and the analysis outcome that triggered the decision must be explicit.
+Additionally, if the results of some preliminary analyses cause revisions to the model itself or a different modelling procedure or approach to be selected than was originally or previously planned, then the trigger for this decision must be explicitly linked to the next version of the preregistration.
+
+# Adaptive Preregistration
+
+In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template format, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by Srivastava [-@Srivastava2018].
+This view of preregistration breaks with the current model of preregistration, wherein the author writes a single deterministic preregistration containing a rule for every decision, and sequentially shifts from ideation, preregistration to execution of the plan.
+Two features define adaptive preregistration.
+It contains 'plans to deploy flexible strategies' [@Srivastava2018] wherein the author can supply a heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses.
+Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed [@Srivastava2018].
+Thus, as the modeller proceeds through the model development process, they will shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed ( @fig-adaptive-preregistration).
+Consequently we propose a methodology for implementing adaptive preregistration that leverages features of GitHub to also facilitate the task of linking the actual study and analysis conducted to the pre-specified plan in the preregistration.
+
+![Adaptive Preregistration for model-based research using GitHub](figures/LivingPreregistration.png){#fig-adaptive-preregistration}

### 2023-09-27 adaptive_preregistration_implementation.qmd
Add and adapt material from adaptive preregistration guide

@@
 title: Implementing Adaptive Preregistration with Git and Github
+bibliography: data/references.bib
 ---
+
+## Adaptive Preregistration using GitHub
+
+We propose leveraging both the version-control and the project management features of GitHub ([www.github.com](www.github.com)) as a tool for implementing adaptive preregistration.
+GitHub is a web-interface that utilises the version-control system, 'git', to store files, track changes, and enable collaboration on computer code [@braga2023].
+When a document is version-controlled using Git and GitHub, each version of a document can be time-stamped, and is assigned its own unique identifier, or commit-hash[^1] so that chronological records of changes to a collection of files, stored in a 'repository', are made transparent [@braga2023; @Bryan:2008en].
+
+Why GitHub for Adaptive Preregistration?
+
+-   Because GitHub documents are 'time-stamped' the genesis of the preregistration from one version to the next is apparent.
+-   Moreover, using the diff view of a document on GitHub, exactly how and where a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted.
+-   The project management and communication features of GitHub, such as GitHub Issues,
+-   tagging and release
+
+GitHub users can collaborate by contributing, modifying and discussing existing code, reporting bugs and other problems, discover new code and data, and publish new code.
+
+GitHub also integrates several communication features, such as Github Issues, Github Discussions, Github Pages, which allows users to engage in discussions, to plan and collaborate on code, and publish information to a webpage (see Box 1)." ([Braga et al., 2023, p.
+1365](zotero://select/library/items/CQU6GUZL)) ([pdf](zotero://open-pdf/library/items/ML8SB436?page=2&annotation=GPAINXUB))
+
+```{r}
+#| label: fig-git-diff
+#| fig-cap: "'Diff View' of changes made to a preregistration in process."
+#| fig-width: 10
+#| fig-height: 5
+#| echo: false
+magick::image_read(path = "_supp-mat/DiffView-PRT.png")
+```
+
+The preregistration document should be saved within the project repository that contains the data and modelling and analysis code, such that any results from analyses that cause the analysis plan to change can be linked and referenced in the updated version of the plan -- thus facilitating the process of model checking.
+By providing a method for contrasting the actual analysis undertaken against the preregistered analysis, this provides a mechanism for allowing reviewers to properly evaluate the preregistration against the reported analysis.
+It thus provides a way of marking preregistered parts of a report from non-preregistered parts of the analysis.
+
+## How it works...
+
+GitHub is used to track amendments to the preregistration --- each time an additional template item is answered, or an alteration is made to the preregistration, the change should be *committed* to the repository.
+*Semantic versioning* is used to document each version of the preregistration.
+Semantic versioning[^2] is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it [@Kitzes2018vn].
+We have created our own semantic versioning scheme to be used to specifically document the incremental version of the preregistration, in tandem with GitHub's tag and release feature, each interim preregistration is tagged and released with its semantic version number.
+All template items that can be completed should be completed before any data analysis or modelling proceeds.
+The preregistration should be completed until a) no further detail can be supplied or b) no additional template item can be answered until preliminary investigations or analyses are undertaken (Fig 3, steps 1 - 3).
+
+```{r}
+#| label: fig-initial-prereg-zoom
+#| fig-cap: "Initial Preregistration: All template items that can be answered should be completed prior to data analysis or modelling."
+#| fig-width: 10
+#| fig-height: 5
+#| echo: false
+magick::image_read(path = "_supp-mat/initial_prereg-closeup.png")
+```
+
+### ...Shifting from planning to execution...
+
+```{r}
+#| label: fig-document-initial-prereg
+#| fig-cap: "Document: After the initial preregistration, document the initial preregistration using GitHub issues."
+#| fig-width: 10
+#| fig-height: 5
+#| echo: false
+magick::image_read(path = "_supp-mat/document-initial-prereg.png")
+```
+
+Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding GitHub issue[^3] created (Figure 3, step 3, 'Document').
+Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an issue's thread.
+Each issue is automatically assigned a unique number by GitHub.
+Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the issue's thread (Fig 3, step 4, 'Do').
+The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new release.
+
+### ... And back again.
+
+```{r}
+#| label: fig-when-plans-change
+#| fig-cap: "Adaptive Preregistration - When Plans Change"
+#| fig-width: 10
+#| fig-height: 5
+#| echo: false
+
+magick::image_read(path = "_supp-mat/when-plans-change.png")
+```
+
+As the modelling proceeds, further detail on later phases of the analysis can be iteratively updated and preregistered.
+Where the results of model evaluation and analysis reveal that there are problems with the model, plans can be changed, and again the next phase can be preregistered.
+For example, if a researcher finds that assumptions are violated or other unexpected results occur that force a change to the planned analysis, the preregistration can be updated based on the findings of those analyses.
+When this occurs, the major version number of the preregistration document should be incremented (Fig 3 steps 5 - 6).
+Since the findings of the analyses are tracked within the relevant issue for that template item, and the issue is recorded in the preregistration itself, the trigger for the change in the plan is made explicitly clear.
+
+For each altered or new preregistration item, the issue thread should either be updated, or a new issue created respectively (Fig 3, step 6).
+Those analyses are then either reconnected based on the revised plan or the new investigatory analyses are then conducted (Fig 3, step 7).
+
+This process continues until the preregistration is complete, and the researcher can continue to execute the plan as it has been fully described in the final version of the preregistration document (Fig 3, step 8 - 9).
+
+### Flexible Strategies, Preliminary Analyses or Investigations
+
+When a researcher needs to conduct preliminary analyses or check parts of the data before committing to a particular decision about the model or analysis, they can specify a flexible strategy in their response to that preregistration item by:
+
+-   stating what needs to be known to move forward with the modelling and analysis and why the analysis is necessary
+-   explaining how the researcher will test this, and what parts of the data will be used in that analysis
+-   describing how the results will be interpreted, listing each potential decision and the analysis result that will trigger that decision, where possible.
+
+# A step-by-step guide to implementing Adaptive Preregistration with GitHub
+
+## The initial preregistration phase
+
+### The first iteration of preregistration
+
+1.  Edit the document with your responses to the template items
+
+-   View the document in GitHub [here](https://github.com/egouldo/VEFMAP_VEG_Stage6/blob/master/analysis/preregistration_template.Rmd)
+-   Click on The 'edit document' button: ![Screen Shot 2020-05-10 at 6 29 33 pm](https://user-images.githubusercontent.com/8400682/81494459-791b7080-92ec-11ea-979b-900317b4a9b6.png)
+-   You will be able to edit the document directly in GitHub, Edit away!
+
+2.  Commit changes
+
+-   If you are finished, or ready to save your changes by 'committing' them, commit by scrolling underneath the document editing window.
+There will be a box called "commit changes", enter in the commit message box the following:
+-   Your message should contain the following: `Initial Preregistration #IssueNumber`. Tag the Preregistration GitHub issue number in the commit message (For this repository, it's `#17`).
+-   Fill out the additional description box briefly describing the changes you just made
+-   Click "commit"
+
+![Screen Shot 2020-05-18 at 12 19 29 pm](https://user-images.githubusercontent.com/8400682/82168463-2b14f700-9902-11ea-95a0-21fe385629bb.png)
+
+## Incrementally Updating the Preregistration Plan
+
+### Moving from general to specific: Incrementally updating the preregistration
+
+You will likely find that some items on the template cannot be answered just yet.
+One reason being that the level of detail required to adequately answer the template item just isn't clear yet, and more thought and time is needed, or you need to be further along the modelling process to be able to adequately answer the question.
+Another reason is that downstream decisions might depend on some preliminary investigations of the data -- for example, you might need to explore the shape of a measured variable before you can properly specify the full model.
+
+In the latter situation, where possible, you can preregister your preliminary analysis: describing what the preliminary analysis entails, providing a short and relevant list of plausible outcomes, and describing for each of those possible outcomes how the dependent or down-stream decision-point will change depending on the outcome.
+
+### Using GitHub to update the plan
+
+We will use GitHub to track amendments to the preregistration for the two situations described above ( a) adding more detail, answering incomplete questions; b) Responding to findings of a preliminary analysis already described in your preregistration).
+
+To make the changes to the preregistration document, follow the same instructions as for creating the initial preregistration above - simply edit the preregistration in GitHub, and commit the changes.
+
+As for the situation where you are responding to findings of a preliminary analysis that you have already described in your preregistration, see the below sections on `Moving from planning to doing` and `Shifting back to planning from doing`.
+
+## Moving from planning to doing
+
+Before you begin to do or start anything you've planned within your preregistration, you **must** create a GitHub issue for that item, phase, or step in the modelling process, follow these steps:
+
+1.
+View the preregistration document in Github, and click on the line-numbers on the left to highlight the PRT item text, as well as your response to the item.
+For example, let's say you want to start working the model checking, you would highlight the following in the document.
+Note that once you've highlighted some lines, a little box with an ellipsis appears on top of the lines numbers.
+![Screen Shot 2020-05-13 at 1 06 25 pm](https://user-images.githubusercontent.com/8400682/81767208-2d9ad980-951b-11ea-8f59-6d6f583eb155.png)
+2.
+Click on the tiny box with the ellipsis, and a pop-up dialogue box will appear with 4 options: `Copy lines`, `Copy permalink`, `View git blame`, `References in new issue`.
+Select the last option `Reference in new issue.` ![Screen Shot 2020-05-13 at 1 06 58 pm](https://user-images.githubusercontent.com/8400682/81767344-80749100-951b-11ea-8b76-1e829528973a.png)
+3.
+This will open up a new window with a new GitHub Issue for you to complete.
+Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted: ![Screen Shot 2020-05-13 at 1 07 30 pm](https://user-images.githubusercontent.com/8400682/81767518-ef51ea00-951b-11ea-9398-18fffdcbbeae.png) Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this: ![Screen Shot 2020-05-13 at 1 17 21 pm](https://user-images.githubusercontent.com/8400682/81767585-1c9e9800-951c-11ea-9d2c-46d2faf6368a.png)
+
+Add any further information or technical details about going forward and working on this task, as necessary.
+Submit the issue to save.
+Any work, and subsequent commits on this topic should be tagged with the issue number.
+We now have a mechanism for linking your actual analysis and actions to the planned analysis in the preregistration document.
+
+### Dealing with unexpected outcomes, and changing plans: Shifting back to planning from doing
+
+Sometimes assumptions are violated, or there are other unexpected outcomes of preliminary analyses, and you will need to change plans.
+How do we deal with this?
+
+-   Add the URL to the relevant GitHub issue under the corresponding Preregistration Item.
+-   Add any the URL for any linked commits to the Preregistration Item.
+-   revise the Preregistration, where necessary by editing the Preregistration document in GitHub, as per previous instructions.
+-   Commit the changes with the following commit message header:
+-   Describe the trigger for the Preregistration revisions in brief detail within the commit message body, adding the link to the relevant GitHub issue containing the tagged commits that triggered the revision.
+
+## 'Living preregistration' - tracking changes to the Preregistration Draft
+
+We will leverage [GitHub's tagging and release feature](https://help.github.com/en/github/administering-a-repository/releasing-projects-on-github) in conjunction with [Semantic Versioning]((https://semver.org/)) to document and track changes to the preregistration document.
+
+GitHub releases mark software iterations that you can release and deploy[^4].
+Within the context of adaptive preregistration, we can use releases to mark interim versions of the preregistration, as well as to take a snapshot of the entire repository at that moment in time.
+
+Note that anything up until the complete preregistration should be additionally tagged as a 'pre-release' to communicate that the initial preregistration is not yet complete.
+
+| Preregistration Change or Action                                                                                                                                                                                                                                                                                       | GitHub Mechanism                                                                                                                                                                                                                                                        | Semantic Versioning                                                      |
+|-----------------------------|-------------------------|-------------------|
+| **Changes preceding initial completion of preregistration**                                                                                                                                                                                                                                                            | Tag and Mark as 'pre-release'                                                                                                                                                                                                                                           | `Minor` increment.
+E.g.
+`0.2.0` The `Major` value should not exceed `0`.
+|
+| **Incremental Change**, e.g.
+respond to new question.
+| Tag and Release                                                                                                                                                                                                                                                         | `Minor` increment.
+E.g.
+`0.2.0` The `Major` value should not exceed `0`.
+|
+| **Initial Preregistration complete**.
+Note that 'complete' needs to be decided.
+Is complete after all preliminary and investigatory analyses have been done and the final Preregistration is complete, or is it the first iteration, where no more items can be filled out until some initial analyses are undertaken?
+| Tag and Release                                                                                                                                                                                                                                                         | `Major` increment to `1.0.0`                                             |
+| **Major change to a complete Preregistration** -- no looking or analysing at data                                                                                                                                                                                                                                      | Tag and Release                                                                                                                                                                                                                                                         | `Major` increment `>1.0.0`                                               |
+| **Linking implemented work to the Preregistration Item**                                                                                                                                                                                                                                                               | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
+Tag and Release that preregistration commit.
+| `Minor` increment.
+|
+| **Data-dependent changes with registered flexibility** Adapting the Preregistration on the basis of results from preliminary or investigatory analyses                                                                                                                                                                 | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
+Respond to dependent incomplete Preregistration Items or Update existing / already completed Preregistration Items + Commit.
+Tag and Release | `Major` increment.
+|
+| **Minor mistakes**                                                                                                                                                                                                                                                                                                     | Changes including small typos, or minor elaborations on existing specified plans                                                                                                                                                                                        | `Patch` increment.
+E.g.
+`0.1.2`.
+|
+
+: Types of changes to the preregistration and corresponding actions and versioning adjustments {#tbl-preregistration-versioning}
+
+### Using the [`fledge::`](https://fledge.cynkra.com) R package for semantic versioning and to document preregistration changes
+
+# References
+
+::: {#refs}
+:::
+
+[^1]: [https://docs.github.com/en/free-pro-team\@latest/github/getting-started-with-github/github-glossary#commit-id](https://docs.github.com/en/free-pro-team@latest/github/getting-started-with-github/github-glossary#commit-id){.uri}
+
+[^2]: The semantic version is a 3-part numeric identifier separated by points, e.g.: `0.0.1`, wherein the first number sequences a major version of the preregistration, the second number sequences a minor change, and the third number sequences small changes or patches, yielding: `MAJOR.MINOR.PATCH`.
+    See <https://semver.org/> for further information.
+    Further detail on how to use semantic versioning with adaptive preregistration is summarised in @tbl-preregistration-versioning.
+
+[^3]: <https://guides.github.com/features/issues/>
+
+[^4]: Instructions for creating releases are here: <https://help.github.com/en/github/administering-a-repository/managing-releases-in-a-repository>

### 2023-09-27 preregistration_modelling.qmd
Add notes on adaptive preregistration

---
title: Adaptive Preregistration for Modelling
sidebar: false
---

How does adaptive preregistration differ from existing preregistration?

-   interim preregistrations for the iterative cycle of model development
-   registrered flexibility for when planned analyses need to change based on the outcomes of previous decisions
-   moving back and forth between different preregistration phases

Communicating deviations, additional analyses, outcomes of registered flexibility

-   fledge, News.md

Interim Preregistrations:

-   GitHub releases, semantic versioning, fledge

### 2023-09-27 _supp-mat/AdaptivePreregistration.Rmd
Formatting edits, remove unnecessary backslash on some symbols

--- 2020-10-13 supp-mat/AdaptivePreregistration.Rmd
@@
 As such, the process of preregistration is completely distinct from, and precedes the implementation of the analysis plan laid out in the preregistration.
-In fact, preregistration has been defined as “the action of confirming an unalterable version of one’s research plan prior to collecting data” (Nosek et al.
+In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" (Nosek et al.
 2018) or prior to analysing the data [@Mertens2019].
@@
 Existing pre-registration formats hinder checking, because the prevailing format of a single static text-based document is designed for authors to very quickly input the required information, rather than to facilitate a comparison of the reported analysis and results against the planned analysis within the preregistration.
-In fact, Pu et al.
-[-@Pu2019] go so far as to say that this format “verge[s] on being write-only media” and are therefore completely ill-suited to the checking task.
+In fact, Pu et al. [-@Pu2019] go so far as to say that this format "verge[s] on being write-only media" and are therefore completely ill-suited to the checking task.
 The task of model checking is further complicated by the non-linear and iterative nature of model development, and by our proposed adaptive preregistration methodology because in most modelling studies there will likely be interim preregistrations, such that there are multiple versions of the preregistration to check against the completed analysis.
@@
 
-In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template format, we take an ‘expanded view of preregistration’, that fits with the concept of ‘adaptive preregistration’ proposed by Srivastava [-@Srivastava2018].
+In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template format, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by Srivastava [-@Srivastava2018].
 This view of preregistration breaks with the current model of preregistration, wherein the author writes a single deterministic preregistration containing a rule for every decision, and sequentially shifts from ideation, preregistration to execution of the plan.
 Two features define adaptive preregistration.
-It contains ‘plans to deploy flexible strategies’ [@Srivastava2018] wherein the author can supply a heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses.
+It contains 'plans to deploy flexible strategies' [@Srivastava2018] wherein the author can supply a heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses.
 Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed [@Srivastava2018].
@@
 We propose leveraging both the version-control and the project management features of GitHub ([www.github.com](www.github.com)) as a tool for implementing adaptive preregistration.
-When a document is version-controlled using git and GitHub, each version of a document is time-stamped, and is assigned its own unique identifier, or commit-hash ^[https://docs.github.com/en/free-pro-team@latest/github/getting-started-with-github/github-glossary#commit-id].
-Because GitHub documents are ‘time-stamped’ the genesis of the preregistration from one version to the next is apparent.
+When a document is version-controlled using git and GitHub, each version of a document is time-stamped, and is assigned its own unique identifier, or commit-hash [^1].
+Because GitHub documents are 'time-stamped' the genesis of the preregistration from one version to the next is apparent.
 Moreover, using the diff view of a document on GitHub, exactly how and where a document has changed between versions is clear - additions are coloured green, deletions are coloured red, and changes within a line are highlighted.
@@
 
-GitHub is used to track amendments to the preregistration — each time an additional template item is answered, or an alteration is made to the preregistration, the change should be committed to the repository.
+GitHub is used to track amendments to the preregistration --- each time an additional template item is answered, or an alteration is made to the preregistration, the change should be committed to the repository.
 *Semantic versioning* is used to document each version of the preregistration.
-Semantic versioning^[The semantic version is a 3-part numeric identifier separated by points, e.g.: `0.0.1`, wherein the first number sequences a major version of the preregistration, the second number sequences a minor change, and the third number sequences small changes or patches, yielding: `MAJOR.MINOR.PATCH`.
-See https://semver.org/ for further information.
-We provide more details on how to use semantic versioning with adaptive preregistration below.] is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it [@Kitzes2018vn].
-We have created our own semantic versioning scheme to be used to specifically document the incremental version of the preregistration, in tandem with GitHub’s tag and release feature, each interim preregistration is tagged and released with its semantic version number.
+Semantic versioning[^2] is an open-source software engineering practice used to document and and communicate the development of software, in a way that allows others to depend on it [@Kitzes2018vn].
+We have created our own semantic versioning scheme to be used to specifically document the incremental version of the preregistration, in tandem with GitHub's tag and release feature, each interim preregistration is tagged and released with its semantic version number.
 All questions that can be answered should be completed before any data analysis or modelling proceeds.
@@
 
-### …Shifting from planning to execution…
+### ...Shifting from planning to execution...
 
@@
 
-Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding GitHub issue^[https://guides.github.com/features/issues/] created (Figure 3, step 3, ‘Document’).
-Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an issue’s thread.
+Before shifting from planning to the execution of the interim preregistration, each complete response in the interim preregistration should have a corresponding GitHub issue[^3] created (Figure 3, step 3, 'Document').
+Relevant discussion between analysts about how to proceed with the intended analysis or interpreting preliminary results are tracked within an issue's thread.
 Each issue is automatically assigned a unique number by GitHub.
-Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the issue’s thread (Fig 3, step 4, ‘Do’).
+Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the issue's thread (Fig 3, step 4, 'Do').
 The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new release.
 
-### … And back again.
+### ... And back again.
 
@@
 -   View the document in GitHub [here](https://github.com/egouldo/VEFMAP_VEG_Stage6/blob/master/analysis/preregistration_template.Rmd)
-- Click on The 'edit document' button:
-![Screen Shot 2020-05-10 at 6 29 33 pm](https://user-images.githubusercontent.com/8400682/81494459-791b7080-92ec-11ea-979b-900317b4a9b6.png)
+-   Click on The 'edit document' button: ![Screen Shot 2020-05-10 at 6 29 33 pm](https://user-images.githubusercontent.com/8400682/81494459-791b7080-92ec-11ea-979b-900317b4a9b6.png)
 -   You will be able to edit the document directly in GitHub, Edit away!
@@
 Click on the tiny box with the ellipsis, and a pop-up dialogue box will appear with 4 options: `Copy lines`, `Copy permalink`, `View git blame`, `References in new issue`.
-Select the last option `Reference in new issue.`
-![Screen Shot 2020-05-13 at 1 06 58 pm](https://user-images.githubusercontent.com/8400682/81767344-80749100-951b-11ea-8b76-1e829528973a.png)
+Select the last option `Reference in new issue.` ![Screen Shot 2020-05-13 at 1 06 58 pm](https://user-images.githubusercontent.com/8400682/81767344-80749100-951b-11ea-8b76-1e829528973a.png)
 3.
 This will open up a new window with a new GitHub Issue for you to complete.
-Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted:
-![Screen Shot 2020-05-13 at 1 07 30 pm](https://user-images.githubusercontent.com/8400682/81767518-ef51ea00-951b-11ea-9398-18fffdcbbeae.png)
-Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this:
-![Screen Shot 2020-05-13 at 1 17 21 pm](https://user-images.githubusercontent.com/8400682/81767585-1c9e9800-951c-11ea-9d2c-46d2faf6368a.png)
+Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted: ![Screen Shot 2020-05-13 at 1 07 30 pm](https://user-images.githubusercontent.com/8400682/81767518-ef51ea00-951b-11ea-9398-18fffdcbbeae.png) Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this: ![Screen Shot 2020-05-13 at 1 17 21 pm](https://user-images.githubusercontent.com/8400682/81767585-1c9e9800-951c-11ea-9d2c-46d2faf6368a.png)
 
@@
 
-Sometimes assumptions are violated, or there are other unexpected outcomes of preliminary analyses, and you will need to change plans. How do we deal with this?
+Sometimes assumptions are violated, or there are other unexpected outcomes of preliminary analyses, and you will need to change plans.
+How do we deal with this?
 
@@
 
-GitHub releases mark software iterations that you can release and deploy^[Instructions for creating releases are here: [https://help.github.com/en/github/administering-a-repository/managing-releases-in-a-repository](https://help.github.com/en/github/administering-a-repository/managing-releases-in-a-repository)].
+GitHub releases mark software iterations that you can release and deploy[^4].
 Within the context of adaptive preregistration, we can use releases to mark interim versions of the preregistration, as well as to take a snapshot of the entire repository at that moment in time.
@@
 
-Preregistration Change or Action | GitHub Mechanism | Semantic Versioning|
-------|-------|--------|
-**Changes preceding initial completion of preregistration** | Tag and Mark as 'pre-release' | `Minor` increment. E.g. `0.2.0` The `Major` value should not exceed `0`. |
-**Incremental Change**, e.g. respond to new question. | Tag and Release | `Minor` increment. E.g. `0.2.0` The `Major` value should not exceed `0`. |
-**Initial Preregistration complete**.
+| Preregistration Change or Action                                                                                                                                                                                                                                                                                       | GitHub Mechanism                                                                                                                                                                                                                                                        | Semantic Versioning                                                      |
+|-------------------------------|--------------------------|---------------|
+| **Changes preceding initial completion of preregistration**                                                                                                                                                                                                                                                            | Tag and Mark as 'pre-release'                                                                                                                                                                                                                                           | `Minor` increment.
+E.g.
+`0.2.0` The `Major` value should not exceed `0`.
+|
+| **Incremental Change**, e.g.
+respond to new question.
+| Tag and Release                                                                                                                                                                                                                                                         | `Minor` increment.
+E.g.
+`0.2.0` The `Major` value should not exceed `0`.
+|
+| **Initial Preregistration complete**.
 Note that 'complete' needs to be decided.
@@
 | Tag and Release                                                                                                                                                                                                                                                         | `Major` increment to `1.0.0`                                             |
-**Major change to a complete Preregistration** -- no looking or analysing at data | Tag and Release | `Major` increment `>1.0.0` |
-**Linking implemented work to the Preregistration Item** | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
+| **Major change to a complete Preregistration** -- no looking or analysing at data                                                                                                                                                                                                                                      | Tag and Release                                                                                                                                                                                                                                                         | `Major` increment `>1.0.0`                                               |
+| **Linking implemented work to the Preregistration Item**                                                                                                                                                                                                                                                               | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
 Tag and Release that preregistration commit.
@@
 |
-**When plans change** Adapting the Preregistration on the basis of results from preliminary or investigatory analyses |  Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
+| **When plans change** Adapting the Preregistration on the basis of results from preliminary or investigatory analyses                                                                                                                                                                                                  | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
 Respond to dependent incomplete Preregistration Items or Update existing / already completed Preregistration Items + Commit.
 Tag and Release | `Major` increment.
-**Minor mistakes** | Changes including small typos, or minor elaborations on existing specified plans | `Patch` increment. E.g. `0.1.2`. |
+|
+| **Minor mistakes**                                                                                                                                                                                                                                                                                                     | Changes including small typos, or minor elaborations on existing specified plans                                                                                                                                                                                        | `Patch` increment.
+E.g.
+`0.1.2`.
+|
+
+[^1]: [https://docs.github.com/en/free-pro-team\@latest/github/getting-started-with-github/github-glossary#commit-id](https://docs.github.com/en/free-pro-team@latest/github/getting-started-with-github/github-glossary#commit-id){.uri}
+
+[^2]: The semantic version is a 3-part numeric identifier separated by points, e.g.: `0.0.1`, wherein the first number sequences a major version of the preregistration, the second number sequences a minor change, and the third number sequences small changes or patches, yielding: `MAJOR.MINOR.PATCH`.
+    See <https://semver.org/> for further information.
+    We provide more details on how to use semantic versioning with adaptive preregistration below.
+
+[^3]: <https://guides.github.com/features/issues/>
+
+[^4]: Instructions for creating releases are here: <https://help.github.com/en/github/administering-a-repository/managing-releases-in-a-repository>

### 2023-10-04 RDF_QRP.qmd
Turn off sidebar
Write page describing RDF's, QRP's and the role of preregistration in mitigating them.

@@
 title: From Researcher Degrees of Freedom to Questionable Research Practices
+sidebar: false
+bibliography: references.bib
 ---
+
+## Researcher Degrees of Freedom
+
+Throughout the course of model development and analysis researchers make numerous decisions.
+These alternative decisions are termed 'researcher degrees of freedom' [@Wicherts2016; @Simmons:2011iw].
+These decisions are usually subjective and arbitrary yet methodologically and substantively acceptable [@Wicherts2016], however may collectively significantly influence the final 'analysis strategy' [*sensu* @Hoffmann2021] or 'specification' [*sensu* @Simonsohn2020], and consequently the research findings and conclusions drawn from that analysis [@Desbureaux2021].
+
+Alternative decisions about model assumptions and implementation can lead researchers down very different analytical paths that result in 'parallel' or 'adjacent' worlds with different interpretations, and potentially conflicting conclusions originating from a single underlying dataset [@Young2021; @Liu2020].
+The set of all (plausible or reasonable) possible decisions that a researcher could make over the entire analysis from beginning to end has been termed collectively as the 'garden of forking paths' [@Gelman2013] or the 'multiverse', which consists of multiple universes of plausible analysis strategies [@Steegen2016].
+
+Researcher degrees of freedom are not inherently problematic, but rather are a product of the uncertainty that besets model development and analysis, whether due to incomplete theoretical understanding of underlying ecological processes, or because modelling is an act of simplification and abstraction; there is always some degree of choice about which ecological processes and phenomena should be included in the model, how they should be represented [@Babel2019; @Getz:2017ic].
+Further uncertainty and multiplicity of analysis strategies stems from uncertainty in data preprocessing, model, parameter, and method uncertainty [@Hoffmann2021].
+
+## Questionable Research Practices
+
+When researchers are faced with ambiguous decisions where there is no single decision or strategy acceptable according to scientific standards and consensus, we are likely to select the decision that results in the most 'favourable' finding with "convincing self-justification" [@Simmons:2011iw; @Hoffmann2021].
+Researchers routinely execute and explore alternative analysis pathways in service of identifying the most favourable (i.e.
+publishable) result, selectively reporting only one or several preferred results [@Liu2020; @Simmons:2011iw].
+
+Importantly, as Gelman and Loken note [@Gelman2013] this behaviour is not a consciously nefarious 'fishing expedition', rather, because researchers are motivated actors in a 'publish or perish' culture who are susceptible to cognitive biases and are willing to change their modelling assumptions when the results conflict with desired outcomes -- sincerely and self-convincingly believing that the more favourable model is superior [@Young2021].
+
+Thus researcher degrees of freedom precipitate questionable research practices when they remain both undisclosed and exploited opportunistically [@Bakker2018].
+Undisclosed and opportunistic use of researcher degrees of freedom artificially increases the probability of false positive results, inflates effect size estimates and ultimately reduces the replicability of published findings while encouraging overconfidence in the precision of results and exascerbating the risk of accepting and propagating false facts [@Wicherts2016; @Hoffmann2021; @Roettger2019; @Nissen2016].
+Emerging evidence suggests that rates of QRPs in ecology and evolutionary biology are similar to rates observed in other disciplines beset by reproducibility crises [@Fraser:2018cl].
+
+## Preregistration mitigates against researcher degrees of freedom
+
+Preregistration is an open-science practice adopted in scientific disciplines that have begun to confront the 'reproducibility crisis', and aims to improve research transparency and mitigate questionable research practices within a study by distinguishing between what is a genuine a priori planned analysis, and what is not [@Parker2019].
+Preregistration requires the methods and analysis plan of a future study are registered in a secure and publicly accessible platform prior to any data collection, and after submission it can no longer be altered.
+
+Despite enthusiastic uptake of preregistration in some scientific disciplines, there has been little uptake of pre-registration in ecology and conservation, and many modellers and researchers engaged in non-hypothesis testing research both within and outside of ecology have eschewed preregistration on the grounds that preregistration and existing templates are irrelevant because they are focused on null-hypothesis significance testing [e.g.
+@McDermott2021; @Prosperi2019].
+
+Instead of throwing out the baby with the bathwater, this guide is premised on the idea that preregistration can and should be used in model-based research in ecology and conservation, but its particular form should reflect the norms and practice of the research context in which it is applied, in order to adequately restrict researcher degrees of freedom that may uniquely emerge within a domain-specific methodology.
+The work presented in this guide aims to translate preregistration into model-based research within ecology and conservation by proposing an expanded view of preregistration we term [*Adaptive Preregistration*](adaptive_preregistration_background.qmd).
+We also present a template for preregistering model-based research in ecology and conservation.

### 2023-10-04 adaptive_preregistration_background.qmd
Complete draft adaptive preregistration background page

--- 2023-09-27 adaptive_preregistration_background.qmd
@@
 ---
-title: A primer on Adaptive Preregistration
+title: A primer on *Adaptive Preregistration*
+date: today
+link-citations: true
+bibliography: data/references.bib
 ---
+
+## Background & Motivation
+
+```{r}
+#| label: fig-traditional-prereg
+#| out-width: 700px
+#| out-height: 100x
+#| echo: false
+#| column: body
+#| fig-cap: "Conventional preregistration process for registered reports, as illustrated by [@Pu2019]."
+magick::image_read(path = "_supp-mat/Pu_et_al_fig_prereg.png")
+```
+
+The iterative nature of model development is a significant barrier to adopting preregistration for model-based research.
+Translating the preregistration process to a modelling research context is challenging because the iterative nature of model development conflicts with the inner logic of existing preregistration templates that presume a linear research workflow.
+The process of preregistration as it is currently implemented is tailored towards the hypothetico-deductive model of the scientific method.
+As such, the process of preregistration is completely distinct from, and precedes the implementation of the analysis plan laid out in the preregistration -- analytic decisions are completely independent from the data.
+
+In fact, preregistration has been defined as "the action of confirming an unalterable version of one's research plan prior to collecting data" [@Pu2019] or prior to analysing the data [@Mertens2019].
+In this process, researcher sequentially shifts from ideation, to study/analysis design and preregistration, to collecting and analysing the data, writing the report or manuscript, to documenting the research and publishing (@fig-traditional-prereg).
+The iterative nature of model development, however, inherently precludes this model of preregistration from being applied to model-based research in ecology and conservation.
+
+## Iterative Cycle of Model Development and Preregistration
+
+Current preregistration practices based on NHST-focussed accounts of questionable research parctices prescribe the principle of 'decision-independence', where analysis decisions are made independently of and *a priori* to analysing the data [@Srivastava2018; @Gelman2013].
+However, the modelling process may genuinely violate the principle of 'decision-independence' because it is non-linear, iterative and usually generates many interim versions of the model preceding publication.
+During modelling a researcher will almost always engage in state-dependent decision-making --- that is, decisions about whether to alter the model in a subsequent round of model development depend on the results of model testing, evaluation and analysis at the end of the previous iteration of model development.
+These practices aren't necessarily questionable in and of themselves, but do have the *potential* to become questionable if they are exploited in service of artificially inflating the accuracy or precision of the model, its predictions or evaluation tests to the effect that the model is perceived to be more credible than it would be if that practice did not occur -- essentially data-dependent decisions that lead to consumers of the model placing unsubstantiated belief in the reliability, validity and utility of the model and its outputs are questionable [@gould] and can be mitigated through the practice of preregistration.
+
+Very often in modelling, there are decisions that are necessarily dependent on the outcome of previous analytic decisions in the modelling workflow [@Liu2020].
+Firstly, preliminary or investigatory analyses might need to be conducted before being able to specify future decision-steps in the analysis plan.
+For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation.
+Secondly, data-dependent decisions from model checking may justifiably result in changes to either the analysis or to interpretations of the mode [@Srivastava2018].
+For instance, inspection of the residuals and other model checking tests may force the modeller to return to earlier decision-points and change the planned analysis or model.
+Other times, the planned analysis and specified model may simply not converge, or the model is saturated and runs out of variation to apportion such that the model and or the model fitting algorithm must be respecified and re-implemented.
+Other times, the modelling process itself may generate knowledge and learning about the system, and the model structure changes throughout the process of model development.
+
+Finally, sometimes there are some decisions that are too difficult to anticipate, or simply cannot be made in advance [@Srivastava2018].
+This is particularly true for decisions occurring at the later phases of the modelling construction and development process (Figure 2, TBC), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses.
+Some decisions might not be able to be specified on the first attempt at writing the preregistration until the model is fully or at least close to fully specified, for example, specifying precisely what and how sensitivity analyses or uncertainty analyses will be conducted.
+
+As illustrated above, the adaptive nature of analysis during the model development process contravenes the central principle of preregistration -- decision-independence and is fundamentally incompatible with the sequential, linear process of traditional preregistration where the analysis plan is fully specified prior to observing and analysing the data.
+
+## Adaptive Preregistration
+
+In order to address the disjuncture between the iterative nature of model development, and the single-use, deterministic preregistration template format, we take an 'expanded view of preregistration', that fits with the concept of 'adaptive preregistration' proposed by Srivastava [-@Srivastava2018].
+The modeller proceeds through the model development process, they will switch they will shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, implementing implementing parts of the analysis plan that were specified in the previous preregistration step, and letting the outcomes in that analysis step inform the specification of the next analysis ( @fig-adaptive-preregistration).
+
+This view of preregistration breaks with the current model of preregistration, wherein the author writes a single deterministic preregistration containing a rule for every decision, and sequentially shifts from ideation, preregistration to execution of the plan.
+Two features define adaptive preregistration:
+
+1.
+It contains 'plans to deploy flexible strategies' [@Srivastava2018] wherein the author can supply heuristic consisting of multiple different analysis or modelling strategies whose execution depends on the outcome of previous decision-points or analyses, for example, decision-trees could be preregistered prior to analysing the data [@Baldwin2022a].
+2.  The preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed [@Srivastava2018].
+
+### Preregistering Flexibility
+
+When preliminary analyses, model checking or other data-dependent decisions need to be made, the modeller can specify a 'flexible strategy' within their preregistration by:
+
+-   stating what quantity or outcome needs to be known to move forward with the modelling and analysis,
+-   explaining what test or analysis will be performed to obtain this quantity or outcome, and what parts of the data will be used in this analysis, and
+-   describing how the results will be interpreted, listing each potential decision and its trigger, where possible.
+
+```{r}
+#| label: fig-registering-flexibility
+#| echo: false
+#| fig-cap: "A general heuristic in the form of a decision-tree for registering flexibile analyses."
+magick::image_read_pdf(path = "figures/registering_flexibility.pdf", density = 500)
+```
+
+### Interim Preregistrations
+
+The modeller follows an iterative process of preregistration.
+As they proceed through the model development process, they can shift from ideation, preregistration and execution of the analysis plan, to preregistration again.
+Depending on observed outcomes of the pre-specified decision-trees or registered flexible heuristics, interim preregistrations are created at multiple points in the model development process, each time there is an addition or amendment to the analysis plan.
+
+```{r}
+#| label: fig-interim-preregistrations
+#| echo: false
+#| fig-cap: Adaptive preregistration can cycle through the specification, execution and ideation phases multiple times in adaptive preregistration, with interim versions of the preregistration being generated at each preregistration point.
+magick::image_read_pdf(path = "figures/interim_preregistrations.pdf", density = 300)
+```
+
+## Documenting adaptive preregistration transparently
+
+### The *Checking Problem*
+
+Following publication of the final study, authors, reviewers, editors and readers may all engage in the process of checking a manuscript or published paper against its preregistration to verify that the study and analyses were conducted as specified in the preregistration.
+The process of preregistration is already cumbersome, because the dominant format of the preregistration is a single, static text-based document designed with the study authors in mind to very quickly insert the required analysis plan information, rather than to facilitate a comparison of the reported analysis and results against the planned analysis in the preregistration.
+In their current form, they are almost exclusively exist as "write-only media" [@Pu2019].
+The task of model checking is further complicated by the non-linear and iterative nature of model development, and by our proposed adaptive preregistration methodology, because in most modelling studies there will likely be interim preregistrations, such that there are multiple versions of the preregistration to check against the completed analysis.
+
+## Implementing *Adaptive Preregistration* with git and GitHub
+
+Given the iterative nature of adaptive preregistration then, the current convention of the single-use, deterministic and static text-based document format for preregistration is inadequate.
+Interim versions of a preregistration must be time-stamped and ideally version-controlled, such that it is clear not just that two or more versions differ from each other, and in what order they were created, but so that it is explicitly clear in how and why they differ.
+To facilitate preregistration checking of flexible modelling and analysis strategies, we must link the results of both final and preliminary analyses back to the specified strategy: the link between a particular triggered analysis decision, and the analysis outcome that triggered the decision must be explicit.
+Additionally, if the results of some preliminary analyses cause revisions to the model itself or a different modelling procedure or approach to be selected than was originally or previously planned, then the trigger for this decision must be explicitly linked to the next version of the preregistration.
+
+We propose leveraging the features of [git](https://git-scm.com) and [GitHub](www.github.com) to facilitate implementation both adaptive preregistration in a transparently documented manner (@fig-adaptive-preregistration), and provide a [step-by-step guide](adaptive_preregistration_implementation.qmd).
+
+```{r}
+#| label: fig-adaptive-preregistration
+#| column: screen
+#| fig-cap: "Adaptive Preregistration for model-based research using GitHub."
+#| echo: false
+magick::image_read(path = "figures/LivingPreregistration.png", density = 600)
+```

### 2023-10-04 index.qmd
Add links to home page outline

@@
 2.  Adaptive Preregistration
-    1.  *Background*: What is it and what are its key features
-    2.  *A practical guide*: How to implement adaptive preregistration with git and github.
+    1.  *Background*: What is adaptive preregistration, why do we need it for registering modeling analyses, and what are its defining features?
+    2.  *A practical guide*: How to implement adaptive preregistration using [git](https://git-scm.com) and [GitHub](www.github.com).
 3.

### 2023-10-04 preregistration_modelling.qmd
Begin adding modelling preregistration template

@@
 ---
-title: Adaptive Preregistration for Modelling
+title: "Adaptive Preregistration for Modelling"
+subtitle: "A template suitable for use in Ecology and Conservation with accompanying rationale and explanation."
+description: "Note: Black boxes provide an explanation or justification of a particular section or preregistration item. Text contained in 'preregistration item' boxes requires a response."
 sidebar: false
+bibliography: data/references.bib
+toc-location: left
+toc-expand: true
+number-sections: true
 ---
 
-How does adaptive preregistration differ from existing preregistration?
+# 1. Problem Formulation
 
--   interim preregistrations for the iterative cycle of model development
--   registrered flexibility for when planned analyses need to change based on the outcomes of previous decisions
--   moving back and forth between different preregistration phases
+::: callout-note
+## Rationale & Explanation
 
-Communicating deviations, additional analyses, outcomes of registered flexibility
+This section specifies the decision-making context in which the model will be used or the intended scope and context of conclusions.
+Important components include the decision maker and stakeholders (including experts) and their view on: i) the nature of the problem or decision addressed and how the scope of the modelling tool fits within the (broader) context (i.e.
+model purpose; ii) the spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they foresee unacceptable outcomes that need to be represented in the model (i.e.
+as constraints), and; vi) what future scenarios does the model need to account for (noting this may be revised later).
+It should also provide a summary of the domain of applicability of the model, and reasonable extrapolation limits [@Grimm:2014es].
+:::
 
--   fledge, News.md
+## 1.1 Model Purpose
 
-Interim Preregistrations:
+::: callout-note
+## Rationale & Explanation
 
--   GitHub releases, semantic versioning, fledge
+Defining the purpose of the model is critical because the model purpose influences choices at later stages of model development [@Jakeman:2006ii].
+Common model purposes in ecology include: gaining a better qualitative understanding of the target system, synthesising and reviewing knowledge, and providing guidance for management and decision-making [@Jakeman:2006ii].
+Note that modelling objectives are distinct from the analytical objectives of the model.
+:::
+
+### 1.1.1 Define model purpose and problem context
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Provide a clear statement of the modelling objectives and problem that the model seeks to illuminate.
+What is the purpose of the model/s?
+Ensure that you specify any focal taxa and study objectives, as well as any clients for whom the model is developed.
+Briefly outline the ecological problem and the decision-problem, including the decision-trigger and any regulatory frameworks relevant to the problem.
+:::
+
+## 1.2 Specify modelling context
+
+::: callout-note
+## Rationale & Explanation
+
+The scope of the model, including temporal and spatial resolutions are defined here [@Mahmoud2009], and any limitations on model development analysis and flexibility should be outlined in this section [@Jakeman:2006ii].
+Note that the modelling context is different from the problem context which are described in 1.3.1.
+:::
+
+### 1.2.1 Identify model users / stakeholders
+
+::: callout-caution
+## Preregistration Item
+
+Identify interest groups in the modelling context.
+This includes clients, and end-users of the model.
+Who is the model for, who is involved in formulating the model?
+Who needs buy in?
+Describe the decision-making context in which the model will be used.
+:::
+
+### 1.2.2 Model Scope and Scale
+
+*Determine the temporal and spatial scope of the model.
+Where is the boundary of the modelled system?
+Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al.
+2006)* **HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me.
+Also, is this question more to do with describing the scope and limitations of the data rather then the model?**\*.
+The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?
+
+Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
+(Should this last step go here or within define conceptual framework?. ***HF comment. I think it belongs with the conceptual framework***).
+
+What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)?
+This is relevant to model transferability...
+(similar to a COG statement..)\* A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation.
+I'm just thinking about how I might fill this out for an imagined study.
+The literal boundaries of the study are usually straightforward, but the theoretic boundaries are a different and also big thing to try to conceptualise in one section.
+You could have a section for the spatial/temporal boundaries of the study, and another section on how far the results can be extrapolated based on the study design.
+This makes sense to me in how I design experiments, if I have too many confounding variables and not enough spatial and environmental replication then I'm less likely to make any broad claims.
+However, if I have controlled or experimentally manipulated factors then I feel safer making bigger generalisations about results.
+
+### 1.2.3 Logistical Constraints
+
+*Time-frame - When must the model by completed, e.g. to help make a decision?*
+
+*What effort and resources are available for both developing the model and operating the model?* ***HF comment; as written this might be slightly confronting because you could interpret this to include writing explictly about budget constraints which might be confidential.
+Maybe this could be alleviated with with an example?***
+
+*What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?*
+
+## 1.3.1 Explain analytical objectives
+
+*How will the model be analysed, what analytical questions will the model be used to answer?
+Examples from ecological decision-making include: to compare the performance of alternative management actions under budget constraint [@Fraser:2017jf] to search for robust decisions under uncertainty [@McDonald-Madden2008], to choose the conservation policy that minimises uncertainty \[insert ref\].
+See other examples in [@Moallemi2019].*
+
+### 1.3.2 - Identify how the model will tell you about your analytic objectives
+
+*The outcomes should speak directly to the analytical objectives identified in 1.3.1.
+Outcomes could be qualitative or quantitative.
+For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.* To be honest, I don't know how I might apply this to my work or any future works.
+Maybe I do not have enough knowledge of this type of research to apply it to my basic ecology research.
+But this isn't to say it's not important, I just wonder if there's a way to make this accessible to non-decision or modelling peeps.
+Like, how would I apply it to the woodland condition analysis?
+
+## 1.4 Define candidate decisions and method of identifying relevant management actions
+
+*Candidate decisions should be investigated and are specified a priori.
+Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019].
+Describe the method used to identify relevant management actions.*
+
+## 1.5 Specify scenarios
+
+*Describe what processes you will use to elicit and identify relevant scenarios, e.g.
+literature review, structured workshops with stakeholders or decision-makers.
+Specify scenarios under which decisions are investigated.
+Scenarios should be set a priori [i.e.
+before the model is built, @Moallemi2019] and may be stakeholder-defined or driven by the judgement of the modeller or other experts [@Mahmoud2009].*
+
+# 2.0 Define Conceptual Model
+
+*Conceptual models underpin the formal or quantitative model*
+
+*(Cartwright et al.
+2016).
+The conceptual model describes the biological mechanisms relevant to the ecological problem and should capture basic premises about how the target system works, including any prior knowledge and assumptions about system processes.
+Conceptual models may be representeed in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables [@Jakeman:2006ii].*
+
+## 2.1 Choose elicitation and representation method
+
+*Describe what method you will use to elicit or identify the conceptual model.
+Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks (Moon et al.
+2019).
+(Libby, to provide link to any structured expert elicitation methods?).While it is difficult to decide and justify which method is most appropriate, however Moon et al.
+(2019) provide guidance addressing this methodological question.
+Finally, how do you intend on representing the final conceptual model?
+This will likely depend on the method chosen to elicit the conceptual model.*
+
+## 2.2.1 Explain Critical Conceptual Design Decisions
+
+*This step should list and explain the critical conceptual design decisions, including: spatial and temporal scales, selection of entities and processes, representation of stochasticity and heterogeneity, consideration of local versus global interactions, environmental drivers, etc.
+[@Grimm:2014es].
+The influence of particular theories, concepts, or importantly, earlier models, should be explained and justified against alternative conceptual design decisions that might lead to alternative model structures [@Grimm:2014es].*
+
+## 2.2.2 Specify key assumptions and uncertainties underlying the model's design.
+
+-   Describe how uncertainty and variation will be represented in this model.
+This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].\*
+
+## 2.3 Identify predictor and response variables
+
+*The identification and definition of primary model input variables should be driven by scenario definitions, and by the scope of the model described in the problem formulation phase [@Mahmoud2009] -- Identify and define system variables structures:*
+
+-   *What variables would support taking this action or making this decision?*
+-   *What additional variables may interact with this system (things we can't control, but can hopefully measure)?*
+-   *What variables have not been measured, but may interact with the system (often occurs in field or observational studies)?*
+-   *What variables are indice or surrogate measures of variables that we cannot or have not measured?*
+-   *In what ways do we expect these variables to interact (model structures)?*
+
+## 2.4 Define prior knowledge, data specification and evaluation
+
+*Collect, process and prepare data available for parameterisation, determining model structure, and for scenario analysis.*
+
+### 2.4.1 Collate available data sources that could be used to parameterise or structure the model
+
+*Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model. Describe how the data is arranged, in terms of replicates and covariates.*
+
+**For pre-existing data: -- delete as appropriate.**
+
+-   *For each dataset, is the data open or publically available? (Y/N)*
+-   *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrctions on the use of data.*
+-   *Date of download, access, or future access:*
+-   *Describe the source of the data - what entity originally collected this data?
+(National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
+-   *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
+-   *Prior work based on this dataset - Have you published / presented any previous work based on this dataset?
+Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based on this dataset you have worked on.* HF- Love this question
+-   *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
+-   *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study.
+For example, have you read any reports or publications about this data?*
+
+**Sampling Plan (For data you will collect) -- delete as appropriate.**
+
+-   Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged.
+Describe any inclusion or exclusion rules, and the study timeline.
+-   Sample Size - Describe the sample size of your study.
+-   Sample Size Rationale - Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel.
+-   If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?
+
+### 2.4.2 Data Processing and Preparation
+
+*Describe any data preparation and processing steps, including manipulation of environmental layers, e.g. standardisation and geographic projection.*
+
+*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.*
+
+### 2.4.3 Describe any data exploration or preliminary data analyses.
+
+*In most modelling cases, it is necessary to perform preliminary analyses to understand the data and check that assumptions and requirements of the chosen modelling procedures are met.
+Data exploration prior to model fitting or development may include exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers, or the need for transformations [@Yates2018].
+Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.
+Specify what you intend to find out or hope to learn from each data exploration analysis.*
+
+### 2.4.4 Data Evaluation
+
+*Describe how you will determine how reliable the data is for the given purpose.
+Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].* \### 2.4.5 Data Evaluation: Data reliability reporting \### 2.4.5.
+Data Evaluation: data reliability reporting
+
+*Document any issues with data reliability.
+This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].*
+
+### 2.4.6. Data evaluation: data exclusion
+
+*How will you determine what data, if any, will be excluded from your analyses?
+How will outliers be handled?
+Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.*
+
+### 2.4.7 Data Evaluation: Missing Data
+
+*How will you identify and deal with incomplete or missing data?*
+
+## 2.5 Conceptual model evaluation
+
+*Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.*
+
+*How will any simplifying assumptions be critically assessed [@Augusiak:2014gz]?*
+
+*Explain whether this process will include consultation or feedback from a client, manager, or model user.*
+
+# 3.0 Formalise and Specify Model
+
+*In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.*
+
+## 3.1 Choose model class, framework and approach
+
+*Describe what class or approach of model you will use and¬†explain how the choice of model class was informed by the analytical objectives of the model.
+Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based¬† [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].*
+
+## 3.2 Choose model features and family
+
+*All modelling approaches require the selection of model features, which conform with the conceptual model and data specified in previous steps (Jakeman 2006).
+The choice of model are determined in conjunction with features are selected.
+Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006).
+However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.* HF- it's not clear to me what a model feature is and I find this a little confusing.
+I wonder if we need the text describing the 3.2 heading when it's stepped out in the subsections.
+
+### 3.2.1 Explain how you will operationalise response variable(s)
+
+*Specify how you will operationalise the response variables in the model.
+This should relate directly to the analytical and or management objectives specified during the problem formulation phase.
+Operationalisations could include:- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, - qualitative change, such as a direction of change or- the frequency, location, or probability of some event occuring.
+Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).* HF- this is great.
+Only slight concern is the deceptively difficult task that last sentence represents
+
+### 3.2.2 Choose model family
+
+*Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions.
+Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.*
+
+*Include in your rational for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*
+
+### 3.2.3 Choose model features
+
+*Specify which variables are included in the model, and the nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).*
+
+*Specify model structural features, including: - the functional form of interactions, - data structures - measures used to specify links, - any bins or discretisation of continuous variables.*
+
+## 3.3 Choose approach for identifying model structure and parameters
+
+*This refers to the process of determining the best/most efficient/parsimonious representation of the system at the appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives specified in the problem formulation phase.
+Approaches to finding model structure and parameters may be knowledge-supported, or data-driven [@Boets:2015gl].
+Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods (Barnard et al.
+2019).
+Ensemble modelling procedures might aim to derive a single model, or a multi-model average [@Yates2018].
+Refining actions to develop a model could include iteratively dropping parameters or adding them, or aggregating / disaggregating system descriptors, such as dimensionality and processes (Jakeman, 2006).
+Specify what approach and methods you will use to identify model structure and parameters.*
+
+## 3.4 Choose estimation technique and performance criteria
+
+*Before calibrating the model to the data, the performance criteria on which the calibration is judged are chosen.
+These criteria and their underlying assumptions should reflect the desired properties of the parameter estimates / structure [@Jakeman:2006ii].
+For example, modellers might seek parameter estimates that are robust to outliers, unbiased, and yield appropriate predictive performance.
+Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are appropriate to the problem at hand.
+For integrated or sub-divided models, other considerations might include choices about where to disaggregate the model for parameter estimation; e.g.
+spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii].
+Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.* HF- I think this paragraph is only relevant to performance criteria- doesn't really explore estimation technique.
+As above though, I'm not sure that this paragraph is neessary given that it's stepped out in the subheadings below
+
+### 3.4.1 Choice of performance criteria
+
+*Specify which suite of performance criteria you will use to judge the performance of the model.
+Examples include correlation scores, coefficient of determination, specificity, sensitivity, AUC, etcetera (Yates et al.
+2018).
+Relate any underlying assumptions of each criterion to the desired properties of the model, and justify the choice of performance metric in relation into whether it is sensitive to the problem at hand.
+Explain how you will identify which model features or components are significant or meaningful.*
+
+### 3.4.2 Choice of parameter estimation technique
+
+HF- I think that it might be double dipping to include structure here as well as in 3.3 *Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g.
+distributed boundary conditions).
+For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006).
+Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge.
+Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*
+
+## 3.5 Specify model assumptions and uncertainties
+
+*Specify all assumptions and key uncertainties in the formal model.
+Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpretation of the model outputs, and what implications are there on evaluating model-output to inform decisions?*
+
+## 3.6 Specify formal model
+
+*Once critical decisions have been made about the approach and method of model specification,translate the conceptual model into the quantitative model.
+For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters.
+(Should this go here or in model calibration?)*
+
+# Model Calibration, Model Fitting & Checking
+
+## Model calibration and validation scheme
+
+*Describe how you will validate and check the calibration of the model.
+The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) [@Yates2018; @Barnard2019].
+Justify your choices.*
+
+### Describe validation data
+
+*If partitioning data for cross-validation: describe the approach specifying the number of folds that will be created and the relative size of each.
+Describe how will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced.*
+
+*If using external data: describe any known differences between the training and validation datasets and the size of the validation dataset.
+Describe how will you document and/or share the external data such that any subsequent modelling can be computationally reproduced.*
+
+## Implementation verification
+
+*What Quality Assurance measures will you take to verify the model has been correctly implemented?
+Tests could include syntax checking of code, and code reviews by peers.
+Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as dictated by the model description [@Grimm:2014es].
+Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging.*
+
+## Model checking
+
+::: callout-note
+## Rationale & Explanation
+
+"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification" ), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd].
+Model checking aids in diagnosing assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd].
+Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].
+:::
+
+### Quantitative Model Checking
+
+::: callout-note
+## Preregistration Item
+
+*During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences.*
+
+-   [ ] Specify any diagnostics or tests you will use during model checking to assess a model's performance in representing the system of interest.
+-   [ ] For each test, specify the criteria that will you use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.
+:::
+
+### Qualitative Model Checking
+
+::: callout-note
+## Explanation
+
+This step is largely informal and case-specific, but requires‚ 'face validation' with model users / clients / managers who aren't involved in the development of the model to assess whether the interactions and outcomes of the model are feasible an defensible [@Grimm:2014es].
+This process is sometimes called a ‚"laugh test" or a "pub test" and in addition to checking the model's believability, it builds the client's confidence in the model [@Jakeman:2006ii].
+Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.
+:::
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Briefly explain how you will qualitatively check the model, and whether and how you will include users and clients in the process.
+:::
+
+### Assumption Violation Checks
+
+::: callout-caution
+The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019].
+
+-   Explain how you will demonstrate robustness to model assumptions and check for violations of model assumptions.
+
+-   If you cannot perform quantitative assumption checks, describe what theoretical justifications would justify a lack of violation of or robustness to model assumptions.
+
+-   If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.
+
+-   If assumption violations cannot be avoided, explain how you will explore the consequences of assumption violations on the interpretation of results (To be completed in interim iterations of the preregistration, only if there are departures from assumptions as demonstrated in the planned tests above).
+:::
+
+# 5.0 Model Validation and Evaluation
+
+HF- I think that this is covered in section 4 with descriptions of validation and checking.
+I'm not sure how this section is different.
+
+*Model Validation & Evaluation - This phase consists of a suite of analyses that collectively inform the decision and whether and when a model is suitable to meet its intended purpose (Augusiak, Van den Brink and Grimm 2014).
+Errors in design and implementation of the model and their implication on the model output are assessed.
+Ideally independent data is used against the model outputs to assess whether the model output behaviour exhibits the required accuracy for the model's intended purpose.
+The outcomes of these analyses build confidence in the model applications and increase understandingof model strengths and limitations.
+Model evaluation including model analysis and other testing should complement model checking.
+It should evaluate model checking, and consider over-fitting and extrapolation.
+The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly \[‚Ä¶\] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm).
+Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al.
+2014).*
+
+## 5.1 Model output corroboration
+
+*Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model.
+Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step \<2.4\>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al.
+2014.
+Independent predictions can instead be tested on submodels.
+Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014).
+If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.*
+
+## 5.2 Choose performance metrics and criteria
+
+*Model performance can be quantified by a range of tests, including measures of agreement between predictions and independent observations, or estimates of accuracy, bias, calibration, discrimination refinement, resolution and skill [@Araujo2019].
+Specify what performance measures you will use to evaluate the model and briefly explain how each test relates to different desired properties of a model's performance.Spatial, temporal and environmental pattern of errors and variance can change the interpretation of model predictions and conservation decisions [@Araujo2019], where relevant and possible, describe how you will characterise and report the spatial, temporal and environmental pattern of errors and variance.
+If comparing alternative models, specify what measures of model comparison or out-of-sample performance metrics will you use to find support for alternative models or else to optimise predictive ability.
+State what numerical threshold or qualities you will use for each of these metrics.*
+
+## Model analysis
+
+::: callout-note
+## Rationale & Explanation
+
+Uncertainty in models arises due to incomplete system understanding (which processes to include, or which interact), from imprecise, finite and sparese data measurements, and from uncertainty in input conditions and scenarios for model simulations or runs [@Jakeman:2006ii].
+Non-technical uncertainties can also be introduced throughout the modellign process, such as uncertainties arising from issues in problem-framing, indeterminicies, and modeller / client values [@Jakeman:2006ii].
+
+The purpose of model analysis is to prevent blind trust in the model by understanding how model outputs have emerged, and to 'challenge' the model by verifying whether the model is still believable and fit for purpose if one or more parameters are changed [@Grimm:2014es].
+
+Model analysis should increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system.
+Model analysis typically consists of sensitivity analyses preceded by uncertainty analyses [@Saltelli2019], and a suite of other simulation or other computational experiments.
+The aim of such computational experiments is to increase understanding of the model behaviour by identifying which processes and process interactions explain characteristic behaviours of the model system [@Grimm:2014es].
+Uncertainty analyses and sensitivity analyses augment one another to draw conclusions about model uncertainty.
+
+Because the results from a full suite of sensitivity analysis and uncertainty analysis can be difficult to interpret due to the number and complexity of causal relations examined [@Jakeman:2006ii], it is useful for the analyst to relate the choice of analysis to the modelling context, purpose and analytical objectives defined in the problem formulation phase, in tandem with any critical uncertainties that have emerged during model development and testing prior to this point.
+:::
+
+### Uncertainty Analyses
+
+::: callout-note
+## Explanation
+
+Uncertainty can arise from different modelling techniques, response data and predictor variables [@Araujo2019].
+Uncertainty analyses characterise the uncertainty in model outputs, and identify how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es; @Saltelli2019].
+Uncertainty analyses can include propagating known uncertainties through the model, or by investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], for example.
+It could also include characterising the output distribution, such as through empirical construction using model output data points.
+It could also include extracting summary statistics like the mean, median and variance from this distribution, and perhaps constructing confidence intervals on the mean [@Saltelli2019].
+:::
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Please describe how you will characterise model and data uncertainties, e.g.
+propagating known uncertainties through the model, investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], or empirically constructing model distributions from model output data points, and extracting summary statistics, including the mean, median, variance, and constructing confidence intervals [@Saltelli2019].
+
+-   [ ] Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase.
+    For instance ‚discrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impacts of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].
+
+-   [ ] Briefly describe how you will summarise the results of these in silico experiments with graphical, tabular, or other devices, such as summary statistics.
+
+-   [ ] If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic parameter covariance, [@Jakeman:2006ii]), specify which measure of uncertainty you will use.
+:::
+
+### Sensitivity Analyses
+
+::: callout-note
+## Explanation
+
+*Sensitivity analysis examines how uncertainty in model outputs can be apportioned to different sources of uncertainty in model input [@Saltelli2019].*
+:::
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Describe the sensitivity analysis approach you will take: deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).
+
+-   [ ] Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameters will be varied.
+
+-   [ ] State the primary objective of each sensitivity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or alternatively to identify which variables or factors contribute little to overall model outputs, and so can be 'dropped' from future iterations of the model [@Saltelli2019].
+:::
+
+### Model application or scenario analysis
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario specified in section 1.
+
+-   [ ] Describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections.
+
+-   [ ] What raw and transformed model outputs will you extract from the model simulations or projections, and how will you map, plot, or otherwise display and synthesise the results of scenario and model analyses.
+
+-   [ ] Explain how you will analyse the outputs to answer your analytical objectives.
+    For instance, describe any trade-off or robustness analyses you will undertake to help evaluate and choose between different alternatives in consultation with experts or decision-makers.
+:::
+
+### Other simulation experiments / robustness analyses
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Describe any other simulation experiments, robustness analyses or other analyses you will perform on the model, including any metrics and their criteria / thresholds for interpreting the results of the analysis.
+:::
+
+# References

### 2023-10-04 _analysis/preregistration_template.Rmd
Delete 'about the template' from previous draft of fully described preregistration template.

--- 2020-06-03 analysis/preregistration_template.Rmd
@@
 
-# About this template
-
-This template has been collated from a combination of literature reviews of the 'Good Modelling Practice' and other modelling literature, as well as a collaborative workshop at DELWP.
-Chris Jones has kindly volunteered one component of the VEFMAP project investigating vegetation responses to environmental flows as a testing ground for the use of preregistration templates in applied ecological modelling research.
-This template will be used to preregister this research, and we will use the VEFMAP project to both develop and evaluate the template 'in situ'.
-
-The template is largely structured around 6 "phases" in the process of developing, evaluating and analysing a model or models in ecological applications: `Problem Formulation`, `Define Conceptual Model`, `Formalise and Specify Model`,  `Model Validation and Evaluation`, `Model Analysis`, `Model Application`.
-These have been defined and described to some extent throughout the template.
-
-## iterative model development and a 'living preregistration'
-
-One of the major challenges in designing a template for modelling is that the process of model development is inherently iterative.
-Typically, the process of preregistration is completely distinct from and precedes the implementation of the analysis plan laid out in the preregistration.
-However, translating the preregistration process to a modelling research context is not so clear cut because some decisions are inherently dependent on the outcomes and results of previous decisions and analyses.
-For example, which sensitivity analyses will be conducted will depend on the relationships between variables found during earlier earlier stages in the modelling process.
-
-One method around this is to accept that the current model of preregistration is not going to work for iterative model development.
-Again, the more complex the model, the more likely this is to be the case.
-Consequently I propose an 'adaptive preregistration' model  that I think will work for modeling research contexts.
-
-In this model of preregistration, all questions that can be answered, should be answered before any data analysis or modelling proceeds.
-For some items, particularly during model evaluation and analysis, perhaps only responses of a general nature can be provided.
-That is OK.
-As the modelling proceeds, further detail can be provided and the next phase of analysis can be preregistered.
-Where the results of model evaluation and analysis reveal that there are problems with the model, or perhaps assumption violations, plans can be changed, and again the next phase of analysis can be preregistered.
-
-A static document for this type of process doesn't really make sense.
-I propose leveraging the code and project management tools within the GitHub platform as a way of hosting an 'adaptive' preregistration.
-Because GitHub documents are 'time-stamped', the genesis of the preregistration from one version to the next can be traced and tagged with versions.
-Since the document would exist in the repository with the code itself, any results from analyses that cause the analysis plan to change can be linked and referenced in the updated plan.
-
---------------------------------
-
 # 1. Problem Formulation
-***HF comment: I think this paragraph makes things more rather than less confusing - and I wonder if its necessary given that each point is stepped through below. I would recommend removing it***
 
 *Specifies the decision-making context in which the model will be used or the intended scope and context of conclusions.
-Important components include the decision maker and stakeholders (including experts) and their view on: i) the nature of the problem or decision addressed and how the scope of the modelling tool fits within the (broader) context (i.e.
-model purpose; ii) the spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they forsee unnacceptable outcomes that need to be represented in the model (i.e.
+Important components include the decision maker and stakeholders (including experts) and their view on: i) the nature of the problem or decision addressed and how the scope of the modeling tool fits within the (broader) context (i.e.
+model purpose; ii) the spatial and temporal scales relevant to the decision context; iii) specified desired outputs; iv) role and inclusion in model development and testing; v) whether they foresee unacceptable outcomes that need to be represented in the model (i.e.
 as constraints), and; vi) what future scenarios does the model need to account for (noting this may be revised later).
@@
 Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al.
-2006) ***HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me.
-Also, is this question more to do with describing the scope and limitations of the data rather then the model?***.
+2006)* **HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me.
+Also, is this question more to do with describing the scope and limitations of the data rather then the model?**\*.
 The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?
@@
 Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
-(Should this last step go here or within define conceptual framework?.
-***HF comment.
-I think it belongs with the conceptual framework***).
+(Should this last step go here or within define conceptual framework?. ***HF comment. I think it belongs with the conceptual framework***).
 
-What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)? This is relevant to model transferability... (similar to a COG statement..)* 
-A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation.
+What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)?
+This is relevant to model transferability...
+(similar to a COG statement..)\* A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation.
 I'm just thinking about how I might fill this out for an imagined study.
@@
 Outcomes could be qualitative or quantitative.
-For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.*
-To be honest, I don't know how I might apply this to my work or any future works.
+For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.* To be honest, I don't know how I might apply this to my work or any future works.
 Maybe I do not have enough knowledge of this type of research to apply it to my basic ecology research.
@@
 
-* Describe how uncertainty and variation will be represented in this model.
-This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].*
+-   Describe how uncertainty and variation will be represented in this model.
+This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].\*
 
@@
 -   *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study.
-For example, have you read any reports or publications about this data?
-*
+For example, have you read any reports or publications about this data?*
 
@@
 *Describe how you will determine how reliable the data is for the given purpose.
-Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].*
-### 2.4.5 Data Evaluation: Data reliability reporting
-### 2.4.5. Data Evaluation: data reliability reporting
+Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].* \### 2.4.5 Data Evaluation: Data reliability reporting \### 2.4.5.
+Data Evaluation: data reliability reporting
 
@@
 Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006).
-However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.*
-HF- it's not clear to me what a model feature is and I find this a little confusing. I wonder if we need the text describing the 3.2 heading when it's stepped out in the subsections.
+However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.* HF- it's not clear to me what a model feature is and I find this a little confusing.
+I wonder if we need the text describing the 3.2 heading when it's stepped out in the subsections.
 
@@
 Operationalisations could include:- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, - qualitative change, such as a direction of change or- the frequency, location, or probability of some event occuring.
-Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).*
-HF- this is great. Only slight concern is the deceptively difficult task that last sentence represents
+Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).* HF- this is great.
+Only slight concern is the deceptively difficult task that last sentence represents
 
@@
 
-*Specify model structural features, including:
-- the functional form of interactions,
-- data structures
-- measures used to specify links,
-- any bins or discretisation of continuous variables.*
+*Specify model structural features, including: - the functional form of interactions, - data structures - measures used to specify links, - any bins or discretisation of continuous variables.*
 
@@
 spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii].
-Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.*
-HF- I think this paragraph is only relevant to performance criteria- doesn't really explore estimation technique.
+Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.* HF- I think this paragraph is only relevant to performance criteria- doesn't really explore estimation technique.
 As above though, I'm not sure that this paragraph is neessary given that it's stepped out in the subheadings below
@@
 ### 3.4.2 Choice of parameter estimation technique
-HF- I think that it might be double dipping to include structure here as well as in 3.3
-*Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g.
+
+HF- I think that it might be double dipping to include structure here as well as in 3.3 *Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g.
 distributed boundary conditions).
@@
 *Describe how you will validate and check the calibration of the model.
-The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation)  [@Yates2018;@Barnard2019].
+The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) [@Yates2018; @Barnard2019].
 Justify your choice*
@@
 
-*"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification" <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd].
+*"Model Checking" goes by many names ("conditional verification", "quantitative verification", "model output verification"* <cite>), and refers to a series of analyses that assess a model's performance in representing the system of interest [@Conn:2018hd].
 Model checking therefore heps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system [@Conn:2018hd].
-Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].*
+Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation [@Jakeman:2006ii].
 
@@
 # 5.0 Model Validation and Evaluation
-HF- I think that this is covered in section 4 with descriptions of validation and checking. I'm not sure how this section is different.
+
+HF- I think that this is covered in section 4 with descriptions of validation and checking.
+I'm not sure how this section is different.
 
@@
 ## 5.1 Model output corroboration
+
 *Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model.
-Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step <2.4>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al.
+Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step \<2.4\>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al.
 2014.
@@
 ## 5.3 Evaluate model assumptions
-HF- I would put this up in section 4 and get rid of section 5
-*Explain how you will demonstrate a lack of violation of or robustness to model assumptions.
+
+HF- I would put this up in section 4 and get rid of section 5 *Explain how you will demonstrate a lack of violation of or robustness to model assumptions.
 If you cannot perform quantitative tests, specify what theoretical justifications would justify a lack of violation of or robustness to model assumptions.
@@
 ## 5.4 Model analysis
-HF- I would put 5.4 and it's sbsections up in section 4 and get rid of section 5. Also, as above, remove the paragraph under 5.4
-*Uncertainty in models arises due to incomplete system understanding (which processes to include, or which interact), from imprecise, finite and sparese data measurements, and from uncertainty in input conditions and scenarios for model simulations or runs [@Jakeman:2006ii].
+
+HF- I would put 5.4 and it's sbsections up in section 4 and get rid of section 5.
+Also, as above, remove the paragraph under 5.4 *Uncertainty in models arises due to incomplete system understanding (which processes to include, or which interact), from imprecise, finite and sparese data measurements, and from uncertainty in input conditions and scenarios for model simulations or runs [@Jakeman:2006ii].
 Non-technical uncertainties can also be introduced throughout the modellign process, such as uncertainties arising from issues in problem-framing, indeterminicies, and modeller / client values [@Jakeman:2006ii].The purpose of model analysis is to prevent blind trust in the model by understanding how model outputs have emerged, and to ‚Äòchallenge' the model by verifying whether the model is still believeable and fit for purpose if one or more parameters are changed [@Grimm:2014es].
@@
 *Uncertainty can arise from different modeling techniques, response data and predictor variables [@Araujo2019].
-Uncertainty analyses characterise the uncertainty in model outputs, and identify how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es;@Saltelli2019].
+Uncertainty analyses characterise the uncertainty in model outputs, and identify how uncertainty in model parameters affects uncertainty in model output, but does not identify which model assumptions are driving this behaviour [@Grimm:2014es; @Saltelli2019].
 Uncertainty analyses can include propagating known uncertainties through the model, or by investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], for example.
@@
 # 6.0 Model Application
-HF- I feel that this belongs at the end of section 1
-*Explain how you will operationalise any scenarios identified during the problem formulation phase, if not already done so.¬†For example, how will you operationalise any qualitative changes of interest, such as ‚Äòdeterioration' or ‚Äòimprovement'?Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario.
+
+HF- I feel that this belongs at the end of section 1 *Explain how you will operationalise any scenarios identified during the problem formulation phase, if not already done so.¬†For example, how will you operationalise any qualitative changes of interest, such as ‚Äòdeterioration' or ‚Äòimprovement'?Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario.
 Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections.
@@
 ## 6.1 Choose performance criteria for scenario analysis
-HF- I feel that this belongs at the end of section 1
-*Describe how you will evaluate and distinguish the performance of alternative scenario outcomes.
+
+HF- I feel that this belongs at the end of section 1 *Describe how you will evaluate and distinguish the performance of alternative scenario outcomes.
 These measures should link back to the analytical objectives specified during the problem formulation phase, and might include for example, descriptive statistic measures, such as a model in the top 10th percentile of worst conditions, satisficing measures such as a minimum performance threshold for achieving some management objective, or perhaps regret measures, such as choosing a decision with minimum regret [@Moallemi2019]."Justify or otherwise explain how you chose these measures, and determined performance criteria in relation to the analytical objectives, modelling context, such as the risk attitudes of decision-makers and stakeholders within this system.*
@@
 ## 6.2 Synthesis of model application results
+
 *Specify how you will map, plot or otherwise display and synthesise the results of scenario and model analyses.
@@
 ## 6.3 Consultation with Experts, Clients, Stakeholders
+
 HF- I'm not sure how possible people will find doing this...

### 2023-10-04 index.qmd
Add page links to index page

--- 2023-09-27 index.qmd
@@
 ---
-title: "quarto"
+title: "Adaptive Preregistration for Model-Based Research"
 ---
 
-This is a Quarto website.
+This site provides a guide to implementing *Adaptive Preregistration for Model-Based Research*, with a focus on ecological modelling as it is used in both basic and applied contexts.
+This site covers the following topics:
 
-To learn more about Quarto websites visit <https://quarto.org/docs/websites>.
-
-```{r}
-1 + 1
-```
+1.  [Background on researcher degrees of freedom](RDF_QRP.qmd), the risk of questionable research practices (QRPs), and introduction to preregistration as a tool for mitigating QRPs.
+2.  Adaptive Preregistration
+    1.  [*Background*](adaptive_preregistration_background.qmd): What is adaptive preregistration, why do we need it for registering modeling analyses, and what are its defining features?
+    2.  [*A practical guide*](adaptive_preregistration_implementation.qmd): How to implement adaptive preregistration using [git](https://git-scm.com) and [GitHub](www.github.com).
+3.  [Preregistration template](preregistration_modelling.qmd) for model-based research in ecology and conservation.
+4.  TBC An overview of the sorts of things that we should preregister when modelling, what practices to look out for, what things should be registered as decision-trees.

### 2023-10-10 preregistration_modelling.qmd
Update preregistration template

@@
 
-## 1.1 Model Purpose
+## Model Context and Purpose
 
@@
 Note that modelling objectives are distinct from the analytical objectives of the model.
+
+The scope of the model includes temporal and spatial resolutions, which should also be defined here [@Mahmoud2009].
+Any external limitations on model development, analysis and flexibility should also be outlined in this section [@Jakeman:2006ii].
 :::
 
-### 1.1.1 Define model purpose and problem context
+### Key stakeholders and model users
 
@@
 
--   [ ] Provide a clear statement of the modelling objectives and problem that the model seeks to illuminate.
-What is the purpose of the model/s?
-Ensure that you specify any focal taxa and study objectives, as well as any clients for whom the model is developed.
-Briefly outline the ecological problem and the decision-problem, including the decision-trigger and any regulatory frameworks relevant to the problem.
+Identify relevant interest groups:
+
+-   [ ] Who is the model for?
+-   [ ] Who is involved in formulating the model?
+-   [ ] How will key stakeholders be involved in model development?
+-   [ ] Describe the decision-making context in which the model will be used (if relevant).
 :::
 
-## 1.2 Specify modelling context
-
-::: callout-note
-## Rationale & Explanation
-
-The scope of the model, including temporal and spatial resolutions are defined here [@Mahmoud2009], and any limitations on model development analysis and flexibility should be outlined in this section [@Jakeman:2006ii].
-Note that the modelling context is different from the problem context which are described in 1.3.1.
-:::
-
-### 1.2.1 Identify model users / stakeholders
+### Model purpose, context and problem context
 
@@
 
-Identify interest groups in the modelling context.
-This includes clients, and end-users of the model.
-Who is the model for, who is involved in formulating the model?
-Who needs buy in?
-Describe the decision-making context in which the model will be used.
+Briefly outline
+
+-   [ ] the ecological problem,
+-   [ ] the decision problem (if relevant), including the decision-trigger and any regulatory frameworks relevant to the problem,
+-   [ ] how the model will address the problem, being clear about the scope of the model i.e.
+is the model addressing the whole problem, or part of it?
+Are there any linked problems that your model should consider?\
+-   [ ] Ensure that you specify any focal taxa and study objectives,
+-   [ ] Briefly outline the ecological problem and the decision-problem, including the decision-trigger and any regulatory frameworks relevant to the problem.
 :::
 
-### 1.2.2 Model Scope and Scale
+### Model Scope and Scale
 
@@
 Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al.
-2006)* **HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me.
+2006)*
+
+**HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me.
 Also, is this question more to do with describing the scope and limitations of the data rather then the model?**\*.
 The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?
-
-Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
-(Should this last step go here or within define conceptual framework?. ***HF comment. I think it belongs with the conceptual framework***).
 
@@
 
-### 1.2.3 Logistical Constraints
+### Logistical Constraints
 
-*Time-frame - When must the model by completed, e.g. to help make a decision?*
+::: callout-caution
+## Preregistration Item
 
-*What effort and resources are available for both developing the model and operating the model?* ***HF comment; as written this might be slightly confronting because you could interpret this to include writing explictly about budget constraints which might be confidential.
-Maybe this could be alleviated with with an example?***
+-   [ ] Time-frame: When must the model be completed by, e.g. to help make a decision?
+-   [ ] What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?
+:::
 
-*What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?*
+## Explain analytical objectives
 
-## 1.3.1 Explain analytical objectives
+::: callout-note
+## Rationale & Explanation
 
-*How will the model be analysed, what analytical questions will the model be used to answer?
+How will the model be analysed, what analytical questions will the model be used to answer?
 Examples from ecological decision-making include: to compare the performance of alternative management actions under budget constraint [@Fraser:2017jf] to search for robust decisions under uncertainty [@McDonald-Madden2008], to choose the conservation policy that minimises uncertainty \[insert ref\].
-See other examples in [@Moallemi2019].*
+See other examples in [@Moallemi2019].
+:::
 
-### 1.3.2 - Identify how the model will tell you about your analytic objectives
+### Identify how the model will tell you about your analytic objectives
 
@@
 Outcomes could be qualitative or quantitative.
-For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.* To be honest, I don't know how I might apply this to my work or any future works.
+For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.*
+
+To be honest, I don't know how I might apply this to my work or any future works.
 Maybe I do not have enough knowledge of this type of research to apply it to my basic ecology research.
@@
 *Candidate decisions should be investigated and are specified a priori.
-Depending on the modelling context, they may be specified by stakeholders, model users or the analyst[@Moallemi2019].
+Depending on the modelling context, they may be specified by stakeholders, model users or the analyst [@Moallemi2019].
 Describe the method used to identify relevant management actions.*
@@
 This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].\*
+-   Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
+(Should this last step go here or within define conceptual framework?.
+***HF comment.
+I think it belongs with the conceptual framework***).
 
@@
 
-# 3.0 Formalise and Specify Model
+# Formalise and Specify Model
 
@@
 
-## 3.1 Choose model class, framework and approach
+## Model class, modelling framework and approach
 
-*Describe what class or approach of model you will use and¬†explain how the choice of model class was informed by the analytical objectives of the model.
-Modelling approaches lie on a spectrum from correlative or phenomenological to mechanistic or process-based¬† [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].*
+::: callout-note
+## Explanation
+
+Modelling approaches can be described as occurring on a spectrum from correlative or phenomenological to mechanistic or process-based [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].
+A model \'class,\' \'family\'\' or \'type\' is often used to describe a set of models each of which has a distinct but related sampling distribution [@Liu2008].
+The model family is driven by choices about the types of variables covered and the nature of their treatment, as well as structural features of the model, such as link functions, spatial and temporal scales of processes and their interactions [@Jakeman:2006ii].
+:::
+
+::: callout-caution
+## Preregistration Item
+
+-   Describe what modelling framework, approach or class of model you will use to implement your model and relate your choice to the model purpose and analytical objectives described in 1.1.2 and 1.1.3.
+:::
 
@@
 
-## 3.3 Choose approach for identifying model structure and parameters
+## Describe *approach* for identifying model structure
 
-*This refers to the process of determining the best/most efficient/parsimonious representation of the system at the appropriate scale of concern (Jakeman, 2006) that best meets the analytical objectives specified in the problem formulation phase.
+::: callout-note
+## Explanation
+
+This section relates to the process of determining the best/most efficient/parsimonious representation of the system at the appropriate scale of concern [@Jakeman:2006ii] that best meets the analytical objectives specified in the problem formulation phase.
+Model structure refers to the choice of variables included in the model, and the nature of the relationship among those variables.
 Approaches to finding model structure and parameters may be knowledge-supported, or data-driven [@Boets:2015gl].
-Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods (Barnard et al.
-2019).
+Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods \[@Barnard2019\].
 Ensemble modelling procedures might aim to derive a single model, or a multi-model average [@Yates2018].
-Refining actions to develop a model could include iteratively dropping parameters or adding them, or aggregating / disaggregating system descriptors, such as dimensionality and processes (Jakeman, 2006).
-Specify what approach and methods you will use to identify model structure and parameters.*
+Refining actions to develop a model could include iteratively dropping parameters or adding them, or aggregating / disaggregating system descriptors, such as dimensionality and processes [@Jakeman:2006ii].
+:::
 
-## 3.4 Choose estimation technique and performance criteria
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Specify what approach and methods you will use to identify model structure and parameters.
+-   [ ] If using a knowledge-supported approach to deriving model strucure (either in whole or in part), specify model structural features, including:
+    -   the functional form of interactions (if any)
+    -   data structures,
+    -   measures used to specify links,
+    -   any bins or discretisation of continuous variables [@Jakeman:2006ii],
+    -   any other relevant features of the model structure.
+:::
+
+## Describe parameter estimation technique and performance criteria
 
@@
 For example, modellers might seek parameter estimates that are robust to outliers, unbiased, and yield appropriate predictive performance.
-Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are appropriate to the problem at hand.
+Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are suited to the problem at hand.
 For integrated or sub-divided models, other considerations might include choices about where to disaggregate the model for parameter estimation; e.g.
-spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii].
-Specifying performance criteria a priori is important because it involves pre-specifying how the fitted or quantitative model will be interpreted in advance, avoiding biases like confirmation bias, HARKing, and cherry-picking some performance tests from the full suite of tests undertaken.* HF- I think this paragraph is only relevant to performance criteria- doesn't really explore estimation technique.
-As above though, I'm not sure that this paragraph is neessary given that it's stepped out in the subheadings below
+spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii].*
+
+NOTE THIS DOESN"T SAY ANYTHING ABOUT PARAMETER ESTIMATION! ONLY PErformacne criteria
+
+### 3.4.2 Choice of parameter estimation technique
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g.
+distributed boundary conditions).
+For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination \[@Jakeman:2006ii\].
+-   [ ] Identify which variables will be parameterised directly, such as by expert elicitation or prior knowledge.
+-   [ ] Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].
+:::
 
@@
 
-### 3.4.2 Choice of parameter estimation technique
-
-HF- I think that it might be double dipping to include structure here as well as in 3.3 *Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g.
-distributed boundary conditions).
-For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006).
-Identify which variables were parameterised directly, such as by expert elicitation or prior knowledge.
-Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms [@Liu2018b].*
-
 ## 3.5 Specify model assumptions and uncertainties
@@
 
-## 3.6 Specify formal model
+## 3.6 Specify formal model(s)
 
-*Once critical decisions have been made about the approach and method of model specification,translate the conceptual model into the quantitative model.
+*Once critical decisions have been made about the approach and method of model specification, translate the conceptual model into the quantitative model.
 For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters.
@@
 
-# Model Calibration, Model Fitting & Checking
+# Model Calibration, Validation & Checking
 
@@
 
-*Describe how you will validate and check the calibration of the model.
-The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) [@Yates2018; @Barnard2019].
-Justify your choices.*
+::: callout-note
+## Explanation
 
-### Describe validation data
+This section pertains to any data calibration, validation or testing schemes that will be implemented.
+For example, the model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) [@Yates2018; @Barnard2019].
+For some types of models, hyper-parameters are estimated from data, and may be tuned on further independent holdouts of the training data, ("validation data").
+:::
 
-*If partitioning data for cross-validation: describe the approach specifying the number of folds that will be created and the relative size of each.
-Describe how will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced.*
+::: callout-caution
+## Preregistration Item
 
-*If using external data: describe any known differences between the training and validation datasets and the size of the validation dataset.
-Describe how will you document and/or share the external data such that any subsequent modelling can be computationally reproduced.*
+-   [ ] Describe any data calibration, validation and testing scheme you will implement, including any procedures for tuning or estimating model hyper-parameters (if any).
+:::
+
+### Describe calibration/validation data
+
+::: callout-note
+## Explanation & Rationale
+
+The following items pertain to properties of the *datasets* used for calibration (training), validation, and testing.
+
+## Preregistration Item
+
+-   If partitioning data for cross-validation or similar approach:
+
+    -   [ ] Describe the approach specifying the number of folds that will be created, the relative size of each fold, and any stratification methods used for ensuring evenness of groups between folds and between calibration / validation data?
+
+-   If using external / independent holdout data for model testing and evaluation:
+
+    -   [ ] Which data will be used as a the testing data?
+        What method will you be used for generating training / test data subsets?
+
+    -   [ ] Describe any known differences between the training/validation and testing datasets, the relative size of each, as well as any stratification methods used for ensuring evenness of groups between data sets?
+
+    -   [ ] It is preferable that any independent data used for model testing remains unknown to modellers during the process of model development, please describe the relationship modellers have to model validation data, will independent datasets be known or accessible to any modeller or analyst?
+:::
 
@@
 
-*What Quality Assurance measures will you take to verify the model has been correctly implemented?
-Tests could include syntax checking of code, and code reviews by peers.
-Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as dictated by the model description [@Grimm:2014es].
-Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging.*
+::: callout-caution
+## Preregistration Item
+
+-   [ ] What Quality Assurance measures will you take to verify the model has been correctly implemented?
+    Specifying a priori quality assurance tests for implementation verification may help to avoid selective debugging and silent errors.
+
+    -   Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as described by the model description *\[@Grimm:2014es\].*
+
+    -   Qualitative tests could include syntax checking of code, and peer-code review [@ivimey2023].
+        Technical measures include using unit tests, or in-built checks within functions to prevent potential errors.
+:::
 
@@
 Ideally independent data is used against the model outputs to assess whether the model output behaviour exhibits the required accuracy for the model's intended purpose.
-The outcomes of these analyses build confidence in the model applications and increase understandingof model strengths and limitations.
+The outcomes of these analyses build confidence in the model applications and increase understanding of model strengths and limitations.
 Model evaluation including model analysis and other testing should complement model checking.

### 2023-10-13 preregistration_modelling.qmd
Continue updating preregistration template, inc. adding callouts and explanations

@@
 description: "Note: Black boxes provide an explanation or justification of a particular section or preregistration item. Text contained in 'preregistration item' boxes requires a response."
+authors:
+  - Elliot Gould
+  - Hannah Fraser
+  - Libby Rumpff
+  - Megan Good
+  - Chris Jones
 sidebar: false
@@
 
-# 1. Problem Formulation
+# Problem Formulation
 
@@
 is the model addressing the whole problem, or part of it?
-Are there any linked problems that your model should consider?\
+Are there any linked problems that your model should consider?
 -   [ ] Ensure that you specify any focal taxa and study objectives,
@@
 
-### Model Scope and Scale
+### Analytical objectives {#analyticalobjectives}
 
-*Determine the temporal and spatial scope of the model.
-Where is the boundary of the modelled system?
-Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs (observed, or not, Jakeman et al.
-2006)*
+::: callout-note
+## Explanation
 
-**HF comment I dont understand the preceeding sentence, the concept of boundaries of the model is foreign to me.
-Also, is this question more to do with describing the scope and limitations of the data rather then the model?**\*.
-The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model (Jakeman et al., 2006) - what is the intended scale and resolution of the model (temporal, spatial or otherwise)?
+How will the model be analysed, what analytical questions will the model be used to answer?
+For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.
+Other examples from ecological decision-making include: to compare the performance of alternative management actions under budget constraint [@Fraser:2017jf] to search for robust decisions under uncertainty [@McDonald-Madden2008], to choose the conservation policy that minimises uncertainty \[insert ref\].
+See other examples in [@Moallemi2019].
+:::
 
-What is the intended domain of applicability of the model, what is the extent of acceptable extrapolations (Grimm)?
-This is relevant to model transferability...
-(similar to a COG statement..)\* A comment:This section feels a bit too big and confusing from my perspective and I do think you might need to move some of it to a different section to reduce the risk of misinterpretation.
-I'm just thinking about how I might fill this out for an imagined study.
-The literal boundaries of the study are usually straightforward, but the theoretic boundaries are a different and also big thing to try to conceptualise in one section.
-You could have a section for the spatial/temporal boundaries of the study, and another section on how far the results can be extrapolated based on the study design.
-This makes sense to me in how I design experiments, if I have too many confounding variables and not enough spatial and environmental replication then I'm less likely to make any broad claims.
-However, if I have controlled or experimentally manipulated factors then I feel safer making bigger generalisations about results.
+::: callout-caution
+## Preregistration Item
+
+*Provide detail on the analytical purpose and scope of the model:*
+
+-   [ ] How will the model be analysed and what analytical questions will the model be used to answer? 
+-   [ ] Candidate decisions should be investigated and are specified a priori. Depending on the modelling context, they may be specified by stakeholders, model users or the analyst [@Moallemi2019].
+    -   [ ] Describe the method used to identify relevant management actions and
+    -   [ ] specify management actions to be considered included in the model.
+    -   [ ] Are there potentially unacceptable management or policy outcomes identified by stakeholders that should be captured in the model, i.e. as constraints?
+-   [ ] Are there scenarios that model inputs or outputs that must accommodated?
+Scenarios should be set a priori, [i.e.
+before the model is built, @Moallemi2019] and may be stakeholder-defined or driven by the judgement of the modeller or other experts [@Mahmoud2009].
+    -   [ ] If relevant, describe what processes you will use to elicit and identify relevant scenarios, e.g. literature review, structured workshops with stakeholders or decision-makers. 
+    -   [ ]  Specify scenarios under which decisions are investigated.
+:::
 
@@
 ::: callout-caution
+-   [ ] What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?
+-   [ ] Are there any limitations on model development analysis and flexibility, such as time or budget constraints, for example, does a model need to be deployed rapidly?
+    -   [ ] When must the model be completed by, e.g. to help make a decision?
+:::
+
+### Model Scope, Scale and Resolution
+
+::: callout-caution
 ## Preregistration Item
 
--   [ ] Time-frame: When must the model be completed by, e.g. to help make a decision?
--   [ ] What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?
+-   [ ] The choice of a model's boundaries is closely linked to the choice of how finely to aggregate the behaviour within the model [@Jakeman:2006ii] - what is the intended scale, and resolution of the model (temporal, spatial or otherwise)?
+-   [ ] Where is the boundary of the modelled system?
+Everything outside beyond the boundary and not crossing it is to be ignored within the domain of the model, and everything crossing the boundary is to be treated as external forcing (known/unknown), or else as model outputs [observed, or not, @Jakeman:2006ii].
 :::
 
-## Explain analytical objectives
+### Intended application of results
 
 ::: callout-note
-## Rationale & Explanation
+## Explanation
 
-How will the model be analysed, what analytical questions will the model be used to answer?
-Examples from ecological decision-making include: to compare the performance of alternative management actions under budget constraint [@Fraser:2017jf] to search for robust decisions under uncertainty [@McDonald-Madden2008], to choose the conservation policy that minimises uncertainty \[insert ref\].
-See other examples in [@Moallemi2019].
+Preregistration Items in this section are relevant to model transferability [@Yates2018] and constraints on generality in model analysis interpretation.
+How far do can the results be extrapolated based on the study design (data + model + analysis)?
+For instance, if there are many confounding variables and not enough spatial / environmental replication, then making broader more general claims beyond the stated boundaries of the model (@analyticalobjectives) may not be warranted.
+However, larger generalisations about results may be acceptable if the data comes from experimentally manipulated or controlled systems.
 :::
 
-### Identify how the model will tell you about your analytic objectives
+::: callout-caution
+## Preregistration Item
 
-*The outcomes should speak directly to the analytical objectives identified in 1.3.1.
-Outcomes could be qualitative or quantitative.
-For example, you might be using your model in a scenario analysis to determine which management decision is associated with minimum regret or the highest likelihood of improvement.*
+-   [ ] What is the intended domain in which the model is to be applied? Are there any reasonable extrapolation limits beyond which you expect the model should not be applied [@Grimm:2014es]?
+:::
 
-To be honest, I don't know how I might apply this to my work or any future works.
-Maybe I do not have enough knowledge of this type of research to apply it to my basic ecology research.
-But this isn't to say it's not important, I just wonder if there's a way to make this accessible to non-decision or modelling peeps.
-Like, how would I apply it to the woodland condition analysis?
+## Scenario Analysis Operationalisation
 
-## 1.4 Define candidate decisions and method of identifying relevant management actions
+::: callout-caution
+## Preregistration Item (delete as necessary)
 
-*Candidate decisions should be investigated and are specified a priori.
-Depending on the modelling context, they may be specified by stakeholders, model users or the analyst [@Moallemi2019].
-Describe the method used to identify relevant management actions.*
+-   How will you operationalise any scenarios identified in 1.1.3?
+    For example, how will you operationalise any qualitative changes of interest, such as ‚ \'deterioration\' or \'improvement\'?
 
-## 1.5 Specify scenarios
+-   Describe how you will evaluate and distinguish the performance of alternative scenario outcomes
 
-*Describe what processes you will use to elicit and identify relevant scenarios, e.g.
-literature review, structured workshops with stakeholders or decision-makers.
-Specify scenarios under which decisions are investigated.
-Scenarios should be set a priori [i.e.
-before the model is built, @Moallemi2019] and may be stakeholder-defined or driven by the judgement of the modeller or other experts [@Mahmoud2009].*
+-   Justify or otherwise explain how you chose these measures and determined performance criteria in relation to the analytical objectives, model purpose and modelling context, such as the risk attitudes of decision-makers and stakeholders within this system
+:::
 
-# 2.0 Define Conceptual Model
+# Define Conceptual Model
 
-*Conceptual models underpin the formal or quantitative model*
+::: callout-note
+## Explanation
 
-*(Cartwright et al.
-2016).
+Conceptual models underpin the formal or quantitative model [@Cartwright:2016kr].
 The conceptual model describes the biological mechanisms relevant to the ecological problem and should capture basic premises about how the target system works, including any prior knowledge and assumptions about system processes.
-Conceptual models may be representeed in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables [@Jakeman:2006ii].*
+Conceptual models may be represented in a variety of formats, such as influence diagrams, linguistic model block diagram or bond graphs, and these illustrate how model drivers are linked to both outputs or observed responses, and internal (state) variables [@Jakeman:2006ii].
+:::
 
-## 2.1 Choose elicitation and representation method
+## Choose elicitation and representation method
 
-*Describe what method you will use to elicit or identify the conceptual model.
-Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks (Moon et al.
-2019).
-(Libby, to provide link to any structured expert elicitation methods?).While it is difficult to decide and justify which method is most appropriate, however Moon et al.
-(2019) provide guidance addressing this methodological question.
-Finally, how do you intend on representing the final conceptual model?
-This will likely depend on the method chosen to elicit the conceptual model.*
+::: callout-caution
+## Preregistration Item
 
-## 2.2.1 Explain Critical Conceptual Design Decisions
+-   [ ] Describe what method you will use to elicit or identify the conceptual model.
+Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks \[@Moon2019\].
+It is difficult to decide and justify which method is most appropriate, see Moon et al.
+[-@Moon2019] for guidance addressing this methodological question.
+-   [ ] Finally, how do you intend on representing the final conceptual model? This will likely depend on the method chosen to elicit the conceptual model.
+:::
 
-*This step should list and explain the critical conceptual design decisions, including: spatial and temporal scales, selection of entities and processes, representation of stochasticity and heterogeneity, consideration of local versus global interactions, environmental drivers, etc.
-[@Grimm:2014es].
-The influence of particular theories, concepts, or importantly, earlier models, should be explained and justified against alternative conceptual design decisions that might lead to alternative model structures [@Grimm:2014es].*
+## Explain Critical Conceptual Design Decisions
 
-## 2.2.2 Specify key assumptions and uncertainties underlying the model's design.
+::: callout-caution
+## Preregistration Item
 
--   Describe how uncertainty and variation will be represented in this model.
-This includes both exogenous uncertainties affecting the system, parametric uncertainty in input data and structural / conceptual nonparametric uncertainty in the model [@Moallemi2019].\*
--   Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
+List and explain critical conceptual design decisions [@Grimm:2014es], including:
+
+-   [ ] spatial and temporal scales,
+
+-   [ ] selection of entities and processes,
+
+-   [ ] representation of stochasticity and heterogeneity,
+
+-   [ ] consideration of local versus global interactions, environmental drivers, etc.
+
+-   [ ] Explain and justify the influence of particular theories, concepts, or earlier models against alternative conceptual design decisions that might lead to alternative model structures.
+:::
+
+## Model assumptions and uncertainties
+
+::: callout-caution
+## Preregistration Item
+
+Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be represented in the model [@Moallemi2019].
+Sources of uncertainty may include:
+
+-   [ ] exogenous uncertainties affecting the system,
+
+```{=html}
+<!-- -->
+```
+-   [ ] parametric uncertainty in input data and
+
+-   [ ] structural / conceptual nonparametric uncertainty in the model.
+:::
+
+## Identify predictor and response variables
+
+::: callout-note
+## Explanation
+
+The identification and definition of primary model input variables should be driven by scenario definitions, and by the scope of the model described in the problem formulation phase [@Mahmoud2009].
+:::
+
+::: callout-caution
+## Preregistration Item
+
+Identify and define system system variables and structures, referencing scenario definitions, and the scope of the model as described within problem formulation (1.1.3):
+
+-   [ ] What variables would support taking this action or making this decision?
+-   [ ] What additional variables may interact with this system (things we can't control, but can hopefully measure)?
+-   [ ] What variables have not been measured, but may interact with the system (often occurs in field or observational studies)?
+-   [ ] What variables are index or surrogate measures of variables that we cannot or have not measured?
+-   [ ] In what ways do we expect these variables to interact (model structures)?
+-   [ ] Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
 (Should this last step go here or within define conceptual framework?.
-***HF comment.
-I think it belongs with the conceptual framework***).
+:::
 
-## 2.3 Identify predictor and response variables
+## Define prior knowledge, data specification and evaluation
 
-*The identification and definition of primary model input variables should be driven by scenario definitions, and by the scope of the model described in the problem formulation phase [@Mahmoud2009] -- Identify and define system variables structures:*
+::: callout-note
+## Explanation
 
--   *What variables would support taking this action or making this decision?*
--   *What additional variables may interact with this system (things we can't control, but can hopefully measure)?*
--   *What variables have not been measured, but may interact with the system (often occurs in field or observational studies)?*
--   *What variables are indice or surrogate measures of variables that we cannot or have not measured?*
--   *In what ways do we expect these variables to interact (model structures)?*
+*This section specifies the plan for collecting, processing and preparing data available for parameterisation, determining model structure, and for scenario analysis.*
+:::
 
-## 2.4 Define prior knowledge, data specification and evaluation
+### Collate available data sources that could be used to parameterise or structure the model
 
-*Collect, process and prepare data available for parameterisation, determining model structure, and for scenario analysis.*
+::: callout-caution
+## Preregistration Item
 
-### 2.4.1 Collate available data sources that could be used to parameterise or structure the model
+**For pre-existing data (delete as appropriate):**
 
-*Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model. Describe how the data is arranged, in terms of replicates and covariates.*
+-   [ ] Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model.
+-   [ ] For each dataset, is the data open or publicly available?
+-   [ ] How can the data be accessed? Provide a link or contact as appropriate, indicating any restrictions on the use of data.
+-   [ ] Date of download, access, or future access.
+-   [ ] Describe the source of the data - what entity originally collected this data?
+(National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).
+-   [ ] Codebook and meta-data. If a codebook or other meta-data is available, link to it here and / or upload the document(s).
+-   [ ] Prior work based on this dataset - Have you published / presented any previous work based on this dataset?
+Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based on this dataset you have worked on.
+-   [ ] Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.
+-   [ ] Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study.
+For example, have you read any reports or publications about this data?
+-   [ ] Describe how the data is arranged, in terms of replicates and covariates.
 
-**For pre-existing data: -- delete as appropriate.**
+**Sampling Plan (for data you will collect, delete as appropriate):**
 
--   *For each dataset, is the data open or publically available? (Y/N)*
--   *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrctions on the use of data.*
--   *Date of download, access, or future access:*
--   *Describe the source of the data - what entity originally collected this data?
-(National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
--   *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
--   *Prior work based on this dataset - Have you published / presented any previous work based on this dataset?
-Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based on this dataset you have worked on.* HF- Love this question
--   *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
--   *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study.
-For example, have you read any reports or publications about this data?*
+-   [ ] Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged.
+Describe any inclusion or exclusion rules, and the study timeline.
+-   [ ] Sample Size - Describe the sample size of your study.
+-   [ ] Sample Size Rationale - Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel.
+-   [ ] If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?
+:::
 
-**Sampling Plan (For data you will collect) -- delete as appropriate.**
+### Data Processing and Preparation
 
--   Data collection procedures - Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged.
-Describe any inclusion or exclusion rules, and the study timeline.
--   Sample Size - Describe the sample size of your study.
--   Sample Size Rationale - Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel.
--   If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?
+::: callout-caution
+## Preregistration Item
 
-### 2.4.2 Data Processing and Preparation
+-   [ ] Describe any data preparation and processing steps, including manipulation of environmental layers (e.g.
+standardisation and geographic projection) or variable construction (e.g.
+Principal Component Analysis).
+:::
 
-*Describe any data preparation and processing steps, including manipulation of environmental layers, e.g. standardisation and geographic projection.*
+### Describe any data exploration or preliminary data analyses.
 
-*Describe how you will separate and distinguish between raw data, manipulated data, and outputs from modelling or any analyses of the model.*
+::: callout-note
+## Explanation
 
-### 2.4.3 Describe any data exploration or preliminary data analyses.
+In most modelling cases, it is necessary to perform preliminary analyses to understand the data and check that assumptions and requirements of the chosen modelling procedures are met.
+Data exploration prior to model fitting or development may include exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers, or the need for transformations [@Yates2018].
+:::
 
-*In most modelling cases, it is necessary to perform preliminary analyses to understand the data and check that assumptions and requirements of the chosen modelling procedures are met.
-Data exploration prior to model fitting or development may include exploratory analyses to check for collinearity, spatial and temporal coverage, quality and resolution, outliers, or the need for transformations [@Yates2018].
-Describe how you will summarise and explore your data, and explain what method you will use (graphical, tabular or otherwise) to represent your data and any analyses.
-Specify what you intend to find out or hope to learn from each data exploration analysis.*
+::: callout-caution
+## Preregistration Item
 
-### 2.4.4 Data Evaluation
+For each separate preliminary or investigatory analysis:
 
-*Describe how you will determine how reliable the data is for the given purpose.
-Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].* \### 2.4.5 Data Evaluation: Data reliability reporting \### 2.4.5.
-Data Evaluation: data reliability reporting
+-   [ ] State what needs to be known to proceed with further decision-making about the modelling procedure, and why the analysis is necessary.
 
-*Document any issues with data reliability.
-This is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].*
+-   [ ] Explain how you will implement this analysis, as well as any techniques you will use to summarise and explore your data.
 
-### 2.4.6. Data evaluation: data exclusion
+-   [ ] What method will you use to represent this analysis (graphical, tabular, or otherwise, describe)
 
-*How will you determine what data, if any, will be excluded from your analyses?
-How will outliers be handled?
-Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.*
+-   [ ] Specify exactly which parts of the data will be used
 
-### 2.4.7 Data Evaluation: Missing Data
+-   [ ] Describe how the results will be interpreted, listing each potential analytic decision, as well as the analysis finding that will trigger each decision, where possible.
+:::
 
-*How will you identify and deal with incomplete or missing data?*
+### Data evaluation, exclusion and missing data
 
-## 2.5 Conceptual model evaluation
+::: callout-note
+## Explanation
 
-*Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.*
+Documenting issues with reliability is important because data quality and ecological relevance might be constrained by measurement error, inappropriate experimental design, and heterogeneity and variability inherent in ecological systems [@Grimm:2014es].
+Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand [@Mahmoud2009].
+:::
 
-*How will any simplifying assumptions be critically assessed [@Augusiak:2014gz]?*
+::: callout-caution
+## Preregistration Item
 
-*Explain whether this process will include consultation or feedback from a client, manager, or model user.*
+-   [ ] Describe how you will determine how reliable the data is for the given model purpose.
+    Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand
+
+-   [ ] Document any issues with data reliability.
+
+-   [ ] How will you determine what data, if any, will be excluded from your analyses?
+
+-   [ ] How will outliers be handled?
+    Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.
+
+-   [ ] How will you identify and deal with incomplete or missing data?
+:::
+
+## Conceptual model evaluation
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.
+-   [ ] How will you critically assess any simplifying assumptions [@Augusiak:2014gz]?
+-   [ ] Will this process will include consultation or feedback from a client, manager, or model user.
+:::
 
@@
 
-*In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.*
+::: callout-note
+## Explanation
+
+In this section describe what quantitative methods you will use to build the model/s, explain how they are relevant to the client/manager/user's purpose.
+:::
 
@@
 Modelling approaches can be described as occurring on a spectrum from correlative or phenomenological to mechanistic or process-based [@Yates2018]; where correlative models use mathematical functions fitted to data to describe underlying processes, and mechanistic models explicitly represent processes and details of component parts of a biological system that are expected to give rise to the data [@White2019a].
-A model \'class,\' \'family\'\' or \'type\' is often used to describe a set of models each of which has a distinct but related sampling distribution [@Liu2008].
+A model 'class,' 'family'' or 'type' is often used to describe a set of models each of which has a distinct but related sampling distribution [@Liu2008].
 The model family is driven by choices about the types of variables covered and the nature of their treatment, as well as structural features of the model, such as link functions, spatial and temporal scales of processes and their interactions [@Jakeman:2006ii].
@@
 Approaches to finding model structure and parameters may be knowledge-supported, or data-driven [@Boets:2015gl].
-Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods \[@Barnard2019\].
+Model selection methods can include traditional inferential approaches such as unconstrained searches of a dataset for patterns that explain variations in the response variable, or use of ensemble-modelling methods [@Barnard2019].
 Ensemble modelling procedures might aim to derive a single model, or a multi-model average [@Yates2018].
@@
 distributed boundary conditions).
-For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination \[@Jakeman:2006ii\].
+For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination [@Jakeman:2006ii].
 -   [ ] Identify which variables will be parameterised directly, such as by expert elicitation or prior knowledge.
@@
 
-    -   Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as described by the model description *\[@Grimm:2014es\].*
+    -   Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as described by the model description [@Grimm:2014es]*.*
 
@@
 
-# References
+# References {.unnumbered}

### 2023-10-15 preregistration_modelling.qmd
Complete formatted preregistration and updating with explanatory text.

@@
 -   How will you operationalise any scenarios identified in 1.1.3?
-    For example, how will you operationalise any qualitative changes of interest, such as ‚ \'deterioration\' or \'improvement\'?
+    For example, how will you operationalise any qualitative changes of interest, such as ‚ 'deterioration' or 'improvement'?
 
@@
 -   [ ] Describe what method you will use to elicit or identify the conceptual model.
-Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks \[@Moon2019\].
+Some common methods include interviews, drawings, and mapping techniques including influence diagrams, cognitive maps and Bayesian belief networks [@Moon2019].
 It is difficult to decide and justify which method is most appropriate, see Moon et al.
@@
 -   [ ] exogenous uncertainties affecting the system,
-
-```{=html}
-<!-- -->
-```
 -   [ ] parametric uncertainty in input data and
-
 -   [ ] structural / conceptual nonparametric uncertainty in the model.
@@
 
--   Describe what modelling framework, approach or class of model you will use to implement your model and relate your choice to the model purpose and analytical objectives described in 1.1.2 and 1.1.3.
+-   [ ] Describe what modelling framework, approach or class of model you will use to implement your model and relate your choice to the model purpose and analytical objectives described in 1.1.2 and 1.1.3.
 :::
 
-## 3.2 Choose model features and family
+## Choose model features and family
 
-*All modelling approaches require the selection of model features, which conform with the conceptual model and data specified in previous steps (Jakeman 2006).
+::: callout-note
+## Explanation
+
+All modelling approaches require the selection of model features, which conform with the conceptual model and data specified in previous steps [@Jakeman:2006ii].
 The choice of model are determined in conjunction with features are selected.
-Usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller (Jakeman, 2006).
-However, if changes to these fundamental aspects of the model do need to change, please document how and why these choices were made.* HF- it's not clear to me what a model feature is and I find this a little confusing.
-I wonder if we need the text describing the 3.2 heading when it's stepped out in the subsections.
+Model features include elements such as the functional form of interactions, data structures, measures used to specify links, any bins or discretisation of continuous variables.
+It is usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller \[@Jakeman:2006ii\].
+However, if changes to these fundamental aspects of the model do need to change, document how and why these choices were made, including any results used to support any changes in the model.
+:::
 
-### 3.2.1 Explain how you will operationalise response variable(s)
+### Explain how you will operationalise response and predictor variables, and covariates
 
-*Specify how you will operationalise the response variables in the model.
-This should relate directly to the analytical and or management objectives specified during the problem formulation phase.
-Operationalisations could include:- the extent of a response- an extreme value- a trend- a long-term mean- a probability distribution- a spatial pattern- a time-series, - qualitative change, such as a direction of change or- the frequency, location, or probability of some event occuring.
-Provide a rationale for your choices, including why plausible alternatives were not chosen (Jakeman, 2006).* HF- this is great.
-Only slight concern is the deceptively difficult task that last sentence represents
+::: callout-caution
+## Preregistration Item
 
-### 3.2.2 Choose model family
+-   [ ] For each response, predictor, and covariate, *specify how these variables will be operationalised in the model.
+This should relate directly to the analytical and/or management objectives specified during the problem formulation phase.
+Operationalisations could include: the extent of a response, an extreme value, a trend, a long-term mean, a probability distribution, a spatial pattern, a time-series, qualitative change, such as a direction of change or, the frequency, location, or probability of some event occuring.
+Specify any treatment of model variables, including whether they are lumped / distributed, lienar / non-linear, stochastic / deterministic [@Jakeman:2006ii]* .
+-   [ ] *Provide a rationale for your choices, including why plausible alternatives under consideration were not chosen, and relate your justification bacj to the purpose, objectives, prior knowledge and or logistical constraints specified in the problem formulation phase \[@Jakeman:2006ii\].*
+:::
 
-*Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions.
-Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints (Jakeman, 2006) specified in the problem formulation phase.*
+### Choose model family
 
-*Include in your rational for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*
+::: callout-caution
+## Preregistration Item
 
-### 3.2.3 Choose model features
-
-*Specify which variables are included in the model, and the nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).*
-
-*Specify model structural features, including: - the functional form of interactions, - data structures - measures used to specify links, - any bins or discretisation of continuous variables.*
+-   [ ] Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions.
+-   [ ] Include in your rational for selection, detail about which variables the model outputs are likely sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.
+:::
 
@@
 -   [ ] Specify what approach and methods you will use to identify model structure and parameters.
--   [ ] If using a knowledge-supported approach to deriving model strucure (either in whole or in part), specify model structural features, including:
+-   [ ] If using a knowledge-supported approach to deriving model structure (either in whole or in part), specify model structural features, including:
     -   the functional form of interactions (if any)
@@
 
-*Before calibrating the model to the data, the performance criteria on which the calibration is judged are chosen.
+::: callout-note
+## Explanation
+
+Before calibrating the model to the data, the performance criteria on which the calibration is judged are specified.
 These criteria and their underlying assumptions should reflect the desired properties of the parameter estimates / structure [@Jakeman:2006ii].
 For example, modellers might seek parameter estimates that are robust to outliers, unbiased, and yield appropriate predictive performance.
-Modellers will need to consider whether the assumptions of the estimation technique yielding those desired criteria are suited to the problem at hand.
+Modellers will need to consider whether the assumptions of the estimation technique yielding those desired properties are suited to the problem at hand.
 For integrated or sub-divided models, other considerations might include choices about where to disaggregate the model for parameter estimation; e.g.
-spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii].*
+spatial sectioning (streams into reaches) and temporal sectioning (piece-wise linear models) [@Jakeman:2006ii].
 
-NOTE THIS DOESN"T SAY ANYTHING ABOUT PARAMETER ESTIMATION! ONLY PErformacne criteria
+<!-- #TODO NOTE THIS DOESN"T SAY ANYTHING ABOUT PARAMETER ESTIMATION! ONLY PErformacne criteria -->
+:::
 
-### 3.4.2 Choice of parameter estimation technique
+### Parameter estimation technique
 
@@
 
-### 3.4.1 Choice of performance criteria
+### Performance Criteria
 
-*Specify which suite of performance criteria you will use to judge the performance of the model.
-Examples include correlation scores, coefficient of determination, specificity, sensitivity, AUC, etcetera (Yates et al.
-2018).
-Relate any underlying assumptions of each criterion to the desired properties of the model, and justify the choice of performance metric in relation into whether it is sensitive to the problem at hand.
-Explain how you will identify which model features or components are significant or meaningful.*
+::: callout-caution
+## Preregistration Item
 
-## 3.5 Specify model assumptions and uncertainties
+-   [ ] Specify which suite of performance criteria you will use to judge the performance of the model.
+Examples include correlation scores, coefficient of determination, specificity, sensitivity, AUC, etcetera [@Yates2018].
+-   [ ] Relate any underlying assumptions of each criterion to the desired properties of the model, and justify the choice of performance metric in relation
+-   [ ] Explain how you will identify which model features or components are significant or meaningful.
+:::
 
-*Specify all assumptions and key uncertainties in the formal model.
-Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpretation of the model outputs, and what implications are there on evaluating model-output to inform decisions?*
+## Model assumptions and uncertainties
 
-## 3.6 Specify formal model(s)
+::: callout-caution
+## Preregistration Item
 
-*Once critical decisions have been made about the approach and method of model specification, translate the conceptual model into the quantitative model.
-For data-driven and model-selection approaches that determine model structure and parameters, describe any initial model specifications and parameterisations, including for any tune-in parameters.
-(Should this go here or in model calibration?)*
+Specify assumptions and key uncertainties in the formal model.
+Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpretation of the model outputs, and what implications are there for evaluating model-output to inform inferences or decisions?
+:::
+
+## Specify formal model(s)
+
+::: callout-caution
+## Preregistration Item
+
+Once critical decisions have been made about the modelling approach and method of model specification, the conceptual model is translated into the quantitative model.
+
+-   [ ] Specify all formal models
+    -   Note, For data-driven approaches to determining model structure and or parameterisation, it will not be able to complete this preregistration item.
+-   [ ] For quantitative model selection approaches, including ensemble modelling, specify each model used in the candidate set, including any null or full/global model.
+:::
 
@@
 
-# 5.0 Model Validation and Evaluation
+# Model Validation and Evaluation
 
-HF- I think that this is covered in section 4 with descriptions of validation and checking.
-I'm not sure how this section is different.
+::: callout-note
+## Explanation
 
-*Model Validation & Evaluation - This phase consists of a suite of analyses that collectively inform the decision and whether and when a model is suitable to meet its intended purpose (Augusiak, Van den Brink and Grimm 2014).
+The model validation & evaluation phase comprises a suite of analyses that collectively inform inferences about whether, and under what conditions, a model is suitable to meet its intended purpose \[@Augusiak:2014gz\].
 Errors in design and implementation of the model and their implication on the model output are assessed.
@@
 The outcomes of these analyses build confidence in the model applications and increase understanding of model strengths and limitations.
-Model evaluation including model analysis and other testing should complement model checking.
+Model evaluation including, model analysis, should complement model checking.
 It should evaluate model checking, and consider over-fitting and extrapolation.
-The higher the proportion of calibrated, or uncertain parameters, the greater the risk that the model seems to work correctly \[‚Ä¶\] but for the wrong reasons" Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mecahnisms of the system (Grimm).
-Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model (Grimm et al.
-2014).*
+The higher the proportion of calibrated, or uncertain parameters, "the greater the risk that the model seems to work correctly, but for the wrong reasons" (citaiton).
+Evaluation thus complements model checking because we can rule out the chance that the model fits the calibration data well, but has not captured the relevant ecological mechanisms of the system pertinent to the research question or the decision problem underpinning the model [@Grimm:2014es].
+Evaluation of model outputs against external data in conjunction with the results from model checking provide information about the structural realism and therefore credibility of the model [@Grimm2016].
 
-## 5.1 Model output corroboration
+<!-- #NOTE, explain how model-checking is different to model validation / evaluation.
+What am i defining as model analysis / application??
+the phases of the modelling process underlying this template need exposition somewhere -->
+:::
 
-*Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model.
-Testing against a dataset of response and predictor variables that are sptially and/or temporally independent from the training dataset¬†minimises the risk of artificially inflating model performance measures [@Araujo2019].State whether you will corroborate the model outputs on external data, and please document any independent validation data in step \<2.4\>.It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modllers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system ), model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames ‚Äî hence, why ecologists often model in the first place (Grimm, et al.
-2014.
-Independent predictions can instead be tested on submodels.
-Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014).
-If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.*
+## Model output corroboration
 
-## 5.2 Choose performance metrics and criteria
+::: callout-note
+## Explanation
 
-*Model performance can be quantified by a range of tests, including measures of agreement between predictions and independent observations, or estimates of accuracy, bias, calibration, discrimination refinement, resolution and skill [@Araujo2019].
-Specify what performance measures you will use to evaluate the model and briefly explain how each test relates to different desired properties of a model's performance.Spatial, temporal and environmental pattern of errors and variance can change the interpretation of model predictions and conservation decisions [@Araujo2019], where relevant and possible, describe how you will characterise and report the spatial, temporal and environmental pattern of errors and variance.
-If comparing alternative models, specify what measures of model comparison or out-of-sample performance metrics will you use to find support for alternative models or else to optimise predictive ability.
-State what numerical threshold or qualities you will use for each of these metrics.*
+Ideally, model outputs or predictions are compared to independent data and patterns that were not used to develop, parameterise, or verify the model.
+Testing against a dataset of response and predictor variables that are spatially and/or temporally independent from the training dataset minimises the risk of artificially inflating model performance measures [@Araujo2019].
+Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system, model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames, hence, why ecologists often model in the first place \[@Grimm:2014es\].
+Independent predictions might instead be tested on sub-models.
+Alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model output is [@Grimm:2014es].
+:::
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] State whether you will corroborate the model outputs on external data, and document any independent validation data in step.
+-   [ ] It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building [@Dwork2015], describe the relationship modellers have to model validation data, e.g.
+will independent datasets be known to any modeller or analyst involved in the model building process?
+-   [ ] If unable to evaluate the model outputs against independent data, explain why and explain what steps you will take to interrogate the model.
+:::
+
+## Choose performance metrics and criteria
+
+::: callout-note
+## Explanation
+
+Model performance can be quantified by a range of tests, including measures of agreement between predictions and independent observations, or estimates of accuracy, bias, calibration, discrimination refinement, resolution and skill [@Araujo2019].
+Note that the performance metrics and criteria in this section are used for evaluating the structured and parameterised models (ideally) on independent holdout data, so this step is additional to any performance criteria used for determining model structure or parameterisation in section X.
+:::
+
+::: callout-caution
+## Preregistration Item
+
+-   [ ] Specify what performance measures you will use to evaluate the model and briefly explain how each test relates to different desired properties of a model's performance.
+-   [ ] Spatial, temporal and environmental pattern of errors and variance can change the interpretation of model predictions and conservation decisions [@Araujo2019], where relevant and possible, describe how you will characterise and report the spatial, temporal and environmental pattern of errors and variance.
+-   [ ] If comparing alternative models, specify what measures of model comparison or out-of-sample performance metrics will you use to find support for alternative models or else to optimise predictive ability.
+State what numerical threshold or qualities you will use for each of these metrics.
+:::
 

### 2023-10-15 adaptive_preregistration_implementation.qmd
Add fledge:: info to adaptive prereg implementation page

@@
 GitHub releases mark software iterations that you can release and deploy[^4].
-Within the context of adaptive preregistration, we can use releases to mark interim versions of the preregistration, as well as to take a snapshot of the entire repository at that moment in time.
+Within the context of adaptive preregistration, we can use releases to mark interim versions of the preregistration, as well as to take a snapshot of the state of the entire repository at that moment in time.
 
+A summary of the types of preregistration changes, and the corresponding required version and GitHub features are presented in @tbl-preregistration-versioning.
 Note that anything up until the complete preregistration should be additionally tagged as a 'pre-release' to communicate that the initial preregistration is not yet complete.
 
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
 | Preregistration Change or Action                                                                                                                                                                                                                                                                                       | GitHub Mechanism                                                                                                                                                                                                                                                         | Semantic Versioning                                                                                      |
-|-----------------------------|-------------------------|-------------------|
++========================================================================================================================================================================================================================================================================================================================+==========================================================================================================================================================================================================================================================================+==========================================================================================================+
 | **Changes preceding initial completion of preregistration**                                                                                                                                                                                                                                                            | Tag and Mark as 'pre-release'                                                                                                                                                                                                                                            | `Minor` increment.
 E.g.
-`0.2.0` The `Major` value should not exceed `0`.
-|
+`0.2.0` The `Major` value should not exceed `0`.\                                |
+|                                                                                                                                                                                                                                                                                                                        |                                                                                                                                                                                                                                                                          | `fledge::bump_version("minor")`                                                                          |
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
 | **Incremental Change**, e.g.
@@
 `0.2.0` The `Major` value should not exceed `0`.
-|
+`fledge::bump_version("minor")` |
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
 | **Initial Preregistration complete**.
@@
 Is complete after all preliminary and investigatory analyses have been done and the final Preregistration is complete, or is it the first iteration, where no more items can be filled out until some initial analyses are undertaken?
-| Tag and Release                                                                                                                                                                                                                                                         | `Major` increment to `1.0.0`                                             |
-| **Major change to a complete Preregistration** -- no looking or analysing at data                                                                                                                                                                                                                                      | Tag and Release                                                                                                                                                                                                                                                         | `Major` increment `>1.0.0`                                               |
-| **Linking implemented work to the Preregistration Item**                                                                                                                                                                                                                                                               | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
+| Tag and Release                                                                                                                                                                                                                                                          | `Major` increment to `1.0.0`.
+`fledge::bump_version("major"); fledge::finalize_version()`.
+|
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
+| **Major change to a complete Preregistration** -- no looking or analysing at data.
+| Tag and Release                                                                                                                                                                                                                                                          | `Major` increment `>1.0.0`.
+`fledge::bump_version("major"); fledge::finalize_version()`.
+|
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
+| **Linking implemented work to the Preregistration Item**                                                                                                                                                                                                                                                               | Update Preregistration with GH Issue URL and all relevant commit hashes pertaining to that Preregistration Item + Commit.
 Tag and Release that preregistration commit.
 | `Minor` increment.
+`fledge::bump_version("minor")`.
 |
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
 | **Data-dependent changes with registered flexibility** Adapting the Preregistration on the basis of results from preliminary or investigatory analyses                                                                                                                                                                 | Update Preregistration with GH Issue Link and all relevant commit hashes pertaining to that Preregistration Item + Commit.
 Respond to dependent incomplete Preregistration Items or Update existing / already completed Preregistration Items + Commit.
-Tag and Release | `Major` increment.
+Tag and Release.
+| `Major` increment.
+`fledge::bump_version("major")`.
 |
-| **Minor mistakes**                                                                                                                                                                                                                                                                                                     | Changes including small typos, or minor elaborations on existing specified plans                                                                                                                                                                                        | `Patch` increment.
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
+| **Minor mistakes or changes**                                                                                                                                                                                                                                                                                          | Changes including small typos, or minor elaborations on existing specified plans                                                                                                                                                                                         | `Patch` increment.
 E.g.
 `0.1.2`.
-|
+`fledge::bump_version("patch")`                                         |
++------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+----------------------------------------------------------------------------------------------------------+
 
@@
 ### Using the [`fledge::`](https://fledge.cynkra.com) R package for semantic versioning and to document preregistration changes
+
+The package `fledge::` can be used right from the R console to bump the version number, make releases, and auto catalogue a chronological history of changes recorded a `NEWS.md` file that is updated and published with every release.
+A detailed guide to using `fledge::` can be found here <https://fledge.cynkra.com/articles/fledge.html>.
+
+This package is not a mandatory component of the process, however can be a useful tool for implementing semantic versioning, tagging and releasing.
+All of these actions can be implemented in the GitHub browser with a point and click interface.
 

### 2023-10-16 adaptive_preregistration_implementation.qmd
update old images and reflect current github workflow

@@
 -   View the document in GitHub [here](https://github.com/egouldo/VEFMAP_VEG_Stage6/blob/master/analysis/preregistration_template.Rmd)
--   Click on The 'edit document' button: ![Screen Shot 2020-05-10 at 6 29 33 pm](https://user-images.githubusercontent.com/8400682/81494459-791b7080-92ec-11ea-979b-900317b4a9b6.png)
+-   Click on The 'edit document' button: ![](https://github.com/egouldo/VEFMAP_VEG_Stage6/assets/8400682/dd6ee5b6-9595-46e0-8576-7d2a3c7e2c9a)
 -   You will be able to edit the document directly in GitHub, Edit away!
@@
 
-![Screen Shot 2020-05-18 at 12 19 29 pm](https://user-images.githubusercontent.com/8400682/82168463-2b14f700-9902-11ea-95a0-21fe385629bb.png)
+![](https://github.com/egouldo/VEFMAP_VEG_Stage6/assets/8400682/55fe090c-b72c-45f4-808f-039960796dc0.png)
 

### 2023-10-16 preregistration_modelling.qmd
unescape references

@@
 Model features include elements such as the functional form of interactions, data structures, measures used to specify links, any bins or discretisation of continuous variables.
-It is usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller \[@Jakeman:2006ii\].
+It is usually difficult to change fundamental features of a model beyond an early stage of model development, so careful thought and planning here is useful to the modeller [@Jakeman:2006ii].
 However, if changes to these fundamental aspects of the model do need to change, document how and why these choices were made, including any results used to support any changes in the model.
@@
 Specify any treatment of model variables, including whether they are lumped / distributed, lienar / non-linear, stochastic / deterministic [@Jakeman:2006ii]* .
--   [ ] *Provide a rationale for your choices, including why plausible alternatives under consideration were not chosen, and relate your justification bacj to the purpose, objectives, prior knowledge and or logistical constraints specified in the problem formulation phase \[@Jakeman:2006ii\].*
+-   [ ] *Provide a rationale for your choices, including why plausible alternatives under consideration were not chosen, and relate your justification bacj to the purpose, objectives, prior knowledge and or logistical constraints specified in the problem formulation phase [@Jakeman:2006ii].*
 :::
@@
 
-The model validation & evaluation phase comprises a suite of analyses that collectively inform inferences about whether, and under what conditions, a model is suitable to meet its intended purpose \[@Augusiak:2014gz\].
+The model validation & evaluation phase comprises a suite of analyses that collectively inform inferences about whether, and under what conditions, a model is suitable to meet its intended purpose [@Augusiak:2014gz].
 Errors in design and implementation of the model and their implication on the model output are assessed.
@@
 Testing against a dataset of response and predictor variables that are spatially and/or temporally independent from the training dataset minimises the risk of artificially inflating model performance measures [@Araujo2019].
-Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system, model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames, hence, why ecologists often model in the first place \[@Grimm:2014es\].
+Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system, model validation is not always possible because empirical experiments are infeasible or model users are working on rapid-response time-frames, hence, why ecologists often model in the first place [@Grimm:2014es].
 Independent predictions might instead be tested on sub-models.

### 2023-10-16 adaptive_preregistration_implementation.qmd
update image links

@@
 -   View the document in GitHub [here](https://github.com/egouldo/VEFMAP_VEG_Stage6/blob/master/analysis/preregistration_template.Rmd)
--   Click on The 'edit document' button: ![](https://github.com/egouldo/VEFMAP_VEG_Stage6/assets/8400682/dd6ee5b6-9595-46e0-8576-7d2a3c7e2c9a)
+-   Click on The 'edit document' button: ![](https://private-user-images.githubusercontent.com/8400682/275362145-dd6ee5b6-9595-46e0-8576-7d2a3c7e2c9a.png?jwt=eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJpc3MiOiJnaXRodWIuY29tIiwiYXVkIjoicmF3LmdpdGh1YnVzZXJjb250ZW50LmNvbSIsImtleSI6ImtleTEiLCJleHAiOjE2OTc0MTYwNTMsIm5iZiI6MTY5NzQxNTc1MywicGF0aCI6Ii84NDAwNjgyLzI3NTM2MjE0NS1kZDZlZTViNi05NTk1LTQ2ZTAtODU3Ni03ZDJhM2M3ZTJjOWEucG5nP1gtQW16LUFsZ29yaXRobT1BV1M0LUhNQUMtU0hBMjU2JlgtQW16LUNyZWRlbnRpYWw9QUtJQUlXTkpZQVg0Q1NWRUg1M0ElMkYyMDIzMTAxNiUyRnVzLWVhc3QtMSUyRnMzJTJGYXdzNF9yZXF1ZXN0JlgtQW16LURhdGU9MjAyMzEwMTZUMDAyMjMzWiZYLUFtei1FeHBpcmVzPTMwMCZYLUFtei1TaWduYXR1cmU9MmYwZWI5MDVjYjhhMzJmYTdiNTcwNDY4ZjAwZWEyNTcwNDg4NThiMTY2YjdjY2FjZTliZWY1MmFkNzJmNmVkNSZYLUFtei1TaWduZWRIZWFkZXJzPWhvc3QmYWN0b3JfaWQ9MCZrZXlfaWQ9MCZyZXBvX2lkPTAifQ.kwCBfZTeofYDbLAPZqW2YsBcs12bQUdntdqcFnB_UJw)
 -   You will be able to edit the document directly in GitHub, Edit away!
@@
 
--   If you are finished, or ready to save your changes by 'committing' them, commit by scrolling underneath the document editing window.
-There will be a box called "commit changes", enter in the commit message box the following:
--   Your message should contain the following: `Initial Preregistration #IssueNumber`. Tag the Preregistration GitHub issue number in the commit message (For this repository, it's `#17`).
+-   If you are finished, or ready to save your changes by 'committing' them.
+There is a box called "commit changes" at the top right of the window, a pop-up window will appear, enter in the commit message box the following:
+    -    `Initial Preregistration #IssueNumber`. Tag the Preregistration GitHub issue number in the commit message (For this repository, it's `#17`).
 -   Fill out the additional description box briefly describing the changes you just made
--   Click "commit"
+-   Ensure that the radio button for creating a new branch is selected, an auto-generated branch name will appear, e.g. `egould-patch-1`.
+-   Click "Propose Changes" to start a [pull request](https://docs.github.com/en/get-started/quickstart/github-flow):
 
-![](https://github.com/egouldo/VEFMAP_VEG_Stage6/assets/8400682/55fe090c-b72c-45f4-808f-039960796dc0.png)
+![](https://private-user-images.githubusercontent.com/8400682/275362469-55fe090c-b72c-45f4-808f-039960796dc0.png?jwt=eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJpc3MiOiJnaXRodWIuY29tIiwiYXVkIjoicmF3LmdpdGh1YnVzZXJjb250ZW50LmNvbSIsImtleSI6ImtleTEiLCJleHAiOjE2OTc0MTYwNTMsIm5iZiI6MTY5NzQxNTc1MywicGF0aCI6Ii84NDAwNjgyLzI3NTM2MjQ2OS01NWZlMDkwYy1iNzJjLTQ1ZjQtODA4Zi0wMzk5NjA3OTZkYzAucG5nP1gtQW16LUFsZ29yaXRobT1BV1M0LUhNQUMtU0hBMjU2JlgtQW16LUNyZWRlbnRpYWw9QUtJQUlXTkpZQVg0Q1NWRUg1M0ElMkYyMDIzMTAxNiUyRnVzLWVhc3QtMSUyRnMzJTJGYXdzNF9yZXF1ZXN0JlgtQW16LURhdGU9MjAyMzEwMTZUMDAyMjMzWiZYLUFtei1FeHBpcmVzPTMwMCZYLUFtei1TaWduYXR1cmU9YjdiNzE0N2Q3NmNlZGIzYWEyM2E0OGM2NzZkZTk4NTNhNTFlODI5MmI0YzhjMDBlYzIwZjk4OWQxZTYzNmY2MSZYLUFtei1TaWduZWRIZWFkZXJzPWhvc3QmYWN0b3JfaWQ9MCZrZXlfaWQ9MCZyZXBvX2lkPTAifQ.kCUf1yhQJRgKtqYUtj4SBV9LULgDUohqeSj-GSQXQ3o)
 
@@
 Note that once you've highlighted some lines, a little box with an ellipsis appears on top of the lines numbers.
-![Screen Shot 2020-05-13 at 1 06 25 pm](https://user-images.githubusercontent.com/8400682/81767208-2d9ad980-951b-11ea-8f59-6d6f583eb155.png)
+![Screen Shot 2020-05-13 at 1 06 25 pm](https://user-images.githubusercontent.com/8400682/81767208-2d9ad980-951b-11ea-8f59-6d6https://private-user-images.githubusercontent.com/8400682/275363216-f0c6b90a-bda6-47d7-b91a-fd1d493153de.png?jwt=eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJpc3MiOiJnaXRodWIuY29tIiwiYXVkIjoicmF3LmdpdGh1YnVzZXJjb250ZW50LmNvbSIsImtleSI6ImtleTEiLCJleHAiOjE2OTc0MTYwNTMsIm5iZiI6MTY5NzQxNTc1MywicGF0aCI6Ii84NDAwNjgyLzI3NTM2MzIxNi1mMGM2YjkwYS1iZGE2LTQ3ZDctYjkxYS1mZDFkNDkzMTUzZGUucG5nP1gtQW16LUFsZ29yaXRobT1BV1M0LUhNQUMtU0hBMjU2JlgtQW16LUNyZWRlbnRpYWw9QUtJQUlXTkpZQVg0Q1NWRUg1M0ElMkYyMDIzMTAxNiUyRnVzLWVhc3QtMSUyRnMzJTJGYXdzNF9yZXF1ZXN0JlgtQW16LURhdGU9MjAyMzEwMTZUMDAyMjMzWiZYLUFtei1FeHBpcmVzPTMwMCZYLUFtei1TaWduYXR1cmU9ODA2NGFjNGFmZmYwODQxM2I3MTczMDJhOTViMWI3MTJkZjBiZDkzNzgxNDYzNDFiNzFhNGE4NTBjNDNkODBjMSZYLUFtei1TaWduZWRIZWFkZXJzPWhvc3QmYWN0b3JfaWQ9MCZrZXlfaWQ9MCZyZXBvX2lkPTAifQ.coX8tF1h--QoECLhsP-TtRES3p3SHabBnR2W1JEz2jgf583eb155.png)
 2.
 Click on the tiny box with the ellipsis, and a pop-up dialogue box will appear with 4 options: `Copy lines`, `Copy permalink`, `View git blame`, `References in new issue`.
-Select the last option `Reference in new issue.` ![Screen Shot 2020-05-13 at 1 06 58 pm](https://user-images.githubusercontent.com/8400682/81767344-80749100-951b-11ea-8b76-1e829528973a.png)
+Select the last option `Reference in new issue.` ![Screen Shot 2020-05-13 at 1 06 58 pm](https://private-user-images.githubusercontent.com/8400682/275363343-9d10ecac-bcf6-4048-a1be-8ea09ac2e64e.png?jwt=eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJpc3MiOiJnaXRodWIuY29tIiwiYXVkIjoicmF3LmdpdGh1YnVzZXJjb250ZW50LmNvbSIsImtleSI6ImtleTEiLCJleHAiOjE2OTc0MTYwNTMsIm5iZiI6MTY5NzQxNTc1MywicGF0aCI6Ii84NDAwNjgyLzI3NTM2MzM0My05ZDEwZWNhYy1iY2Y2LTQwNDgtYTFiZS04ZWEwOWFjMmU2NGUucG5nP1gtQW16LUFsZ29yaXRobT1BV1M0LUhNQUMtU0hBMjU2JlgtQW16LUNyZWRlbnRpYWw9QUtJQUlXTkpZQVg0Q1NWRUg1M0ElMkYyMDIzMTAxNiUyRnVzLWVhc3QtMSUyRnMzJTJGYXdzNF9yZXF1ZXN0JlgtQW16LURhdGU9MjAyMzEwMTZUMDAyMjMzWiZYLUFtei1FeHBpcmVzPTMwMCZYLUFtei1TaWduYXR1cmU9MmRiYjcxM2MwYjZmMWEyNDVlZmY2M2ZhZjg4YWFmMjU2Yzk2NjdkYmJiZjA4MjRkYjY4Y2EyODU2ZTEyNjU4ZCZYLUFtei1TaWduZWRIZWFkZXJzPWhvc3QmYWN0b3JfaWQ9MCZrZXlfaWQ9MCZyZXBvX2lkPTAifQ.yoaxWIHOaaE5bTN8ClYogrHKU6bQYk3n0qjpkJb2Ksw)
 3.
 This will open up a new window with a new GitHub Issue for you to complete.
-Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted: ![Screen Shot 2020-05-13 at 1 07 30 pm](https://user-images.githubusercontent.com/8400682/81767518-ef51ea00-951b-11ea-9398-18fffdcbbeae.png) Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this: ![Screen Shot 2020-05-13 at 1 17 21 pm](https://user-images.githubusercontent.com/8400682/81767585-1c9e9800-951c-11ea-9d2c-46d2faf6368a.png)
+Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted: ![Screen Shot 2020-05-13 at 1 07 30 pm](https://private-user-images.githubusercontent.com/8400682/275363562-b9daee6c-ec56-44c0-a848-9b4412bf8adc.png?jwt=eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJpc3MiOiJnaXRodWIuY29tIiwiYXVkIjoicmF3LmdpdGh1YnVzZXJjb250ZW50LmNvbSIsImtleSI6ImtleTEiLCJleHAiOjE2OTc0MTYwNTMsIm5iZiI6MTY5NzQxNTc1MywicGF0aCI6Ii84NDAwNjgyLzI3NTM2MzU2Mi1iOWRhZWU2Yy1lYzU2LTQ0YzAtYTg0OC05YjQ0MTJiZjhhZGMucG5nP1gtQW16LUFsZ29yaXRobT1BV1M0LUhNQUMtU0hBMjU2JlgtQW16LUNyZWRlbnRpYWw9QUtJQUlXTkpZQVg0Q1NWRUg1M0ElMkYyMDIzMTAxNiUyRnVzLWVhc3QtMSUyRnMzJTJGYXdzNF9yZXF1ZXN0JlgtQW16LURhdGU9MjAyMzEwMTZUMDAyMjMzWiZYLUFtei1FeHBpcmVzPTMwMCZYLUFtei1TaWduYXR1cmU9YzQyNTgyOTE0ZmQ0MmM3Zjk3YzA1ZTRkNTMzNmI1ZmE5Mjc5ODFmNTMzNWRkYmY3YTYyYmQ1N2NlMzM4NzIxZiZYLUFtei1TaWduZWRIZWFkZXJzPWhvc3QmYWN0b3JfaWQ9MCZrZXlfaWQ9MCZyZXBvX2lkPTAifQ.hwuMj_cYQgK1Kzqaf4lBlU4kJHxG8QWaXjLEKJ5_wZI) Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this: ![Screen Shot 2020-05-13 at 1 17 21 pm](https://private-user-images.githubusercontent.com/8400682/275363581-3729b261-d492-4b09-ac6a-6392aa4dfc25.png?jwt=eyJhbGciOiJIUzI1NiIsInR5cCI6IkpXVCJ9.eyJpc3MiOiJnaXRodWIuY29tIiwiYXVkIjoicmF3LmdpdGh1YnVzZXJjb250ZW50LmNvbSIsImtleSI6ImtleTEiLCJleHAiOjE2OTc0MTYwNTMsIm5iZiI6MTY5NzQxNTc1MywicGF0aCI6Ii84NDAwNjgyLzI3NTM2MzU4MS0zNzI5YjI2MS1kNDkyLTRiMDktYWM2YS02MzkyYWE0ZGZjMjUucG5nP1gtQW16LUFsZ29yaXRobT1BV1M0LUhNQUMtU0hBMjU2JlgtQW16LUNyZWRlbnRpYWw9QUtJQUlXTkpZQVg0Q1NWRUg1M0ElMkYyMDIzMTAxNiUyRnVzLWVhc3QtMSUyRnMzJTJGYXdzNF9yZXF1ZXN0JlgtQW16LURhdGU9MjAyMzEwMTZUMDAyMjMzWiZYLUFtei1FeHBpcmVzPTMwMCZYLUFtei1TaWduYXR1cmU9NzBlM2I4NzFiODVmNTZjZDY4NTFhZGZjOGFkY2NiNmY3Mjg4ZTA1MDZjZjhlMjVlMjFlZDFiOTEwZmViZWVmYyZYLUFtei1TaWduZWRIZWFkZXJzPWhvc3QmYWN0b3JfaWQ9MCZrZXlfaWQ9MCZyZXBvX2lkPTAifQ.ROGUCBIOafjv1U6wlaubKkTtowNzpr7rxgeSQjYgdGk)
 

### 2023-10-16 adaptive_preregistration_implementation.qmd
replace images with local files bc links not rendering. rerender

--- 2023-10-16 adaptive_preregistration_implementation.qmd
@@
 -   View the document in GitHub [here](https://github.com/egouldo/VEFMAP_VEG_Stage6/blob/master/analysis/preregistration_template.Rmd)
--   Click on The 'edit document' button: ![](https://github.com/egouldo/VEFMAP_VEG_Stage6/assets/8400682/dd6ee5b6-9595-46e0-8576-7d2a3c7e2c9a)
+-   Click on The 'edit document' button: ![](figures/edit-document.png)
 -   You will be able to edit the document directly in GitHub, Edit away!
@@
 
--   If you are finished, or ready to save your changes by 'committing' them, commit by scrolling underneath the document editing window.
-There will be a box called "commit changes", enter in the commit message box the following:
--   Your message should contain the following: `Initial Preregistration #IssueNumber`. Tag the Preregistration GitHub issue number in the commit message (For this repository, it's `#17`).
+-   If you are finished, or ready to save your changes by 'committing' them.
+There is a box called "commit changes" at the top right of the window, a pop-up window will appear, enter in the commit message box the following:
+    -    `Initial Preregistration #IssueNumber`. Tag the Preregistration GitHub issue number in the commit message (For this repository, it's `#17`).
 -   Fill out the additional description box briefly describing the changes you just made
--   Click "commit"
+-   Ensure that the radio button for creating a new branch is selected, an auto-generated branch name will appear, e.g. `egould-patch-1`.
+-   Click "Propose Changes" to start a [pull request](https://docs.github.com/en/get-started/quickstart/github-flow):
 
-![](https://github.com/egouldo/VEFMAP_VEG_Stage6/assets/8400682/55fe090c-b72c-45f4-808f-039960796dc0.png)
+![](figures/propose-changes.png)
 
@@
 Note that once you've highlighted some lines, a little box with an ellipsis appears on top of the lines numbers.
-![Screen Shot 2020-05-13 at 1 06 25 pm](https://user-images.githubusercontent.com/8400682/81767208-2d9ad980-951b-11ea-8f59-6d6f583eb155.png)
+![](figures/highlight-lines.png)
 2.
 Click on the tiny box with the ellipsis, and a pop-up dialogue box will appear with 4 options: `Copy lines`, `Copy permalink`, `View git blame`, `References in new issue`.
-Select the last option `Reference in new issue.` ![Screen Shot 2020-05-13 at 1 06 58 pm](https://user-images.githubusercontent.com/8400682/81767344-80749100-951b-11ea-8b76-1e829528973a.png)
+Select the last option `Reference in new issue.` ![Screen Shot 2020-05-13 at 1 06 58 pm](figures/reference-in-new-issue.png)
 3.
 This will open up a new window with a new GitHub Issue for you to complete.
-Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted: ![Screen Shot 2020-05-13 at 1 07 30 pm](https://user-images.githubusercontent.com/8400682/81767518-ef51ea00-951b-11ea-9398-18fffdcbbeae.png) Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this: ![Screen Shot 2020-05-13 at 1 17 21 pm](https://user-images.githubusercontent.com/8400682/81767585-1c9e9800-951c-11ea-9d2c-46d2faf6368a.png)
+Note that part of the issue description has already been filled out for you, this is a reference to the lines you just highlighted: ![](figures/new-issue.png) Which is cool, because if you open the `Preview` tab to view the preview of your issue, it will look like this: ![](figures/preview-new-issue.png)
 

### 2024-01-24 preregistration_modelling.qmd

@@
 
--   How will you operationalise any scenarios identified in 1.1.3?
-    For example, how will you operationalise any qualitative changes of interest, such as ‚ 'deterioration' or 'improvement'?
+-   How will you operationalise any scenarios identified in 1.1.3? For example, how will you operationalise any qualitative changes of interest, such as ‚ 'deterioration' or 'improvement'?
 
@@
 
-Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be represented in the model [@Moallemi2019].
-Sources of uncertainty may include:
+Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be represented in the model [@Moallemi2019]. Sources of uncertainty may include:
 
@@
 
--   [ ] How will outliers be handled?
-    Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.
+-   [ ] How will outliers be handled? Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.
 
@@
 
-    -   [ ] Which data will be used as a the testing data?
-        What method will you be used for generating training / test data subsets?
+    -   [ ] Which data will be used as a the testing data? What method will you be used for generating training / test data subsets?
 
@@
 
-::: callout-note
+::: callout-caution
 ## Preregistration Item
 
-*During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences.*
+*During this process, observed data, or data and patterns that guided model design and calibration, are compared to model output in order to identify if and where there are any systematic differences.*
 

### 2024-03-12 index.qmd
Update index.qmd
Change name of adaptive preregistration from background to overview

@@
 2.  Adaptive Preregistration
-    1.  [*Background*](adaptive_preregistration_background.qmd): What is adaptive preregistration, why do we need it for registering modeling analyses, and what are its defining features?
+    1.  [*Overview*](adaptive_preregistration_background.qmd): What is adaptive preregistration, why do we need it for registering modeling analyses, and what are its defining features?
     2.  [*A practical guide*](adaptive_preregistration_implementation.qmd): How to implement adaptive preregistration using [git](https://git-scm.com) and [GitHub](www.github.com).

### 2024-03-12 preregistration_modelling.qmd
template typo and formatting updates
Signed-off-by: egouldo <egouldo@users.noreply.github.com>

@@
 ::: callout-caution
+## Preregistration Item
+
 -   [ ] What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?
@@
 
-### Explain how you will operationalise response and predictor variables, and covariates
+### Operationalising Model Variables
 
@@
 
+:::
+
+::: callout-caution
+
 ## Preregistration Item
@@
 ::: callout-caution
+## Preregistration Item
+
 The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019].

### 2024-03-12 preregistration_modelling.qmd
Add Jian as author to template

@@
   - Libby Rumpff
+  - Jian Yen
   - Megan Good
@@
 The following items pertain to properties of the *datasets* used for calibration (training), validation, and testing.
-
 :::
@@
 ::: callout-caution
-
 ## Preregistration Item

### 2024-03-14 _analysis/ms.qmd
- Add elsevier template and paper metadata

---
title: "But I can't preregister my research: Improving the reproducibility and transparency of ecology and conservation with adaptive preregistration for model-based research"
author:
  - id: aut1
    name:
      given: Elliot
      family: Gould
    email: elliot.gould@unimelb.edu.au
    orcid: 0000-0002-6585-538X
    attributes:
      corresponding: true
    roles:
      - conceptualization
      - data curation
      - formal analysis
      - funding acquisition
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - validation
      - visualization
      - writing - original draft
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut2
    name:
      given: Chris
      family: Jones
    email: chris.jones@delwp.vic.gov.au
    orcid: 0000-0003-2833-0194
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - data curation
      - funding acquisition
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - supervision
      - validation
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut3
    name:
      given: Jian
      family: Yen
    email: jian.yen@delwp.vic.gov.au
    orcid: 0000-0001-7964-923X
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - supervision
      - validation
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut4
    name:
      given: Hannah
      family: Fraser
    email: hannahsfraser@gmail.com
    orcid: 0000-0003-2443-4463
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - investigation
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut5
    name:
      given: Henry
      family: Wootton
    email: henry.wootton@delwp.vic.gov.au
    orcid: 0000-0001-6506-0248
    attributes:
      corresponding: false
    roles:
      - data curation
      - formal analysis
      - investigation
      - methodology
      - resources
      - software
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut6
    name:
      given: Lyndsey
      family: Vivian
    email: lyndsey.vivian@delwp.vic.gov.au
    orcid: 0000-0003-1927-723X
    attributes:
      corresponding: false
    roles:
      - data curation
      - investigation
      - methodology
      - resources
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut7
    name:
      given: Megan
      family: Good
    email: goodm@unimelb.edu.au
    orcid: 0000-0002-6908-1633
    attributes:
      corresponding: false
    roles:
      - investigation
      - methodology
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut8
    name:
      given: David
      family: Duncan
    email: david.duncan@unimelb.edu.au
    orcid: 0000-0003-4411-8214
    attributes:
      corresponding: false
    roles:
      - investigation
      - methodology
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut9
    name:
      given: Libby
      family: Rumpff
    email: lrumpff@unimelb.edu.au
    orcid: 0000-0001-9400-8086
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - investigation
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut10
    name:
      given: Fiona
      family: Fidler
    email: fidlerfm@unimelb.edu.au
    orcid: 0000-0002-2700-2562
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
abstract: |
  **1.** Preregistration is an open-science practice adopted in scientific disciplines that have begun to confront the 'reproducibility crisis', which aims to improve research transparency and mitigate questionable research practices within a study. There has been little uptake of preregistration in ecology and conservation, and many modellers and researchers engaged in non-hypothesis testing research both within and outside of ecology have eschewed preregistration on the grounds that existing templates are irrelevant because they are focussed on null-hypothesis significance testing.
  **2.** We advocate that preregistration can and should be used in model-based research in ecology and conservation but that preregistration should reflect the norms and practice of the research context in which it is applied in order to adequately restrict researcher degrees of freedom that are unique to a domain-specific methodology. We took a user-centred design approach to the task of translating preregistration into model-based research contexts for ecology and conservation.
  **3.** We identified a generalised 'modelling workflow' that captures critical analysis decisions in the process of the model development cycle. The workflow informed the content and structure of a preregistration template to be used in model-based research. The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the current format and linear process of preregistration. To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose a methodology for implementing 'adaptive preregistration' leveraging git and GitHub. We tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
  **4.** This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings, and will facilitate the wider uptake of preregistration in ecology. Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model-based research within science more broadly. Modellers need no longer cry "but, I can't preregister my research."
keywords:
  - preregistration
  - transparency
  - reproducibility
  - modelling
  - ecology
  - conservation
  - open-science
  - research software engineering
affiliations:
  - id: aff1
    name: University of Melbourne
  - id: aff2
    name: Arthur Rylah Institute for Environmental Research
format:
  pdf: default
  elsevier-pdf:
    journal:
      name: Journal Name
      formatting: preprint #review or doubleblind
      model: 3p
      layout: onecolumn
      cite-style: authoryear
      # graphical-abstract: "![](abstract.png)"
      # highlights:
        # - Highlight 1
        # - Highlight 2 
        # - Highlight 3
    keep-tex: true
---

# Introduction

# Materials and Methods

# Results

# Discussion & Conclusions

# Data & Software Availability

# Author Contributions

# Supplementary Material

### 2024-03-14 _analysis/ms/ms.qmd
- Add elsevier template and paper metadata

--- 2024-03-14 _analysis/ms.qmd
@@
 ---
-title: "But I can't preregister my research: Improving the reproducibility and transparency of ecology and conservation with adaptive preregistration for model-based research"
+title: "But I can't preregister my research"
+subtitle: "Improving the reproducibility and transparency of ecology and conservation with adaptive preregistration for model-based research"
 author:
@@
       - ref: aff1
+affiliations:
+  - id: aff1
+    name: MetaMelb Research, University of Melbourne
+  - id: aff2
+    name: Arthur Rylah Institute for Environmental Research
 abstract: |
@@
   - research software engineering
-affiliations:
-  - id: aff1
-    name: University of Melbourne
-  - id: aff2
-    name: Arthur Rylah Institute for Environmental Research
+date: last-modified
+bibliography: ../../references.bib
 format:
-  pdf: default
   elsevier-pdf:
+    keep-tex: true
     journal:
       name: Journal Name
-      formatting: preprint #review or doubleblind
+      formatting: preprint
       model: 3p
-      layout: onecolumn
-      cite-style: authoryear
-      # graphical-abstract: "![](abstract.png)"
-      # highlights:
-        # - Highlight 1
-        # - Highlight 2 
-        # - Highlight 3
-    keep-tex: true
+      cite-style: super
 ---
 
-# Introduction
+Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal.
+It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
 
-# Materials and Methods
+For detailed instructions regarding the elsevier article class, see <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>
 
-# Results
+# Bibliography styles
 
-# Discussion & Conclusions
+Here are two sample references: @Feynman1963118 @Dirac1953888.
 
-# Data & Software Availability
+By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. You can sets extra options with `natbiboptions` variable in YAML header. Example
 
-# Author Contributions
+```         
+natbiboptions: longnamesfirst,angle,semicolon
+```
 
-# Supplementary Material
+There are various more specific bibliography styles available at <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>.
+To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.
+
+## Using CSL
+
+If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`.
+In this case, the `csl` option is used to format the references.
+By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>.
+These can be downloaded and stored locally, or the url can be used as in the example header.
+
+# Equations
+
+Here is an equation: $$ 
+  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
+  \left(\frac{x}{\beta}\right)^{\alpha-1}
+  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
+  \alpha,\beta,x > 0 .
+$$
+
+Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$
+
+# Figures and tables
+
+@fig-meaningless is generated using an R chunk.
+
+```{r}
+#| label: fig-meaningless
+#| fig-cap: A meaningless scatterplot
+#| fig-width: 5
+#| fig-height: 5
+#| fig-align: center
+#| out-width: 50%
+#| echo: false
+plot(runif(25), runif(25))
+```
+
+# Tables coming from R
+
+Tables can also be generated using R chunks, as shown in @tbl-simple example.
+
+```{r}
+#| label: tbl-simple
+#| tbl-cap: Caption centered above table
+#| echo: true
+knitr::kable(head(mtcars)[,1:4])
+```
+
+# References {.unnumbered}

### 2024-03-14 _analysis/ms/ms.tex
- Add elsevier template and paper metadata

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={But I can't preregister my research},
  pdfauthor={Elliot Gould; Chris Jones; Jian Yen; Hannah Fraser; Henry Wootton; Lyndsey Vivian; Megan Good; David Duncan; Libby Rumpff; Fiona Fidler},
  pdfkeywords={preregistration, transparency, reproducibility, modelling, ecology, conservation, open-science, research
software engineering},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{But I can't preregister my research \\\large{Improving the
reproducibility and transparency of ecology and conservation with
adaptive preregistration for model-based research} }
\author[1]{Elliot Gould%
\corref{cor1}%
}
 \ead{elliot.gould@unimelb.edu.au} 
\author[2]{Chris Jones%
%
}
 \ead{chris.jones@delwp.vic.gov.au} 
\author[2]{Jian Yen%
%
}
 \ead{jian.yen@delwp.vic.gov.au} 
\author[1]{Hannah Fraser%
%
}
 \ead{hannahsfraser@gmail.com} 
\author[2]{Henry Wootton%
%
}
 \ead{henry.wootton@delwp.vic.gov.au} 
\author[2]{Lyndsey Vivian%
%
}
 \ead{lyndsey.vivian@delwp.vic.gov.au} 
\author[1]{Megan Good%
%
}
 \ead{goodm@unimelb.edu.au} 
\author[1]{David Duncan%
%
}
 \ead{david.duncan@unimelb.edu.au} 
\author[1]{Libby Rumpff%
%
}
 \ead{lrumpff@unimelb.edu.au} 
\author[1]{Fiona Fidler%
%
}
 \ead{fidlerfm@unimelb.edu.au}

\affiliation[1]{organization={MetaMelb Research, University of
Melbourne},,postcodesep={}}
\affiliation[2]{organization={Arthur Rylah Institute for Environmental
Research},,postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
\textbf{1.} Preregistration is an open-science practice adopted in
scientific disciplines that have begun to confront the `reproducibility
crisis', which aims to improve research transparency and mitigate
questionable research practices within a study. There has been little
uptake of preregistration in ecology and conservation, and many
modellers and researchers engaged in non-hypothesis testing research
both within and outside of ecology have eschewed preregistration on the
grounds that existing templates are irrelevant because they are focussed
on null-hypothesis significance testing. \textbf{2.} We advocate that
preregistration can and should be used in model-based research in
ecology and conservation but that preregistration should reflect the
norms and practice of the research context in which it is applied in
order to adequately restrict researcher degrees of freedom that are
unique to a domain-specific methodology. We took a user-centred design
approach to the task of translating preregistration into model-based
research contexts for ecology and conservation. \textbf{3.} We
identified a generalised `modelling workflow' that captures critical
analysis decisions in the process of the model development cycle. The
workflow informed the content and structure of a preregistration
template to be used in model-based research. The iterative nature of
model development was identified as a key barrier to implementing
preregistration for modelling research because it conflicts with the
current format and linear process of preregistration. To better align
the internal logic of preregistration with the iterative and non-linear
process of ecological modelling, we propose a methodology for
implementing `adaptive preregistration' leveraging git and GitHub. We
tested and evaluated the preregistration template and methodology using
a case-study of environmental flows modelling for decision-support in
Victoria, Australia. \textbf{4.} This research provides a template and
methodology for implementing adaptive preregistration of ecological
models in either pure or applied settings, and will facilitate the wider
uptake of preregistration in ecology. Although we focus on ecology and
conservation in this paper, the template is extensible to alternative
fields and methods, and the proposed methodology for implementing
adaptive preregistration can be applied in model-based research within
science more broadly. Modellers need no longer cry ``but, I can't
preregister my research.''
\end{abstract}

\begin{keyword}
    preregistration \sep transparency \sep reproducibility \sep modelling \sep ecology \sep conservation \sep open-science \sep 
    research software engineering
\end{keyword}
\end{frontmatter}

Please make sure that your manuscript follows the guidelines in the
Guide for Authors of the relevant journal. It is not necessary to
typeset your manuscript in exactly the same way as an article, unless
you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see
\url{https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions}

\section{Bibliography styles}\label{bibliography-styles}

Here are two sample references: \citet{Feynman1963118}
\citet{Dirac1953888}.

By default, natbib will be used with the \texttt{authoryear} style, set
in \texttt{classoption} variable in YAML. You can sets extra options
with \texttt{natbiboptions} variable in YAML header. Example

\begin{verbatim}
natbiboptions: longnamesfirst,angle,semicolon
\end{verbatim}

There are various more specific bibliography styles available at
\url{https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files}.
To use one of these, add it in the header using, for example,
\texttt{biblio-style:\ model1-num-names}.

\subsection{Using CSL}\label{using-csl}

If \texttt{cite-method} is set to \texttt{citeproc} in
\texttt{elsevier\_article()}, then pandoc is used for citations instead
of \texttt{natbib}. In this case, the \texttt{csl} option is used to
format the references. By default, this template will provide an
appropriate style, but alternative \texttt{csl} files are available from
\url{https://www.zotero.org/styles?q=elsevier}. These can be downloaded
and stored locally, or the url can be used as in the example header.

\section{Equations}\label{equations}

Here is an equation: \[ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
\]

Inline equations work as well: \(\sum_{i = 2}^\infty\{\alpha_i^\beta\}\)

\section{Figures and tables}\label{figures-and-tables}

Figure~\ref{fig-meaningless} is generated using an R chunk.

\begin{figure}

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{ms_files/figure-pdf/fig-meaningless-1.pdf}

}

\caption{\label{fig-meaningless}A meaningless scatterplot}

\end{figure}%

\section{Tables coming from R}\label{tables-coming-from-r}

Tables can also be generated using R chunks, as shown in
Table~\ref{tbl-simple} example.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{kable}\NormalTok{(}\FunctionTok{head}\NormalTok{(mtcars)[,}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{])}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-simple}Caption centered above table}

\tabularnewline

\toprule\noalign{}
& mpg & cyl & disp & hp \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Mazda RX4 & 21.0 & 6 & 160 & 110 \\
Mazda RX4 Wag & 21.0 & 6 & 160 & 110 \\
Datsun 710 & 22.8 & 4 & 108 & 93 \\
Hornet 4 Drive & 21.4 & 6 & 258 & 110 \\
Hornet Sportabout & 18.7 & 8 & 360 & 175 \\
Valiant & 18.1 & 6 & 225 & 105 \\

\end{longtable}

\renewcommand\refname{References}
  \bibliography{../../references.bib}

\end{document}

### 2024-03-27 _analysis/EcoConsPreReg-supervisor_draft-Sep-2020.Rmd
reformat manuscript draft .rmd
- fix rel filepath to references.bib

@@
 date: "`r Sys.Date()`"
-bibliography: ./data/references.bib
+bibliography: ../data/references.bib
 output: #
@@
 
-@TODO: add argument from Berna's formal theory paper about the need for testing and evaluating new reforms.
-While only a qualitative assessment... still useful information, such as XYZ.
+@TODO: add argument from Berna's formal theory paper about the need for testing and evaluating new reforms. While only a qualitative assessment... still useful information, such as XYZ.
 
-Mention testing / eval here?
-Yes need to mention a little bit, BUT perhaps could save Berna's quote for discussion.
+Mention testing / eval here? Yes need to mention a little bit, BUT perhaps could save Berna's quote for discussion.
 
@@
 This template was prepared earlier (\<specific OSF link to the document\>), and consisted of several key 'phases' of the scientific process, with decision-steps grouped under each phase.
-The development of the draft modelling workflow was derived from previous work by Gould et al. *in prep.,* whilst the field sampling workflow was created specifically for this exercise.
+The development of the draft modelling workflow was derived from previous work by Gould et al.
+*in prep.,* whilst the field sampling workflow was created specifically for this exercise.
 We sorted and categorised each decision-step as belonging to a particular workflow phase.
@@
 We conducted two lots of structured discussions, one for each of the field and modelling template groups.
-Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature \<cite / link to OSF, or even discuss??\>
-.
+Each discussion was structured using a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature \<cite / link to OSF, or even discuss??\> .
 These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in activity 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.
@@
 A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on 'good modelling practice' and structured decision-making literature in ecological management.
-Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], \<insert link to appendices or just present the final?\>
-.
+Steps in the workflow were coded and broken down into 'phases' consisting of 'steps' and 'substeps', in line with the ontology we specified for translating a modelling workflow into a preregistration template [section link], \<insert link to appendices or just present the final?\> .
 Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and substeps.
@@
 
-1.  Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases.
-    \<Note, that these are being scrapped from the template and are turned into tool-tips\>
+1.  Decision Phase - each has a "title" and "description" or "definition", steps are grouped under phases. \<Note, that these are being scrapped from the template and are turned into tool-tips\>
 
-2.  Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow.
+2.
+Decision Step - each has a "title" and "description" or "definition" choices are grouped into 'steps', which corresponds to an activity or set of related activities in a scientific workflow.
 \<again, these are being scrapped and turned into tool-tips.\>
 
-3.  Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
+3.
+Choices ---\> each 'choice' corresponds to a uniquely numbered item on the PRT requiring a response from user.
 The style should be in the form of a directive or a question.
-    E.g. "explain how you will do xyzzy".
+E.g.
+"explain how you will do xyzzy".
 Or "What performance measure will you use for assessing 'goodness of fit'?"
@@
 
-@TODO - anything about *applied contexts??* Aligning with management?
-(In terms of the structure and steps above).
+@TODO - anything about *applied contexts??* Aligning with management? (In terms of the structure and steps above).
 
@@
 Secondly, the preregistration may be iterative and consist of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observed \<Srivastava\>.
-Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, \<x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?i
-d=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0\>\>.
+Thus, as the modeller proceeds through the model development process, the researcher can shift from ideation, preregistration, execution of the analysis and back to ideation and preregistration again, depending on the observed outcomes as each interim preregistration is executed \<Link to Figure, GitHub as a platform for adaptive preregistration, \<x-scrivener-item:///Users/egould/Dropbox/Apps/Scrivener/PhD_thesis.scriv?i d=32FCFBDA-7CEE-4FD9-B8DD-F3073309C0D0\>\>.
 
@@
 Each issue is assigned a unique number.
-Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the *issue's* thread (Fig X, step 4, Do) \<screen-cap?\>
-.
+Any code or analysis outputs, such as figures, tables or other files, should be committed and tagged with the issue number, so that all analysis addressing that task is tracked within the *issue's* thread (Fig X, step 4, Do) \<screen-cap?\> .
 The URL for the GitHub issue should then be added to the relevant preregistration item, the preregistration document should then be committed with the changes, and the version number of the preregistration updated in a new *release*.
@@
 We considered removing in-template text describing and defining key concepts, phases and steps, as well as reducing the number of decision-points in a model-based research workflow captured by the template by removing items within the template.
-\<@TODO - did we end up removing any items by the end of the preregistration and evaluation with Chris' study?\>
-While we 'tweaked' the template items and descriptive text during the in situ evaluation of the templates for the environmental flows preregistration, we decided not to alter the template structure and content significantly, but to instead implement a new format and interface for working with the template.
+\<@TODO - did we end up removing any items by the end of the preregistration and evaluation with Chris' study?\> While we 'tweaked' the template items and descriptive text during the in situ evaluation of the templates for the environmental flows preregistration, we decided not to alter the template structure and content significantly, but to instead implement a new format and interface for working with the template.
 We outline these reasons below.
 
-1.  Opinionated Design - norms around modelling practice.
-    'Good Modelling Practice' Literature.
-    This also has the benefit of aiding new modellers.
+1.  Opinionated Design - norms around modelling practice. 'Good Modelling Practice' Literature. This also has the benefit of aiding new modellers.
 
-2.  A related point: linguistic ambiguity and other problems with modelling concepts and terms.
+2.
+A related point: linguistic ambiguity and other problems with modelling concepts and terms.
 Contested terms in the ecological modelling literature about what each different step is, their activities, etc.
 
-3.  *research transparency*.
+3.
+*research transparency*.
 We felt that the decision-points captured in the template are critical to increasing reporting transparency sufficiently for others to properly evaluate the modelling and analysis.
@@
 
-Berna Devezer et al. argue that most open science reforms propose them and are never formally tested.
-In this research, we provide a qualitative assessment of our new method only.
+Berna Devezer et al. argue that most open science reforms propose them and are never formally tested. In this research, we provide a qualitative assessment of our new method only.
 
@@
 
-All questions are on the table.
-We approached this goal with a 'proof is in the pudding' mentality - what are the limits of preregistration?
-Does it slow down the research process?
+All questions are on the table. We approached this goal with a 'proof is in the pudding' mentality - what are the limits of preregistration? Does it slow down the research process?
 
@@
 
-**"**We fear that preregistration will lessen the engagement of analysts with data and that it represents a step back toward the punch card style of analysis." MacEachern, S. N., & Van Zandt, T.
+**"**We fear that preregistration will lessen the engagement of analysts with data and that it represents a step back toward the punch card style of analysis." MacEachern, S.
+N., & Van Zandt, T.
 (2019).
@@
 
-It's, such as t proposed by Feng et al.
-The decisions and details are recorded within the preregistration.
-Along with any deviation from the initial plan.
+It's, such as t proposed by Feng et al. The decisions and details are recorded within the preregistration. Along with any deviation from the initial plan.
 

### 2024-04-30 _analysis/ms.qmd
First Draft Manuscript - add content from draft obsidian doc - mv elsevier quarto extension - commit PDF in case there are rendering issues down the line :|

@@
 abstract: |
-  **1.** Preregistration is an open-science practice adopted in scientific disciplines that have begun to confront the 'reproducibility crisis', which aims to improve research transparency and mitigate questionable research practices within a study.
-There has been little uptake of preregistration in ecology and conservation, and many modellers and researchers engaged in non-hypothesis testing research both within and outside of ecology have eschewed preregistration on the grounds that existing templates are irrelevant because they are focussed on null-hypothesis significance testing.
-  **2.** We advocate that preregistration can and should be used in model-based research in ecology and conservation but that preregistration should reflect the norms and practice of the research context in which it is applied in order to adequately restrict researcher degrees of freedom that are unique to a domain-specific methodology.
-We took a user-centred design approach to the task of translating preregistration into model-based research contexts for ecology and conservation.
-  **3.** We identified a generalised 'modelling workflow' that captures critical analysis decisions in the process of the model development cycle.
+  1.
+Preregistration is an open-science practice adopted in scientific disciplines that have begun to confront the ‘reproducibility crisis’, which aims to improve research transparency and mitigate questionable research practices within a study.
+There has been little uptake of preregistration in ecology and conservation, and many modellers and researchers engaged in non-hypothesis testing research both within and outside of ecology have eschewed pre- registration on the grounds that existing templates are irrelevant because they are focused on null-hypothesis significance testing.
+  2.
+We advocate that preregistration can and should be used in model-based research in ecology and conservation, but that preregistration should reflect the norms and practice of the research context in which it is applied, in order to adequately restrict researcher degrees of freedom that are unique to a domain-specific methodology.
+We took a user-centred design approach to the task of translating preregistration into model- based research contexts for ecology and conservation.
+  3.
+We identified a generalised ‘modelling workflow’ that captures critical analysis decisions in the process of the model development cycle.
 The workflow informed the content and structure of a preregistration template to be used in model-based research.
 The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the current format and linear process of preregistration.
-To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose a methodology for implementing 'adaptive preregistration' leveraging git and GitHub.
-We tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
-  **4.** This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings, and will facilitate the wider uptake of preregistration in ecology.
-Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model-based research within science more broadly.
-Modellers need no longer cry "but, I can't preregister my research."
+To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose a methodology for implementing ‘adaptive preregistration’ leveraging git and GitHub.
+We tested and evaluated the preregistration template and methodology using a case- study of environmental flows modelling for decision-support in Victoria, Australia.
+  4.
+This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings and will facilitate the wider uptake of preregistration in ecology.
+Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model- based re- search within science more broadly.
+Modellers need no longer cry “but, I can’t preregister my research.”
 keywords:
@@
     keep-tex: true
+bibliography: references_obsidian.bib
 ---
@@
 
-# Materials and Methods
+Science's replication crisis, and the subsequent recognition of problems with reliability of scientific research findings, has triggered a 'credibility revolution' spawning a suite of proposed methodological and policy reforms, including *preregistration* [@parker2019].
+Preregistration is an open-science practice that aims to distinguish between planned and ad hoc analyses  [@Parker2019] and aims to combat the risk of questionable research practices by requiring researchers to register their methods and analysis plan in a secure and publicly accessible platform prior to collecting and/or analysing data, which cannot be altered after submission ( #todo #citation).
+The purpose of preregistration is to prevent biased research findings by guarding against questionable research practices (such as selective reporting, p-hacking, HARKing, among others #todo #citation ) by distinguishing what is a genuine *a priori* planned analysis, and what is not, thereby preventing the opportunistic misuse of researcher degrees of freedom, or those many analytic decisions researchers make during the planning, data collection, data analysis and reporting of results in scientific studies (See Box 1 for background).
+QRPs have been labelled as one of the primary cause of the replication crisis (Shrout and Rodgers, 2018 in [@Fife2022]) and serve to XYZ bias the overall literature, etc ( #todo finish this point).
+
+Despite enthusiastic uptake in some scientific disciplines, such as psychology and medicine ( #todo #citation ), the use of registered reports and preregistration in ecology and conservation more broadly is new and limited [@Desbureaux2021], particularly for model-based research ( #todo Joel Pick preprint?)**.
+Ecologists have brushed preregistration off on the basis that existing templates are irrelevant since they are focussed on null-hypothesis significance testing**, whereas in applied ecology and conservation, we don't usually explicitly test hypotheses [@Betts2021], but instead use explanatory models, or increasingly predictive models that generate anticipatory predictions, to inform management decisions [@Wood2020].
+Model-fitting and evaluation, rather than falsification of hypotheses about natural phenomena, are increasingly prevalent approaches to scientific inquiry in ecology [@Connolly2017]
+
+Preregistration in model-based research has been eschewed on several grounds by modellers and researchers engaged in model-based research both within and outside of ecology.
+Much of the literature advocating for the use of preregistration assumes a hypothetico-deductive model of scientific inquiry, and the templates used to write preregistrations have been accused of being focussed on null-hypothesis significance testing [e.g.
+@McDermott2021; @Prosperi2019].
+Approaches and advocacy for preregistration of MBR across all scientific disciplines are limited, and focussed have focused on the phases of the modelling process and modelling approaches that are more confirmatory ( #todo see @ioannidis2022).
+For example, @Cruwell2019b present a template for preregistration of MBR where the model construction and development phase is explicitly and intentionally excluded from the preregistration process, while much of the discussion around preregistration for models has centred on machine learning [@Hildebrandt2018] or predictive modelling [@hofman2023] ( #todo see \[preregistration for modelling\] for more refs).
+
+A second argument against preregistration for model-based research is based on the mischaracterisation of the model development process as a intrinsically exploratory process [@McDermott2021] , and since only the preregistration of confirmatory research, such as hypothesis testing, is considered necessary [@Prosperi2019], there is no need to pre-register model-based research ( #todo #citation).
+The description of modelling as a wholly exploratory process is a mischaracterisation based on a misunderstanding of sensemaking and reasoning within the modelling process stemming from a lack of clear formal definitions distinguishing confirmatory research from exploratory research [@Flournoy2020; @Devezer2020; @alspaugh2019].
+Although modelling includes some tasks and activities that are exploratory in nature, it can be distinguished from wholly *exploratory analysis* in that tends be goal-driven, directed analysis that is less open-ended [@alspaugh2019].
+For model-based research, the exploratory/confirmatory divide is an idealised and simplifying dichotomy that, in practice, is more complicated and nuanced and can be considered as a spectrum on which exploratory analysis sits at the analysis is open-ended, high-level and opportunistic end, to goal-driven, low-level and precise analyses at the other [@alspaugh2019].
+In practice the divide is often blurred, and researchers will move back and forth between different types of reasoning and inference throughout the modelling process, switching from activity at one end of the spectrum to the other and back again [@alspaugh2019], with deductions leading to abductions as the analysis process proceeds [@alspaugh2019; @Prosperi2019], or even occuring together in a hybridized manner [@Prosperi2019].
+
+When critics of preregistration for MBR describe modelling as intrinsically 'exploratory', what is seemingly being referred to is the iterative, adaptive and non-linear nature of model-development.
+In practice, modelling proceeds in an adaptive manner by which models are iteratively developed and refined [@Augusiak:2014gz; @Babel2019; @Bennett2013; @Bodner2020].
+New analyses are often chosen based on the basis of data exploration and previous analyses of the same data [@Dwork2015], For example, modellers may perform exploratory analyses to validate distributional assumptions are met, such as no over-dispersion or zero-inflation in count-data to ensure valid inference when using particular models [@Campbell2021] ( #todo also see @alspaugh2019 for list of examples).
+Such practices violate a key tenet of preregistration, that is 'data-independent' analytic decision-making ( #todo #citation).
+Consequently, the iterative nature of model development, wherein models are continually refined through analysing and reanalysing a dataset and the model, is considered to be in conflict with the practice of preregistration; by requiring researchers to commit to a single static plan that prevents changes or deviations from the original plan it is seen to be forcing adaptive analysis to be *non-adaptive* through 'defining the entire analysis protocol ahead of time' [@Dwork2015; @Benning2019; @MacEachern2019; @McDermott2021].
+In short, we can't write off preregistration for modelling on the basis that it is exploratory, because in practice it employs a range of modes of reasoning from induction, deduction, to abduction, with analysts switching between those types of reasoning or even hybridising them.
+
+What *is* important to consider when considering the application of preregistration within MBR is whether the reliability of MBR may be potentially threatened by the potential for Questionable Research Practices (Box 1).
+Evidence from ecology more broadly has revealed that the rate of QRPs is similar to those in other disciplines [@Fraser:2018cl] and while there has been little study of QRPs in MBR (except @Gould:2020ebg #todo preprint doc & cross-ref here), preliminary evidence reveals that modellers do engage in QRPs such as selective reporting, for example, they may may omit explored analysis paths from their reported analyses if they did not corroborate their conclusions [@Liu2020].
+Furthermore, QRPs are likely to arise in conditions where there is incomplete and opaque reporting practices [@Fidler:2018ho], and poor model-reporting practices and an overall lack of transparency about the modelling process are endemic in MBR in ecology.
+Concerns about the state of model reporting in ecological modelling have been raised repeatedly, arguing that current practices are neither transparent or comprehensive [@Schmolke:2010fd; @deVos:2011uk].
+Despite a number of attempts to promote thorough and transparent reporting, (e.g.
+@Jakeman:2006ii; @Alexandrov:2011iv; @Bodner2020) as a critical component of *Good Modelling Practice*, a culture of adequate research transparency in modelling is yet to be established and remains limited [@deVos:2011uk; @Grimm:2014es].
+The lack of transparency in describing ecological models and reporting the modelling process and results provides significant opportunity for undisclosed researcher degrees of freedom, and hence, the potential for Questionable Research Practices.
+
+Although the purpose of preregistration is predominantly conceptualised in terms of distinguishing exploratory from confirmatory research (e.g.
+@Nosek:2018eg, @munafo2017 and #todo insert other refs), we we adopt a more general stance towards the purpose of preregistration, i.e. that preregistration may mitigate against questionable research practices by reducing the potential for cognitive biases to cause opportunistic misuse of researcher degrees of freedom.
+In this paper, we advocate that preregistration can and should be used in model-based research in ecology and conservation, but its particular form should reflect the norms and practices of the research context in which it is applied, in order to adequately restrict researcher degrees of freedom that may uniquely emerge within this domain-specific methodology.
+
+The goal of this study is to rethink preregistration to better suit the domain-specific challenges and methodological particularities of model-based research in ecology and conservation.
+Preregistration cannot simply be ported over to other disciplines and adopted as-is.
+Different research methodologies, and their domain-specific application across different domains and fields of study follow different analytical choices and are therefore susceptible to different QRPs.
+Consequently, domain-specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom as they uniquely emerge within a domain or field-specific application of a particular research methodology.
+If we wish to adopt preregistration for ecological modelling, both the content required to be preregistered, as well as the procedure for creating a preregistration must be adapted to suit model-based research as it is undertaken in ecology.
+
+We first aimed to build a preregistration template for ecological modelling.
+We identified a generalised ‘modelling workflow’ that captured critical analysis decisions in the process of the model development cycle.
+The workflow informed the content and structure of a preregistration template to be used in model-based research.
+The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the current format and linear process of preregistration.
+To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we devised a methodology for implementing an expanded view of preregistration that we term *Adaptive Preregistration* [sensu @Srivastava2018].
+Adaptive preregistration allows for the iterative cycle of model development to be reflected in the preregistration and better accounts for the inherently adaptive nature of model development than current approaches to preregistration in which researchers are committed to a static, unalterable and deterministic view of preregistration.
+Finally, to assess the feasibility and practicality of preregistration for MBR, we tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
+
+------------------------------------------------------------------------
+
+# Box 1: How Researcher Degrees of Freedom become Questionable Research Practices in MBR
+
+Modellers are faced with numerous 'researcher degrees of freedom' [@Wicherts2016] or subjective decisions during the course of modelling that may substantially impact the chosen analysis strategy, and therefore the research findings and conclusions drawn from that analysis [@Hoffmann2021; @Simonsohn2020; @Desbureaux2021; @gouldSameDataDifferent2023].
+The set of all plausible or reasonable potential decisions that a researcher could take from the beginning to end of analysis has been described as the 'garden of forking paths' [@Gelman2013].
+'Many analyst'-style studies, whereby multiple separate teams of researchers independently analyse the same dataset and same research question, demonstrate that different pathways in this 'garden of forking paths' can lead to potentially conflicting conclusions [@Young2021; @Liu2020; @gouldSameDataDifferent2023].
+
+Researcher degrees of freedom arise because there is inherent analytic uncertainty in model-based research whether due to incomplete theoretical understanding of underlying ecological processes, or because modelling is an act of simplification and abstraction; there is always some degree of choice about which ecological processes and phenomena should be included in the model, how they should be represented [@Babel2019; @Getz:2017ic].
+Further analytic uncertainty may arise due to uncertainty in data processing or choice about which analysis method is best [@Hoffmann2021].
+
+-   *Analytic Uncertainty* [@Walker2003] talks about uncertainty specifically in terms of *model based* decision-making. #todo
+
+When researchers are faced with ambiguous decisions where there is no single decision or strategy acceptable according to scientific standards and consensus, we are likely to select the decision that results in the most 'favourable' finding with "convincing self-justification" [@Simmons:2011iw; @Hoffmann2021].
+Researchers routinely execute and explore alternative analysis pathways in service of identifying the most favourable (i.e.
+publishable) result, selectively reporting only one or several preferred results [@Liu2020; @Simmons:2011iw].
+Thus researcher degrees of freedom precipitate 'questionable' research practices, or QRPs, when they remain both undisclosed and exploited opportunistically [@Bakker2018].
+Such 'questionable' research practices have been demonstrated to be surprisingly common upon admission among ecologists, with comparable rates to other scientific disciplines [@Fraser:2018cl].
+Questionable research practices are categorically distinct from fraudulent practices, arising not from consciously nefarious 'fishing expedition\[s\]' [@Gelman2013], but rather because researchers are motivated actors working within a 'publish or perish' culture who are susceptible to cognitive biases and are willing to change their modelling assumptions when the results conflict with desired outcomes, for example, sincerely and self-convincingly believing that the more favourable model is superior [@Young2021].
+
+QRPs may increase the probability of false positive results, may artificially inflate effect size estimates, and ultimately reduce the replicability of published findings, while encouraging overconfidence in the precision of results and exacerbating the risk of accepting and propagating false facts [@Wicherts2016; @Hoffmann2021; @Roettger2019; @Nissen2016].
+Good intentions, and simply being 'aware' of the potential for researcher degrees of freedom to lead to questionable research practices cannot sufficiently mitigate their risk [@Nosek:2018eg].
+
+------------------------------------------------------------------------
+
+# Methods
+
+We approached the task of translating preregistration into model-based research contexts for ecology and conservation with a user-centred design (UCD) framework.
+UCD is a philosophy for designing processes, services or products that account for user requirements at all stages of the design process so that the resource generated by this approach is effective and usable for specific people in specific contexts [@Pavelin2012].
+Because it is difficult to predict how users will actually engage with a resource [@Pavelin2012] and a designer's intention for preregistration may conflict with the user's purpose and undermine the overarching goal of preregistration [@Pu2019], we heed the advice of Pu et al.
+[-@Pu2019], who recommend treating preregistration as a UCD problem so that preregistration can effectively meet its intended objectives, despite UCD being typically used within software engineering domains, rather than academic domains.
+
+We borrow from the methodology proposed by Pavelin et al.
+[-@Pavelin2012] to design, test and refine a preregistration template suitable for ecological modelling, and a methodology for preregistering model-based research (@fig-UCD).
+The methods description is structured following the key steps illustrated in @fig-UCD.
+
+![Insert caption.
+Number each step so i can refer to it in the headings and text below.
+Make it clear in the figure that we're not doing the build interface and conduct usability testing component of implement, nor the evaluate and iterate.
+we throw that out to the community of practice to do that part.](images/Pasted image 20240314123346.png){#fig-UCD}
+
+## Define Vision & Objectives
+
+> Define the goals of the research under development and identify the audience
+
+The first step in a UCD approach is to articulate the goals, objectives and target audience of the resource, consequently we define the intended purpose of the preregistration template and methodology, the community of users, as well as their research contexts and use-cases.
+Finally, we explicate the aspirations and desirable attributes of the template and methodology.
+Much like in 'Good Modelling Practice', articulating the overarching purpose, intended audience, as well as the objectives and aspirations will guide the development and evaluation of preregistration for model-based research in ecology and conservation, with important implications for the design of the template -- from its structure, content, user-interface and broader workflow in which it is designed to be used.
+
+### Articulating Intended purpose of preregistration template & methodology
+
+We define the purposes of the template as:
+
+1\.
+*To delimit researcher degrees of freedom* \> when conducting model-based research in applied ecology and conservation so as to mitigate the risk of questionable research practices in order to improve the reliability and credibility of this research.
+
+2\.
+*Increase research transparency* \> of model-based research in ecology and conservation.
+*Research transparency* refers to both production transparency — which includes research artefacts like open-access data and materials, or data collection procedures — and analytic transparency — which describes a complete account of how the analytic conclusions are drawn from the data [@Lupia2014].
+
+### Target user community and their research-contexts
+
+We aim for the template and preregistration process developed in this study to be used for model-based research, which we define as any research, empirical or applied, that uses quantitative modelling to answer its research question [@Brudvig2017], where *quantitative modelling* may be defined as any modelling that "describes and/or forecast(s) the behaviour of a system using mathematical and statistical concepts, and whose parameters and their relationships are expressed as quantities" [@Garcia-Diaz2019].
+
+We define model-based research loosely in order to capture the breadth of different modelling goals (e.g.
+exploration, inference and prediction; @Tredennick2021) and type of research contributions (e.g.
+development of a new method, etc #todo see @Moallemi2019 and across both pure and applied ecological modelling.
+
+### Specifying Preregistration use-cases
+
+Pu et al.
+[-@Pu2019] describe two key use-cases of preregistration, depending on the role assumed in interacting with the preregistration.
+Users have different requirements of preregistration depending on the role that they assume, that is, study authors and modellers have a different use-case than reviewers, editors and readers, however this latter use-case is often not explicitly designed and accounted for in current platforms/infrastructure/processes [@Pu2019].
+Reviewers and editors are usually tasked with the process of 'checking' the preregistration, comparing the final manuscript to its preregistration and verifying that the preregistered analyses are conducted and reported as specified [@Pu2019].
+However existing formats are designed primarily for the use-case of the author - they are text-based static documents that are ill-suited to checking the reported study against the preregistration.
+
+Although preregistration authorship is the primary target use-case of the preregistration template and methodology developed in this study, we aim to consider and incorporate features into the preregistration process, such as template formatting, that will facilitate preregistration checking to also address non-authorship use-cases.
+
+### Articulating desirable attributes of the preregistration template and adaptive preregistration methodology
+
+The preregistration template should capture critical decision-points that may influence the overall model and analysis as well as any recommendations or decisions drawn from analysis of the model, thus delimiting researcher degrees of freedom at these junctures in the modelling decision process.
+The structure and content of the template is necessarily opinionated, encouraging Good Modelling Practices, while reflecting community norms and practices in model-development while still meeting the objective of mitigating the risk of Questionable Research Practices by delimiting researcher degrees of freedom.
+The template needs to be generalisable enough to apply to a range of common modelling purposes, methods and research contributions, and flexible enough that if a particular decision is not applicable for the planned study, that those items may be removed or altered, and other items may be added onto the template.
+
+Secondary to the aim of mitigating QRPs, is that the the template should also enhance the research transparency of the preregistered study, ensuring that analysis choices are documented with sufficient transparency to facilitate the reader of the preregistered completed study in making informed assessments about the credibility and reliability of the model and any associated analysis and recommended interventions.
+Borrowing from @Bodner2020's framework to increase the accuracy, reliability and transparency of predictive models, in order to increase transparency of model-reporting, the template should explicitly outline critical model choices and their assumptions, steps in the modelling process, and explicit expectations about the outputs and analysis of the model.
+
+**Adaptive Preregistration Methodology**
+
+The Adaptive Preregistration Methodology aims to capture the entire modelling process, including model development and construction, rather than beginning at the point of analysing a pre-existing and parameterised model for deductive inference / prediction.
+Critically, the methodology must not preclude the iterative and adaptive nature of model construction, while simultaneously restricting the types of data-dependent analytic decision-making that may compromise the reliability and validity of the model and its contingent analysis and conclusions.
+
+## User Research & Analysis
+
+> Characterise users and their needs
+
+On the 4th and 5th of March 2020, we facilitated a series of structured activities at the Arthur Rylah Institute for Environmental Research which aimed to inform the development of two preregistration templates, one for field-work and sampling design, the other for modelling and data-analysis.
+We split the activities into two groups, to better focus discussion for each of these templates.
+
+-   [ ] <mark style="background: #FFF3A3A6;">Note: might need to reframe this and omit the field study stuff, but should probably state that we did this but decided not to pursue?</mark>
+
+The objectives of the collaborative workshop activities were to aid in: 1.
+Understanding how modelling is done in practice to inform and tailor the preregistration templates to adequately capture key decision-points in the scientific modelling process; 2.
+Understanding how science in practice -- whether due to logistical, technical, feasibility or normative constraints -- differs from some idealised practice, in order to: 3.
+Identifying challenges and roadblocks to the use and uptake of preregistration in among ecological modellers, both in empirical and applied research contexts; and to 4.
+Using the above knowledge to inform the preregistration template and process design so they meet their intended purpose, and are adopted by our target user community for their particular use-cases.
+
+### Workshop 1: Identifying Ecological Modelling Workflows, Norms, and Practice
+
+The objective of this workshop was to identify modelling workflows that capture critical decision-steps in the model development cycle that should be included in a preregistration template.
+By "critical", we mean that without these decision-steps being included, the template would not adequately constrain researcher degrees of freedom nor b) transparently document the analytic decision-process.
+The activity consisted of two-parts, first, we individually reflected on our scientific process for a recent or memorable research project, listing the steps we undertook from the beginning to end.
+In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto pre-prepared workflow template (insert specific OSF link to the document, Supp Mat 1).
+The template consisted of several key 'phases' of the modelling process, with commonly implemented decision-steps grouped under each phase.
+The development of this modelling workflow sheet was derived from previous work by Gould et al.
+*in prep* (insert link to paper preprint) and is outlined in Appendix 1 ( #todo insert ).
+We sorted and categorised each decision-step in our personal workflows as belonging to a particular workflow phase.
+Next we grouped similar decision-steps that could be included as a single item on a preregistration template.
+At the end of this process we reviewed the suggested phases of the workflow to assess whether phases needed further discretisation, aggregation, removal, or amendment.
+
+### Workshop 2: Understanding ecology- and modelling-specific challenges to adopting preregistration in research practice
+
+This second activity aimed to identify critical challenges and roadblocks to undertaking preregistration in ecology and conservation in general, and for model-based research in particular.
+We facilitated a discussion structured around a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature ( #todo Supp Mat 1).
+These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in Workshop 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.
+Discussion points included the following: - What should a research workflow look like for a modeller in an applied ecological setting who wishes to preregister their work?
+- What barriers are there for preregistering model-based research?
+- How do we accommodate the iterative cycle of model development into the preregistration process?
+- At what stage in a modelling or structured decision-making process should preregistration begin?
+- should we and how can we change the medium of the template and/or archiving platform to accomodate any roadblocks?
+- How could we change the medium of both the template and the archiving platform to accommodate these issues?
+- What procedural and or technical solutions could we implement to address these roadblocks to preregistration?
+- What decision-tools could be used to aid in completing a preregistration?
+- What sort of parsimony should we aim for in terms of the resolution of the templates and therefore their applicability across different methodologies?
+
+### Analysis of Workshop Outputs & Design
+
+> Analyse outputs of user-research workshops to inform the design of resource mockups
+
+**1. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
+
+The goal of this analysis task was to describe a generalised workflow of the modelling cycle that will underpin the structure and content of the preregistration template.
+We used a combination of inductive and deductive coding ( #todo is this what it is called and #citation) using the qualitative coding software ATLAS.ti [-@atlasTI] to analyse the workshop outputs and generate a model of scientific workflow for model-based research that would constitute the presumed scientific workflow that would be used to guide the structure of the preregistration template.
+We collated and coded the workshop outputs using.
+All workshop outputs can be found here ( #todo Link to Workshop Outputs OSF repository), and the coded analysis can be found here ( #todo link to analysis files / screeen caps).
+
+A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on ‘good modelling practice’ and structured decision-making literature in ecological management.
+Steps in the workflow were coded and broken down into ‘phases’ consisting of ‘steps’ and ‘sub-steps’, in line with the ontology we specified earlier for translating a modelling workflow into a preregistration template (section link), <mark style="background: #FF5582A6;">(insert link to appendices or just present the final?)</mark>.
+Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and sub-steps.
+We revised the coding structure (inductively is this the word?), according to patterns and themes identified across multiple personal workflows.
+This workflow is used to provide structure and content to the preregistration template ( #todo insert link to this workflow).
+
+**2. Translating Workflows into Preregistration Templates**
+
+We translated the model-based workflow identified in part one of workshop analysis into a text-based preregistration template.
+Reflecting on existing templates ( #todo link to OSF repository of templates page, and others, namely secondary data analysis, see Olmo's links), we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part 1 of workshop outputs analysis.
+Following the features of a good guideline identified by Moallemi et al.
+[-@Moallemi2019], we provide examples of choices or illustrations to clarify each step of the preregistration, especially for phases or steps in the the modelling process that are semantically, epistemically and/or philosophically contentious or ambiguous (link to paper describing contentious issues in modelling history)\_.
+We synthesised the following ontology with which to build our preregistration template: 1.        Decision Phase - each has a “title” and “description” or “definition”, steps are grouped under phases.
+(Note, that these are being scrapped from the template and are turned into tool-tips for formatting) 2.        Decision Step - each has a “title” and “description” or “definition” choices are grouped into ‘steps’, which corresponds to an activity or set of related activities in a scientific workflow.
+(again, these are being scrapped and turned into tool-tips.
+CHECK THIS, it's changed) 3.        Choices —\> each ‘choice’ corresponds to a uniquely numbered item on the PRT requiring a response from user.
+The style should be in the form of a directive or a question.
+E.g.
+“explain how you will do xyzzy”.
+Or “What performance measure will you use for assessing ‘goodness of fit’?”
+
+**3. Identifying challenges and solutions for implementing preregistration in model-based research**
+
+We analysed the outputs of Activity 2 in the collaborative workshop identifying barriers and solutions to undertaking preregistration in model-based research and to broader adoption of preregistration in the modelling community of practice.
+The analysis of Activity 2 in the collaborative workshop was used to inform the design of a suitable preregistration procedure for preregistering MBR in the domain of ecology.
+
+## User Testing
+
+> Build mock-ups of the product and test with users
+
+While user-testing is an essential step in the UCD framework, we also concur with @Devezer2020 that open-science reforms should be evaluated against their stated objectives.
+We therefore preregistered a real-world model-based research problem as a case-study for evaluating and testing the preregistration templates and proposed adaptive preregistration methodology.
+
+### Case Study
+
+We preregistered a real-world study of vegetation responses to environmental flows in Victoria, Australia.
+This study is one component of a larger set of studies within the Stage 6 Victorian Environmental Flows Monitoring and Assessment Program (VEFMAP), managed by Department of Energy, Environment and Climate Action (DEECA) and delivered through the Arthur Rylah Institute for Environmental Research.
+The case-study analysis aimed to evaluate the effectiveness of environmental-flows on riparian vegetation with the following objectives: - Quantify short and medium-term vegetation responses to individual flow events and regimes.  - Describe detailed vegetation trends associated with season, flow, rainfall and other factors such as livestock grazing and exotic vegetation  - Quantify the effect of various factors on vegetation in order to guide future management expectations and planning.
+
+Details of the study background, methods and results are described in Appendix A4.
+
+### Evaluation Criteria & Rationale
+
+-   [ ] *Evaluation Criteria and Rationale (should this be moved to earlier aims section?* Seems like there's a little bit of overlap between the two...
+
+The content of the template should be 'parsimonious' in that the included items within the template should adequately a) constrain researcher degrees of freedom and by being specific, precise and exhaustive [@Wicherts2016] b) facilitate transparent documentation of the modelling process and analytic decisions during model development and analysis, while c) not being too detailed to the point that the preregistration is too difficult and time-consuming to complete.
+
+Many modellers [e.g.
+@MacEachern2019] #todo #citation are concerned about preregistration making modelling a more difficult task than it already is, and hampering the ability to meet project deadlines, especially in applied settings in which models directly underpin decisions.
+Modellers already struggle with sufficiently documenting models during publication and reporting , and there is a resource cost to increasing transparency of model-reporting [@deVos:2011uk].
+Consequently we will also assess the preregistration template and process on its perceived feasibility and practicality.
+
+### Evaluation Method
+
+Rather than attempting to reconstruct the process of creating the preregistration through using interview questions alone and digging back through the version history of a completed preregistration [e.g.
+@Pu2019], we aimed to capture the analysts 'thinking in the moment' by utilising the version-control and collaborative project management features of Git and GitHub (Ben, Smith, and Bailey 2018; Blischak, Davenport, and Wilson 2016) as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to ‘live-develop’ the template template.
+For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration, the suggested change, and its justification would be both recorded in GitHub and linked to the actual preregistration.
+
+Finally, we supplemented this ‘in situ’ testing and evaluation process with follow-up semi-structured interviews on 9th May 2024 in order to capture more detailed reflections of the analysts.
+Lead analysts of the case study were often under strict time constraints for meeting deliverables on the project, and the process of preregistration was a new and unfamiliar task, so capturing reflections on the broader preregistration process during preregistration and analysis of the case-study was not always possible.
+
+## Implementation
+
+> Implement the product, write technical implementations and build the product
+
+We wrote the initial template as a text-based document in Rmarkdown.
+The initial template used to preregister the case-study underwent a series of changes throughout the preregistration process, including changes to the organisation, content, help-text and visual appearance.
+The final preregistration template is made available as a quarto markdown extension ( #citation & link to doc) that can be installed and used to populate a new preregistration document for a new modelling study, and as a form that can be used with the `preregr` R package [@Gjalt-Jorn2023].
+
+To aid the case-study analysts in using the adaptive preregistration methodology to preregister their study, we wrote a detailed user-guide outlining the key principles of adaptive preregistration and a protocol for implementing adaptive preregistration using git and GitHub.
+The guide is presented as a publicly available website and also includes the template.
 
@@
 
-# Discussion & Conclusions
+## User-research and analysis
 
-# Data & Software Availability
+The outputs and analysis for both Workshop 1 and 2 is archived at (insert OSF link).
+The modelling workflow generated from the analysis of Workshop 1 that informed the development of the initial preregistration template is archived at the same link.
+During user-testing with our case-study, the preregistration template structure, and therefore its underlying modelling workflow on which it is based underwent, a series of changes.
+Consequently we present the revised workflow alongside the refined preregistration template in section X below.
 
-# Author Contributions
+Discussions from Workshop Activity 2 focussing on challenges to preregistering MBR revealed that the iterative nature of model development is a significant barrier to adopting preregistration for model-based research.
+Modellers felt that adopting preregistration in their workflows was impossible because preregistration was understood to be a process that precludes iterative model development because its inner logic presumes a linear research workflow premised on hypothesis-testing research, and because it requires researchers to confirm an unalterable version of the analysis plan prior to collecting or analysing the data [@Pu2019; @mertens2019].
 
-# Supplementary Material
+The key tenets of preregistration are based on NHST-focused accounts of questionable research practices that prescribe the principle of 'decision-independence' where analysis decisions are made independently of and *a priori* to analysing the data [@Srivastava2018; @Gelman2013].
+However, analysts reflected on their personal modelling workflows and described situations in the modelling process that may violate the principle of 'decision-independence' that could not be avoided and were considered to be an intrinsic feature of the model-development process.
+During modelling a researcher will almost always engage in state-dependent or data-dependent decision-making -- that is, decisions about whether to alter the model in a subsequent round of model development depend on the results of model testing, evaluation and analysis at the end of the previous iteration of model development.
+
+Very often in modelling, there are decisions that are necessarily dependent on the outcome of previous analytic decisions in the modelling workflow [@Liu2020].
+Firstly, preliminary or investigatory analyses might need to be conducted before being able to specify future decision-steps in the analysis plan.
+For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation.
+Secondly, data-dependent decisions from model checking may justifiably result in changes to either the analysis or to interpretations of the model [@Srivastava2018].
+For instance, inspection of the residuals and other model checking tests may force the modeller to return to earlier decision-points and change the planned analysis or model.
+Other times, the planned analysis and specified model may simply not converge, or the model is saturated and runs out of variation to apportion such that the model and or the model fitting algorithm must be respecified and re-implemented.
+Other times, the modelling process itself may generate knowledge and learning about the system, and the model structure changes throughout the process of model development.
+
+Data-dependent decision-making is not necessarily intrinsically *questionable*, but it does have the *potential* to reduce the reliability of the model and its findings if they are exploited in service of artificially inflating the accuracy or precision of the model, its predictions or evaluation tests to the effect that the model is perceived to be more credible than it would be if that practice did not occur -- essentially data-dependent decisions that lead to consumers of the model placing unsubstantiated belief in the reliability, validity and utility of the model and its outputs are questionable [@Gould:2019ebg] and can be mitigated through the practice of preregistration.
+
+The concept of preregistering all modelling choices *a priori* was seemingly overwhelming for analysts.
+They described situations where decisions are too difficult to anticipate, or cannot be made in advance, particularly for decisions occurring at the later phases of model construction and development process (Figure 2, TBC), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses.
+Some decisions might not be able to be specified on the first attempt at writing the preregistration until the model is fully or at least close to fully specified, for example, specifying precisely what and how sensitivity analyses or uncertainty analyses will be conducted.
+
+-   Finally, sometimes there are some decisions that are too difficult to anticipate, or simply cannot be made in advance [@Srivastava2018].
+-   I think Pu or Pavelin describe the same thing... they coin a name for these types of decisions too but I can't find it.
+
+## Design
+
+### Adaptive Preregistration: A methodology for preregistering model-based research
+
+In order to address the disjuncture between the iterative nature of model development and the single-use, deterministic preregistration procedure, we propose an expanded view of preregistration, *adaptive reregistration*, which allows for some degree of flexibility and adaptivity in the analysis workflow, while still mitigating against unconstrained and undisclosed researcher degrees of freedom.
+Adaptive preregistration breaks with the current preregistration workflow, wherein the author writes a single deterministic preregistration containing a single deterministic choice for every decision-step and the research shifts from ideation, to preregistration to execution of the plan (fig or #citation to ref?).
+Instead, in adaptive preregistration, as the modeller proceeds through the model development process, they will transition from ideation, preregistration, execution, and back to preregistration and execution again.
+The modeller may repeatedly cycle through the preregistration process rather than proceeding in a linear fashion.
+Two features facilitate this process: 1.
+An *adaptive preregistration* contains 'plans to deploy flexible strategies' [@Srivastava2018] whereby the preregistration author can supply a heuristic consisting of different analysis or modelling strategies whose execution depends on the outcome of the previous decision-point or analysis, for example, decision-trees for certain points in the analysis or items in the preregistration template could be preregistered prior to analysing [@Baldwin2022a].
+2.
+The preregistration may be iterative and consist of interim preregistrations that mark the transition between phases of modelling and analysis as different parts of the data are observed [@Srivastava2018].
+
+-   [ ] Other refs to chase up... Benning, Schuwirth, Latihan, hamalainen, hofman, reusable holdout paper, Dwork?
+
+#### Preregistering Flexibility
+
+When preliminary analyses, model checking or other data-dependent analytic decisions are to be preregistered, the modeller can specify a 'flexible strategy' within their preregistration.
+This is particularly relevant for when it is uncertain as to whether the planned analyses will be appropriate for the data [@Baldwin2022a], which is a commonly encountered issue in ecological modelling, because often-times the model we might ideally want to fit that best represents our conceptual model of the target phenomena cannot be fitted due to a paucity of data ( #todo citaiton for this?
+was a motivating factor for the VEFMAP project).
+Like existing approaches to preregistration, the registered heuristic is still a deterministic one, but there there may be multiple potential choices rather than one, and their realisation depends on the outcome of an earlier decision step or analysis.
+
+To preregister flexibility, there are some critical requirements: 1.
+State what quantity or outcome needs to be known in order to move forward with the modelling and analysis, and 2.
+explain what test or analysis will be performed to obtain this quantity or outcome, and which parts of the data will be used in this analysis, and 3.
+describe how the results will be interpreted, linking each potential decision to its outcome trigger, where possible.
+
+-   [ ] Might want to give an example of this...?
+Or could give example in reporting the case study and cross-ref that?
+Alternatively, could put an example in the guide, and send readers there?
+I think second option is best.
+-   [ ] SEE ALSO OTHER refs in this vault, the idea of registering flexibility e.g. using decision-trees is not new.. which paper was it?- e.g. BENNING?
+
+#### Interim Preregistrations
+
+Just as model development proceeds in an iterative process of preregistration, so too does the process of adaptive preregistration.
+As the modeller proceeds, they will implement parts of the analysis plan that were specified in the previous preregistration step, letting the outcomes of that preregistered analysis inform either the specification or the realisation of the pre-registered specification in the next step of the analysis workflow (@fig-adaptive-preregistration).
+
+The initial preregistration should have as many items as possible, if not all items, in the preregistration template answered.
+If an item cannot be answered with a single decision because that decision is data-dependent, then the analyst should register flexibility for that item following the aforementioned requirements.
+Depending on observed outcomes of the registered decision-trees or flexible heuristics, interim preregistrations are created at multiple points in the model development process, each time there is an addition or amendment to the plan.
+
+### Integrating additional data handling and modelling practices
+
+We propose additional data handling and analysis techniques be embedded into the both the modelling and adaptive preregistration processes to help avoid the pitfalls of data-dependent decisions, such as partitioning the data to create a holdout dataset or using blinded datasets for data exploration.
+
+Partitioning data into train/test splits is a modelling technique that has been suggested for incorporation into preregistration for predictive modelling [@hofman2023] or other non-hypothesis testing research [@Baldwin2022a; @Dwork2015].
+We suggest that activities considered as typically 'exploratory' partition the data into 'training' and 'holdout' datasets where the training dataset can be used to inform potential models or analyses that will be preregistered in later versions.
+The holdout set may then be used for components of the modelling process that align more closely with confirmatory modes of inquiry, such as model evaluation or testing, forecasting, hind-casting and other predictive methods [@lewis2023; @Pu2018; @hofman2023] #todo add pred modelling citations, e.g.
+@Houlahan:2016fl etc.
+see memo for Lewis et al.
+).
+However, partitioning ecological data is not always feasible due to small datasets, studies that are limited in duration or spatial extent compared to the target phenomena, or data patchiness due to the logistical and financial challenges to data acquisition in ecology [@Wood2020].
+
+In these cases where preregistration of flexibility and data partitioning is not feasible or where the analyst may need to check the data in order to pre-specify a decision, e.g.
+checking for data distributions or missingness, blinded datasets, such as those generated by data scrambling [@Baldwin2022a].
+Data blinding can prevent data-peeking that may influence downstream modelling decisions while aiding the modeller in making informed analytic decisions that are appropriate for the dataset on hand.
+
+Finally, oftentimes interpretation of the model and its outputs are based on subjective assessments, such as visually interpreting plots of model fit quality for indications of issues such as heteroscedasticity, interpreting the results of model performance where there is no clear decision threshold, or interpreting model performance across a suite of different metrics and indicators.
+We encourage analysts to follow Baldwin et al.'s [-@Baldwin2022a] suggestion whereby both the research questions and the conditions for interpretation are preregistered.
+We have embedded this approach to preregistering MBR into our template as well, by including prompts within help and explanatory text that direct the researcher to ensure that the con
+
+### Implementing Adaptive Preregistration with GitHub
+
+-   [ ] #todo insert summarised version of guide
+
+### Draft Template & Modelling Workflow
+
+We developed a draft template for preregistering model-based research in ecology and conservation that was tested and further refined using our case-study of adaptive preregistration ( #todo Appendix A2).
+We report the final preregistration template refined through the case-study, as well as the refined description of the modelling workflow underlying the preregistration template.
+
+-   [ ] How do I present this in the manuscript?? Do I just refer to this in the supp mat? Feels important to present but this manuscript is already LONG!
+-   [ ] Also need to either make fig or table based on preregistration template for the underlying model development workflow underpinning it.
+
+## User Testing
+
+Model development and adaptive preregistration processes occurred in tandem, and were implemented using git and GitHub, Leveraging GitHub's version control features.
+This approach is summarised in the guide <https://egouldo.github.io/EcoConsPreReg/adaptive_preregistration_implementation.html,> [@gould2024].
+The GitHub repository is located at <https://github.com/egouldo/VEFMAP_VEG_Stage6,> but is currently embargoed as of today (30 April 2024).
+
+-   [ ] #todo consider whether we explain the GitHub procedure here...
+in the 'Design' section, in an appendix section, OR leave out and just reference the website and zenodo link.
+I DO think however, that it has some key features that map onto the adaptive preregistration process that I want to highlight.
+So perhaps we can present in the 'Design' section.
+
+### Final Template Version & Changes
+
+The template used in the VEFMAP study underwent a series of changes to both its content and visual formatting during the analysis to better suit the needs of the modellers.
+The preregistration template was written as an *Rmarkdown* document saved within the project GitHub repository (INSERT LINK TO DOCUMENT #todo).
+Using knitr (CITE) and pandoc (CITE), this document is converted to a markdown document using RStudio's `github_document` format, which is automatically rendered into HTML for viewing in the web browser by GitHub (<https://rmarkdown.rstudio.com/github_document_format.html).>
+
+Improvements to template formatting include the use of dropdown boxes to store template text.
+Template text was divided into a) text that describes the rationale for the inclusion of a particular preregistration section or item within the template, as well as any explanation of key terms and definitions of some modelling practice / process terms, and b) text that contains specific questions with check-boxes for researchers to address in their preregistration.
+Each text-type is given a different icon to help differentiate between them.
+The use of drop-down text-boxes with icons better differentiates template prompts and text from authors' responses to preregistration items and the VEFMAP modellers noted that these changes greatly improved readability of the preregistration document <https://github.com/egouldo/VEFMAP_VEG_Stage6/pull/75#pullrequestreview-1959017213.>
+
+In terms of substantive changes to the structure and content of the preregistration template, significant changes were made to the structure of the document (reorganized different section), and to descriptions of item rationale and explanations, as well as the preregistration items themselves.
+Some items were omitted, some were added, most were reworded for improved expression, and some were re-structured into different sections.
+
+-   [ ] Collate comments, pull requests and reviews and issues describing problems with the changes see: <https://github.com/egouldo/VEFMAP_VEG_Stage6/issues/29> #chrisjones and also other comments from #jdyen
+
+Modellers think about the modelling process differently, and use different terms for the same or similar practices and modelling concepts, depending on their particular modelling background, and at times the language used in the original version of the template was confusing and unclear to the VEFMAP modellers.
+
+-    \[ \] #todo consider as #discussion point in this manuscript #writing
+
+### Adaptive Preregistration of the Case Study
+
+We tested and refined both the preregistration template and adaptive preregistration protocol as outlined in Version X of the guide using the VEFMAP modelling study as a case study.
+We did not preregister this study on an existing official preregistration platform that meets the three core requirements of preregistration outlined by Haroz [-@haroz2022] because existing platforms are unable to accommodate adaptive preregistration where there are multiple and interim versions of the preregistration.
+Consequently tracked changes to the preregistration using version control with git and GitHub, and used GitHub's tag and release features to demarcate major changes and interim versions of the preregistration.
+
+The preregistration process was marked by three main phases of analysis and modelling, each marked with an interim preregistration:
+
+1\. V1 - Data analysis and cleaning, specification of problem context and background preregistered (insert link to supp mat)
+
+2\. V2 - Pilot Analysis preregistered (insert link to supp mat)
+
+3\. V3 - Main analysis including decision tree for informing model selection based on results of pilot analysis (insert link to supp mat)
+
+**V1: Preregistration of problem context, data analysis and cleaning**
+
+Prior to preregistration and analysis, data for the VEFMAP study had already been collected, as this study is a small component of a broader project that has been running for a number of years.
+The preregistration was iteratively completed and updated prior to beginning any analysis or modelling with the VEFMAP dataset, however, data cleaning and wrangling was undertaken in tandem with completion of the initial preregistration document by different researchers on different git branches due to project time-constraints.
+
+**V2: Preregistration of Pilot Analysis**
+
+There was a great deal of uncertainty about whether the dataset would support fitting the desired candidate models due to temporal and spatial patchiness and density of the data.
+The dataset consisted of over 2 million observations across different waterbodies which were missing data during some years, while other sites had been surveyed more times than others (insert table from meeting with DEECA into supp.
+mat and cross-ref \[\[2024-03-27_Meeting_Notes\]\]).
+Some preliminary analysis and exploration of the dataset was needed in order to make appropriate decisions about how best to operationalise the models given the patterns in the underlying dataset and the spatio-temporal variation in the density / patchiness of the dataset.
+For example, simple tests of data distributions within hierarchies needed to be implemented to check for data spread and presence of zero-inflation, and checks for collinearity among candidate predictor variables were needed to reduce overfitting and to ensure the relevant interactions and captured in the models.
+Consequently, we decided to implement the use of a holdout dataset by partitioning the data into a smaller 'pilot analysis' dataset and 'main analysis' dataset.
+The pilot analysis dataset consisted of data from only a single site that had good temporal data coverage, while the dataset for the main analysis consisted of the entire dataset, including the pilot dataset.
+
+The pilot analysis was exploratory in nature and aimed to resolve critical uncertainties about how best to operationalise the models given data constraints in order to identify a candidate set of models that could feasibly be fitted using the full dataset.
+In addition to the preliminary analyses described above, two initial models for the two response variables, cover and richness, were preregistered.
+The same model checking procedure was preregistered for both the pilot and main analyses.
+The pilot analysis was preregistered in section "2.5.3 Describe any data exploration or preliminary data analyses" of the template.
+
+*Pilot Analysis Outcomes*
+
+The pilot analysis resolved some uncertainties, while some further analytic uncertainty about the best model operationalisation remained.
+Firstly, for models of both response variables, the data were unable to support a single model containing separate terms representing both flow regimes and flow events.
+Consequently two separate models, one with each term operationalising flow, were fitted to the pilot dataset instead.
+Secondly, separate models for each plant functional group were desired to help understand how each group responds to environmental flows.
+However, there was not enough data for each functional groups split across both native and introduced species to get these models to converge so plant functional group was included as a random effect instead and fitted to the pilot data.
+
+Model checking revealed both high levels of over-dispersion and over-inflation.
+To account for different proportions of zeros among plant functional groupings, the pilot analysis allowed for zero-inflation to differ among plant functional groups, however did not account for over-dispersion.
+During the pilot analysis, models were refitted using negative-binomial distributions but failed to converge.
+
+-   [ ] #todo Report models here!
+
+-   [ ] #todo Report Decision tree here!
+
+-   [ ] See notes from the report back meeting here: \[\[2024-03-25_Meeting_Notes\]\] and \[\[2024-03-27_Meeting_Notes\]\]
+
+**V3: Preregistration of Main Analysis and Decision Tree**
+
+Although the full models with complex three-way interactions failed to converge on the smaller pilot dataset, it was hoped that they may still converge on the full dataset in the main analysis.
+Consequently, the full models were also specified in the main analysis, but with the two simplified 'flow events' and 'flow regime' models preregistered also in the event that the full models were unable to converge and/or provide reliable parameter estimates.
+
+Given that zero-inflated Poisson models both converged, and reliably captured the proportion of zeros within the pilot dataset, this model family was preregistered for all models fitted to the full dataset, including simplified models.
+However, modellers were concerned that there may be unexpected issues with these approaches due to actual data distributions in the full dataset not matching expectations informed by the pilot analysis, such as unaccounted over-dispersion of the data.
+Consequently a second flexible strategy was preregistered to account for the case where over-dispersion was identified.
+
+First, the two key vegetation response variables were to be modelled as zero-inflated Poisson distributions with a log-link function.
+However, in the case that model-checking identified over-dispersion in any of the models, the models were to be refitted using a negative binomial distribution.
+However, a rule was preregistered that the models re-fitted using the negative binomial distribution would only be accepted over the Poisson models if they *both* improved over-dispersion *and* did not decline in model performance, particularly in terms of model fit, zero-inflation, and posterior predictive checks.
+
+The final decision on which model family should be selected based on model checking in the instance that overdispersion is identified is a fuzzy and subjective one in which explicit weightings between the different model performance criteria cannot be expressed or preregistered.
+Instead, modellers attempted to provide an explanation and description of the conditions for interpretation informing decisions about the model family:
+
+> While we are unable to *a priori* precisely weight these criteria in determining the final distribution, the final decision will be guided by the model’s overall ability to capture key associations reliably.
+This is especially likely to occur when there is no ‘perfect model’ and there is no dominant alternative choice of model distribution.
+
+**Analysis following final preregistration**
+
+TBD, briefly summarise what outcomes from the preregistered decision heuristics were realised in the final analysis.
+
+## Implementation: User Guides and Documentation
+
+We have developed a guide for modellers on how to implement adaptive preregistration for model-based research packaged as a website at <https://egouldo.github.io/EcoConsPreReg/> [@gould2024].
+The guide first provides background material on the risk of researcher degrees of freedom to transparency and reliability of model-based research in order to motivate the use of preregistration in model-based research.
+The core materials include an overview of adaptive preregistration, as well as a practical step-by-step guide to implementing adaptive preregistration using [git](https://git-scm.com) and [GitHub](www.github.com), two commonly used tools by modellers for version-controlling modelling code during model development [@braga2023].
+The preregistration template for model-based research in ecology and conservation is also hosted on this website.
+Components of the guide to adaptive preregistration have been repackaged and incorporated into this manuscript.
+
+**Continuous Improvement and Feedback**
+
+One of the key features of user-centred design is the principle of ‘continuous feedback’.
+We therefore encourage modellers to preregister their research using our template and adaptive preregistration protocol outlined in this manuscript, and to provide feedback and suggestions for improvement.
+Suggested changes may be submitted in the form of a pull-request to the repository ( #todo insert link to how to make a pull request to the new template repository), or by posting a new issue ( #todo link to creating a new issue on the repo and setup template).
+The repository and issues are publicly available, so any future discussion will be transparent and open to the ecological modelling community to participate in.
+
+# Discussion
+
+-   intro para summarising problem / objective:
+    -   Preregistration of model-based research has been rejected both within ecology and in other disciplines.
+        -   But the problems that preregistration addresses, i.e. RDFs / QRPs, are likely to exist for MBR too.
+    -   There is a lack of discipline-specific preregistration templates and this impedes the uptake of open-science tools, such as preregistration, within ecology.
+    -   This study aimed to expand preregistration for use within ecological modelling and modelling more broadly by developing a preregistration template for ecological modelling research, and a methdology for preregistering model-based research more generally.
+
+We took a user-centred design approach to the task of developing a preregistration template and methodology that accounts for the domain and methodology-specific nuances of ecological modelling.
+We did this by: 1.
+Identifying idealised norms and norms of practices to develop ‘idealised’ / ‘generalised’ research workflows that describe sets of commonly undertaken tasks (and some sort of order/sequence to reflect the workflow, both idealised and as practically conducted), within an ecological modelling study.
+2.
+We then transformed these workflows into a ‘preregistration template’, by distilling a set of decision-items that, should a) adequately constrain researcher degrees of freedom to mitigate the risk of Questionable Research Practices in an ecological modelling exercise and b) adequately facilitates research/analytic transparency.
+We transformed the decision-steps (the highest resolution) into template items posed as directives or questions for preregistration users to complete, in accordance with the particularities of the study.
+3.
+We qualitatively evaluated the template and the adaptive preregistration methodology for its feasibility of use in MBR using a case-study of environmental flows modelling in Victoria, Australia...
+The VEFMAP modelling study faced particular challenges to preregistration in terms of data complexity and patchiness that made this study a good candidate for trialling adaptive preregistration...
+
+## Preregistration Template
+
+Blah what to say?
+
+## Adaptive Preregistration Methodology
+
+User research revealed critical features of the model development process, namely the inherent iterative nature of model development, and requirements for model checking that forced us to reconsider the procedure for completing a preregistration suitable to ecological modelling.
+Consequently, we took an ‘expanded view of preregistration’, “Adaptive Preregistration”.
+Key features: - plans for preregistering and deploying flexible analysis strategies: - Iterative, cyclical preregistration rather than linear process that maps onto the iterative cycles of model development generating - interim preregistrations marking key points in the analysis
+
+We also propose combining adaptive preregistration with previously proposed technical strategies for reducing the risk of overfitting and bias in MBR.
+E.g.
+using data-partitioning with a exploratory/training vs.
+test/holdout dataset / data scrambling for preliminary or exploratory analyses.
+We depart from existing research, e.g.
+Baldwin, by proposing that these preliminary analyses also be pre-registered, even if preregistration is not yet complete.
+The results of the preliminary or exploratory analyses may be used to inform analytic decision-making that can then be preregistered.
+
+Presently there are no existing preregistration repository services that meet the critical requirements of a preregistration platform outlined by Haroz [-@haroz2022] that can be used for adaptive preregistration.
+We proposed and refined an implementation of the Adaptive Preregistration methodology that leverages features of git and GitHub for transparently tracking the evolution of the preregistration document as the modelling progresses.
+
+### Reflections (heading TBD, can we make a water pun given case study and concluding sentence or is that lame)
+
+We approached the application and evaluation of both the template and adaptive preregistration and sought to evaluate the feasibility of preregistration for MBR.
+We genuinely did not know what to expect when we began preregistering the case study, but approached this goal with a 'proof is in the pudding mentality' to ask what are the limits of preregistration....
+Here we reflect on the parts that worked, and those that did not.
+
+**Time & Timing**
+
+-   “A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
+This trade-off was something we considered important during the design phase, and deciding what was in and out of scope for the preregistration template was particularly challenging.
+
+-   However, as Pu et al. note, some participants who were both authors and reviewers didn’t think that the time-costs of preregistration in research were significantly greater than not preregistering research.
+Instead they thought of the preregistration process invoking a ‘time shift’, that brought ‘decision points’ before data collection instead of during analysis.” P.19.
+In addition, “some reviewers thought that preregistration saves them the time they might have spent wondering about undisclosed flexibilities when working on an un-preregistered paper.”
+
+-   Even in adaptive preregistration where the process is iterative, modellers reflected on the front-heavy reserach timeline forced by the process of adaptive preregistration.
+Shifting decision-making to the start of the project timeline delayed the start of the analysis, but modellers felt that this resulted in better decision-making because the decision-making was more carefully reasoned.
+Additionally, the case study researchers felt that the level of detail reported in the preregistration aided in efficiently writing the internal report.
+
+    -   This relates to what Pu et al. found where participants ‘felt their research workflow improved by having preregistration as a “forcing function” \[…\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers’ degree of freedmen advance.’ (Page 13).
+For reviewers ‘participant reports that preregistration makes the review process “easier”.’
+
+-   Mention difficulties I had with the internal deadlines of DEECA, and that we could not preregister everything we wanted to (i..e the analysis of the model and its outputs used for presentation in the report, which were extremely vaguely specified, aside from model checking).
+Preregistration could be challenging for modelling in such settings where there extremely tight timelines.
+BUT, as we are doing for the stage 2 report preregistration of the second study now, could be used during early consultation / scoping.
+So the document could begin to be filled out much earlier than we did here (data already collected).
+
+**Resolution of the template**
+
+-   Competing objectives?
+Although the purposes of preregistration are not mutually exclusive - as we have decided the purpose of preregistration we have defined here is to both delimit/state flexibility, *and* be flexible.
+However, “the preference of one purpose over another can have tangible effects on the design of preregistration” p15..
+Pu.
+
+-   Compared to Hofman [-@hofman2023], a template for preregistering predictive modelling research, our template was much more detailed and contained many more items to be preregistered.
+
+-   Get Jian / Chris' thoughts on template.
+I suspect it's a bit long / overwhelming.
+But potentially that's because of the visual presentation / limitations of a static text based medium for the document.
+
+-   However, the benefit of a more exhaustive, specific and detailed preregistration is increased transparency (link back to intro on transparency in ecological modelling).
+By adopting preregistration for MBR we can also improve the transparency of ecological modelling, which is done terribly!!!
+
+-   cost to completeness, and specificity / precision is that it is time-consuming.
+But aside from increasing transparency, it is also critical for appropriately restraining unmitigated and undisclosed RDFs by preventing loose interpretation / operationalisation of the preregistration in instances where the preregistration is vague.
+
+    -   We can see this in the disjuncture between all of the modelling outputs conducted and reported for the internal report, but that weren't fully specified in the preregistration due to time-constraints on the modellers.
+
+**Restricting Researcher Degrees of Freedom, not creativity!**
+
+Counter arguments to both preregistration and preregistration for modelling express fear that preregistation will restrict creativity and lead to 'punch card style analysis' [@MacEachern2019].
+However, modellers in our preregistration of the case study found the opposite, that it prompted critical thinking about analytic choices during modelling, that might have only ever been tacit and never explicitly recorded (cite paper saying many modelling decisions never thought of as decisions and therefore never reported).
+
+Combining the use of data partitioning, and splitting the analysis into a 'pilot' and 'main' analysis aided the modellers in understanding what models could feasibly be preregistered given constraints on the data, while also helping inform the preregistration of flexible analyses.
+The addition of the pilot study components to the template, and into the modelling process was good.
+And I think we should recommend doing this where possible.
+Link to that arxiv paper on preregistering machine learning studies.
+Might seem less natural for ecological modelling studies since we are data poor!
+BUT, in machine learning and other computer-science focussed modelling contexts, this type of practice is the norm.
+One issue to flag is the potential for \[\[information leakage\]\], however.
+Because part of the data *is* used again in the full dataset.
+However, given that this study is situated within an ongoing program of monitoring and decision-making, I think it's ok.
+The modelling can be built into an adaptive monitoring / modelling context where future modelling can be used to test / evaluate and then update the model.
+
+\- How do we think the process actually went in terms of restricting researcher degrees of freedom / improving reliability / reproducibility? **What other reflections are there?**
+
+\- Tricky to anticipate??
+But can certainly say that all of the exploratory analysis undertaken on the pilot dataset would have surely been applied to the full dataset had we not partitioned the dataset and the preregistration process.
+
+## Future Implementation & Testing
+
+We did not undertake the 'usability testing' component of User-Centered design set out in fig 1.
+We offer this template and adaptive preregistration methodology to the ecological modelling community, and modelling communities across disciplines, to use, refine and adapt in their modelling practices moving forward.
+We hope insights gained through trial and error of its application will be used to further refine and improve the preregistration template and adaptive preregistration methodology.
+
+Advocacy and uptake of preregistration is based on the working hypothesis that it is able to mitigate against QRPs by delimiting opportunistic exploitation of researcher degrees of freedom while improving transparency.
+While initial evidence suggests that preregistration may improve the reliability and reproducibility of research (cite study looking at replication rates of preregistered vs not), empirical testing of this working hypothesis is necessary for evaluating whether this new practice meets its intended purpose [@Devezer2020] (and others).
+Future research should evaluate the effectiveness of adaptive preregistration for improving the reliability and transparency of MBR.
+
+Hosting the preregistration template on GitHub, has the added benefit of providing a mechanism for dealing with *uncertain* *norms* in modelling practice [@Pu2019], and for following modelling methods and decision contexts that might not fit the presumed modelling development cycle underpinning the preregistration template.
+Anyone is free to *fork* our GitHub repository, meaning that the entire contents of the repository can be copied to a users own repository, and the template is free to be altered as the user sees fit.
+This provides the opportunity for creating domain-and/or modelling-approach-specific templates [@Cruwell2019b].
+
+-   [ ] #todo need to explain a bit further what the hell i mean by 'uncertain norms'.
+I think I can link the confusion from VEFMAP modellers about some concepts within the preregistration document in here.
+
+In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
+However, I don’t expect users to agree on this process.
+Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
+
+### Using GitHub for Adaptive Preregistration
+
+**Benefits**
+
+-   transparency
+-   existing tool already used by many modellers for both version control and collaborative project management of computational code for models
+
+**Challenges**
+
+Separate branches...
+Under significant time pressure (internal reporting) to do the analysis.
+Data cleaning and analysis of pilot data was happening on one branch while the preregistration was happening on another!
+Also meant that the fledge release system outlined in the guide did not quite work because you had to wait until each branch had been merged back into the main / master branch in order to bump the version.
+
+\*\*Critiques of GitHub as a platform for preregistration. and combining other tools to meet the minimum requirements of preregistration: timestamps, persistence, registry\*
+
+While there are some proponents of the use of version control systems for preregistration [e.g.
+@braga2023; @hofman2023], [@haroz2022] make the compelling argument that GitHub (and potentially alternative version control systems) are inappropriate as preregistration platforms because it doesn't meet any of the three core requirements for a preregistration: timestamps, an indexed registry, and persistence.
+While we have described the use of GitHub timestamps in this study, [@haroz2022] illustrate that timestamps can be manipulated on GitHub.
+Moreover, currently GitHub does not have a formal registry for indexing or searching for preregistrations hosted on GitHub, which means that there is the potential for *\[\[preregistration hacking\]\]*, wherein preregistrations may be deleted changed and reuploaded [@haroz2022].
+Finally, GitHub does not meet the *persistence* requirement, there is no immutable unique identifier, like a digital object identifier (DOI), and GitHub is not guaranteed as an archive service in perpetuity \[\[data sharing\]\] / \[\[data curation\]\].
+
+While the use of GitHub in this study was to test it as a platform for implementing *Adaptive Preregistration* rather than *Preregistration* per se, given these shortcomings, we recommend pairing GitHub with other platforms for implementing Adaptive Preregistration.
+
+Here we tested it as a system for managing and implementing *Adaptive Preregistration* rather than just preregistration in general.
+However, given these shortcomings, we recommend pairing the use of GitHub for Adaptive Preregistration with some additional tools to ensure that these three minimum requirements of preregistration are met, we outline two complementary approaches here that meet the three core requisites of preregistration:
+
+1.
+Archive the GitHub repository using Zenodo (CITE) at the first draft preregistration (V.1.0) and combine with GitHub tag and release strategy, such that each interim preregistration is recorded as a new version / release on both GitHub and Zenodo .
+Deals with both the timestamped issue and the persistence issue by giving it a digital object identifier [@smithSoftwareCitationPrinciples2016], Furthermore, Because it is given a DOI, it is indexed and searchable [@hofman2023].
+    -   [ ] Add links to other papers suggesting Zenodo be combined with GitHub for data / code sharing #todo
+2.
+But not a specifically preregistration registry.
+How to deal with the registry component?
+Can we add as component to OSF repo??
+and then preregister?
+Yes.
+In fact, an alternative solution has been outlined by [@vanlissa2021], where they have used a similar approach to us with a version-controlled plain-text preregistration document containing the pre-analysis plan, and tagging and releasing this document on GitHub.
+However, they suggest a secondary optional step of uploding the preregistration to a dedicated preregistration server, without specifying what that server may be.
+We could use an 'open-ended Registration' template on the OSF [@hofman2023], and then link the document from GitHub component on the OSF project to the registry.
+Instructions for doing this are given in [@hofman2023].
+    -   [ ] WHAT ABOUT DIFFERENT VERSIONS OF THE PREREGISTRATION..
+Can we make changes using the OSF??
+DOUBLE CHECK THIS!
+Might not be able to use the registry here.
+#todo I think #MarkCall said there is the option to make changes now.
+Could we test with their staging platform??
+Put the different versions on there??
+But the template has changed....
+Could just try it anyway as proof of concept.
+And note this as a limitation of the OSF templates approach (not being modular or flexible, as [@hofman2023] outline ).
+
+-   [ ] #todo CHECK THIS: Embedding the repository within an OSF repository will fix the timestamp issue from GitHub that [@hofman2023] outline??
+
+# Conclusions
+
+Instead of throwing the baby out with the bathwater, in this study, we advocate that preregistration can and should be used isn model-based research in ecology and conservation, but its particular form should reflect the norms and practices of the research context in which it is applied, in order to  in order to adequately restrict researcher degrees of freedom that may uniquely emerge within a domain-specific methodology.
+
+This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings and will facilitate the wider uptake of preregistration in ecology.
+Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model-based re-search within science more broadly.
+Modellers need no longer cry “but, I can’t preregister my research.”
+
+## Software & Data Availability Statement
+
+## Author Contributions
+
+CRediT
+
+# Supp Material
+
+-   [ ] Check out journal guidelines or papers #todo
+
+## SM1: Workshop Materials and Process Design
+
+-   Design process for the modelling workflow template for activity 1. The actual materials are in A1.
+-   Activity 2 structured discussion with prepared questions and talking points informed by recent debates about preregistration in the literature.
+Any literature review leading to this.
+Actual talking points in A1.
+-   Mention that we also ran the above in separate groups for designing a field-study preregistration template, but decided not to pursue further here.
+
+## SM?: Analysis of Workshop Outputs
+
+"We revised the coding structure (inductively is this the word), according to patterns and themes identified across multiple personal workflows:" For example, we added the final phase (name for last phase), which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions.
+In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both (idealised norms and norms of practice, Pu et al., see other philosophy of sci refs).
+
+## A2: Preregistration Template
+
+## A3: Adaptive Preregistration Guide
+
+## A4: Environmental Flows Modelling Case Study
+
+**Study Background**
+
+Riparian vegetation is a critical ecosystem component for so many reasons (biodiversity, habitat, food, stability, filter, carbon source, aesthetic, etc.) but it is also severely degraded in many waterways all over the world.
+Vegetation clearing, livestock grazing, pollution, exotic species (flora and fauna) and alteration of flows are some of the most severe threats.
+
+The impacts of river regulation on vegetation depend on the type and extent of flow alteration.
+Flows can increase or decrease total flow, be more or less variable, or have altered season flow timing.
+Environmental flows are designed to be manipulations of flow (usually flow releases from a reservoir) that aim to have specific and direct environmental benefits.
+
+Evaluating the effects of environmental flows on waterways is difficult because of the need to evaluate the isolated impacts of specific flow events (short term response) as well as the impacts of successive flows (environmental or otherwise) within a broader flow regime (medium to long term response).
+Data collection at large temporal scales (annual or greater) prevents evaluation of any single flow event or component from the regime.
+Evaluating the effects of environmental flows in most cases is therefore only possible with repeated sampling at relatively short intervals.
+Evaluation also needs to be done with consideration of other major drivers of vegetation, such as rainfall, other flow attributes, livestock grazing, soil properties, exotic species and previous disturbance.
+
+In south-eastern Australia, river regulation (including water extraction), livestock grazing and exotic species continue to have serious negative impacts on all major waterways, despite major government investment into waterway management.
+However, increased understanding of the impacts of major threats and the effectiveness of management is resulting in greatly improved management decisions.
+While many broad trends are known, critical details remain elusive and require concerted effort to address.
+
+**Study Aims**
+
+### Study Methods
+
+#### Data Collection
+
+-   [ ] 1 - 2 sentence overview, then point readers to details in \[\[project-VEFMAP\|VEFMAP\]\] manuscript #todo #writing
+
+#### Modelling Dataset
+
+| Variable Name                            | Unit                  | Definition                                                                                                                                     | Variable Type | Variable Group                                                         |
+|-----------|-----------|----------------------------|-----------|-----------|
+| Species Vegetation Cover                 | Cover                 | Cover per-species, per sub-transect                                                                                                            | Response      | Vegetation Cover                                                       |
+| Plant Functional Group Cover             | Cover                 | Cover per-plant functional group, per-sub-transect                                                                                             | Response      | Vegetation Cover                                                       |
+| Sub-transect Vegetation Cover            | Cover                 | Total cover (sum all species), per sub-transect                                                                                                | Response      | Vegetation Cover                                                       |
+| Species Richness                         | Counts                | Number unique Species Functional Group, per sub-transect                                                                                       | Response      | Species Richness                                                       |
+| Days above baseflow / springfresh levels | Counts, Z-scale       | Number of days per year (averaged over prev 3 water years)<br>discharge \> baseflow or springfresh levels,<br>Z-scaled                         | Predictor     | Flow Regime                                                            |
+| Sub-transect zone                        | Factor                | Zone in which a sub-transect occurred:<br>- below baseflow level, <br>- between baseflow and spring-fresh levels<br>- Above spring-fresh level | Predictor     | Flow Regime                                                            |
+| Flow event                               | Factor                | Timing of sampling:<br>- before spring fresh<br>- Between spring and summer fresh<br>- After summer fresh                                      | Predictor     | Flow Event                                                             |
+| Site Grazing Status                      | Factor                | Grazing status of site:<br>- Grazed<br>- Ungrazed                                                                                              | Predictor     | Other factors influencing spatio-temporal variation in cover/abundance |
+| Species Functional Group                 | Factor                | For each Origin group (native or exotic):<br>- aquatic<br>- emergent<br>- riparian<br>- terrestrial                                            | Predictor     | Other factors influencing spatio-temporal variation in cover/abundance |
+| Site Identity                            | Factor                | Name / Identity of the site                                                                                                                    | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
+| Transect Identity                        | Factor                | Identity of Nested within sites                                                                                                                | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
+| Sub-transect height                      | Metres above baseflow | Height of the sub-transect along bank                                                                                                          | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
+| Survey Year                              | Date, YYYY            | Year in which the sample was recorded                                                                                                          | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
+| Waterbody                                | Factor                | Waterbody in which the site is located                                                                                                         | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
+
+#### Pilot Study
+
+The final data set comprised XYZ observations, which presented a significant computational challenge.
+In addition, exploratory data analysis revealed considerable zero-inflation and potential over-dispersion of the response variables, particularly vegetation cover.
+We conducted a pilot study to determine the feasibility of different model structures and identify likely decision points given these challenges.
+
+-   [ ] when did the EDA occur... during the pilot study? On a different data set? #help-wanted #jdyen
+
+To address this, we defined a maximal model structure and tested this structure with a pilot data set (all observations from the Campaspe River from 2017–2020) to identify feasible model structures and likely decision points.
+We developed separate analyses for vegetation cover and species richness, with the same maximal model structure for both.
+
+Initial model fitting assumed a zero-inflated Poisson distribution for both vegetation cover and species richness, with zero-inflated negative binomial distributions designated in cases where over-dispersion was severe.
+This approach does not introduce an upper bound on estimates of cover (e.g., specifying a maximum cover level of 100%).
+We used this approach for three reasons.
+First, overlapping species may generate cover estimates that exceed 100%.
+Second, parameters of bounded distributions are challenging to estimate when many values occur near the bounds.
+Third, bounded distributions require link functions (e.g., the logit function) that complicate the specification and interpretation of autoregressive models.
+We note that the Poisson distribution approximates the bounded binomial distribution in the limit of many trials (REF).
+The extent of zero-inflation and over-dispersion (and a model’s capacity to capture these patterns) was assessed using posterior predictive checks, which use values simulated from the fitted model to determine whether the model structure captures the observed distribution of the response variable (REF).
+\### Results
+
+#### Pilot Study Results
+
+The final maximal model was as follows:
+
+``` r
+response (cover or species richness) ~ cover in previous survey +
+                                        functional group × origin × flow metrics +
+                                        functional group × origin × zone × sampling period +
+                                        grazing +
+                                        (1 | site / transect) + (1 | sub-transect height) + (1 | survey year) + (1 | waterbody) +
+                                        offset(number of points).
+```
+
+Vegetation cover in the previous survey was included only for analyses of cover and meant that models of vegetation cover examined how flow influenced the change in cover through time rather than absolute vegetation cover (akin to models of population growth rate used in the VEFMAP long-term fish monitoring project).
+Flow metrics are the (z-scaled) days above baseflow level and days above spring-fresh level in the three years preceding a survey and included both linear and quadratic associations[\[JDLY(1\]](#_msocom_1) .
+Zone is the location of a sub-transect relative to baseflow or spring-fresh levels and sampling period is the timing of a survey relative to the delivery of spring and summer freshes.
+Sub-transect height records the level of a sub-transect in metres above baseflow level.
+Waterbody was excluded from the pilot study given its focus on a single waterway (the Campaspe River).
+
+Results of the pilot study inform the subsequent analyses and are presented here.
+Models of vegetation cover at the species level did not converge in any instance.
+Models of vegetation cover aggregated to watering-response plant functional groups converged but required separate models of two-way interactions instead of the three- and four-way interactions included in the maximal model structure.
+Models of vegetation cover aggregated to structural groups (aquatic, emergent, riparian, terrestrial) converged but required removal of interactions with species origin and separate analyses of flow metrics and for each of the categorical flow terms (zone and sampling period).
+Models of vegetation species richness within watering-response functional groups converged but required separate models of two-way interactions instead of the three- and four-way interactions included in the maximal model structure.
+Models of vegetation species richness within structural groups converged but required removal of interactions with species origin and separate models for the two categorical flow variables (zone and sampling period).
+Therefore, subsequent analyses of vegetation cover and species richness focused on structural groups because these supported model structures closest to the maximal structure.
+The included interactions were re-assessed in models of the full data set because the inclusion of additional data was predicted to either support more-complex model structures or introduce additional heterogeneity (and, therefore, support less-complex model structures) than in the pilot study.
+This did not include re-assessment of the functional group classifications.
+
+Posterior predictive checks indicated that a zero-inflated Poisson distribution reliably captured the distribution of observed vegetation cover and a Poisson distribution (without zero-inflation) reliably captured observed distributions of species richness (FIGS).
+Models of cover had excess over-dispersion, but a zero-inflated negative binomial distribution did not substantially reduce unmodelled over-dispersion and resulted in worse characterisation of zero-inflation and the observed distribution of cover data (FIGS?).
+Therefore, subsequent models of vegetation cover and species richness assumed that data were derived from a Poisson distribution, with additional parameters for zero-inflation specific to each functional group in the case of vegetation cover.
+
+### Full Dataset
+
+# Appendices
+
+## A1 : Workshop Materials
+
+-   Activity 1 Template, and instructions / materials provided to workshop attendees. (insert link to OSF)
+-   Activity 2 structured discussion with prepared questions and talking points informed by recent debates about preregistration in the literature (insert link to OSF)
+
+## A2: Draft Preregistration Template
+
+-   [ ] Insert draft template
+
+## A3: Final preregistration Template
+
+-   [ ] Insert template

### 2024-05-05 _analysis/ms.qmd
Add docx output format to ms
update discussion section of ms after discussion with libby

@@
 format:
+  docx: default
   pdf: default
@@
 Make it clear in the figure that we're not doing the build interface and conduct usability testing component of implement, nor the evaluate and iterate.
-we throw that out to the community of practice to do that part.](images/Pasted image 20240314123346.png){#fig-UCD}
+we throw that out to the community of practice to do that part.](images/Pasted%20image%2020240314123346.png){#fig-UCD}
 
@@
 
--    \[ \] #todo consider as #discussion point in this manuscript #writing
+-   [ ] #todo consider as #discussion point in this manuscript #writing
 
@@
 
-## Preregistration Template
-
-Blah what to say?
-
-## Adaptive Preregistration Methodology
-
-User research revealed critical features of the model development process, namely the inherent iterative nature of model development, and requirements for model checking that forced us to reconsider the procedure for completing a preregistration suitable to ecological modelling.
-Consequently, we took an ‘expanded view of preregistration’, “Adaptive Preregistration”.
-Key features: - plans for preregistering and deploying flexible analysis strategies: - Iterative, cyclical preregistration rather than linear process that maps onto the iterative cycles of model development generating - interim preregistrations marking key points in the analysis
+## Benefits, challenges, Limitations of Adaptive Preregistration
+
+User research revealed inherent features of the model development process, namely its iterative workflow and adaptive decision making, were critical barriers to modellers adopting preregistration in their modelling practice.
+In the case of the VEFMAP modelling study, there were critical uncertainties about the the appropriate model specification given limitations of the underlying dataset which prevented the researchers from being able to pre-register fully specified models without undertaking some exploratory analysis.
+Such practices would preclude the application of preregistration to this study because these decisions would violate preregistration's core principle of *data-independent* decision-making.
+
+*What would have happened without preregistration / adaptive preregistration??* (insert into last sentence of above para)
+
+-   In the absence of any preregistration, adaptive or not, (#todo check this with Jian, because maybe his practices are different), modellers would typically explore the entire dataset
+-   Explain what is wrong with this, what the risk of this is.
+
+*Next para:*
+
+To mitigate the risk of researcher degrees of freedom and to try to preserve some degree of data-independence in decision-making we combined data-partitioning into the adaptive preregistration process and split the preregistration process into two phases, resulting in three interim preregistrations.
+We first preregistered a preliminary or 'pilot' analysis, in which the researchers under took directed exploration of a subset of the dataset.
+The pilot analysis helped the researchers to understand what models could feasibly be fitted for their two outcome variables given uneven spatio-temporal density and patchiness of the dataset.
+Although the pilot analysis was 'exploratory' in nature, by preregistering the preliminary analysis, any exploration occurring here was still *directed* in that the scope of the pilot analysis was bounded and determined by transparently explicated analysis goals and articulartion of critical uncertainties.
+Following preregistration of the pilot analysis, the researchers then switched back to the *analysis* phase of *adaptive preregistration*.
+Carrying out the preregistered pilot analysis.
+The results of the pilot analysis aided in determining the *main analysis* which would be conducted on the full dataset and preregistered using the EcoEvoModelling template.
+
+Researchers preregistered flexibility into their main analysis, because despite the pilot analysis, some critical uncertainties still reamined cf. limitations on the data.
+
+-   outline the uncertainties: model family, and something else. see results.
+
+Explain the different models, and the different
+
+-   loosely preregistered (Baldwin) grounds for interpretation made explicit
+
+Consequently, we took an ‘expanded view of preregistration’, “Adaptive Preregistration”. Key features:
+
+\- plans for preregistering and deploying flexible analysis strategies:
+
+\- Iterative, cyclical preregistration rather than linear process that maps onto the iterative cycles of model development generating
+
+\- interim preregistrations marking key points in the analysis
 
@@
 Here we reflect on the parts that worked, and those that did not.
-
-**Time & Timing**
-
--   “A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
-This trade-off was something we considered important during the design phase, and deciding what was in and out of scope for the preregistration template was particularly challenging.
-
--   However, as Pu et al. note, some participants who were both authors and reviewers didn’t think that the time-costs of preregistration in research were significantly greater than not preregistering research.
-Instead they thought of the preregistration process invoking a ‘time shift’, that brought ‘decision points’ before data collection instead of during analysis.” P.19.
-In addition, “some reviewers thought that preregistration saves them the time they might have spent wondering about undisclosed flexibilities when working on an un-preregistered paper.”
-
--   Even in adaptive preregistration where the process is iterative, modellers reflected on the front-heavy reserach timeline forced by the process of adaptive preregistration.
-Shifting decision-making to the start of the project timeline delayed the start of the analysis, but modellers felt that this resulted in better decision-making because the decision-making was more carefully reasoned.
-Additionally, the case study researchers felt that the level of detail reported in the preregistration aided in efficiently writing the internal report.
-
-    -   This relates to what Pu et al. found where participants ‘felt their research workflow improved by having preregistration as a “forcing function” \[…\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers’ degree of freedmen advance.’ (Page 13).
-For reviewers ‘participant reports that preregistration makes the review process “easier”.’
-
--   Mention difficulties I had with the internal deadlines of DEECA, and that we could not preregister everything we wanted to (i..e the analysis of the model and its outputs used for presentation in the report, which were extremely vaguely specified, aside from model checking).
-Preregistration could be challenging for modelling in such settings where there extremely tight timelines.
-BUT, as we are doing for the stage 2 report preregistration of the second study now, could be used during early consultation / scoping.
-So the document could begin to be filled out much earlier than we did here (data already collected).
-
-**Resolution of the template**
-
--   Competing objectives?
-Although the purposes of preregistration are not mutually exclusive - as we have decided the purpose of preregistration we have defined here is to both delimit/state flexibility, *and* be flexible.
-However, “the preference of one purpose over another can have tangible effects on the design of preregistration” p15..
-Pu.
-
--   Compared to Hofman [-@hofman2023], a template for preregistering predictive modelling research, our template was much more detailed and contained many more items to be preregistered.
-
--   Get Jian / Chris' thoughts on template.
-I suspect it's a bit long / overwhelming.
-But potentially that's because of the visual presentation / limitations of a static text based medium for the document.
-
--   However, the benefit of a more exhaustive, specific and detailed preregistration is increased transparency (link back to intro on transparency in ecological modelling).
-By adopting preregistration for MBR we can also improve the transparency of ecological modelling, which is done terribly!!!
-
--   cost to completeness, and specificity / precision is that it is time-consuming.
-But aside from increasing transparency, it is also critical for appropriately restraining unmitigated and undisclosed RDFs by preventing loose interpretation / operationalisation of the preregistration in instances where the preregistration is vague.
-
-    -   We can see this in the disjuncture between all of the modelling outputs conducted and reported for the internal report, but that weren't fully specified in the preregistration due to time-constraints on the modellers.
 
@@
 But can certainly say that all of the exploratory analysis undertaken on the pilot dataset would have surely been applied to the full dataset had we not partitioned the dataset and the preregistration process.
+
+**Resolution of the template - tradeoff between meeting preregistration objectives, and burden and benefits to the researcher.**
+
+-   Competing objectives?
+Although the purposes of preregistration are not mutually exclusive - as we have decided the purpose of preregistration we have defined here is to both delimit/state flexibility, *and* be flexible.
+However, “the preference of one purpose over another can have tangible effects on the design of preregistration” p15..
+Pu.
+
+-   Compared to Hofman [-@hofman2023], a template for preregistering predictive modelling research, our template was much more detailed and contained many more items to be preregistered.
+
+-   Get Jian / Chris' thoughts on template.
+I suspect it's a bit long / overwhelming.
+But potentially that's because of the visual presentation / limitations of a static text based medium for the document.
+
+-   However, the benefit of a more exhaustive, specific and detailed preregistration is increased transparency (link back to intro on transparency in ecological modelling).
+By adopting preregistration for MBR we can also improve the transparency of ecological modelling, which is done terribly!!!
+
+-   cost to completeness, and specificity / precision is that it is time-consuming.
+But aside from increasing transparency, it is also critical for appropriately restraining unmitigated and undisclosed RDFs by preventing loose interpretation / operationalisation of the preregistration in instances where the preregistration is vague.
+
+    -   We can see this in the disjuncture between all of the modelling outputs conducted and reported for the internal report, but that weren't fully specified in the preregistration due to time-constraints on the modellers.
+
+*This will lead in nicely to the final criticism*
+
+**Time & Timing**
+
+-   “A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
+This trade-off was something we considered important during the design phase, and deciding what was in and out of scope for the preregistration template was particularly challenging.
+
+-   However, as Pu et al. note, some participants who were both authors and reviewers didn’t think that the time-costs of preregistration in research were significantly greater than not preregistering research.
+Instead they thought of the preregistration process invoking a ‘time shift’, that brought ‘decision points’ before data collection instead of during analysis.” P.19.
+In addition, “some reviewers thought that preregistration saves them the time they might have spent wondering about undisclosed flexibilities when working on an un-preregistered paper.”
+
+-   Even in adaptive preregistration where the process is iterative, modellers reflected on the front-heavy reserach timeline forced by the process of adaptive preregistration.
+Shifting decision-making to the start of the project timeline delayed the start of the analysis, but modellers felt that this resulted in better decision-making because the decision-making was more carefully reasoned.
+Additionally, the case study researchers felt that the level of detail reported in the preregistration aided in efficiently writing the internal report.
+
+    -   This relates to what Pu et al. found where participants ‘felt their research workflow improved by having preregistration as a “forcing function” \[…\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers’ degree of freedmen advance.’ (Page 13).
+For reviewers ‘participant reports that preregistration makes the review process “easier”.’
+
+-   Mention difficulties I had with the internal deadlines of DEECA, and that we could not preregister everything we wanted to (i..e the analysis of the model and its outputs used for presentation in the report, which were extremely vaguely specified, aside from model checking).
+Preregistration could be challenging for modelling in such settings where there extremely tight timelines.
+BUT, as we are doing for the stage 2 report preregistration of the second study now, could be used during early consultation / scoping.
+So the document could begin to be filled out much earlier than we did here (data already collected).
 
@@
 | Variable Name                            | Unit                  | Definition                                                                                                                                     | Variable Type | Variable Group                                                         |
-|-----------|-----------|----------------------------|-----------|-----------|
+|-------------|-------------|----------------------|-------------|-------------|
 | Species Vegetation Cover                 | Cover                 | Cover per-species, per sub-transect                                                                                                            | Response      | Vegetation Cover                                                       |

### 2024-09-22 preregistration_modelling.qmd
Reformat callouts for consistency,  including checkboxes and italicisation. Fix typos.

@@
 
-Identify relevant interest groups:
+*Identify relevant interest groups:*
 
@@
 
-Briefly outline
+*Briefly outline:*
 
@@
 
--   How will you operationalise any scenarios identified in 1.1.3? For example, how will you operationalise any qualitative changes of interest, such as ‚ 'deterioration' or 'improvement'?
-
--   Describe how you will evaluate and distinguish the performance of alternative scenario outcomes
-
--   Justify or otherwise explain how you chose these measures and determined performance criteria in relation to the analytical objectives, model purpose and modelling context, such as the risk attitudes of decision-makers and stakeholders within this system
+-  [ ] How will you operationalise any scenarios identified in 1.1.3? For example, how will you operationalise any qualitative changes of interest, such as ‚ 'deterioration' or 'improvement'?
+- [ ]  Describe how you will evaluate and distinguish the performance of alternative scenario outcomes
+- [ ]  Justify or otherwise explain how you chose these measures and determined performance criteria in relation to the analytical objectives, model purpose and modelling context, such as the risk attitudes of decision-makers and stakeholders within this system
 :::
@@
 
-List and explain critical conceptual design decisions [@Grimm:2014es], including:
+*List and explain critical conceptual design decisions [@Grimm:2014es], including:*
 
 -   [ ] spatial and temporal scales,
-
 -   [ ] selection of entities and processes,
-
 -   [ ] representation of stochasticity and heterogeneity,
-
 -   [ ] consideration of local versus global interactions, environmental drivers, etc.
-
 -   [ ] Explain and justify the influence of particular theories, concepts, or earlier models against alternative conceptual design decisions that might lead to alternative model structures.
@@
 
-Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be represented in the model [@Moallemi2019]. Sources of uncertainty may include:
+*Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be represented in the model [@Moallemi2019]. Sources of uncertainty may include:*
 
@@
 
-Identify and define system system variables and structures, referencing scenario definitions, and the scope of the model as described within problem formulation (1.1.3):
+*Identify and define system system variables and structures, referencing scenario definitions, and the scope of the model as described within problem formulation:*
 
@@
 -   [ ] Explain how any key concepts or terms within problem or decision-making contexts, such as regulatory terms, will be operationalised and defined in a biologically meaningful way to answer the research question appropriately?
-(Should this last step go here or within define conceptual framework?.
 :::
@@
 
-*This section specifies the plan for collecting, processing and preparing data available for parameterisation, determining model structure, and for scenario analysis.*
+*This section specifies the plan for collecting, processing and preparing data available for parameterisation, determining model structure, and for scenario analysis.
+It also allows the researchers to disclose any prior interaction with the data.*
 :::
@@
 
-For each separate preliminary or investigatory analysis:
+*For each separate preliminary or investigatory analysis:*
 
 -   [ ] State what needs to be known to proceed with further decision-making about the modelling procedure, and why the analysis is necessary.
-
 -   [ ] Explain how you will implement this analysis, as well as any techniques you will use to summarise and explore your data.
-
 -   [ ] What method will you use to represent this analysis (graphical, tabular, or otherwise, describe)
-
 -   [ ] Specify exactly which parts of the data will be used
-
 -   [ ] Describe how the results will be interpreted, listing each potential analytic decision, as well as the analysis finding that will trigger each decision, where possible.
@@
 Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand
-
 -   [ ] Document any issues with data reliability.
-
 -   [ ] How will you determine what data, if any, will be excluded from your analyses?
-
 -   [ ] How will outliers be handled? Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.
-
 -   [ ] How will you identify and deal with incomplete or missing data?
@@
 
-### Performance Criteria
+### Parameter estimation / model fit performance criteria
 
@@
 
-Specify assumptions and key uncertainties in the formal model.
+- [ ] Specify assumptions and key uncertainties in the formal model.
 Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpretation of the model outputs, and what implications are there for evaluating model-output to inform inferences or decisions?
@@
 
+::: callout-note
+## Explanation
+
+Once critical decisions have been made about the modelling approach and method of model specification, the conceptual model is translated into the quantitative model.
+:::
+
 ::: callout-caution
@@
 
-Once critical decisions have been made about the modelling approach and method of model specification, the conceptual model is translated into the quantitative model.
-
 -   [ ] Specify all formal models
-    -   Note, For data-driven approaches to determining model structure and or parameterisation, it will not be able to complete this preregistration item.
+    - [ ]  Note, For data-driven approaches to determining model structure and or parameterisation, it may not be possible to respond to this preregistration item.
+In such cases, explain why this is the case, and how you will document the model(s) used in the final analysis.
 -   [ ] For quantitative model selection approaches, including ensemble modelling, specify each model used in the candidate set, including any null or full/global model.
@@
 
--   If partitioning data for cross-validation or similar approach:
+*If partitioning data for cross-validation or similar approach (delete as needed):*
 
@@
 
--   If using external / independent holdout data for model testing and evaluation:
+*If using external / independent holdout data for model testing and evaluation (delete as needed):*
 
 -   [ ] Which data will be used as a the testing data? What method will you be used for generating training / test data subsets?
-
 -   [ ] Describe any known differences between the training/validation and testing datasets, the relative size of each, as well as any stratification methods used for ensuring evenness of groups between data sets?
-
 -   [ ] It is preferable that any independent data used for model testing remains unknown to modellers during the process of model development, please describe the relationship modellers have to model validation data, will independent datasets be known or accessible to any modeller or analyst?
@@
 ## Implementation verification
+
+::: callout-note
+## Explanation & Examples
+
+Model implementation verification is the process of ensuring that the model has been correctly implemented, and that the model performs as described by the model description [@Grimm:2014es].
+This process is distinct from model checking, which assesses the model's performance in representing the system of interest [@Conn:2018hd].
+
+-   Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as described by the model description [@Grimm:2014es].
+-   Qualitative tests could include syntax checking of code, and peer-code review [@ivimey2023].
+Technical measures include using unit tests, or in-built checks within functions to prevent potential errors.
+:::
 
@@
 Specifying a priori quality assurance tests for implementation verification may help to avoid selective debugging and silent errors.
-
-    -   Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as described by the model description [@Grimm:2014es]*.*
-
-    -   Qualitative tests could include syntax checking of code, and peer-code review [@ivimey2023].
-Technical measures include using unit tests, or in-built checks within functions to prevent potential errors.
 :::
@@
 
-### Quantitative Model Checking
+### Quantitative model checking
 
@@
 
-### Qualitative Model Checking
+### Qualitative model checking
 
@@
 
-The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019].
+*The consequences of assumption violations on the interpretation of results should be assessed [@Araujo2019].*
 
--   Explain how you will demonstrate robustness to model assumptions and check for violations of model assumptions.
-
--   If you cannot perform quantitative assumption checks, describe what theoretical justifications would justify a lack of violation of or robustness to model assumptions.
-
--   If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.
-
--   If assumption violations cannot be avoided, explain how you will explore the consequences of assumption violations on the interpretation of results (To be completed in interim iterations of the preregistration, only if there are departures from assumptions as demonstrated in the planned tests above).
+- [ ] Explain how you will demonstrate robustness to model assumptions and check for violations of model assumptions.
+- [ ] If you cannot perform quantitative assumption checks, describe what theoretical justifications would justify a lack of violation of or robustness to model assumptions.
+- [ ] If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.
+- [ ] If assumption violations cannot be avoided, explain how you will explore the consequences of assumption violations on the interpretation of results (To be completed in interim iterations of the preregistration, only if there are departures from assumptions as demonstrated in the planned tests above).
 :::
@@
 <!-- #NOTE, explain how model-checking is different to model validation / evaluation.
-What am i defining as model analysis / application??
+What am I defining as model analysis / application??
 the phases of the modelling process underlying this template need exposition somewhere -->
@@
 propagating known uncertainties through the model, investigating the effect of different model scenarios with different parameters and modelling technique combinations [@Araujo2019], or empirically constructing model distributions from model output data points, and extracting summary statistics, including the mean, median, variance, and constructing confidence intervals [@Saltelli2019].
-
 - [ ] Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase.
-For instance ‚discrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impacts of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].
-
+For instance ‚ discrepancies between model output and observed output may be important for forecasting models, where cost, benefit, an risk over a substantial period must be gauged, but much less critical for decision-making or management models where the user may be satisfied with knowing that the predicted ranking order of impacts of alternative scenarios or management options is likely to be correct, with only a rough indication of their sizes" [@Jakeman:2006ii].
 - [ ] Briefly describe how you will summarise the results of these in silico experiments with graphical, tabular, or other devices, such as summary statistics.
-
 - [ ] If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic parameter covariance, [@Jakeman:2006ii]), specify which measure of uncertainty you will use.
@@
 
-### Sensitivity Analyses
+### Sensitivity analyses
 
@@
 -   [ ] Describe the sensitivity analysis approach you will take: deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).
-
 -   [ ] Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameters will be varied.
-
 -   [ ] State the primary objective of each sensitivity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or alternatively to identify which variables or factors contribute little to overall model outputs, and so can be 'dropped' from future iterations of the model [@Saltelli2019].
@@
 -   [ ] Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario specified in section 1.
-
 -   [ ] Describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections.
-
 -   [ ] What raw and transformed model outputs will you extract from the model simulations or projections, and how will you map, plot, or otherwise display and synthesise the results of scenario and model analyses.
-
 -   [ ] Explain how you will analyse the outputs to answer your analytical objectives.

### 2024-09-22 preregistration_modelling.qmd
Remove duplicate preregistration item
qualify preregistration with explanatory clause

@@
 Are there any linked problems that your model should consider?
--   [ ] Ensure that you specify any focal taxa and study objectives,
--   [ ] Briefly outline the ecological problem and the decision-problem, including the decision-trigger and any regulatory frameworks relevant to the problem.
+-   [ ] Ensure that you specify any focal taxa and study objectives.
 :::
@@
 -   [ ] How can the data be accessed? Provide a link or contact as appropriate, indicating any restrictions on the use of data.
--   [ ] Date of download, access, or future access.
+-   [ ] Date of download, access, or expected timing offuture access.
 -   [ ] Describe the source of the data - what entity originally collected this data?

### 2024-09-22 preregistration_modelling.qmd
add todo
Add qualifying clause for explanation and remove todo comment

@@
 
-Before calibrating the model to the data, the performance criteria on which the calibration is judged are specified.
+Before calibrating the model to the data, the performance criteria for judging the calibration (or model fit) are specified.
 These criteria and their underlying assumptions should reflect the desired properties of the parameter estimates / structure [@Jakeman:2006ii].
@@
 
-<!-- #TODO NOTE THIS DOESN"T SAY ANYTHING ABOUT PARAMETER ESTIMATION! ONLY PErformacne criteria -->
 :::
@@
 Note that the performance metrics and criteria in this section are used for evaluating the structured and parameterised models (ideally) on independent holdout data, so this step is additional to any performance criteria used for determining model structure or parameterisation in section X.
+
+<!-- #TODO add section-cross referencing -->
 :::

### 2024-09-22 _analysis/ms.qmd
Add gt table with quarto template items

@@
 
+```{r}
+#| label: tbl-preregistration-template-final
+#| tbl-cap: "Final version of preregistation template developed after user-testing."
+
+library(tidyverse)
+library(gt)
+prereg_tbl_data <- 
+  readr::read_csv(
+    here::here("outputs/quarto_preregistration_template.csv")
+) %>% 
+    drop_na(phase) %>% 
+    nest(.by = phase) %>% 
+    mutate(phase_id = row_number()) %>% 
+    unnest(data) %>% 
+    nest(.by = c(phase, 
+                 step)) %>% 
+    group_by(phase) %>%  
+    mutate(step_id = row_number()) %>% 
+    unnest(data) %>%  
+    ungroup %>% 
+    nest(.by = c(phase, 
+                 step, 
+                 substep)) %>% 
+    group_by(phase, step) %>% 
+    mutate(substep_id = row_number()) %>% 
+    unnest(data) %>% 
+    ungroup %>% 
+    nest(.by = c(phase, 
+                 step, 
+                 substep, 
+                 preregistration_item)) %>% 
+    group_by(phase, step, substep) %>% 
+    mutate(preregistration_item_id = row_number()) %>% 
+    unnest(data) %>% ungroup %>% 
+    nest(.by = c(phase, 
+                 step, 
+                 substep, 
+                 preregistration_item, 
+                 preregistration_subitem)) %>% 
+    group_by(phase, step, substep, preregistration_item) %>% 
+  mutate(preregistration_subitem_id = row_number()) %>% 
+    unnest(data) %>% 
+  ungroup %>%  
+    mutate(preregistration_item = 
+             case_when(preregistration_subitem_id != 1 ~ NA, 
+                       .default = preregistration_item),) %>% 
+    mutate(substep = 
+             case_when(substep_id != row_number() ~ NA, 
+                               .default = substep),
+           substep_description = 
+             case_when(substep_id != row_number() ~ NA, 
+                       .default = substep_description), 
+           .by = substep) %>% 
+    mutate(step = 
+             case_when(step_id != row_number() ~ NA, 
+                            .default = step),
+           step_description = 
+             case_when(substep_id != row_number() ~ NA, 
+                                        .default = step_description),
+           .by = step)  %>% 
+    mutate(across(everything(), ~ stringr::str_remove_all(.x, "\\*")),
+           phase = gluedown::md_bold(phase),
+           phase_description = gluedown::md_italic(phase_description),
+           across(ends_with("_id"), as.numeric))
+
+prereg_tbl_data %>% 
+    select(starts_with("phase"), starts_with("step"), starts_with("substep"), ends_with("_item"), ends_with("subitem"), everything()) %>% 
+    mutate(step = case_when(!is.na(step) ~ gluedown::md_bold(step), .default = step),
+           step_description = case_when(!is.na(step_description) ~ gluedown::md_italic(step_description), .default = step_description)) %>% 
+    gt::gt(groupname_col = c("phase", "phase_description"),
+           row_group.sep = ": ") %>% 
+    gt::cols_hide(ends_with("_id")) %>% 
+    gt::sub_missing(missing_text = "") %>% 
+    gt::text_transform(fn = function(x) map(x, gt::md), 
+                       locations = gt::cells_row_groups()) %>% 
+   gt::text_transform(fn = gt::md, 
+                       locations = gt::cells_body()) %>% 
+    gt::cols_merge(columns = c("step", "step_description"), pattern = "{1} {2}") %>% 
+    gt::cols_merge(columns = c("substep", "substep_description"), pattern = "{1} {2}") 
+```
+
+gt
+
 The template used in the VEFMAP study underwent a series of changes to both its content and visual formatting during the analysis to better suit the needs of the modellers.

### 2024-09-25 _analysis/ms.qmd
Add quarto preprint output extension and update YAML
update prereg template table organisation

--- 2024-05-05 _analysis/ms.qmd
@@
 format:
+  preprint-typst:
+    fontsize: 11pt
+    mainfont: Libertinus Serif 
   docx: default
-  pdf: default
+  preprint-docx:
+    toc: false
+    number-sections: true
+    highlight-style: github
   elsevier-pdf:
@@
 ### Final Template Version & Changes
+
+```{r}
+#| label: preregistration-template-final
+#| eval: true
+#| echo: false
+#| message: false
+
+library(tidyverse)
+library(gt)
+library(gtExtras)
+
+prereg_tbl_data <- 
+  readr::read_csv(
+    here::here("outputs/quarto_preregistration_template.csv")) %>% 
+  drop_na(phase) %>% 
+  mutate(across(everything(), ~ stringr::str_remove_all(.x, "\\*")),
+         phase = forcats::as_factor(phase), 
+         step = forcats::as_factor(step), 
+         substep = forcats::as_factor(substep)) %>% 
+  mutate(phase_id = cur_group_id(), .by = phase) %>% 
+  group_by(pick(starts_with("phase"), 
+                starts_with("step"), 
+                starts_with("substep"),
+                -ends_with("_id")), 
+           preregistration_item) %>% 
+  reframe(preregistration_subitem = list(preregistration_subitem), 
+          subitem_na = map_lgl(preregistration_subitem, rlang::is_na)) %>% 
+  group_by(pick(starts_with("phase"), 
+                starts_with("step"), 
+                starts_with("substep"),
+                -ends_with("_id"), 
+                "subitem_na")) %>% 
+  reframe(preregistration_item = 
+            case_when(subitem_na ~ list(preregistration_item), 
+                      .default = map(preregistration_item, list)),  
+          preregistration_subitem = 
+            case_when(subitem_na ~ list(NA), 
+                      .default = map(preregistration_subitem, 
+                                     ~ gluedown::md_list(.x, type = "bullet", marker = "-")))) %>% 
+  rowwise() %>% 
+  mutate(preregistration_item = 
+           list(gluedown::md_list(preregistration_item, type = "bullet", marker = "-"))) %>% 
+  arrange(phase, step, substep) %>% 
+  ungroup %>%  
+  mutate(step_id = 1:n(), 
+         step = case_when(step_id == 1 ~ step, .default = NA),
+         step_description = case_when(step_id == 1 ~ step_description, .default = NA),
+         .by = c(phase, step)) %>% 
+  mutate(substep_id = 1:n(), 
+         substep = 
+           case_when(substep_id == 1 ~ substep, 
+                     .default = NA),
+         substep_description = 
+           case_when(substep_id == 1 ~ substep_description, 
+                     .default = NA),
+         .by = c(phase, step, substep)) %>% 
+  mutate( phase = gluedown::md_bold(phase),
+          phase_description = gluedown::md_italic(phase_description),
+          across(ends_with("_id"), as.numeric)) %>% 
+  rowwise() %>% 
+  mutate(step = 
+           forcats::fct_relabel(step, 
+                                .fun = ~ if(!rlang::is_na(.x)) gluedown::md_bold(.x)),
+         substep = 
+           forcats::fct_relabel(substep, 
+                                .fun = ~ if(!rlang::is_na(.x)) gluedown::md_bold(.x)),
+         step_description = 
+           forcats::fct_relabel(step_description, 
+                                .fun = ~ if(!rlang::is_na(.x)) gluedown::md_italic(.x)),
+         substep_description = 
+           forcats::fct_relabel(substep_description, 
+                                .fun = ~ if(!rlang::is_na(.x)) gluedown::md_italic(.x))) %>% 
+  ungroup %>% 
+  select(starts_with("phase"), 
+         starts_with("step"), 
+         starts_with("substep"), 
+         ends_with("_item"), 
+         ends_with("subitem"), 
+         everything(),
+         -subitem_na)
+
+prereg_tbl_data %>% 
+  gt::gt(
+    groupname_col = c("phase", "phase_description"),
+    rowname_col = "step",
+    row_group.sep = ": ") %>% 
+  gt::cols_hide(ends_with("_id")) %>% 
+  gt::sub_missing(missing_text = "") %>% 
+  gt::cols_merge(columns = c("step", "step_description"),
+                 pattern =   "{1}\n {2}",) %>% 
+  gt::cols_merge(columns = c("substep", "substep_description"),
+                 pattern = "{1}\n {2}") %>% 
+  gt::text_transform(fn = gt::md, 
+                     locations = gt::cells_stub()) %>%
+  gt::text_transform(fn = function(x) map(x, gt::md), 
+                     locations = gt::cells_row_groups()) %>% 
+  gt::text_transform(fn = gt::md, gt::cells_body()) %>% 
+  gt::sub_missing(missing_text = "") %>% 
+  gt::cols_align(align = "left") %>% 
+  # gt::cols_label(substep = "Substep") %>% 
+  # gt::tab_spanner(columns = starts_with("preregistration_item"), 
+                  # label = "Preregistration Items", gather = TRUE) %>% 
+  gt::tab_options(column_labels.hidden = TRUE, table.width = 10,table.font.size = 10) %>%
+  gt::opt_stylize(2,color = "cyan") %>% 
+  gt::tab_style(style = gt::cell_text(v_align = "top", whitespace = "break-spaces"), 
+                locations = cells_body()) %>% 
+   gt::tab_style(style = gt::cell_text(v_align = "top", whitespace = "break-spaces"), 
+                locations = cells_stub()) 
+```
 

### 2025-03-03 _analysis/ms.qmd
Update Introduction & Methods

--- 2024-05-05 _analysis/ms.qmd
@@
 abstract: |
-  1.
-Preregistration is an open-science practice adopted in scientific disciplines that have begun to confront the ‘reproducibility crisis’, which aims to improve research transparency and mitigate questionable research practices within a study.
-There has been little uptake of preregistration in ecology and conservation, and many modellers and researchers engaged in non-hypothesis testing research both within and outside of ecology have eschewed pre- registration on the grounds that existing templates are irrelevant because they are focused on null-hypothesis significance testing.
-  2.
-We advocate that preregistration can and should be used in model-based research in ecology and conservation, but that preregistration should reflect the norms and practice of the research context in which it is applied, in order to adequately restrict researcher degrees of freedom that are unique to a domain-specific methodology.
-We took a user-centred design approach to the task of translating preregistration into model- based research contexts for ecology and conservation.
-  3.
-We identified a generalised ‘modelling workflow’ that captures critical analysis decisions in the process of the model development cycle.
-The workflow informed the content and structure of a preregistration template to be used in model-based research.
-The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the current format and linear process of preregistration.
-To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose a methodology for implementing ‘adaptive preregistration’ leveraging git and GitHub.
-We tested and evaluated the preregistration template and methodology using a case- study of environmental flows modelling for decision-support in Victoria, Australia.
-  4.
-This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings and will facilitate the wider uptake of preregistration in ecology.
-Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model- based re- search within science more broadly.
-Modellers need no longer cry “but, I can’t preregister my research.”
+  1.	Preregistration is an open-science practice which aims to improve research transparency and mitigate questionable research practices, like cherry picking results.
+It helps protects against cognitive biases, like hindsight bias, that can influence how study outcomes are interpreted.
+There has been little uptake of preregistration in ecology and conservation, arguably because existing pre-registration templates focus on null-hypothesis significance testing whereas ecology and conservation often rely on different types of statistical modelling.
+  2.	We argue that preregistration in model-based research in ecology and conservation is both possible and beneficial, using templates adapted for domain-specific methodologies.
+We applied a user-centred design approach to translate the concept of preregistration into model-based research practice for ecology and conservation.
+  3.	We developed a generalised ‘modelling workflow’ to capture critical analysis decisions in the model development cycle, informing a preregistration template for model-based research.
+The iterative nature of model development is incompatible with existing, linear preregistration templates, creating a barrier to implementing preregistration for model-based research.
+  4.	To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose, test and evaluate a methodology for ‘adaptive preregistration’, using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
+  5.
+This research provides a template and methodology for implementing adaptive preregistration of ecological models.
+Although we focus on ecology and conservation in this paper, the concept of adaptive preregistration, and the templates developed here, could be applied in model-based research in other scientific disciplines within science more broadly.
+Modellers need no longer cry “but I can’t preregister my research.”
+
 keywords:
@@
 format:
-  docx: default
-  pdf: default
+  # preprint-typst:
+    # fontsize: 11pt
+    # mainfont: Libertinus Serif 
+  # docx: default
+  preprint-docx:
+    toc: false
+    number-sections: true
+    highlight-style: github
   elsevier-pdf:
@@
     keep-tex: true
-bibliography: references_obsidian.bib
+bibliography: "references_obsidian.bib"
+csl: methods-in-ecology-and-evolution.csl
 ---
@@
 
-Science's replication crisis, and the subsequent recognition of problems with reliability of scientific research findings, has triggered a 'credibility revolution' spawning a suite of proposed methodological and policy reforms, including *preregistration* [@parker2019].
-Preregistration is an open-science practice that aims to distinguish between planned and ad hoc analyses  [@Parker2019] and aims to combat the risk of questionable research practices by requiring researchers to register their methods and analysis plan in a secure and publicly accessible platform prior to collecting and/or analysing data, which cannot be altered after submission ( #todo #citation).
-The purpose of preregistration is to prevent biased research findings by guarding against questionable research practices (such as selective reporting, p-hacking, HARKing, among others #todo #citation ) by distinguishing what is a genuine *a priori* planned analysis, and what is not, thereby preventing the opportunistic misuse of researcher degrees of freedom, or those many analytic decisions researchers make during the planning, data collection, data analysis and reporting of results in scientific studies (See Box 1 for background).
-QRPs have been labelled as one of the primary cause of the replication crisis (Shrout and Rodgers, 2018 in [@Fife2022]) and serve to XYZ bias the overall literature, etc ( #todo finish this point).
+Over the past ten years or so, a ‘replication crisis’ in some scientific fields has exposed problems with the reliability of research findings [@Neoh2023; @OpenScienceCollaboration:2015cn; @camererEvaluatingReplicabilitySocial2018; @Ioannidis:2014cm].
+One such problem is the prevalence of Questionable Research Practices (QRPs), including p-hacking, Hypothesising After Results are Known (HARKing), and cherry-picking [@Wicherts2016].
+QRPs increase the probability of false positive results, artificially inflate effect size estimates, and ultimately reduce the replicability of published findings, while encouraging overconfidence in the precision of results and exacerbating the risk of accepting and propagating false facts [@Hoffmann2021; @Roettger2019; @Nissen2016].
+Self-reports from researchers show that QRPs are surprisingly common across disciplines [@John:2012eo; @Agnoli:2017kl; @Makel2023], including in ecology [@Fraser:2018cl; @kimmel2023].
+Although these self-report surveys of QRPs are heavily focused on statistical significance testing research employing p-values, some broader concepts, like selective reporting/cherry picking, have counterparts in other inferential frameworks.
 
-Despite enthusiastic uptake in some scientific disciplines, such as psychology and medicine ( #todo #citation ), the use of registered reports and preregistration in ecology and conservation more broadly is new and limited [@Desbureaux2021], particularly for model-based research ( #todo Joel Pick preprint?)**.
-Ecologists have brushed preregistration off on the basis that existing templates are irrelevant since they are focussed on null-hypothesis significance testing**, whereas in applied ecology and conservation, we don't usually explicitly test hypotheses [@Betts2021], but instead use explanatory models, or increasingly predictive models that generate anticipatory predictions, to inform management decisions [@Wood2020].
-Model-fitting and evaluation, rather than falsification of hypotheses about natural phenomena, are increasingly prevalent approaches to scientific inquiry in ecology [@Connolly2017]
+QRPs are precipitated when researchers opportunistically or systematically misuse ‘researcher degrees of freedom’ without disclosure [@Bakker2018; @Wicherts2016].
+Researcher degrees of freedom refer to the many decisions and alternative analytic choices researchers make throughout the research process [@Simmons:2011iw; @Wicherts2016].
+Analytic decisions are typically subjective but methodologically defensible, however, choices and omissions taken at each ‘fork’ may collectively and significantly influence the final ‘analysis strategy’ [*sensu* @Hoffmann2021], ‘specification’ [*sensu* @Simonsohn2020], or ‘modelling path’ [*sensu* @hamalainen2016].
+The set of all (plausible or reasonable) possible decisions that a researcher could make over the entire analysis from beginning to end has been termed collectively as the ‘garden of forking paths’ [@Gelman2013] or the ‘multiverse’ [@Steegen2016].
+‘Many-studies’ have now demonstrated the consequences of the ‘garden of forking paths’, whereby alternative analytical decisions can lead researchers to make vastly different interpretations of results, with conflicting conclusions, even when investigating the same research question using the same underlying dataset [@Gould2025; @Hoogeveen2022; @Silberzahn2018; @Desbureaux2021; @Young2021].
 
-Preregistration in model-based research has been eschewed on several grounds by modellers and researchers engaged in model-based research both within and outside of ecology.
-Much of the literature advocating for the use of preregistration assumes a hypothetico-deductive model of scientific inquiry, and the templates used to write preregistrations have been accused of being focussed on null-hypothesis significance testing [e.g.
-@McDermott2021; @Prosperi2019].
-Approaches and advocacy for preregistration of MBR across all scientific disciplines are limited, and focussed have focused on the phases of the modelling process and modelling approaches that are more confirmatory ( #todo see @ioannidis2022).
-For example, @Cruwell2019b present a template for preregistration of MBR where the model construction and development phase is explicitly and intentionally excluded from the preregistration process, while much of the discussion around preregistration for models has centred on machine learning [@Hildebrandt2018] or predictive modelling [@hofman2023] ( #todo see \[preregistration for modelling\] for more refs).
+Good intentions and simply being ‘aware’ of the potential for researcher degrees of freedom to lead to QRPs cannot sufficiently mitigate their risk because of the unconscious nature of their origin [@Nosek:2018eg; @Zvereva2021].
+Fortunately, there are some potential solutions to this problem.
+Preregistration is an open-science practice that aims to distinguish between planned and ad hoc analyses [@Parker2019] thereby restricting opportunities for trialling analytical alternatives (‘researcher degrees of freedom’) and making any deviations from the preregistered plan transparent [@Wicherts2016].
+Preregistration requires researchers to register their methods and analysis plan in a secure and publicly accessible platform prior to collecting and/or analysing data, which cannot be altered after submission [@Parker2019].
+Preliminary empirical evidence indicates that studies with preregistered analysis plans are less likely to have positive or significant results and are more transparently reported [@brodeur2024; @toth2021], especially when the preregistration template includes detailed instructions and requirements [@Bakker2018] or the preregistered study is published as a ‘Registered Report’ [@scheel2021] – a new publication format that aims to combat selective reporting and publication bias wherein the peer-review process and decision to publish is made prior to the results are known.
 
-A second argument against preregistration for model-based research is based on the mischaracterisation of the model development process as a intrinsically exploratory process [@McDermott2021] , and since only the preregistration of confirmatory research, such as hypothesis testing, is considered necessary [@Prosperi2019], there is no need to pre-register model-based research ( #todo #citation).
-The description of modelling as a wholly exploratory process is a mischaracterisation based on a misunderstanding of sensemaking and reasoning within the modelling process stemming from a lack of clear formal definitions distinguishing confirmatory research from exploratory research [@Flournoy2020; @Devezer2020; @alspaugh2019].
-Although modelling includes some tasks and activities that are exploratory in nature, it can be distinguished from wholly *exploratory analysis* in that tends be goal-driven, directed analysis that is less open-ended [@alspaugh2019].
-For model-based research, the exploratory/confirmatory divide is an idealised and simplifying dichotomy that, in practice, is more complicated and nuanced and can be considered as a spectrum on which exploratory analysis sits at the analysis is open-ended, high-level and opportunistic end, to goal-driven, low-level and precise analyses at the other [@alspaugh2019].
-In practice the divide is often blurred, and researchers will move back and forth between different types of reasoning and inference throughout the modelling process, switching from activity at one end of the spectrum to the other and back again [@alspaugh2019], with deductions leading to abductions as the analysis process proceeds [@alspaugh2019; @Prosperi2019], or even occuring together in a hybridized manner [@Prosperi2019].
+One potential objection to preregistration for modelling is that model development is an intrinsically exploratory process [@McDermott2021], while preregistration is for confirmatory research [i.e.
+hypothesis or theory testing, @Prosperi2019].
+These conceptions of modelling and preregistration are not entirely wrong, but they are oversimplified.
+In practice, the distinction between exploratory and confirmatory activities is not a sharp one [@Flournoy2020; @Devezer2020; @alspaugh2019].
+Modelling includes some tasks and activities that are exploratory in nature, but it can be distinguished from wholly *exploratory analysis* in that modelling is often goal-driven and directed [@alspaugh2019].
+In practice, researchers move back and forth between different types of reasoning and inference throughout the modelling process, with deductions leading to abductions as the analysis proceeds [@alspaugh2019; @Prosperi2019] or even occurring together in a hybridized manner [@Prosperi2019].
+This bears a strong resemblance to more general philosophical observations about the iterative nature of science [@Chang2004].
 
-When critics of preregistration for MBR describe modelling as intrinsically 'exploratory', what is seemingly being referred to is the iterative, adaptive and non-linear nature of model-development.
-In practice, modelling proceeds in an adaptive manner by which models are iteratively developed and refined [@Augusiak:2014gz; @Babel2019; @Bennett2013; @Bodner2020].
-New analyses are often chosen based on the basis of data exploration and previous analyses of the same data [@Dwork2015], For example, modellers may perform exploratory analyses to validate distributional assumptions are met, such as no over-dispersion or zero-inflation in count-data to ensure valid inference when using particular models [@Campbell2021] ( #todo also see @alspaugh2019 for list of examples).
-Such practices violate a key tenet of preregistration, that is 'data-independent' analytic decision-making ( #todo #citation).
-Consequently, the iterative nature of model development, wherein models are continually refined through analysing and reanalysing a dataset and the model, is considered to be in conflict with the practice of preregistration; by requiring researchers to commit to a single static plan that prevents changes or deviations from the original plan it is seen to be forcing adaptive analysis to be *non-adaptive* through 'defining the entire analysis protocol ahead of time' [@Dwork2015; @Benning2019; @MacEachern2019; @McDermott2021].
-In short, we can't write off preregistration for modelling on the basis that it is exploratory, because in practice it employs a range of modes of reasoning from induction, deduction, to abduction, with analysts switching between those types of reasoning or even hybridising them.
+In this iterative and non-linear process of modelling, initial analyses often inform future analyses of the same data whereby the modeller’s understanding of the problem and the model itself are incrementally adjusted in light of intermediate results from previous decision points [@hamalainen2016; @Schweinsberg2021; @Dwork2015].
+Modellers are typically uncertain about which model is most appropriate for the data [@popovic2024] and are reluctant to risk registering a modelling strategy that may be incompatible with the data [@Roettger2019].
+For example, modellers may perform exploratory analyses to validate distributional assumptions are met [@Campbell2021], which may result in subsequent changes to the model specification.
+Should data-dependent decisions like these be interpreted as violating a key tenet of preregistration, that is, failing to define the entire analysis protocol ahead of time [@Dwork2015; @Benning2019; @MacEachern2019]?
+Our answer to this is ‘no’, and within this paper we outline a methodology for implementing an expanded view of preregistration called *Adaptive Preregistration* [*sensu* @Srivastava2018], which captures decision points and reasoning about modelling choices, but which embraces the data-dependency practices that constitute the modelling development process.
+This goes beyond the work of others who offer discussion and guidance about when and how to deviate from a preregistration [@Lakens2024; @willroth2024], instead providing an approach to preregistration embodying the principles of *registered flexibility* [@Roettger2019; @Benning2019; @Srivastava2018] and iterative or *interim preregistrations* [@ioannidis2022; @hofman2023].
 
-What *is* important to consider when considering the application of preregistration within MBR is whether the reliability of MBR may be potentially threatened by the potential for Questionable Research Practices (Box 1).
-Evidence from ecology more broadly has revealed that the rate of QRPs is similar to those in other disciplines [@Fraser:2018cl] and while there has been little study of QRPs in MBR (except @Gould:2020ebg #todo preprint doc & cross-ref here), preliminary evidence reveals that modellers do engage in QRPs such as selective reporting, for example, they may may omit explored analysis paths from their reported analyses if they did not corroborate their conclusions [@Liu2020].
-Furthermore, QRPs are likely to arise in conditions where there is incomplete and opaque reporting practices [@Fidler:2018ho], and poor model-reporting practices and an overall lack of transparency about the modelling process are endemic in MBR in ecology.
-Concerns about the state of model reporting in ecological modelling have been raised repeatedly, arguing that current practices are neither transparent or comprehensive [@Schmolke:2010fd; @deVos:2011uk].
-Despite a number of attempts to promote thorough and transparent reporting, (e.g.
-@Jakeman:2006ii; @Alexandrov:2011iv; @Bodner2020) as a critical component of *Good Modelling Practice*, a culture of adequate research transparency in modelling is yet to be established and remains limited [@deVos:2011uk; @Grimm:2014es].
-The lack of transparency in describing ecological models and reporting the modelling process and results provides significant opportunity for undisclosed researcher degrees of freedom, and hence, the potential for Questionable Research Practices.
-
-Although the purpose of preregistration is predominantly conceptualised in terms of distinguishing exploratory from confirmatory research (e.g.
-@Nosek:2018eg, @munafo2017 and #todo insert other refs), we we adopt a more general stance towards the purpose of preregistration, i.e. that preregistration may mitigate against questionable research practices by reducing the potential for cognitive biases to cause opportunistic misuse of researcher degrees of freedom.
-In this paper, we advocate that preregistration can and should be used in model-based research in ecology and conservation, but its particular form should reflect the norms and practices of the research context in which it is applied, in order to adequately restrict researcher degrees of freedom that may uniquely emerge within this domain-specific methodology.
-
-The goal of this study is to rethink preregistration to better suit the domain-specific challenges and methodological particularities of model-based research in ecology and conservation.
-Preregistration cannot simply be ported over to other disciplines and adopted as-is.
-Different research methodologies, and their domain-specific application across different domains and fields of study follow different analytical choices and are therefore susceptible to different QRPs.
-Consequently, domain-specific templates are necessary for guiding preregistration in such a way that restricts researcher degrees of freedom as they uniquely emerge within a domain or field-specific application of a particular research methodology.
-If we wish to adopt preregistration for ecological modelling, both the content required to be preregistered, as well as the procedure for creating a preregistration must be adapted to suit model-based research as it is undertaken in ecology.
-
-We first aimed to build a preregistration template for ecological modelling.
-We identified a generalised ‘modelling workflow’ that captured critical analysis decisions in the process of the model development cycle.
-The workflow informed the content and structure of a preregistration template to be used in model-based research.
-The iterative nature of model development was identified as a key barrier to implementing preregistration for modelling research because it conflicts with the current format and linear process of preregistration.
-To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we devised a methodology for implementing an expanded view of preregistration that we term *Adaptive Preregistration* [sensu @Srivastava2018].
-Adaptive preregistration allows for the iterative cycle of model development to be reflected in the preregistration and better accounts for the inherently adaptive nature of model development than current approaches to preregistration in which researchers are committed to a static, unalterable and deterministic view of preregistration.
-Finally, to assess the feasibility and practicality of preregistration for MBR, we tested and evaluated the preregistration template and methodology using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
-
-------------------------------------------------------------------------
-
-# Box 1: How Researcher Degrees of Freedom become Questionable Research Practices in MBR
-
-Modellers are faced with numerous 'researcher degrees of freedom' [@Wicherts2016] or subjective decisions during the course of modelling that may substantially impact the chosen analysis strategy, and therefore the research findings and conclusions drawn from that analysis [@Hoffmann2021; @Simonsohn2020; @Desbureaux2021; @gouldSameDataDifferent2023].
-The set of all plausible or reasonable potential decisions that a researcher could take from the beginning to end of analysis has been described as the 'garden of forking paths' [@Gelman2013].
-'Many analyst'-style studies, whereby multiple separate teams of researchers independently analyse the same dataset and same research question, demonstrate that different pathways in this 'garden of forking paths' can lead to potentially conflicting conclusions [@Young2021; @Liu2020; @gouldSameDataDifferent2023].
-
-Researcher degrees of freedom arise because there is inherent analytic uncertainty in model-based research whether due to incomplete theoretical understanding of underlying ecological processes, or because modelling is an act of simplification and abstraction; there is always some degree of choice about which ecological processes and phenomena should be included in the model, how they should be represented [@Babel2019; @Getz:2017ic].
-Further analytic uncertainty may arise due to uncertainty in data processing or choice about which analysis method is best [@Hoffmann2021].
-
--   *Analytic Uncertainty* [@Walker2003] talks about uncertainty specifically in terms of *model based* decision-making. #todo
-
-When researchers are faced with ambiguous decisions where there is no single decision or strategy acceptable according to scientific standards and consensus, we are likely to select the decision that results in the most 'favourable' finding with "convincing self-justification" [@Simmons:2011iw; @Hoffmann2021].
-Researchers routinely execute and explore alternative analysis pathways in service of identifying the most favourable (i.e.
-publishable) result, selectively reporting only one or several preferred results [@Liu2020; @Simmons:2011iw].
-Thus researcher degrees of freedom precipitate 'questionable' research practices, or QRPs, when they remain both undisclosed and exploited opportunistically [@Bakker2018].
-Such 'questionable' research practices have been demonstrated to be surprisingly common upon admission among ecologists, with comparable rates to other scientific disciplines [@Fraser:2018cl].
-Questionable research practices are categorically distinct from fraudulent practices, arising not from consciously nefarious 'fishing expedition\[s\]' [@Gelman2013], but rather because researchers are motivated actors working within a 'publish or perish' culture who are susceptible to cognitive biases and are willing to change their modelling assumptions when the results conflict with desired outcomes, for example, sincerely and self-convincingly believing that the more favourable model is superior [@Young2021].
-
-QRPs may increase the probability of false positive results, may artificially inflate effect size estimates, and ultimately reduce the replicability of published findings, while encouraging overconfidence in the precision of results and exacerbating the risk of accepting and propagating false facts [@Wicherts2016; @Hoffmann2021; @Roettger2019; @Nissen2016].
-Good intentions, and simply being 'aware' of the potential for researcher degrees of freedom to lead to questionable research practices cannot sufficiently mitigate their risk [@Nosek:2018eg].
-
-------------------------------------------------------------------------
+In this study we aim to translate preregistration for application to model-based research, specifically aimed at ecology, conservation and related disciplines.
+Here, we first identify a generalised ‘modelling workflow’ that captured critical analysis decisions in the model development cycle and use it to build an Adaptive Preregistration template that can work for ecological modelling.
+We then tested and evaluated our template using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
+This is the first attempt at developing a fully-fledged preregistration template designed for model development and evaluation that we are aware of, and we propose it can be extended to work in disciplines outside of ecology.
 
@@
 
-We approached the task of translating preregistration into model-based research contexts for ecology and conservation with a user-centred design (UCD) framework.
-UCD is a philosophy for designing processes, services or products that account for user requirements at all stages of the design process so that the resource generated by this approach is effective and usable for specific people in specific contexts [@Pavelin2012].
-Because it is difficult to predict how users will actually engage with a resource [@Pavelin2012] and a designer's intention for preregistration may conflict with the user's purpose and undermine the overarching goal of preregistration [@Pu2019], we heed the advice of Pu et al.
-[-@Pu2019], who recommend treating preregistration as a UCD problem so that preregistration can effectively meet its intended objectives, despite UCD being typically used within software engineering domains, rather than academic domains.
+We modified a typical preregistration into model-based research contexts with a user-centred design framework, where users were engaged at all stages to produce an implementation of preregistration that is fit for purpose [@Pavelin2012, Fig.
+1].
 
-We borrow from the methodology proposed by Pavelin et al.
-[-@Pavelin2012] to design, test and refine a preregistration template suitable for ecological modelling, and a methodology for preregistering model-based research (@fig-UCD).
-The methods description is structured following the key steps illustrated in @fig-UCD.
+<!--# Insert Figure 1 Here -->
 
-![Insert caption.
-Number each step so i can refer to it in the headings and text below.
-Make it clear in the figure that we're not doing the build interface and conduct usability testing component of implement, nor the evaluate and iterate.
-we throw that out to the community of practice to do that part.](images/Pasted%20image%2020240314123346.png){#fig-UCD}
+Figure 1 Process for translating preregistration to ecological modelling research [adapted from @Pavelin2012].
+Note that the ‘Evaluate and Iterate’ component of the process was not undertaken in this study
 
-## Define Vision & Objectives
+## Problem Formulation: Defining Vision & Objectives for a preregistration template for model-based research
 
-> Define the goals of the research under development and identify the audience
+We first specified the research context, purpose, target users and use-cases, per (Pu et al.
+2019a; Pavelin et al.
+2012).
+Our context is model-based research, defined broadly as any research that uses quantitative modelling to answer its research question (Brudvig 2017).
+The purposes of the preregistration template and methodology are to:
 
-The first step in a UCD approach is to articulate the goals, objectives and target audience of the resource, consequently we define the intended purpose of the preregistration template and methodology, the community of users, as well as their research contexts and use-cases.
-Finally, we explicate the aspirations and desirable attributes of the template and methodology.
-Much like in 'Good Modelling Practice', articulating the overarching purpose, intended audience, as well as the objectives and aspirations will guide the development and evaluation of preregistration for model-based research in ecology and conservation, with important implications for the design of the template -- from its structure, content, user-interface and broader workflow in which it is designed to be used.
+1.  **Delimit 'researcher degrees of freedom'** [@pu2019a], to mitigate the the risk of questionable research practices when conducting model-based research in applied ecology and conservation.
+2.  **Increase ‘research transparency’ of ecological modelling,** including both:
+    1.  *production transparency* — including research artefacts like open-access data and materials, or data collection procedures [@Lupia2014],
+    2.
+*analytic transparency* — a complete account of how analytic conclusions are drawn from the data [@Lupia2014] where model choices, steps in the modelling process, assumptions and expectations about model outputs are clearly articulated [@Bodner2020].
 
-### Articulating Intended purpose of preregistration template & methodology
+We specified two use-cases for our Adaptive Preregistration template:
 
-We define the purposes of the template as:
+1.  Study authors or modellers seeking to preregister their modelling study,
+2.
+Editors, reviewers, and/or readers of the completed preregistration [@Pu2019; @pu2019a], who are tasked with verifying that analyses and findings reported in the final manuscript are consistent with the preregistered analysis plan.
 
-1\.
-*To delimit researcher degrees of freedom* \> when conducting model-based research in applied ecology and conservation so as to mitigate the risk of questionable research practices in order to improve the reliability and credibility of this research.
+Existing preregistration formats are designed primarily for study authors; they are static text-based documents that do not easily facilitate ‘preregistration checking.’ We designed our template to address this.
 
-2\.
-*Increase research transparency* \> of model-based research in ecology and conservation.
-*Research transparency* refers to both production transparency — which includes research artefacts like open-access data and materials, or data collection procedures — and analytic transparency — which describes a complete account of how the analytic conclusions are drawn from the data [@Lupia2014].
+## Designing the General Modelling Workflow
 
-### Target user community and their research-contexts
+We conducted a literature review to identify a modelling workflow suitable for application to a range of different modelling goals, e.g. exploration, inference and prediction, including both explanatory and anticipatory predictions [@Tredennick2021; @Mouquet2015; @Maris2018; @Shmueli:2010ec]; research contributions, e.g.
+development of new methods, or application / extension of existing methods and model transfers [@Yates2018]; types of models, e.g.
+phenomenological vs.
+mechanistic [@Connolly2017; @White2019; @lewis2023], across basic and applied contexts and across fields / disciplines in ecology and conservation [@Carmel2013; @Brudvig2017] while offering prescriptive guidelines based on ‘Good Modelling Practice’ [@Augusiak:2014gz].
+The final output was a workflow reflecting a balance between idealised and actual modelling in practice.
 
-We aim for the template and preregistration process developed in this study to be used for model-based research, which we define as any research, empirical or applied, that uses quantitative modelling to answer its research question [@Brudvig2017], where *quantitative modelling* may be defined as any modelling that "describes and/or forecast(s) the behaviour of a system using mathematical and statistical concepts, and whose parameters and their relationships are expressed as quantities" [@Garcia-Diaz2019].
+We developed an Adaptive Preregistration methodology, so that users have the flexibility to adapt the preregistration template to their own needs and processes [see user guide <https://egouldo.github.io/EcoConsPreReg/>, @gould2024].
+This can be facilitated in two ways: 1).
+The preregistration author can supply analytical or modelling heuristics or strategies that depend on the outcome of the previous decision-point or analysis, for example, decision-trees for certain points in the analysis or items in the preregistration template could be preregistered prior to analysing [@Baldwin2022a].
+2) using interim preregistrations that mark the transition between phases of modelling and analysis as different parts of the data are observed [@Srivastava2018].
 
-We define model-based research loosely in order to capture the breadth of different modelling goals (e.g.
-exploration, inference and prediction; @Tredennick2021) and type of research contributions (e.g.
-development of a new method, etc #todo see @Moallemi2019 and across both pure and applied ecological modelling.
+Modellers can preregister ‘plans to deploy flexible strategies’ [@Srivastava2018] describing heuristics consisting of multiple analysis or modelling strategies, whose execution depends on the outcome of previous decision-points or analyses.
+For example, the modeller can preregister a ‘decision-tree’ [@Baldwin2022a] consisting of pre-defined rules about when a particular modelling strategy or decision should be implemented.
+Modellers can specify a ‘flexible’ strategy within their preregistration when faced with methodological or analytic uncertainty that cannot be resolved without observing parts of the data or by engaging in preliminary analyses, model checking, or other data-dependent decisions.
 
-### Specifying Preregistration use-cases
+The modeller may follow an iterative process of preregistration in parallel with modelling, consisting of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observe [@Srivastava2018].
+As they proceed through the model development process the researcher shifts from ideation, preregistration and execution of the preregistered analysis plan and back again, generating interim preregistrations at different points within the model development process, depending on observed outcomes of the flexible heuristics described in the interim preregistration.
+The analyses executed in practice will depend on the observed outcomes of the registered flexible analyses.
 
-Pu et al.
-[-@Pu2019] describe two key use-cases of preregistration, depending on the role assumed in interacting with the preregistration.
-Users have different requirements of preregistration depending on the role that they assume, that is, study authors and modellers have a different use-case than reviewers, editors and readers, however this latter use-case is often not explicitly designed and accounted for in current platforms/infrastructure/processes [@Pu2019].
-Reviewers and editors are usually tasked with the process of 'checking' the preregistration, comparing the final manuscript to its preregistration and verifying that the preregistered analyses are conducted and reported as specified [@Pu2019].
-However existing formats are designed primarily for the use-case of the author - they are text-based static documents that are ill-suited to checking the reported study against the preregistration.
+Last, to facilitate transparency and preregistration checking of flexible modelling and analysis strategies, the results of any final and preliminary analyses must be linked back to the preregistered flexible strategy such that the link between a particular triggered analysis decision, and the analysis outcome that triggered the decision is explicit.
+Fig.
+2 demonstrates a generalised process for using git and GitHub as the vehicle for transparently documenting Adaptive Preregistration of a modelling study.
+The procedure leverages [GitHub’s tag and release feature](https://help.github.com/en/github/administering-a-repository/releasing-projects-on-github) in conjunction with [Semantic Versioning](https://egouldo.github.io/EcoConsPreReg/(https:/semver.org/)) to track, document and collaborate on changes to the preregistration document.
+The preregistration is stored alongside modelling code and results with GitHub issues [see @braga2023 for explanation] tracking analysis tasks and results for a discrete task described in the preregistration and providing the opportunity to document interpretation of discussion informing the rationale for potential changes or updates to the preregistration that generate interim preregistrations.
+The use GitHub issues is particularly handy for registered flexible analyses: the outcome of the registered flexible analysis is recorded within that issue, linked to the described heuristic, and modellers can use the discussion features to document interpretation of those results, and provide interpretation to explain which registered decision alternative is to be selected in light of those results.
 
-Although preregistration authorship is the primary target use-case of the preregistration template and methodology developed in this study, we aim to consider and incorporate features into the preregistration process, such as template formatting, that will facilitate preregistration checking to also address non-authorship use-cases.
+## Designing the Preregistration Template
 
-### Articulating desirable attributes of the preregistration template and adaptive preregistration methodology
+We facilitated a workshop with 10 ecologists and ecological modellers.
+A key task was to identifycommon modelling workflows and critical decision-steps in model development.
+‘Critical’ decision-steps are points in the modelling process where researchers, or modellers, make analytical decisions that could change the output of the analysis.
+In this activity, workshop participants first individually reflected on their own scientific process for a recent or memorable research project, listing the steps undertaken throughout the process.
+Next, participants collaboratively mapped the decision steps in their own personal modelling workflows onto a pre-prepared modelling workflow template (<https://osf.io/fgd23>), informed by our general workflow.
+We then collaboratively reviewed and refined the modelling workflow into a standard set of key modelling phases and commonly implemented decision steps.
 
-The preregistration template should capture critical decision-points that may influence the overall model and analysis as well as any recommendations or decisions drawn from analysis of the model, thus delimiting researcher degrees of freedom at these junctures in the modelling decision process.
-The structure and content of the template is necessarily opinionated, encouraging Good Modelling Practices, while reflecting community norms and practices in model-development while still meeting the objective of mitigating the risk of Questionable Research Practices by delimiting researcher degrees of freedom.
-The template needs to be generalisable enough to apply to a range of common modelling purposes, methods and research contributions, and flexible enough that if a particular decision is not applicable for the planned study, that those items may be removed or altered, and other items may be added onto the template.
+<!--# Insert Figure 2 - Adaptive Preregistration Workflow -->
 
-Secondary to the aim of mitigating QRPs, is that the the template should also enhance the research transparency of the preregistered study, ensuring that analysis choices are documented with sufficient transparency to facilitate the reader of the preregistered completed study in making informed assessments about the credibility and reliability of the model and any associated analysis and recommended interventions.
-Borrowing from @Bodner2020's framework to increase the accuracy, reliability and transparency of predictive models, in order to increase transparency of model-reporting, the template should explicitly outline critical model choices and their assumptions, steps in the modelling process, and explicit expectations about the outputs and analysis of the model.
-
-**Adaptive Preregistration Methodology**
-
-The Adaptive Preregistration Methodology aims to capture the entire modelling process, including model development and construction, rather than beginning at the point of analysing a pre-existing and parameterised model for deductive inference / prediction.
-Critically, the methodology must not preclude the iterative and adaptive nature of model construction, while simultaneously restricting the types of data-dependent analytic decision-making that may compromise the reliability and validity of the model and its contingent analysis and conclusions.
-
-## User Research & Analysis
-
-> Characterise users and their needs
-
-On the 4th and 5th of March 2020, we facilitated a series of structured activities at the Arthur Rylah Institute for Environmental Research which aimed to inform the development of two preregistration templates, one for field-work and sampling design, the other for modelling and data-analysis.
-We split the activities into two groups, to better focus discussion for each of these templates.
-
--   [ ] <mark style="background: #FFF3A3A6;">Note: might need to reframe this and omit the field study stuff, but should probably state that we did this but decided not to pursue?</mark>
-
-The objectives of the collaborative workshop activities were to aid in: 1.
-Understanding how modelling is done in practice to inform and tailor the preregistration templates to adequately capture key decision-points in the scientific modelling process; 2.
-Understanding how science in practice -- whether due to logistical, technical, feasibility or normative constraints -- differs from some idealised practice, in order to: 3.
-Identifying challenges and roadblocks to the use and uptake of preregistration in among ecological modellers, both in empirical and applied research contexts; and to 4.
-Using the above knowledge to inform the preregistration template and process design so they meet their intended purpose, and are adopted by our target user community for their particular use-cases.
-
-### Workshop 1: Identifying Ecological Modelling Workflows, Norms, and Practice
-
-The objective of this workshop was to identify modelling workflows that capture critical decision-steps in the model development cycle that should be included in a preregistration template.
-By "critical", we mean that without these decision-steps being included, the template would not adequately constrain researcher degrees of freedom nor b) transparently document the analytic decision-process.
-The activity consisted of two-parts, first, we individually reflected on our scientific process for a recent or memorable research project, listing the steps we undertook from the beginning to end.
-In the second part of this exercise, we came together as a group to collate each of the steps in our personal workflows onto pre-prepared workflow template (insert specific OSF link to the document, Supp Mat 1).
-The template consisted of several key 'phases' of the modelling process, with commonly implemented decision-steps grouped under each phase.
-The development of this modelling workflow sheet was derived from previous work by Gould et al.
-*in prep* (insert link to paper preprint) and is outlined in Appendix 1 ( #todo insert ).
-We sorted and categorised each decision-step in our personal workflows as belonging to a particular workflow phase.
-Next we grouped similar decision-steps that could be included as a single item on a preregistration template.
-At the end of this process we reviewed the suggested phases of the workflow to assess whether phases needed further discretisation, aggregation, removal, or amendment.
-
-### Workshop 2: Understanding ecology- and modelling-specific challenges to adopting preregistration in research practice
-
-This second activity aimed to identify critical challenges and roadblocks to undertaking preregistration in ecology and conservation in general, and for model-based research in particular.
-We facilitated a discussion structured around a series of pre-prepared questions and talking points informed by recent debates about preregistration in the literature ( #todo Supp Mat 1).
-These talking points were chosen with the aim of identifying problems in the content and or structuring of the phases and decision steps that were collectively identified in Workshop 1, as well as any conceptual or philosophical issues with preregistration in ecology, conservation, and in particular modelling, that might impact on the template structure, content and procedural implementation of preregistration.
-Discussion points included the following: - What should a research workflow look like for a modeller in an applied ecological setting who wishes to preregister their work?
-- What barriers are there for preregistering model-based research?
-- How do we accommodate the iterative cycle of model development into the preregistration process?
-- At what stage in a modelling or structured decision-making process should preregistration begin?
-- should we and how can we change the medium of the template and/or archiving platform to accomodate any roadblocks?
-- How could we change the medium of both the template and the archiving platform to accommodate these issues?
-- What procedural and or technical solutions could we implement to address these roadblocks to preregistration?
-- What decision-tools could be used to aid in completing a preregistration?
-- What sort of parsimony should we aim for in terms of the resolution of the templates and therefore their applicability across different methodologies?
-
-### Analysis of Workshop Outputs & Design
-
-> Analyse outputs of user-research workshops to inform the design of resource mockups
-
-**1. Identify Scientific Workflows for ecological modelling, both real/in-practice and idealised.**
-
-The goal of this analysis task was to describe a generalised workflow of the modelling cycle that will underpin the structure and content of the preregistration template.
-We used a combination of inductive and deductive coding ( #todo is this what it is called and #citation) using the qualitative coding software ATLAS.ti [-@atlasTI] to analyse the workshop outputs and generate a model of scientific workflow for model-based research that would constitute the presumed scientific workflow that would be used to guide the structure of the preregistration template.
-We collated and coded the workshop outputs using.
-All workshop outputs can be found here ( #todo Link to Workshop Outputs OSF repository), and the coded analysis can be found here ( #todo link to analysis files / screeen caps).
-
-A preliminary idealised scientific workflow for ecological modelling research was synthesised after reviewing the literature on ‘good modelling practice’ and structured decision-making literature in ecological management.
-Steps in the workflow were coded and broken down into ‘phases’ consisting of ‘steps’ and ‘sub-steps’, in line with the ontology we specified earlier for translating a modelling workflow into a preregistration template (section link), <mark style="background: #FF5582A6;">(insert link to appendices or just present the final?)</mark>.
-Each decision-point in the personal modelling workflows generated from activity one of the collaborative workshop were coded as belonging to one of these phases, steps and sub-steps.
-We revised the coding structure (inductively is this the word?), according to patterns and themes identified across multiple personal workflows.
-This workflow is used to provide structure and content to the preregistration template ( #todo insert link to this workflow).
-
-**2. Translating Workflows into Preregistration Templates**
-
-We translated the model-based workflow identified in part one of workshop analysis into a text-based preregistration template.
-Reflecting on existing templates ( #todo link to OSF repository of templates page, and others, namely secondary data analysis, see Olmo's links), we identified common elements / units of preregistration templates that we mapped onto the generalised modelling workflow identified in part 1 of workshop outputs analysis.
-Following the features of a good guideline identified by Moallemi et al.
-[-@Moallemi2019], we provide examples of choices or illustrations to clarify each step of the preregistration, especially for phases or steps in the the modelling process that are semantically, epistemically and/or philosophically contentious or ambiguous (link to paper describing contentious issues in modelling history)\_.
-We synthesised the following ontology with which to build our preregistration template: 1.        Decision Phase - each has a “title” and “description” or “definition”, steps are grouped under phases.
-(Note, that these are being scrapped from the template and are turned into tool-tips for formatting) 2.        Decision Step - each has a “title” and “description” or “definition” choices are grouped into ‘steps’, which corresponds to an activity or set of related activities in a scientific workflow.
-(again, these are being scrapped and turned into tool-tips.
-CHECK THIS, it's changed) 3.        Choices —\> each ‘choice’ corresponds to a uniquely numbered item on the PRT requiring a response from user.
-The style should be in the form of a directive or a question.
-E.g.
-“explain how you will do xyzzy”.
-Or “What performance measure will you use for assessing ‘goodness of fit’?”
-
-**3. Identifying challenges and solutions for implementing preregistration in model-based research**
-
-We analysed the outputs of Activity 2 in the collaborative workshop identifying barriers and solutions to undertaking preregistration in model-based research and to broader adoption of preregistration in the modelling community of practice.
-The analysis of Activity 2 in the collaborative workshop was used to inform the design of a suitable preregistration procedure for preregistering MBR in the domain of ecology.
-
-## User Testing
-
-> Build mock-ups of the product and test with users
-
-While user-testing is an essential step in the UCD framework, we also concur with @Devezer2020 that open-science reforms should be evaluated against their stated objectives.
-We therefore preregistered a real-world model-based research problem as a case-study for evaluating and testing the preregistration templates and proposed adaptive preregistration methodology.
-
-### Case Study
-
-We preregistered a real-world study of vegetation responses to environmental flows in Victoria, Australia.
-This study is one component of a larger set of studies within the Stage 6 Victorian Environmental Flows Monitoring and Assessment Program (VEFMAP), managed by Department of Energy, Environment and Climate Action (DEECA) and delivered through the Arthur Rylah Institute for Environmental Research.
-The case-study analysis aimed to evaluate the effectiveness of environmental-flows on riparian vegetation with the following objectives: - Quantify short and medium-term vegetation responses to individual flow events and regimes.  - Describe detailed vegetation trends associated with season, flow, rainfall and other factors such as livestock grazing and exotic vegetation  - Quantify the effect of various factors on vegetation in order to guide future management expectations and planning.
-
-Details of the study background, methods and results are described in Appendix A4.
-
-### Evaluation Criteria & Rationale
-
--   [ ] *Evaluation Criteria and Rationale (should this be moved to earlier aims section?* Seems like there's a little bit of overlap between the two...
-
-The content of the template should be 'parsimonious' in that the included items within the template should adequately a) constrain researcher degrees of freedom and by being specific, precise and exhaustive [@Wicherts2016] b) facilitate transparent documentation of the modelling process and analytic decisions during model development and analysis, while c) not being too detailed to the point that the preregistration is too difficult and time-consuming to complete.
-
-Many modellers [e.g.
-@MacEachern2019] #todo #citation are concerned about preregistration making modelling a more difficult task than it already is, and hampering the ability to meet project deadlines, especially in applied settings in which models directly underpin decisions.
-Modellers already struggle with sufficiently documenting models during publication and reporting , and there is a resource cost to increasing transparency of model-reporting [@deVos:2011uk].
-Consequently we will also assess the preregistration template and process on its perceived feasibility and practicality.
-
-### Evaluation Method
-
-Rather than attempting to reconstruct the process of creating the preregistration through using interview questions alone and digging back through the version history of a completed preregistration [e.g.
-@Pu2019], we aimed to capture the analysts 'thinking in the moment' by utilising the version-control and collaborative project management features of Git and GitHub (Ben, Smith, and Bailey 2018; Blischak, Davenport, and Wilson 2016) as a tool for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to ‘live-develop’ the template template.
-For example, if it turned out that we had missed an important item, or perhaps the phrasing of an item or the order and structure of the template needed to change whilst in the process of completing the preregistration, the suggested change, and its justification would be both recorded in GitHub and linked to the actual preregistration.
-
-Finally, we supplemented this ‘in situ’ testing and evaluation process with follow-up semi-structured interviews on 9th May 2024 in order to capture more detailed reflections of the analysts.
-Lead analysts of the case study were often under strict time constraints for meeting deliverables on the project, and the process of preregistration was a new and unfamiliar task, so capturing reflections on the broader preregistration process during preregistration and analysis of the case-study was not always possible.
-
-## Implementation
-
-> Implement the product, write technical implementations and build the product
-
-We wrote the initial template as a text-based document in Rmarkdown.
-The initial template used to preregister the case-study underwent a series of changes throughout the preregistration process, including changes to the organisation, content, help-text and visual appearance.
-The final preregistration template is made available as a quarto markdown extension ( #citation & link to doc) that can be installed and used to populate a new preregistration document for a new modelling study, and as a form that can be used with the `preregr` R package [@Gjalt-Jorn2023].
-
-To aid the case-study analysts in using the adaptive preregistration methodology to preregister their study, we wrote a detailed user-guide outlining the key principles of adaptive preregistration and a protocol for implementing adaptive preregistration using git and GitHub.
-The guide is presented as a publicly available website and also includes the template.
-
-# Results
-
-## User-research and analysis
-
-The outputs and analysis for both Workshop 1 and 2 is archived at (insert OSF link).
-The modelling workflow generated from the analysis of Workshop 1 that informed the development of the initial preregistration template is archived at the same link.
-During user-testing with our case-study, the preregistration template structure, and therefore its underlying modelling workflow on which it is based underwent, a series of changes.
-Consequently we present the revised workflow alongside the refined preregistration template in section X below.
-
-Discussions from Workshop Activity 2 focussing on challenges to preregistering MBR revealed that the iterative nature of model development is a significant barrier to adopting preregistration for model-based research.
-Modellers felt that adopting preregistration in their workflows was impossible because preregistration was understood to be a process that precludes iterative model development because its inner logic presumes a linear research workflow premised on hypothesis-testing research, and because it requires researchers to confirm an unalterable version of the analysis plan prior to collecting or analysing the data [@Pu2019; @mertens2019].
-
-The key tenets of preregistration are based on NHST-focused accounts of questionable research practices that prescribe the principle of 'decision-independence' where analysis decisions are made independently of and *a priori* to analysing the data [@Srivastava2018; @Gelman2013].
-However, analysts reflected on their personal modelling workflows and described situations in the modelling process that may violate the principle of 'decision-independence' that could not be avoided and were considered to be an intrinsic feature of the model-development process.
-During modelling a researcher will almost always engage in state-dependent or data-dependent decision-making -- that is, decisions about whether to alter the model in a subsequent round of model development depend on the results of model testing, evaluation and analysis at the end of the previous iteration of model development.
-
-Very often in modelling, there are decisions that are necessarily dependent on the outcome of previous analytic decisions in the modelling workflow [@Liu2020].
-Firstly, preliminary or investigatory analyses might need to be conducted before being able to specify future decision-steps in the analysis plan.
-For example, modellers might need to check the distribution of particular variables in order to determine how to specify the model, perform assumption checks, or check for collinearity or spatial autocorrelation.
-Secondly, data-dependent decisions from model checking may justifiably result in changes to either the analysis or to interpretations of the model [@Srivastava2018].
-For instance, inspection of the residuals and other model checking tests may force the modeller to return to earlier decision-points and change the planned analysis or model.
-Other times, the planned analysis and specified model may simply not converge, or the model is saturated and runs out of variation to apportion such that the model and or the model fitting algorithm must be respecified and re-implemented.
-Other times, the modelling process itself may generate knowledge and learning about the system, and the model structure changes throughout the process of model development.
-
-Data-dependent decision-making is not necessarily intrinsically *questionable*, but it does have the *potential* to reduce the reliability of the model and its findings if they are exploited in service of artificially inflating the accuracy or precision of the model, its predictions or evaluation tests to the effect that the model is perceived to be more credible than it would be if that practice did not occur -- essentially data-dependent decisions that lead to consumers of the model placing unsubstantiated belief in the reliability, validity and utility of the model and its outputs are questionable [@Gould:2019ebg] and can be mitigated through the practice of preregistration.
-
-The concept of preregistering all modelling choices *a priori* was seemingly overwhelming for analysts.
-They described situations where decisions are too difficult to anticipate, or cannot be made in advance, particularly for decisions occurring at the later phases of model construction and development process (Figure 2, TBC), where downstream decisions depend on the outcomes of earlier decisions and outcomes of modelling analyses.
-Some decisions might not be able to be specified on the first attempt at writing the preregistration until the model is fully or at least close to fully specified, for example, specifying precisely what and how sensitivity analyses or uncertainty analyses will be conducted.
-
--   Finally, sometimes there are some decisions that are too difficult to anticipate, or simply cannot be made in advance [@Srivastava2018].
--   I think Pu or Pavelin describe the same thing... they coin a name for these types of decisions too but I can't find it.
-
-## Design
-
-### Adaptive Preregistration: A methodology for preregistering model-based research
-
-In order to address the disjuncture between the iterative nature of model development and the single-use, deterministic preregistration procedure, we propose an expanded view of preregistration, *adaptive reregistration*, which allows for some degree of flexibility and adaptivity in the analysis workflow, while still mitigating against unconstrained and undisclosed researcher degrees of freedom.
-Adaptive preregistration breaks with the current preregistration workflow, wherein the author writes a single deterministic preregistration containing a single deterministic choice for every decision-step and the research shifts from ideation, to preregistration to execution of the plan (fig or #citation to ref?).
-Instead, in adaptive preregistration, as the modeller proceeds through the model development process, they will transition from ideation, preregistration, execution, and back to preregistration and execution again.
-The modeller may repeatedly cycle through the preregistration process rather than proceeding in a linear fashion.
-Two features facilitate this process: 1.
-An *adaptive preregistration* contains 'plans to deploy flexible strategies' [@Srivastava2018] whereby the preregistration author can supply a heuristic consisting of different analysis or modelling strategies whose execution depends on the outcome of the previous decision-point or analysis, for example, decision-trees for certain points in the analysis or items in the preregistration template could be preregistered prior to analysing [@Baldwin2022a].
-2.
-The preregistration may be iterative and consist of interim preregistrations that mark the transition between phases of modelling and analysis as different parts of the data are observed [@Srivastava2018].
-
--   [ ] Other refs to chase up... Benning, Schuwirth, Latihan, hamalainen, hofman, reusable holdout paper, Dwork?
-
-#### Preregistering Flexibility
-
-When preliminary analyses, model checking or other data-dependent analytic decisions are to be preregistered, the modeller can specify a 'flexible strategy' within their preregistration.
-This is particularly relevant for when it is uncertain as to whether the planned analyses will be appropriate for the data [@Baldwin2022a], which is a commonly encountered issue in ecological modelling, because often-times the model we might ideally want to fit that best represents our conceptual model of the target phenomena cannot be fitted due to a paucity of data ( #todo citaiton for this?
-was a motivating factor for the VEFMAP project).
-Like existing approaches to preregistration, the registered heuristic is still a deterministic one, but there there may be multiple potential choices rather than one, and their realisation depends on the outcome of an earlier decision step or analysis.
-
-To preregister flexibility, there are some critical requirements: 1.
-State what quantity or outcome needs to be known in order to move forward with the modelling and analysis, and 2.
-explain what test or analysis will be performed to obtain this quantity or outcome, and which parts of the data will be used in this analysis, and 3.
-describe how the results will be interpreted, linking each potential decision to its outcome trigger, where possible.
-
--   [ ] Might want to give an example of this...?
-Or could give example in reporting the case study and cross-ref that?
-Alternatively, could put an example in the guide, and send readers there?
-I think second option is best.
--   [ ] SEE ALSO OTHER refs in this vault, the idea of registering flexibility e.g. using decision-trees is not new.. which paper was it?- e.g. BENNING?
-
-#### Interim Preregistrations
-
-Just as model development proceeds in an iterative process of preregistration, so too does the process of adaptive preregistration.
-As the modeller proceeds, they will implement parts of the analysis plan that were specified in the previous preregistration step, letting the outcomes of that preregistered analysis inform either the specification or the realisation of the pre-registered specification in the next step of the analysis workflow (@fig-adaptive-preregistration).
-
-The initial preregistration should have as many items as possible, if not all items, in the preregistration template answered.
-If an item cannot be answered with a single decision because that decision is data-dependent, then the analyst should register flexibility for that item following the aforementioned requirements.
-Depending on observed outcomes of the registered decision-trees or flexible heuristics, interim preregistrations are created at multiple points in the model development process, each time there is an addition or amendment to the plan.
-
-### Integrating additional data handling and modelling practices
-
-We propose additional data handling and analysis techniques be embedded into the both the modelling and adaptive preregistration processes to help avoid the pitfalls of data-dependent decisions, such as partitioning the data to create a holdout dataset or using blinded datasets for data exploration.
-
-Partitioning data into train/test splits is a modelling technique that has been suggested for incorporation into preregistration for predictive modelling [@hofman2023] or other non-hypothesis testing research [@Baldwin2022a; @Dwork2015].
-We suggest that activities considered as typically 'exploratory' partition the data into 'training' and 'holdout' datasets where the training dataset can be used to inform potential models or analyses that will be preregistered in later versions.
-The holdout set may then be used for components of the modelling process that align more closely with confirmatory modes of inquiry, such as model evaluation or testing, forecasting, hind-casting and other predictive methods [@lewis2023; @Pu2018; @hofman2023] #todo add pred modelling citations, e.g.
-@Houlahan:2016fl etc.
-see memo for Lewis et al.
-).
-However, partitioning ecological data is not always feasible due to small datasets, studies that are limited in duration or spatial extent compared to the target phenomena, or data patchiness due to the logistical and financial challenges to data acquisition in ecology [@Wood2020].
-
-In these cases where preregistration of flexibility and data partitioning is not feasible or where the analyst may need to check the data in order to pre-specify a decision, e.g.
-checking for data distributions or missingness, blinded datasets, such as those generated by data scrambling [@Baldwin2022a].
-Data blinding can prevent data-peeking that may influence downstream modelling decisions while aiding the modeller in making informed analytic decisions that are appropriate for the dataset on hand.
-
-Finally, oftentimes interpretation of the model and its outputs are based on subjective assessments, such as visually interpreting plots of model fit quality for indications of issues such as heteroscedasticity, interpreting the results of model performance where there is no clear decision threshold, or interpreting model performance across a suite of different metrics and indicators.
-We encourage analysts to follow Baldwin et al.'s [-@Baldwin2022a] suggestion whereby both the research questions and the conditions for interpretation are preregistered.
-We have embedded this approach to preregistering MBR into our template as well, by including prompts within help and explanatory text that direct the researcher to ensure that the con
-
-### Implementing Adaptive Preregistration with GitHub
-
--   [ ] #todo insert summarised version of guide
-
-### Draft Template & Modelling Workflow
-
-We developed a draft template for preregistering model-based research in ecology and conservation that was tested and further refined using our case-study of adaptive preregistration ( #todo Appendix A2).
-We report the final preregistration template refined through the case-study, as well as the refined description of the modelling workflow underlying the preregistration template.
-
--   [ ] How do I present this in the manuscript?? Do I just refer to this in the supp mat? Feels important to present but this manuscript is already LONG!
--   [ ] Also need to either make fig or table based on preregistration template for the underlying model development workflow underpinning it.
-
-## User Testing
-
-Model development and adaptive preregistration processes occurred in tandem, and were implemented using git and GitHub, Leveraging GitHub's version control features.
-This approach is summarised in the guide <https://egouldo.github.io/EcoConsPreReg/adaptive_preregistration_implementation.html,> [@gould2024].
-The GitHub repository is located at <https://github.com/egouldo/VEFMAP_VEG_Stage6,> but is currently embargoed as of today (30 April 2024).
-
--   [ ] #todo consider whether we explain the GitHub procedure here...
-in the 'Design' section, in an appendix section, OR leave out and just reference the website and zenodo link.
-I DO think however, that it has some key features that map onto the adaptive preregistration process that I want to highlight.
-So perhaps we can present in the 'Design' section.
-
-### Final Template Version & Changes
-
-The template used in the VEFMAP study underwent a series of changes to both its content and visual formatting during the analysis to better suit the needs of the modellers.
-The preregistration template was written as an *Rmarkdown* document saved within the project GitHub repository (INSERT LINK TO DOCUMENT #todo).
-Using knitr (CITE) and pandoc (CITE), this document is converted to a markdown document using RStudio's `github_document` format, which is automatically rendered into HTML for viewing in the web browser by GitHub (<https://rmarkdown.rstudio.com/github_document_format.html).>
-
-Improvements to template formatting include the use of dropdown boxes to store template text.
-Template text was divided into a) text that describes the rationale for the inclusion of a particular preregistration section or item within the template, as well as any explanation of key terms and definitions of some modelling practice / process terms, and b) text that contains specific questions with check-boxes for researchers to address in their preregistration.
-Each text-type is given a different icon to help differentiate between them.
-The use of drop-down text-boxes with icons better differentiates template prompts and text from authors' responses to preregistration items and the VEFMAP modellers noted that these changes greatly improved readability of the preregistration document <https://github.com/egouldo/VEFMAP_VEG_Stage6/pull/75#pullrequestreview-1959017213.>
-
-In terms of substantive changes to the structure and content of the preregistration template, significant changes were made to the structure of the document (reorganized different section), and to descriptions of item rationale and explanations, as well as the preregistration items themselves.
-Some items were omitted, some were added, most were reworded for improved expression, and some were re-structured into different sections.
-
--   [ ] Collate comments, pull requests and reviews and issues describing problems with the changes see: <https://github.com/egouldo/VEFMAP_VEG_Stage6/issues/29> #chrisjones and also other comments from #jdyen
-
-Modellers think about the modelling process differently, and use different terms for the same or similar practices and modelling concepts, depending on their particular modelling background, and at times the language used in the original version of the template was confusing and unclear to the VEFMAP modellers.
-
--   [ ] #todo consider as #discussion point in this manuscript #writing
-
-### Adaptive Preregistration of the Case Study
-
-We tested and refined both the preregistration template and adaptive preregistration protocol as outlined in Version X of the guide using the VEFMAP modelling study as a case study.
-We did not preregister this study on an existing official preregistration platform that meets the three core requirements of preregistration outlined by Haroz [-@haroz2022] because existing platforms are unable to accommodate adaptive preregistration where there are multiple and interim versions of the preregistration.
-Consequently tracked changes to the preregistration using version control with git and GitHub, and used GitHub's tag and release features to demarcate major changes and interim versions of the preregistration.
-
-The preregistration process was marked by three main phases of analysis and modelling, each marked with an interim preregistration:
-
-1\. V1 - Data analysis and cleaning, specification of problem context and background preregistered (insert link to supp mat)
-
-2\. V2 - Pilot Analysis preregistered (insert link to supp mat)
-
-3\. V3 - Main analysis including decision tree for informing model selection based on results of pilot analysis (insert link to supp mat)
-
-**V1: Preregistration of problem context, data analysis and cleaning**
-
-Prior to preregistration and analysis, data for the VEFMAP study had already been collected, as this study is a small component of a broader project that has been running for a number of years.
-The preregistration was iteratively completed and updated prior to beginning any analysis or modelling with the VEFMAP dataset, however, data cleaning and wrangling was undertaken in tandem with completion of the initial preregistration document by different researchers on different git branches due to project time-constraints.
-
-**V2: Preregistration of Pilot Analysis**
-
-There was a great deal of uncertainty about whether the dataset would support fitting the desired candidate models due to temporal and spatial patchiness and density of the data.
-The dataset consisted of over 2 million observations across different waterbodies which were missing data during some years, while other sites had been surveyed more times than others (insert table from meeting with DEECA into supp.
-mat and cross-ref \[\[2024-03-27_Meeting_Notes\]\]).
-Some preliminary analysis and exploration of the dataset was needed in order to make appropriate decisions about how best to operationalise the models given the patterns in the underlying dataset and the spatio-temporal variation in the density / patchiness of the dataset.
-For example, simple tests of data distributions within hierarchies needed to be implemented to check for data spread and presence of zero-inflation, and checks for collinearity among candidate predictor variables were needed to reduce overfitting and to ensure the relevant interactions and captured in the models.
-Consequently, we decided to implement the use of a holdout dataset by partitioning the data into a smaller 'pilot analysis' dataset and 'main analysis' dataset.
-The pilot analysis dataset consisted of data from only a single site that had good temporal data coverage, while the dataset for the main analysis consisted of the entire dataset, including the pilot dataset.
-
-The pilot analysis was exploratory in nature and aimed to resolve critical uncertainties about how best to operationalise the models given data constraints in order to identify a candidate set of models that could feasibly be fitted using the full dataset.
-In addition to the preliminary analyses described above, two initial models for the two response variables, cover and richness, were preregistered.
-The same model checking procedure was preregistered for both the pilot and main analyses.
-The pilot analysis was preregistered in section "2.5.3 Describe any data exploration or preliminary data analyses" of the template.
-
-*Pilot Analysis Outcomes*
-
-The pilot analysis resolved some uncertainties, while some further analytic uncertainty about the best model operationalisation remained.
-Firstly, for models of both response variables, the data were unable to support a single model containing separate terms representing both flow regimes and flow events.
-Consequently two separate models, one with each term operationalising flow, were fitted to the pilot dataset instead.
-Secondly, separate models for each plant functional group were desired to help understand how each group responds to environmental flows.
-However, there was not enough data for each functional groups split across both native and introduced species to get these models to converge so plant functional group was included as a random effect instead and fitted to the pilot data.
-
-Model checking revealed both high levels of over-dispersion and over-inflation.
-To account for different proportions of zeros among plant functional groupings, the pilot analysis allowed for zero-inflation to differ among plant functional groups, however did not account for over-dispersion.
-During the pilot analysis, models were refitted using negative-binomial distributions but failed to converge.
-
--   [ ] #todo Report models here!
-
--   [ ] #todo Report Decision tree here!
-
--   [ ] See notes from the report back meeting here: \[\[2024-03-25_Meeting_Notes\]\] and \[\[2024-03-27_Meeting_Notes\]\]
-
-**V3: Preregistration of Main Analysis and Decision Tree**
-
-Although the full models with complex three-way interactions failed to converge on the smaller pilot dataset, it was hoped that they may still converge on the full dataset in the main analysis.
-Consequently, the full models were also specified in the main analysis, but with the two simplified 'flow events' and 'flow regime' models preregistered also in the event that the full models were unable to converge and/or provide reliable parameter estimates.
-
-Given that zero-inflated Poisson models both converged, and reliably captured the proportion of zeros within the pilot dataset, this model family was preregistered for all models fitted to the full dataset, including simplified models.
-However, modellers were concerned that there may be unexpected issues with these approaches due to actual data distributions in the full dataset not matching expectations informed by the pilot analysis, such as unaccounted over-dispersion of the data.
-Consequently a second flexible strategy was preregistered to account for the case where over-dispersion was identified.
-
-First, the two key vegetation response variables were to be modelled as zero-inflated Poisson distributions with a log-link function.
-However, in the case that model-checking identified over-dispersion in any of the models, the models were to be refitted using a negative binomial distribution.
-However, a rule was preregistered that the models re-fitted using the negative binomial distribution would only be accepted over the Poisson models if they *both* improved over-dispersion *and* did not decline in model performance, particularly in terms of model fit, zero-inflation, and posterior predictive checks.
-
-The final decision on which model family should be selected based on model checking in the instance that overdispersion is identified is a fuzzy and subjective one in which explicit weightings between the different model performance criteria cannot be expressed or preregistered.
-Instead, modellers attempted to provide an explanation and description of the conditions for interpretation informing decisions about the model family:
-
-> While we are unable to *a priori* precisely weight these criteria in determining the final distribution, the final decision will be guided by the model’s overall ability to capture key associations reliably.
-This is especially likely to occur when there is no ‘perfect model’ and there is no dominant alternative choice of model distribution.
-
-**Analysis following final preregistration**
-
-TBD, briefly summarise what outcomes from the preregistered decision heuristics were realised in the final analysis.
-
-## Implementation: User Guides and Documentation
-
-We have developed a guide for modellers on how to implement adaptive preregistration for model-based research packaged as a website at <https://egouldo.github.io/EcoConsPreReg/> [@gould2024].
-The guide first provides background material on the risk of researcher degrees of freedom to transparency and reliability of model-based research in order to motivate the use of preregistration in model-based research.
-The core materials include an overview of adaptive preregistration, as well as a practical step-by-step guide to implementing adaptive preregistration using [git](https://git-scm.com) and [GitHub](www.github.com), two commonly used tools by modellers for version-controlling modelling code during model development [@braga2023].
-The preregistration template for model-based research in ecology and conservation is also hosted on this website.
-Components of the guide to adaptive preregistration have been repackaged and incorporated into this manuscript.
-
-**Continuous Improvement and Feedback**
-
-One of the key features of user-centred design is the principle of ‘continuous feedback’.
-We therefore encourage modellers to preregister their research using our template and adaptive preregistration protocol outlined in this manuscript, and to provide feedback and suggestions for improvement.
-Suggested changes may be submitted in the form of a pull-request to the repository ( #todo insert link to how to make a pull request to the new template repository), or by posting a new issue ( #todo link to creating a new issue on the repo and setup template).
-The repository and issues are publicly available, so any future discussion will be transparent and open to the ecological modelling community to participate in.
-
-# Discussion
-
--   intro para summarising problem / objective:
-    -   Preregistration of model-based research has been rejected both within ecology and in other disciplines.
-        -   But the problems that preregistration addresses, i.e. RDFs / QRPs, are likely to exist for MBR too.
-    -   There is a lack of discipline-specific preregistration templates and this impedes the uptake of open-science tools, such as preregistration, within ecology.
-    -   This study aimed to expand preregistration for use within ecological modelling and modelling more broadly by developing a preregistration template for ecological modelling research, and a methdology for preregistering model-based research more generally.
-
-We took a user-centred design approach to the task of developing a preregistration template and methodology that accounts for the domain and methodology-specific nuances of ecological modelling.
-We did this by: 1.
-Identifying idealised norms and norms of practices to develop ‘idealised’ / ‘generalised’ research workflows that describe sets of commonly undertaken tasks (and some sort of order/sequence to reflect the workflow, both idealised and as practically conducted), within an ecological modelling study.
-2.
-We then transformed these workflows into a ‘preregistration template’, by distilling a set of decision-items that, should a) adequately constrain researcher degrees of freedom to mitigate the risk of Questionable Research Practices in an ecological modelling exercise and b) adequately facilitates research/analytic transparency.
-We transformed the decision-steps (the highest resolution) into template items posed as directives or questions for preregistration users to complete, in accordance with the particularities of the study.
-3.
-We qualitatively evaluated the template and the adaptive preregistration methodology for its feasibility of use in MBR using a case-study of environmental flows modelling in Victoria, Australia...
-The VEFMAP modelling study faced particular challenges to preregistration in terms of data complexity and patchiness that made this study a good candidate for trialling adaptive preregistration...
-
-## Benefits, challenges, Limitations of Adaptive Preregistration
-
-User research revealed inherent features of the model development process, namely its iterative workflow and adaptive decision making, were critical barriers to modellers adopting preregistration in their modelling practice.
-In the case of the VEFMAP modelling study, there were critical uncertainties about the the appropriate model specification given limitations of the underlying dataset which prevented the researchers from being able to pre-register fully specified models without undertaking some exploratory analysis.
-Such practices would preclude the application of preregistration to this study because these decisions would violate preregistration's core principle of *data-independent* decision-making.
-
-*What would have happened without preregistration / adaptive preregistration??* (insert into last sentence of above para)
-
--   In the absence of any preregistration, adaptive or not, (#todo check this with Jian, because maybe his practices are different), modellers would typically explore the entire dataset
--   Explain what is wrong with this, what the risk of this is.
-
-*Next para:*
-
-To mitigate the risk of researcher degrees of freedom and to try to preserve some degree of data-independence in decision-making we combined data-partitioning into the adaptive preregistration process and split the preregistration process into two phases, resulting in three interim preregistrations.
-We first preregistered a preliminary or 'pilot' analysis, in which the researchers under took directed exploration of a subset of the dataset.
-The pilot analysis helped the researchers to understand what models could feasibly be fitted for their two outcome variables given uneven spatio-temporal density and patchiness of the dataset.
-Although the pilot analysis was 'exploratory' in nature, by preregistering the preliminary analysis, any exploration occurring here was still *directed* in that the scope of the pilot analysis was bounded and determined by transparently explicated analysis goals and articulartion of critical uncertainties.
-Following preregistration of the pilot analysis, the researchers then switched back to the *analysis* phase of *adaptive preregistration*.
-Carrying out the preregistered pilot analysis.
-The results of the pilot analysis aided in determining the *main analysis* which would be conducted on the full dataset and preregistered using the EcoEvoModelling template.
-
-Researchers preregistered flexibility into their main analysis, because despite the pilot analysis, some critical uncertainties still reamined cf. limitations on the data.
-
--   outline the uncertainties: model family, and something else. see results.
-
-Explain the different models, and the different
-
--   loosely preregistered (Baldwin) grounds for interpretation made explicit
-
-Consequently, we took an ‘expanded view of preregistration’, “Adaptive Preregistration”. Key features:
-
-\- plans for preregistering and deploying flexible analysis strategies:
-
-\- Iterative, cyclical preregistration rather than linear process that maps onto the iterative cycles of model development generating
-
-\- interim preregistrations marking key points in the analysis
-
-We also propose combining adaptive preregistration with previously proposed technical strategies for reducing the risk of overfitting and bias in MBR.
-E.g.
-using data-partitioning with a exploratory/training vs.
-test/holdout dataset / data scrambling for preliminary or exploratory analyses.
-We depart from existing research, e.g.
-Baldwin, by proposing that these preliminary analyses also be pre-registered, even if preregistration is not yet complete.
-The results of the preliminary or exploratory analyses may be used to inform analytic decision-making that can then be preregistered.
-
-Presently there are no existing preregistration repository services that meet the critical requirements of a preregistration platform outlined by Haroz [-@haroz2022] that can be used for adaptive preregistration.
-We proposed and refined an implementation of the Adaptive Preregistration methodology that leverages features of git and GitHub for transparently tracking the evolution of the preregistration document as the modelling progresses.
-
-### Reflections (heading TBD, can we make a water pun given case study and concluding sentence or is that lame)
-
-We approached the application and evaluation of both the template and adaptive preregistration and sought to evaluate the feasibility of preregistration for MBR.
-We genuinely did not know what to expect when we began preregistering the case study, but approached this goal with a 'proof is in the pudding mentality' to ask what are the limits of preregistration....
-Here we reflect on the parts that worked, and those that did not.
-
-**Restricting Researcher Degrees of Freedom, not creativity!**
-
-Counter arguments to both preregistration and preregistration for modelling express fear that preregistation will restrict creativity and lead to 'punch card style analysis' [@MacEachern2019].
-However, modellers in our preregistration of the case study found the opposite, that it prompted critical thinking about analytic choices during modelling, that might have only ever been tacit and never explicitly recorded (cite paper saying many modelling decisions never thought of as decisions and therefore never reported).
-
-Combining the use of data partitioning, and splitting the analysis into a 'pilot' and 'main' analysis aided the modellers in understanding what models could feasibly be preregistered given constraints on the data, while also helping inform the preregistration of flexible analyses.
-The addition of the pilot study components to the template, and into the modelling process was good.
-And I think we should recommend doing this where possible.
-Link to that arxiv paper on preregistering machine learning studies.
-Might seem less natural for ecological modelling studies since we are data poor!
-BUT, in machine learning and other computer-science focussed modelling contexts, this type of practice is the norm.
-One issue to flag is the potential for \[\[information leakage\]\], however.
-Because part of the data *is* used again in the full dataset.
-However, given that this study is situated within an ongoing program of monitoring and decision-making, I think it's ok.
-The modelling can be built into an adaptive monitoring / modelling context where future modelling can be used to test / evaluate and then update the model.
-
-\- How do we think the process actually went in terms of restricting researcher degrees of freedom / improving reliability / reproducibility? **What other reflections are there?**
-
-\- Tricky to anticipate??
-But can certainly say that all of the exploratory analysis undertaken on the pilot dataset would have surely been applied to the full dataset had we not partitioned the dataset and the preregistration process.
-
-**Resolution of the template - tradeoff between meeting preregistration objectives, and burden and benefits to the researcher.**
-
--   Competing objectives?
-Although the purposes of preregistration are not mutually exclusive - as we have decided the purpose of preregistration we have defined here is to both delimit/state flexibility, *and* be flexible.
-However, “the preference of one purpose over another can have tangible effects on the design of preregistration” p15..
-Pu.
-
--   Compared to Hofman [-@hofman2023], a template for preregistering predictive modelling research, our template was much more detailed and contained many more items to be preregistered.
-
--   Get Jian / Chris' thoughts on template.
-I suspect it's a bit long / overwhelming.
-But potentially that's because of the visual presentation / limitations of a static text based medium for the document.
-
--   However, the benefit of a more exhaustive, specific and detailed preregistration is increased transparency (link back to intro on transparency in ecological modelling).
-By adopting preregistration for MBR we can also improve the transparency of ecological modelling, which is done terribly!!!
-
--   cost to completeness, and specificity / precision is that it is time-consuming.
-But aside from increasing transparency, it is also critical for appropriately restraining unmitigated and undisclosed RDFs by preventing loose interpretation / operationalisation of the preregistration in instances where the preregistration is vague.
-
-    -   We can see this in the disjuncture between all of the modelling outputs conducted and reported for the internal report, but that weren't fully specified in the preregistration due to time-constraints on the modellers.
-
-*This will lead in nicely to the final criticism*
-
-**Time & Timing**
-
--   “A natural consequence of increased transparency, is that preregistrations become longer as more possible decisions are made transparent: *increased transparency* implies *increased length*.
-This trade-off was something we considered important during the design phase, and deciding what was in and out of scope for the preregistration template was particularly challenging.
-
--   However, as Pu et al. note, some participants who were both authors and reviewers didn’t think that the time-costs of preregistration in research were significantly greater than not preregistering research.
-Instead they thought of the preregistration process invoking a ‘time shift’, that brought ‘decision points’ before data collection instead of during analysis.” P.19.
-In addition, “some reviewers thought that preregistration saves them the time they might have spent wondering about undisclosed flexibilities when working on an un-preregistered paper.”
-
--   Even in adaptive preregistration where the process is iterative, modellers reflected on the front-heavy reserach timeline forced by the process of adaptive preregistration.
-Shifting decision-making to the start of the project timeline delayed the start of the analysis, but modellers felt that this resulted in better decision-making because the decision-making was more carefully reasoned.
-Additionally, the case study researchers felt that the level of detail reported in the preregistration aided in efficiently writing the internal report.
-
-    -   This relates to what Pu et al. found where participants ‘felt their research workflow improved by having preregistration as a “forcing function” \[…\] For example, writing down study plans ahead of time might prompt them to think more thoroughly and to anticipate flexibilities / researchers’ degree of freedmen advance.’ (Page 13).
-For reviewers ‘participant reports that preregistration makes the review process “easier”.’
-
--   Mention difficulties I had with the internal deadlines of DEECA, and that we could not preregister everything we wanted to (i..e the analysis of the model and its outputs used for presentation in the report, which were extremely vaguely specified, aside from model checking).
-Preregistration could be challenging for modelling in such settings where there extremely tight timelines.
-BUT, as we are doing for the stage 2 report preregistration of the second study now, could be used during early consultation / scoping.
-So the document could begin to be filled out much earlier than we did here (data already collected).
-
-## Future Implementation & Testing
-
-We did not undertake the 'usability testing' component of User-Centered design set out in fig 1.
-We offer this template and adaptive preregistration methodology to the ecological modelling community, and modelling communities across disciplines, to use, refine and adapt in their modelling practices moving forward.
-We hope insights gained through trial and error of its application will be used to further refine and improve the preregistration template and adaptive preregistration methodology.
-
-Advocacy and uptake of preregistration is based on the working hypothesis that it is able to mitigate against QRPs by delimiting opportunistic exploitation of researcher degrees of freedom while improving transparency.
-While initial evidence suggests that preregistration may improve the reliability and reproducibility of research (cite study looking at replication rates of preregistered vs not), empirical testing of this working hypothesis is necessary for evaluating whether this new practice meets its intended purpose [@Devezer2020] (and others).
-Future research should evaluate the effectiveness of adaptive preregistration for improving the reliability and transparency of MBR.
-
-Hosting the preregistration template on GitHub, has the added benefit of providing a mechanism for dealing with *uncertain* *norms* in modelling practice [@Pu2019], and for following modelling methods and decision contexts that might not fit the presumed modelling development cycle underpinning the preregistration template.
-Anyone is free to *fork* our GitHub repository, meaning that the entire contents of the repository can be copied to a users own repository, and the template is free to be altered as the user sees fit.
-This provides the opportunity for creating domain-and/or modelling-approach-specific templates [@Cruwell2019b].
-
--   [ ] #todo need to explain a bit further what the hell i mean by 'uncertain norms'.
-I think I can link the confusion from VEFMAP modellers about some concepts within the preregistration document in here.
-
-In order to address this confusion among norms, I state the expected norms that the template has been designed around here.
-However, I don’t expect users to agree on this process.
-Because we have kept the template in a GitHub repository that can be *forked*, people are free to adapt the template (the items themselves) to suit whatever norm they see fit about how preregistrations should be created and used.
-
-### Using GitHub for Adaptive Preregistration
-
-**Benefits**
-
--   transparency
--   existing tool already used by many modellers for both version control and collaborative project management of computational code for models
-
-**Challenges**
-
-Separate branches...
-Under significant time pressure (internal reporting) to do the analysis.
-Data cleaning and analysis of pilot data was happening on one branch while the preregistration was happening on another!
-Also meant that the fledge release system outlined in the guide did not quite work because you had to wait until each branch had been merged back into the main / master branch in order to bump the version.
-
-\*\*Critiques of GitHub as a platform for preregistration. and combining other tools to meet the minimum requirements of preregistration: timestamps, persistence, registry\*
-
-While there are some proponents of the use of version control systems for preregistration [e.g.
-@braga2023; @hofman2023], [@haroz2022] make the compelling argument that GitHub (and potentially alternative version control systems) are inappropriate as preregistration platforms because it doesn't meet any of the three core requirements for a preregistration: timestamps, an indexed registry, and persistence.
-While we have described the use of GitHub timestamps in this study, [@haroz2022] illustrate that timestamps can be manipulated on GitHub.
-Moreover, currently GitHub does not have a formal registry for indexing or searching for preregistrations hosted on GitHub, which means that there is the potential for *\[\[preregistration hacking\]\]*, wherein preregistrations may be deleted changed and reuploaded [@haroz2022].
-Finally, GitHub does not meet the *persistence* requirement, there is no immutable unique identifier, like a digital object identifier (DOI), and GitHub is not guaranteed as an archive service in perpetuity \[\[data sharing\]\] / \[\[data curation\]\].
-
-While the use of GitHub in this study was to test it as a platform for implementing *Adaptive Preregistration* rather than *Preregistration* per se, given these shortcomings, we recommend pairing GitHub with other platforms for implementing Adaptive Preregistration.
-
-Here we tested it as a system for managing and implementing *Adaptive Preregistration* rather than just preregistration in general.
-However, given these shortcomings, we recommend pairing the use of GitHub for Adaptive Preregistration with some additional tools to ensure that these three minimum requirements of preregistration are met, we outline two complementary approaches here that meet the three core requisites of preregistration:
-
-1.
-Archive the GitHub repository using Zenodo (CITE) at the first draft preregistration (V.1.0) and combine with GitHub tag and release strategy, such that each interim preregistration is recorded as a new version / release on both GitHub and Zenodo .
-Deals with both the timestamped issue and the persistence issue by giving it a digital object identifier [@smithSoftwareCitationPrinciples2016], Furthermore, Because it is given a DOI, it is indexed and searchable [@hofman2023].
-    -   [ ] Add links to other papers suggesting Zenodo be combined with GitHub for data / code sharing #todo
-2.
-But not a specifically preregistration registry.
-How to deal with the registry component?
-Can we add as component to OSF repo??
-and then preregister?
-Yes.
-In fact, an alternative solution has been outlined by [@vanlissa2021], where they have used a similar approach to us with a version-controlled plain-text preregistration document containing the pre-analysis plan, and tagging and releasing this document on GitHub.
-However, they suggest a secondary optional step of uploding the preregistration to a dedicated preregistration server, without specifying what that server may be.
-We could use an 'open-ended Registration' template on the OSF [@hofman2023], and then link the document from GitHub component on the OSF project to the registry.
-Instructions for doing this are given in [@hofman2023].
-    -   [ ] WHAT ABOUT DIFFERENT VERSIONS OF THE PREREGISTRATION..
-Can we make changes using the OSF??
-DOUBLE CHECK THIS!
-Might not be able to use the registry here.
-#todo I think #MarkCall said there is the option to make changes now.
-Could we test with their staging platform??
-Put the different versions on there??
-But the template has changed....
-Could just try it anyway as proof of concept.
-And note this as a limitation of the OSF templates approach (not being modular or flexible, as [@hofman2023] outline ).
-
--   [ ] #todo CHECK THIS: Embedding the repository within an OSF repository will fix the timestamp issue from GitHub that [@hofman2023] outline??
-
-# Conclusions
-
-Instead of throwing the baby out with the bathwater, in this study, we advocate that preregistration can and should be used isn model-based research in ecology and conservation, but its particular form should reflect the norms and practices of the research context in which it is applied, in order to  in order to adequately restrict researcher degrees of freedom that may uniquely emerge within a domain-specific methodology.
-
-This research provides a template and methodology for implementing adaptive preregistration of ecological models in either pure or applied settings and will facilitate the wider uptake of preregistration in ecology.
-Although we focus on ecology and conservation in this paper, the template is extensible to alternative fields and methods, and the proposed methodology for implementing adaptive preregistration can be applied in model-based re-search within science more broadly.
-Modellers need no longer cry “but, I can’t preregister my research.”
-
-## Software & Data Availability Statement
-
-## Author Contributions
-
-CRediT
-
-# Supp Material
-
--   [ ] Check out journal guidelines or papers #todo
-
-## SM1: Workshop Materials and Process Design
-
--   Design process for the modelling workflow template for activity 1. The actual materials are in A1.
--   Activity 2 structured discussion with prepared questions and talking points informed by recent debates about preregistration in the literature.
-Any literature review leading to this.
-Actual talking points in A1.
--   Mention that we also ran the above in separate groups for designing a field-study preregistration template, but decided not to pursue further here.
-
-## SM?: Analysis of Workshop Outputs
-
-"We revised the coding structure (inductively is this the word), according to patterns and themes identified across multiple personal workflows:" For example, we added the final phase (name for last phase), which aimed to capture the fact that analysts usually present the results of their modelling to clients, decision-makers or other stakeholders that must interpret the evidence and are responsible for making management decisions.
-In this way, we identified a common or generalised workflow from different types of applied ecological modelling projects that synthesised both (idealised norms and norms of practice, Pu et al., see other philosophy of sci refs).
-
-## A2: Preregistration Template
-
-## A3: Adaptive Preregistration Guide
-
-## A4: Environmental Flows Modelling Case Study
-
-**Study Background**
-
-Riparian vegetation is a critical ecosystem component for so many reasons (biodiversity, habitat, food, stability, filter, carbon source, aesthetic, etc.) but it is also severely degraded in many waterways all over the world.
-Vegetation clearing, livestock grazing, pollution, exotic species (flora and fauna) and alteration of flows are some of the most severe threats.
-
-The impacts of river regulation on vegetation depend on the type and extent of flow alteration.
-Flows can increase or decrease total flow, be more or less variable, or have altered season flow timing.
-Environmental flows are designed to be manipulations of flow (usually flow releases from a reservoir) that aim to have specific and direct environmental benefits.
-
-Evaluating the effects of environmental flows on waterways is difficult because of the need to evaluate the isolated impacts of specific flow events (short term response) as well as the impacts of successive flows (environmental or otherwise) within a broader flow regime (medium to long term response).
-Data collection at large temporal scales (annual or greater) prevents evaluation of any single flow event or component from the regime.
-Evaluating the effects of environmental flows in most cases is therefore only possible with repeated sampling at relatively short intervals.
-Evaluation also needs to be done with consideration of other major drivers of vegetation, such as rainfall, other flow attributes, livestock grazing, soil properties, exotic species and previous disturbance.
-
-In south-eastern Australia, river regulation (including water extraction), livestock grazing and exotic species continue to have serious negative impacts on all major waterways, despite major government investment into waterway management.
-However, increased understanding of the impacts of major threats and the effectiveness of management is resulting in greatly improved management decisions.
-While many broad trends are known, critical details remain elusive and require concerted effort to address.
-
-**Study Aims**
-
-### Study Methods
-
-#### Data Collection
-
--   [ ] 1 - 2 sentence overview, then point readers to details in \[\[project-VEFMAP\|VEFMAP\]\] manuscript #todo #writing
-
-#### Modelling Dataset
-
-| Variable Name                            | Unit                  | Definition                                                                                                                                     | Variable Type | Variable Group                                                         |
-|-------------|-------------|----------------------|-------------|-------------|
-| Species Vegetation Cover                 | Cover                 | Cover per-species, per sub-transect                                                                                                            | Response      | Vegetation Cover                                                       |
-| Plant Functional Group Cover             | Cover                 | Cover per-plant functional group, per-sub-transect                                                                                             | Response      | Vegetation Cover                                                       |
-| Sub-transect Vegetation Cover            | Cover                 | Total cover (sum all species), per sub-transect                                                                                                | Response      | Vegetation Cover                                                       |
-| Species Richness                         | Counts                | Number unique Species Functional Group, per sub-transect                                                                                       | Response      | Species Richness                                                       |
-| Days above baseflow / springfresh levels | Counts, Z-scale       | Number of days per year (averaged over prev 3 water years)<br>discharge \> baseflow or springfresh levels,<br>Z-scaled                         | Predictor     | Flow Regime                                                            |
-| Sub-transect zone                        | Factor                | Zone in which a sub-transect occurred:<br>- below baseflow level, <br>- between baseflow and spring-fresh levels<br>- Above spring-fresh level | Predictor     | Flow Regime                                                            |
-| Flow event                               | Factor                | Timing of sampling:<br>- before spring fresh<br>- Between spring and summer fresh<br>- After summer fresh                                      | Predictor     | Flow Event                                                             |
-| Site Grazing Status                      | Factor                | Grazing status of site:<br>- Grazed<br>- Ungrazed                                                                                              | Predictor     | Other factors influencing spatio-temporal variation in cover/abundance |
-| Species Functional Group                 | Factor                | For each Origin group (native or exotic):<br>- aquatic<br>- emergent<br>- riparian<br>- terrestrial                                            | Predictor     | Other factors influencing spatio-temporal variation in cover/abundance |
-| Site Identity                            | Factor                | Name / Identity of the site                                                                                                                    | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
-| Transect Identity                        | Factor                | Identity of Nested within sites                                                                                                                | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
-| Sub-transect height                      | Metres above baseflow | Height of the sub-transect along bank                                                                                                          | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
-| Survey Year                              | Date, YYYY            | Year in which the sample was recorded                                                                                                          | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
-| Waterbody                                | Factor                | Waterbody in which the site is located                                                                                                         | Random effect | Other factors influencing spatio-temporal variation in cover/abundance |
-
-#### Pilot Study
-
-The final data set comprised XYZ observations, which presented a significant computational challenge.
-In addition, exploratory data analysis revealed considerable zero-inflation and potential over-dispersion of the response variables, particularly vegetation cover.
-We conducted a pilot study to determine the feasibility of different model structures and identify likely decision points given these challenges.
-
--   [ ] when did the EDA occur... during the pilot study? On a different data set? #help-wanted #jdyen
-
-To address this, we defined a maximal model structure and tested this structure with a pilot data set (all observations from the Campaspe River from 2017–2020) to identify feasible model structures and likely decision points.
-We developed separate analyses for vegetation cover and species richness, with the same maximal model structure for both.
-
-Initial model fitting assumed a zero-inflated Poisson distribution for both vegetation cover and species richness, with zero-inflated negative binomial distributions designated in cases where over-dispersion was severe.
-This approach does not introduce an upper bound on estimates of cover (e.g., specifying a maximum cover level of 100%).
-We used this approach for three reasons.
-First, overlapping species may generate cover estimates that exceed 100%.
-Second, parameters of bounded distributions are challenging to estimate when many values occur near the bounds.
-Third, bounded distributions require link functions (e.g., the logit function) that complicate the specification and interpretation of autoregressive models.
-We note that the Poisson distribution approximates the bounded binomial distribution in the limit of many trials (REF).
-The extent of zero-inflation and over-dispersion (and a model’s capacity to capture these patterns) was assessed using posterior predictive checks, which use values simulated from the fitted model to determine whether the model structure captures the observed distribution of the response variable (REF).
-\### Results
-
-#### Pilot Study Results
-
-The final maximal model was as follows:
-
-``` r
-response (cover or species richness) ~ cover in previous survey +
-                                        functional group × origin × flow metrics +
-                                        functional group × origin × zone × sampling period +
-                                        grazing +
-                                        (1 | site / transect) + (1 | sub-transect height) + (1 | survey year) + (1 | waterbody) +
-                                        offset(number of points).
-```
-
-Vegetation cover in the previous survey was included only for analyses of cover and meant that models of vegetation cover examined how flow influenced the change in cover through time rather than absolute vegetation cover (akin to models of population growth rate used in the VEFMAP long-term fish monitoring project).
-Flow metrics are the (z-scaled) days above baseflow level and days above spring-fresh level in the three years preceding a survey and included both linear and quadratic associations[\[JDLY(1\]](#_msocom_1) .
-Zone is the location of a sub-transect relative to baseflow or spring-fresh levels and sampling period is the timing of a survey relative to the delivery of spring and summer freshes.
-Sub-transect height records the level of a sub-transect in metres above baseflow level.
-Waterbody was excluded from the pilot study given its focus on a single waterway (the Campaspe River).
-
-Results of the pilot study inform the subsequent analyses and are presented here.
-Models of vegetation cover at the species level did not converge in any instance.
-Models of vegetation cover aggregated to watering-response plant functional groups converged but required separate models of two-way interactions instead of the three- and four-way interactions included in the maximal model structure.
-Models of vegetation cover aggregated to structural groups (aquatic, emergent, riparian, terrestrial) converged but required removal of interactions with species origin and separate analyses of flow metrics and for each of the categorical flow terms (zone and sampling period).
-Models of vegetation species richness within watering-response functional groups converged but required separate models of two-way interactions instead of the three- and four-way interactions included in the maximal model structure.
-Models of vegetation species richness within structural groups converged but required removal of interactions with species origin and separate models for the two categorical flow variables (zone and sampling period).
-Therefore, subsequent analyses of vegetation cover and species richness focused on structural groups because these supported model structures closest to the maximal structure.
-The included interactions were re-assessed in models of the full data set because the inclusion of additional data was predicted to either support more-complex model structures or introduce additional heterogeneity (and, therefore, support less-complex model structures) than in the pilot study.
-This did not include re-assessment of the functional group classifications.
-
-Posterior predictive checks indicated that a zero-inflated Poisson distribution reliably captured the distribution of observed vegetation cover and a Poisson distribution (without zero-inflation) reliably captured observed distributions of species richness (FIGS).
-Models of cover had excess over-dispersion, but a zero-inflated negative binomial distribution did not substantially reduce unmodelled over-dispersion and resulted in worse characterisation of zero-inflation and the observed distribution of cover data (FIGS?).
-Therefore, subsequent models of vegetation cover and species richness assumed that data were derived from a Poisson distribution, with additional parameters for zero-inflation specific to each functional group in the case of vegetation cover.
-
-### Full Dataset
-
-# Appendices
-
-## A1 : Workshop Materials
-
--   Activity 1 Template, and instructions / materials provided to workshop attendees. (insert link to OSF)
--   Activity 2 structured discussion with prepared questions and talking points informed by recent debates about preregistration in the literature (insert link to OSF)
-
-## A2: Draft Preregistration Template
-
--   [ ] Insert draft template
-
-## A3: Final preregistration Template
-
--   [ ] Insert template
+Figure 2 - Implementing Adaptive Preregistration.
+The preregistration (blue steps) and modelling and analysis processes (orange steps) operate in parallel.
+Purple tags represent git tags, corresponding to incremental versions of the preregistration.  Any changes to the analysis plan receive a GitHub release and major version number increment.
+Step 3 corresponds to the initial preregistration document, Step 6 corresponds to an interim preregistration, and Step 8 corresponds to the final preregistration.
+The execution of the Final Preregistration occurs in Step 9.

### 2025-03-03 _analysis/ms.qmd
Finish update methods section

@@
 
+## Designing the Preregistration Template
+
+We facilitated a workshop with 10 ecologists and ecological modellers.
+A key task was to identifycommon modelling workflows and critical decision-steps in model development.
+‘Critical’ decision-steps are points in the modelling process where researchers, or modellers, make analytical decisions that could change the output of the analysis.
+In this activity, workshop participants first individually reflected on their own scientific process for a recent or memorable research project, listing the steps undertaken throughout the process.
+Next, participants collaboratively mapped the decision steps in their own personal modelling workflows onto a pre-prepared modelling workflow template (<https://osf.io/fgd23>), informed by our general workflow.
+We then collaboratively reviewed and refined the modelling workflow into a standard set of key modelling phases and commonly implemented decision steps.
+
+## Developing an Adaptive Preregistration Methodology
+
 We developed an Adaptive Preregistration methodology, so that users have the flexibility to adapt the preregistration template to their own needs and processes [see user guide <https://egouldo.github.io/EcoConsPreReg/>, @gould2024].
@@
 
-## Designing the Preregistration Template
-
-We facilitated a workshop with 10 ecologists and ecological modellers.
-A key task was to identifycommon modelling workflows and critical decision-steps in model development.
-‘Critical’ decision-steps are points in the modelling process where researchers, or modellers, make analytical decisions that could change the output of the analysis.
-In this activity, workshop participants first individually reflected on their own scientific process for a recent or memorable research project, listing the steps undertaken throughout the process.
-Next, participants collaboratively mapped the decision steps in their own personal modelling workflows onto a pre-prepared modelling workflow template (<https://osf.io/fgd23>), informed by our general workflow.
-We then collaboratively reviewed and refined the modelling workflow into a standard set of key modelling phases and commonly implemented decision steps.
-
 <!--# Insert Figure 2 - Adaptive Preregistration Workflow -->
@@
 The execution of the Final Preregistration occurs in Step 9.
+
+## Evaluating and Refining Adaptive Preregistration with a Case Study
+
+We preregistered a real-world model-based research problem as a case-study for evaluating and testing the Preregistration Template and proposed Adaptive Preregistration methodology.
+The case study analysed the effectiveness of environmental flows on maintaining riparian vegetation condition in Victoria, Australia [@Tonkin2020].
+We captured the case-study researchers’ feedback about the Preregistration Template and Adaptive Preregistration methodology using the version-control and collaborative project management features of git and GitHub [@BenBondLamberty:2018by; @Blischak:2016kr; @braga2023].
+We used these tools for documenting both the analytic decisions of the researcher, and for using the preregistration process itself to ‘live-develop’ the template.
+For example, if it turned out that we had missed an important item, or the order and structure of the template needed to change in the process of completing the preregistration, the suggested change and its justification would be both recorded in GitHub.
+We also conducted follow-up semi-structured interviews to capture more detailed reflections of the case-study researchers.

### 2025-03-03 _analysis/ms.qmd
Draft manuscript for co-authors
(word count 8401)

@@
 We also conducted follow-up semi-structured interviews to capture more detailed reflections of the case-study researchers.
+
+# Results
+
+## The Preregistration Template
+
+The final Preregistration Template, refined by application of the draft Preregistration Template developed from the user-research workshops to the case study, is hosted at [https://github.com/egouldo/EcoConsModPreReg](https://github.com/egouldo/EcoConsModPreReg.).
+We provide the template as a quarto markdown extension template [@Allaire2024 insert ZENODO DOI] which can be installed and rendered to three different formats, html, pdf, and docx (Supp Mat C).
+Intermediate versions of the template developed through application of the initial template to the case-study are available in [the case-study preregistration document's GitHub commit history](https://github.com/egouldo/VEFMAP_VEG_Stage6/blob/263914d434864679d1e5cacca34caef10fe4d67c/analysis/preregistration_template.Rmd) or in [case-study GitHub repository releases](https://github.com/egouldo/VEFMAP_VEG_Stage6/releases).
+
+## Findings From the Case Study
+
+The application of our preregistration template and Adaptive Preregistration methodology to the environmental flows modelling case study occurred across three distinct stages (Table 1, Fig.
+3), with two interim preregistrations (Version 1 and Version 2) preceding the finalised preregistration (Version 3, Supp Mat E), each corresponding to distinct phases in both the Preregistration and Modelling & Analysis processes.
+While preparing Version 1, the modellers decided that they needed to conduct some analysis to specify a set of appropriate candidate models.
+They preregistered a pilot analysis (Version 2) which used a subset of the full dataset to preserve some degree of data-dependent decision-making while permitting a degree of exploratory analysis.
+Results from the pilot analysis informed subsequent preregistration of the Main Analysis on the full dataset (Version 3, Supp.
+Mat.
+E).
+Several classes of data-dependent decisions (Liu et al.
+2020) emerged during our case study, which were managed using a combination of registered flexibility and interim preregistrations:
+
+-   *Data-flow dependencies*: where the output of one decision is the input to another.
+For example, the models parameterised on the full dataset in the case study’s Main Analysis were subjected to a suite of model checks as outlined in the decision-trees in Preregistration Version 3, with the results of model checking informing *a priori* specified alternative model structures and model functional forms.
+-   *Information dependency*: where one decision informs another.
+For example, outcomes from the exploratory pilot analysis (Interim Preregistration Version 2, Table 1) informed the model structures specified in Preregistration Version 3 (Table 1), as well as the decision heuristics for triggering their fitting and acceptance.
+-   *Procedural dependency*: where downstream decisions would not exist if some alternative upstream decision was made instead.
+For example, if preliminary analysis of the data had not revealed over-dispersion or zero-inflation in the data, then the case study modellers would not have fitted zero-inflated Poisson models, nor would they have trialled accounting for over-dispersion or refitting using negative binomial models (Fig.
+4).
+
+### Interim Preregistration (Versions 1 and 2)
+
+Two initial candidate models were preregistered in Version 1 (Fig.
+3), one for each study response variable, however modellers were unsure if these candidate models would be supported by the dataset due to spatio-temporal patchiness and heterogeneous density in the data.
+Preregistration Version 2 described an exploratory pilot analysis that aimed to resolve critical uncertainties about how to operationalise the models, given data constraints, and how to specify a candidate set of models that could feasibly be fitted to the full dataset (Table 1).
+This involved partitioning the dataset into a ‘Pilot Analysis’ subset for making subsequent data-informed modelling decisions, which were then preregistered in a subsequent preregistration of the ‘Main Analysis’ (Version 3, Table 1 and Fig.
+3).
+The pilot analysis dataset consisted of observations from a single site with good temporal data coverage, while the main analysis dataset comprised the full dataset, including the pilot dataset.
+Preliminary analysis informed how to operationalise the models given the underlying structure of the dataset.
+For example, simple tests of data distributions within hierarchies were required to check for data spread and the presence of zero-inflation, while checks for collinearity among candidate predictor variables were needed to prevent confounding and overfitting, and to ensure relevant interactions were captured in the models.
+The same procedures for evaluating model fit and model checking were preregistered for both the pilot and main analysis (Fig.
+3B), however the purpose of model checking in the Pilot Analysis was to identify appropriate model structures, whereas the purpose for the Main Analysis was to ensure that the fitted models were supported by the data.
+
+### Final Preregistration (Version 3): Main Analysis
+
+In Version 3 of the Preregistration, we derived and preregistered:
+
+1.
+A preferred candidate model or ‘full model’ for each response variable (vegetation richness and vegetation cover), to be fitted to the full dataset (Initial Preregistration, Fig.
+3, Supp.
+Mat.
+E, p13).
+2.
+A candidate set of simplified models that separately captured different flow components for each response variable, with two versions of the flow events model for vegetation cover (Flow Events Models A & B, Fig.
+3, Supp.
+Mat.
+E, pp.20-21).
+3.
+Decision-trees for triggering fitting and acceptance of the candidate simplified models should the preferred full model specifications fail to converge and/or provide appropriate parameter estimates (Fig.
+3).
+4.
+A second flexible strategy, for determining model functional form and to account for potential over-dispersion (Fig.
+3).
+Preregistering this flexibility was motivated by the modellers’ concerns that the actual data distributions in the full dataset did not match expectations informed by the pilot analysis.
+This final decision-point for choosing the model family is subjective and fuzzy, and explicit weightings between the three model performance criteria could not be expressed or preregistered.
+Instead, modellers explained that the final decision will be guided by the model’s overall ability to capture key associations reliably.
+
+| Preregistration Version | Preregistration Process | Modelling and Analysis Process |
+|----|----|----|
+| **Version 1:** (v0.10 – v0.8.1) | Establish problem context, study background and aims, describe data collection and data cleaning process, articulate modelling objectives, and specify candidate models.
+| Data cleaning and preparation undertaken concurrently by lead and supporting researchers due to limited resourcing and short timelines for project deliverables.
+Candidate models addressing different components of the research question were preregistered.
+Identify data structural properties and potential consequences for model fitting |
+| **Version 2:** (v0.9.0-v.0.12.5) | Describe Pilot Analysis on data subset and specify registered flexibility in the form of decision heuristics.
+| Critical uncertainties in modelling decisions were articulated (e.g.
+how do we classify best flow regime based on inundation data?).
+In addition, data partitioning, exploratory data analysis, refined candidate models and model checking procedure preregistered |
+| **Version 3:** (v0.13.0-v.0.20.1) | Describe Main Analysis, including specification of decision tree for model selection process based on results of pilot analysis.
+| Conducted modelling and analysis as outlined in preregistration.
+Reported deviations and rationale as necessary.
+|
+|  |  |  |
+
+: Table 1 - Key stages of the modelling and Adaptive Preregistration process implemented in the environmental flows modelling case study.
+Version numbers in parentheses are GitHub tags, marking snapshots of the repository for each Preregistration Version.
+
+<!--# Insert Figure 3 - case study workflow -->
+
+Figure 3 illustrates the evolution of model development across preregistration iterations in the case study.
+The two Registered Flexibility steps in the Main Analysis (v3) (shown in blue) are further described in Fig.
+4.
+
+<!--# Insert Figure 4 -->
+
+Figure 4 - illustrates how registered flexibility was applied within the case study preregistration, for selecting model structure (A) and choosing an alternative model functional form to the default (zero-inflated Poisson) (B).
+
+## Adaptive Preregistration in Practice: Benefits and Challenges
+
+At the end of user-testing, case study researchers concluded that preregistration is a valuable tool to facilitate Good Modelling Practice, but also acknowledged several difficulties encountered during Adaptive Preregistration (Table 3).
+
+Table 3 Summary of challenges, benefits and added value of and tips to using Adaptive Preregistration for ecological modelling.
+
+# Discussion
+
+The principles and potential benefits of preregistration are clear: preregistration encourages more careful and considered analysis, restricts researcher degrees of freedom, and thereby mitigates the Questionable Research Practices that can derail research.
+Ultimately, preregistration can improve the reproducibility and reliability of scientific research [@koivisto2024; @ODea2021].
+Existing preregistration templates sit squarely in an idealised hypothesis testing research space.
+In contrast, ecologists take a different approach to analysis, relying more on modelling complex relationships than significance testing.
+Modelling is usually a more iterative process, with frequent moves between exploratory and confirmatory practices.
+Although this makes preregistration less approachable for model-based studies, it need not preclude preregistration entirely.
+We have reconceptualised preregistration as an intrinsically adaptive process that can move with researchers through the modelling process.
+We show how Adaptive Preregistration can still encourage model reproducibility, transparency, and facilitate uptake of best practice by taking a structured, principled and flexible approach to preregistration.
+
+Despite fears that preregistration may limit creativity and flexibility [@Pu2019], lessen researchers' engagement with data and result in “punch-card” style analysis [@MacEachern2019], the case-study supported Liu et al’s [-@Liu2020] claim that preregistration can improve both reproducibility of process and research outputs as well as analytic decision-making.
+By enforcing separation between planning and conducting the analysis, case-study modellers found that following a structured approach to analysis (even with preregistered flexibility) helped to keep them on track, maintaining clear links between methodological decisions and the study aims and conceptual models (Table 3).
+
+By encouraging careful analytic decision-making while formally incorporating data exploration into the modelling process through registered flexibility, data-subsetting and interim preregistrations, Adaptive Preregistration traded largely unbounded explorations for more goal-directed analyses, constraining researcher degrees of freedom and potential 'fishing expeditions'.
+In the absence of Adaptive Preregistration, the case study modellers felt that they still would have taken the same modelling approach but would have started with a much larger decision-space in which they would have trialled more variations on model structure and predictor variable operationalisations along the way.
+Although this may have resulted better overall model fit, it would also risk overfitting (Lewis et al.
+2023, Barnard et al.
+2019).
+
+Adaptive Preregistration increases both 'analytic' and 'model transparency’' [@Bodner2020; @Lupia2014] by ensuring that modelling choices, steps, assumptions, and expectations about outputs are clearly articulated and documented in greater detail than they typically would be.
+Interim preregistrations reveal intermediate models, analyses and results, which circumvents reporting only the final models.
+Additionally, registered flexibility exposes the impact of exploratory and data-contingent analyses on modelling choices.
+In our case-study, registered flexibility for choosing between alternative model functional forms revealed all alternative decisions under consideration, and what analytical outcomes would trigger the selection of each alternative.
+This process provided transparent justification for the choice of model functional form.
+Adaptive Preregistration can therefore provide a more transparent account of how analytic conclusions are drawn from the data, which are often poorly reported in ecological modelling studies [@Zurell2020a].
+
+The structure and contents of the preregistration template can encourage Good Modelling Practice by explicitly eliciting information about assumptions, expectations, and rationale behind modelling choices.
+Modelling is considered by many to be an 'art form' in addition to a science [@MacEachern2019; @Smaldino:2020; @Taylor2018], consisting of tacit knowledge and expertise with decision-points that may not always be recognised by modellers [@Babel2019a; @lahtinen2017].
+In our modelling case-study, it took many attempts to properly articulate registered flexibility when selecting model structure and functional form, perhaps because these procedures were so ingrained into the modeller's practice.
+By directing users to specific modelling tasks, preregistration prompted modellers to explicitly recognise, document and justify tacit modelling choices.
+
+## Difficulties encountered using Adaptive Preregistration in our case study
+
+Our case study researchers reported that the most challenging aspects of Adaptive Preregistration involved anticipating analysis decisions in advance, in addition to changes to model management and timelines.
+Registering flexibility requires careful thought and time to adequately capture analytic decision-points and describe the process for resolving those decisions.
+In our case study, while the decision heuristic for choosing model functional form was accurately preregistered, the process for determining model structure was underspecified in the ‘Main Analysis’ preregistration.
+This occurred because the preregistration did not specify that the heuristic should only be applied to some models and not others (i.e.
+Flow event models and not Flow regime models).
+
+The preregistration, while detailed, did not adequately capture the modellers' intended and implemented model selection strategy, where after fitting the set of simpler models identified from the pilot study (preregistration version 2, Table 1), the modellers would iteratively trial different combinations of model terms to make the models closer to the more complex ‘maximal’ models identified in the initial preregistration (preregistration version 1, Table 1).
+The preregistered specification did not explain which combinations of model terms would be trialled, and in what order they would be trialled, rendering components of the model selection process unregistered.
+
+The case-study researchers noted that anticipating potential analysis scenarios and decision pathways was complicated by large uncertainties about the underlying system dynamics in a complex and patchy dataset.
+Tight project turn-around times meant that there was limited time to iteratively review analysis specification decisions, resulting in some decisions and components of the modelling process being omitted from the preregistration.
+Following our Adaptive Preregistration protocol more closely may have avoided underspecificity in the case-study.
+However, because of temporal and resource constraints on the case study project, compounded by the resource costs required to upskill some members of the research team with git and GitHub, the protocol for implementing Adaptive Preregistration with git & GitHub as outlined in our guide was not fully followed in the case-study preregistration.
+
+These aspects combined may feel like Adaptive Preregistration is slowing the research process because preregistration front-loads the decision-making and planning in the project timeline, delaying analysis and implementation of the preregistration [@evans2023], which may be incongruent with existing institutional approaches to the governance, administration and operational support of modelling [i.e.
+"model management" @Arnold2020, p2].
+In hindsight, the case study preregistration process may have been more efficient and better specified if modellers and/or biostatisticians were brought earlier into the project’s planning and decision-making process to allow sufficient time to review, record and better connect analysis decisions to project aims.
+While we recognise that learning and implementing a new approach is unappealing at first, Adaptive Preregistration can save time later on, but requires a rearrangement of project timelines and resources.
+
+We suggest the following considerations when implementing Adaptive Preregistration for the first time:
+
+1.  Adaptive Preregistration is likely to become easier and more efficient as researchers gain experience and familiarity with preregistration.
+
+2.
+Start with simple descriptions or dot points in preregistration responses and build detail and complexity in analysis specification incrementally, as well as in thorough and ongoing collaborative review.
+
+3.  Building a database of decision-points in common modelling methods and approaches may aid in registering flexibility and evolve naturally over time through practice.
+
+4.
+Any attempt to implement Adaptive Preregistration is unlikely to work perfectly the first time, there will be mistakes and details that are omitted.
+Being upfront about this in study reporting is still better than avoiding the process all together.
+
+5.  For projects where using git & GitHub may be too difficult, separate sequential analysis plans could be preregistered for the same study with metadata linking the chain of preregistrations.
+
+## What's the right resolution for a modelling preregistration template?
+
+We aimed to design a preregistration template that was ‘parsimonious’, in that 1) template items should be specific and exhaustive enough to adequately constrain researcher degrees of freedom [@Wicherts2016] so as to; 2) facilitate transparent documentation of the modelling paths and decisions conducted during research, while; 3) not requiring a prohibitive level of detail.
+Increasing the completeness and resolution of the template in capturing decisions throughout the modelling process may improve the ability of preregistration to restrict researcher degrees of freedom and transparently report results, however it also makes the preregistration more challenging to complete.
+Refining the template during user-testing, greatly reduced the perceived difficulty of preregistration.
+However, our case-study researchers conceded that some modellers may still consider the final template to be too cumbersome to invest the time in preregistration.
+
+The writing and reporting phase of the case-study also highlighted the need for ongoing refinement of the template content and structure so as to facilitate the secondary use-case identified for our template, ‘preregistration checking’, or the process of verifying the reported analyses in a study were conducted as preregistered [@pu2019a].
+Information about the registered flexibility for determining alternative model functional forms was spread over multiple sub-sections within the case-study's preregistration template, hampering our ability to conduct preregistration checking post-analysis.
+We simplified the template during testing to ameliorate this issue, but it might be further improved by building modularity into the template for multiple or complex models.
+
+## Future Work
+
+While we believe we have partly addressed the problem of preregistering model-based research with the Adaptive Preregistration protocol, further work remains, particularly around developing cyberinfrastructure to effectively and transparently handle registering and reporting flexible analyses.
+We encourage modellers to preregister their research using our template and Adaptive Preregistration protocol flexibly, adopting elements that work and experimenting with alternative implementations that work better.
+
+New preregistration templates for expanding application of preregistration specific modelling approaches and methods could be developed by integrating our template with existing reporting checklists and guidelines, such as ODMAP for reporting Species Distribution Models [@Fitzpatrick2021].
+Currently, improvements to the template content, structure and visual presentation are limited by the infrastructure at existing registries.
+The registry platform requires a dedicated instrument or protocol for capturing flexible analyses, as well as reporting their outcomes and linking them to future decisions in subsequent interim preregistrations, which cannot be accommodated by existing registries, such as the [Open Science Foundation (OSF)](https://help.osf.io/article/330-welcome-to-registrations), [OSF registries](https://www.cos.io/products/osf-registries), [As Predicted](https://aspredicted.org/), and [EcoEvoArXiv](https://ecoevorxiv.org/).
+
+# Conclusions
+
+Although some challenges remain outstanding in applying preregistration to ecological modelling, it is important to remember the costs and risks of not preregistering modelling studies.
+Ecological modelling has long been plagued by poor transparency [@deVos:2011uk; @Jakeman:2006ii; @Schmolke:2010fd].
+Incomplete reporting of the modelling process risks models being used in inappropriate applications or decision contexts [@Grimm:2014es] and may mask poor model design or serious flaws in the model, resulting in adverse or irreversible outcomes when informing decision-making [@Addison:2013wa].
+Moreover, a lack of transparency in describing ecological models and reporting the modelling process provides significant opportunity for undisclosed researcher degrees of freedom, and hence, the possibility of Questionable Research Practices, with researchers routinely engaging in executing alternatively analyses, and selectively reporting them when they do [@Fraser:2018cl; @Liu2020; @kimmel2023].
+Reproducible, transparent and reliable models are essential for sound conservation decision-making [@Garcia-Diaz2019; @bodner2021] and Adaptive Preregistration can improve documentation of modelling decisions, which is helpful for remembering and explaining the study's method at the time of report-writing, to aid others in replicating and understanding model-based research, and to communicate to funders and stakeholders.
+
+We provide a template and methodology for ‘Adaptive Preregistration’ that we hope will extend the benefits of preregistration to model-based research in a way that strengthens rather than inhibits that research.
+Although we focus on ecological applications in this paper, many elements of the template are relevant across fields, and the template is open source and can be readily adapted to new fields and purposes.
+
+# References

### 2025-03-03 _analysis/ms.qmd
rm superfluous bullet in abstract

@@
 We applied a user-centred design approach to translate the concept of preregistration into model-based research practice for ecology and conservation.
-  3.	We developed a generalised ‘modelling workflow’ to capture critical analysis decisions in the model development cycle, informing a preregistration template for model-based research.
-The iterative nature of model development is incompatible with existing, linear preregistration templates, creating a barrier to implementing preregistration for model-based research.
-  4.	To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose, test and evaluate a methodology for ‘adaptive preregistration’, using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
-  5.
+  3.	To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose, test and evaluate a methodology for ‘adaptive preregistration’, using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
+  4.
 This research provides a template and methodology for implementing adaptive preregistration of ecological models.
@@
 | Preregistration Version | Preregistration Process | Modelling and Analysis Process |
-|----|----|----|
+|------------------------|------------------------|------------------------|
 | **Version 1:** (v0.10 – v0.8.1) | Establish problem context, study background and aims, describe data collection and data cleaning process, articulate modelling objectives, and specify candidate models.

### 2025-03-04 _analysis/ms.qmd
Add table 1 tag and cross-reference

@@
 
-The application of our preregistration template and Adaptive Preregistration methodology to the environmental flows modelling case study occurred across three distinct stages (Table 1, Fig.
+The application of our preregistration template and Adaptive Preregistration methodology to the environmental flows modelling case study occurred across three distinct stages (@tbl-modelling-stages, Fig.
 3), with two interim preregistrations (Version 1 and Version 2) preceding the finalised preregistration (Version 3, Supp Mat E), each corresponding to distinct phases in both the Preregistration and Modelling & Analysis processes.
@@
 -   *Information dependency*: where one decision informs another.
-For example, outcomes from the exploratory pilot analysis (Interim Preregistration Version 2, Table 1) informed the model structures specified in Preregistration Version 3 (Table 1), as well as the decision heuristics for triggering their fitting and acceptance.
+For example, outcomes from the exploratory pilot analysis (Interim Preregistration Version 2, @tbl-modelling-stages) informed the model structures specified in Preregistration Version 3 (@tbl-modelling-stages), as well as the decision heuristics for triggering their fitting and acceptance.
 -   *Procedural dependency*: where downstream decisions would not exist if some alternative upstream decision was made instead.
@@
 3), one for each study response variable, however modellers were unsure if these candidate models would be supported by the dataset due to spatio-temporal patchiness and heterogeneous density in the data.
-Preregistration Version 2 described an exploratory pilot analysis that aimed to resolve critical uncertainties about how to operationalise the models, given data constraints, and how to specify a candidate set of models that could feasibly be fitted to the full dataset (Table 1).
-This involved partitioning the dataset into a ‘Pilot Analysis’ subset for making subsequent data-informed modelling decisions, which were then preregistered in a subsequent preregistration of the ‘Main Analysis’ (Version 3, Table 1 and Fig.
+Preregistration Version 2 described an exploratory pilot analysis that aimed to resolve critical uncertainties about how to operationalise the models, given data constraints, and how to specify a candidate set of models that could feasibly be fitted to the full dataset (@tbl-modelling-stages).
+This involved partitioning the dataset into a ‘Pilot Analysis’ subset for making subsequent data-informed modelling decisions, which were then preregistered in a subsequent preregistration of the ‘Main Analysis’ (Version 3, @tbl-modelling-stages and Fig.
 3).
@@
 
-: Table 1 - Key stages of the modelling and Adaptive Preregistration process implemented in the environmental flows modelling case study.
-Version numbers in parentheses are GitHub tags, marking snapshots of the repository for each Preregistration Version.
+: Key stages of the modelling and Adaptive Preregistration process implemented in the environmental flows modelling case study.
+Version numbers in parentheses are GitHub tags, marking snapshots of the repository for each Preregistration Version {#tbl-modelling-stages}
 
@@
 
-The preregistration, while detailed, did not adequately capture the modellers' intended and implemented model selection strategy, where after fitting the set of simpler models identified from the pilot study (preregistration version 2, Table 1), the modellers would iteratively trial different combinations of model terms to make the models closer to the more complex ‘maximal’ models identified in the initial preregistration (preregistration version 1, Table 1).
+The preregistration, while detailed, did not adequately capture the modellers' intended and implemented model selection strategy, where after fitting the set of simpler models identified from the pilot study (preregistration version 2, @tbl-modelling-stages), the modellers would iteratively trial different combinations of model terms to make the models closer to the more complex ‘maximal’ models identified in the initial preregistration (preregistration version 1, @tbl-modelling-stages).
 The preregistered specification did not explain which combinations of model terms would be trialled, and in what order they would be trialled, rendering components of the model selection process unregistered.

### 2025-03-04 _analysis/ms.qmd
rm email addresses for non-corresponding authors

@@
       family: Jones
-    email: chris.jones@delwp.vic.gov.au
     orcid: 0000-0003-2833-0194
@@
       family: Yen
-    email: jian.yen@delwp.vic.gov.au
     orcid: 0000-0001-7964-923X
@@
     email: hannahsfraser@gmail.com
-    orcid: 0000-0003-2443-4463
     attributes:
@@
       family: Wootton
-    email: henry.wootton@delwp.vic.gov.au
     orcid: 0000-0001-6506-0248
@@
       family: Vivian
-    email: lyndsey.vivian@delwp.vic.gov.au
     orcid: 0000-0003-1927-723X
@@
       family: Good
-    email: goodm@unimelb.edu.au
     orcid: 0000-0002-6908-1633
@@
       family: Duncan
-    email: david.duncan@unimelb.edu.au
     orcid: 0000-0003-4411-8214
@@
       family: Rumpff
-    email: lrumpff@unimelb.edu.au
     orcid: 0000-0001-9400-8086
@@
       family: Fidler
-    email: fidlerfm@unimelb.edu.au
     orcid: 0000-0002-2700-2562
@@
 Modellers need no longer cry “but I can’t preregister my research.”
-
 keywords:
@@
 format:
+  html:
+   toc: true
+   number-sections: true
+   highlight-style: github
   # preprint-typst:
@@
 | Preregistration Version | Preregistration Process | Modelling and Analysis Process |
-|------------------------|------------------------|------------------------|
+|----|----|----|
 | **Version 1:** (v0.10 – v0.8.1) | Establish problem context, study background and aims, describe data collection and data cleaning process, articulate modelling objectives, and specify candidate models.

### 2025-03-04 _analysis/ms.qmd
Add figure 3 case study workflow image file, insert into ms, add caption and update all cross-refs

@@
 
-The application of our preregistration template and Adaptive Preregistration methodology to the environmental flows modelling case study occurred across three distinct stages (@tbl-modelling-stages, Fig.
-3), with two interim preregistrations (Version 1 and Version 2) preceding the finalised preregistration (Version 3, Supp Mat E), each corresponding to distinct phases in both the Preregistration and Modelling & Analysis processes.
+The application of our preregistration template and Adaptive Preregistration methodology to the environmental flows modelling case study occurred across three distinct stages (@tbl-modelling-stages, @fig-case-study-workflow), with two interim preregistrations (Version 1 and Version 2) preceding the finalised preregistration (Version 3, Supp Mat E), each corresponding to distinct phases in both the Preregistration and Modelling & Analysis processes.
 While preparing Version 1, the modellers decided that they needed to conduct some analysis to specify a set of appropriate candidate models.
@@
 
-Two initial candidate models were preregistered in Version 1 (Fig.
-3), one for each study response variable, however modellers were unsure if these candidate models would be supported by the dataset due to spatio-temporal patchiness and heterogeneous density in the data.
+Two initial candidate models were preregistered in Version 1 (@fig-case-study-workflow), one for each study response variable, however modellers were unsure if these candidate models would be supported by the dataset due to spatio-temporal patchiness and heterogeneous density in the data.
 Preregistration Version 2 described an exploratory pilot analysis that aimed to resolve critical uncertainties about how to operationalise the models, given data constraints, and how to specify a candidate set of models that could feasibly be fitted to the full dataset (@tbl-modelling-stages).
-This involved partitioning the dataset into a ‘Pilot Analysis’ subset for making subsequent data-informed modelling decisions, which were then preregistered in a subsequent preregistration of the ‘Main Analysis’ (Version 3, @tbl-modelling-stages and Fig.
-3).
+This involved partitioning the dataset into a ‘Pilot Analysis’ subset for making subsequent data-informed modelling decisions, which were then preregistered in a subsequent preregistration of the ‘Main Analysis’ (Version 3, @tbl-modelling-stages and @fig-case-study-workflow).
 The pilot analysis dataset consisted of observations from a single site with good temporal data coverage, while the main analysis dataset comprised the full dataset, including the pilot dataset.
@@
 The same procedures for evaluating model fit and model checking were preregistered for both the pilot and main analysis (Fig.
-3B), however the purpose of model checking in the Pilot Analysis was to identify appropriate model structures, whereas the purpose for the Main Analysis was to ensure that the fitted models were supported by the data.
+4B), however the purpose of model checking in the Pilot Analysis was to identify appropriate model structures, whereas the purpose for the Main Analysis was to ensure that the fitted models were supported by the data.
 
@@
 1.
-A preferred candidate model or ‘full model’ for each response variable (vegetation richness and vegetation cover), to be fitted to the full dataset (Initial Preregistration, Fig.
-3, Supp.
+A preferred candidate model or ‘full model’ for each response variable (vegetation richness and vegetation cover), to be fitted to the full dataset (Initial Preregistration, @fig-case-study-workflow, Supp.
 Mat.
@@
 2.
-A candidate set of simplified models that separately captured different flow components for each response variable, with two versions of the flow events model for vegetation cover (Flow Events Models A & B, Fig.
-3, Supp.
+A candidate set of simplified models that separately captured different flow components for each response variable, with two versions of the flow events model for vegetation cover (Flow Events Models A & B, @fig-case-study-workflow, Supp.
 Mat.
@@
 3.
-Decision-trees for triggering fitting and acceptance of the candidate simplified models should the preferred full model specifications fail to converge and/or provide appropriate parameter estimates (Fig.
-3).
+Decision-trees for triggering fitting and acceptance of the candidate simplified models should the preferred full model specifications fail to converge and/or provide appropriate parameter estimates (@fig-case-study-workflow).
 4.
-A second flexible strategy, for determining model functional form and to account for potential over-dispersion (Fig.
-3).
+A second flexible strategy, for determining model functional form and to account for potential over-dispersion (@fig-case-study-workflow).
 Preregistering this flexibility was motivated by the modellers’ concerns that the actual data distributions in the full dataset did not match expectations informed by the pilot analysis.
@@
 
-<!--# Insert Figure 3 - case study workflow -->
-
-Figure 3 illustrates the evolution of model development across preregistration iterations in the case study.
+![illustrates the evolution of model development across preregistration iterations in the case study.
 The two Registered Flexibility steps in the Main Analysis (v3) (shown in blue) are further described in Fig.
-4.
+4.](images/Figure_3-Case_Study_Workflow.png){#fig-case-study-workflow width="14cm" height="24cm"}
 

### 2025-03-04 _analysis/ms.qmd
Add fig 4 case study decision trees image and insert, add caption, tag, update cross-refs.

@@
 -   *Procedural dependency*: where downstream decisions would not exist if some alternative upstream decision was made instead.
-For example, if preliminary analysis of the data had not revealed over-dispersion or zero-inflation in the data, then the case study modellers would not have fitted zero-inflated Poisson models, nor would they have trialled accounting for over-dispersion or refitting using negative binomial models (Fig.
-4).
+For example, if preliminary analysis of the data had not revealed over-dispersion or zero-inflation in the data, then the case study modellers would not have fitted zero-inflated Poisson models, nor would they have trialled accounting for over-dispersion or refitting using negative binomial models (fig-case-study-registered-flexibility).
 
@@
 For example, simple tests of data distributions within hierarchies were required to check for data spread and the presence of zero-inflation, while checks for collinearity among candidate predictor variables were needed to prevent confounding and overfitting, and to ensure relevant interactions were captured in the models.
-The same procedures for evaluating model fit and model checking were preregistered for both the pilot and main analysis (Fig.
-4B), however the purpose of model checking in the Pilot Analysis was to identify appropriate model structures, whereas the purpose for the Main Analysis was to ensure that the fitted models were supported by the data.
+The same procedures for evaluating model fit and model checking were preregistered for both the pilot and main analysis (fig-case-study-registered-flexibilityB), however the purpose of model checking in the Pilot Analysis was to identify appropriate model structures, whereas the purpose for the Main Analysis was to ensure that the fitted models were supported by the data.
 
@@
 ![illustrates the evolution of model development across preregistration iterations in the case study.
-The two Registered Flexibility steps in the Main Analysis (v3) (shown in blue) are further described in Fig.
-4.](images/Figure_3-Case_Study_Workflow.png){#fig-case-study-workflow width="14cm" height="24cm"}
+The two Registered Flexibility steps in the Main Analysis (v3) (shown in blue) are further described in @fig-case-study-registered-flexibility .](images/Figure_3-Case_Study_Workflow.png){#fig-case-study-workflow width="14cm" height="24cm"}
 
-<!--# Insert Figure 4 -->
-
-Figure 4 - illustrates how registered flexibility was applied within the case study preregistration, for selecting model structure (A) and choosing an alternative model functional form to the default (zero-inflated Poisson) (B).
+::: landscape
+![illustrates how registered flexibility was applied within the case study preregistration, for selecting model structure (A) and choosing an alternative model functional form to the default (zero-inflated Poisson) (B).](images/Figure_4-case_study_registered_flexibility%20-%20preregistered%20decision%20trees.png){#fig-case-study-registered-flexibility}
+:::
 
